### 2020-09-02 src/uk.Rmd
all but .RData

---
title: "UK"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs")
  })
---

```{r}
library(rmarkdown)
library(data.table)
library(tidyverse)
library(readxl)
library(httr)
library(rprojroot)

library(ggplot2)
#library(geojsonio)
library(rgdal)
library(maptools)
library(sp)
library(sf)
library(rgeos)
#library(classInt)
#library(RColorBrewer)
library(knitr)
#library(viridis)
#library(hrbrthemes)
library(fs)
library(betareg)
library(broom)

library(AER)
library(MASS)
#library(foreign)
library(stargazer)
library(ivpack) 
library(leaflet)
library(rvest)
library(xml2)
library(tmaptools)
library(spdep)
library(spatialreg)
library(spatialEco)

# This is the project path
path <- find_rstudio_root_file()
```

## Merge bua with buasd

```{r}
# Get BUA 
# Some BUA also include sub-division, so the two spatial objects need to be downloaded selerately and merged:

# (i) directly from the web
# buasd <- readOGR("https://opendata.arcgis.com/datasets/bc2dbefc1285410e9fdf800745d1280f_0.geojson") #, layer="OGRGeoJSON")
# bua <- readOGR("https://opendata.arcgis.com/datasets/210f705366044b8c9ca3fa6f4f83fa5d_0.geojson")

# or, (ii) from the geojson saved locally
path.json <- paste(path, "/data/uk/raw/Built-up_Area_Sub_Divisions_(December_2011)_Boundaries.geojson", sep = "")
buasd <- readOGR(path.json)#, layer="OGRGeoJSON")
path.json <- paste(path, "/data/UK/raw/Built-up_Areas_(December_2011)_Boundaries_V2.geojson", sep = "")
bua <- readOGR(path.json)#, layer="OGRGeoJSON")
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
buasd <- spTransform(buasd, CRS("+init=epsg:4326"))
bua <- spTransform(bua, CRS("+init=epsg:4326"))

dim(bua) # 5830
dim(buasd) #1826

# drop the BUA with sub-division to only include the sub-divisions (see rbind() below)
bua <- bua[bua$has_sd=="N",] #5262

names(bua)
names(buasd)

bua$has_sd <- NULL
bua$urban_bua <- NULL
bua$sd_count <- NULL
buasd$buasd_id <- NULL
names(buasd)[3] <- "bua11cd"
names(buasd)[4] <- "bua11nm"
dim(bua)
dim(buasd)
bua$sd_count <- NULL
buasd$sd_type <- NULL

# merge bua with buasd
bua_buasd <- rbind(bua, buasd, makeUniqueIDs = TRUE)
dim(bua_buasd@data)

# factors to characters
bua_buasd@data$bua11cd <- as.character(bua_buasd@data$bua11cd)
bua_buasd@data$bua11nm <- as.character(bua_buasd@data$bua11nm)
```

## Broadband speed from Speedchecker Ltd., 2017

```{r}
# load broadband data
path.data <- paste(path, "/data/uk/raw/broadband.speed/2015-2017/broadbandspeed.csv", sep = "")
bb2017 <- read_csv(path.data, col_names = T)
range(bb2017$DateTimeStamp)

bb2017 <- bb2017 %>% 
  filter(DateTimeStamp > "2017-01-01") %>% 
  filter(DownloadSpeed > 512 & DownloadSpeed< 362000) %>% # Riddlesden and Singleton 2014 + my interpretation
  filter(UploadSpeed > 100 & UploadSpeed < 21000) %>% # Riddlesden and Singleton 2014 + my interpretation
  drop_na()

# convert points to spatial object
coords_bb <- cbind(bb2017$Longitude, bb2017$Latitude)
bb.bua <- SpatialPointsDataFrame(coords_bb, data = data.frame(bb2017))
proj4string(bb.bua) <- CRS("+init=epsg:4326") #define projection

# spatial join to bua_buasd
bb.bua.sp <- over(bb.bua, bua_buasd[, "bua11cd"]) # not a spatial object
bb.bua$bua11cd <- bb.bua.sp$bua11cd

# create dataframe object to analyse (remove lat and lon?)
bb2017 <- bb.bua@data

# drop NAs
sapply(bb2017, function(x) sum(is.na(x)))
bb2017 <- bb2017[!is.na(bb2017$bua11cd),]
#NA bua11cd, coord outside bua, buasd or the UK = 476381 / 1953654 obs (76% of obs retained)

bb2017 <- bb2017 %>%
  group_by(bua11cd) %>% # this also include buasd
  summarize(download = median(DownloadSpeed), upload = median(UploadSpeed), n.tests = n()) 
```

## Broadband speed from Speedchecker Ltd., 2011

```{r}
# load broadband data
path.data <- paste(path, "/data/uk/raw/broadband.speed/2011", sep = "")
files <- dir_ls(path=path.data, pattern="*.csv") # dir_ls() also inludes the paths so it works for .Rmd

bb2011 <- lapply(files, function(x) read_delim(x, delim=",", col_names = F)) %>% 
  bind_rows() #
names(bb2011)[1:6] <- c("DateTimeStamp","DownloadSpeed","UploadSpeed","provider","Latitude", "Longitude")

bb2011 <- bb2011 %>% 
  filter(DownloadSpeed > 512 & DownloadSpeed< 362000) %>% # Riddlesden and Singleton 2014 + my interpretation
  filter(UploadSpeed > 100 & UploadSpeed < 21000) %>% # Riddlesden and Singleton 2014 + my interpretation
  drop_na()

# convert points to spatial object
coords_bb <- cbind(bb2011$Longitude, bb2011$Latitude)
bb.bua <- SpatialPointsDataFrame(coords_bb, data = data.frame(bb2011))
proj4string(bb.bua) <- CRS("+init=epsg:4326") #define projection

# spatial join to bua_buasd
bb.bua.sp <- over(bb.bua, bua_buasd[, "bua11cd"]) # not a spatial object
bb.bua$bua11cd <- bb.bua.sp$bua11cd

# create dataframe object to analyse (remove lat and lon?)
bb2011 <- bb.bua@data

# drop NAs
sapply(bb2011, function(x) sum(is.na(x)))
bb2011 <- bb2011[!is.na(bb2011$bua11cd),]
#NA bua11cd, coord outside bua, buasd or the UK = 269310 / 1307293 obs (80% of obs retained)

bb2011 <- bb2011 %>%
  group_by(bua11cd) %>% # this also include buasd
  summarize(download = median(DownloadSpeed), upload = median(UploadSpeed), n.tests = n()) 
```

## Population, population rank and differences; LHS variable

```{r}
pop <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_2010_1.data.csv?geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&gender=0&c_age=200&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xa7a1e65a29f4c48e3059704c1fe86a40e2c0b670")
# saved also as a wide .csv in /ict_un_us_uk/data/uk/raw/bua_pop.csv

pop <- pop %>% 
  dplyr::select(DATE, GEOGRAPHY_CODE, GEOGRAPHY_NAME, OBS_VALUE) %>%
  spread(DATE, OBS_VALUE) %>% 
  mutate(r2011 = dense_rank(desc(`2011`))) %>% 
  mutate(r2018 = dense_rank(desc(`2018`))) %>%
  mutate(r_diff = r2011 - r2018) %>%
  mutate(r_diff_tr = (r_diff + sum(!is.na(r_diff)))/(2*sum(!is.na(r_diff)))) %>%
  rename(pop2011 = `2011`) %>%
  rename(pop2012 = `2012`) %>%
  rename(pop2013 = `2013`) %>%
  rename(pop2014 = `2014`) %>%
  rename(pop2015 = `2015`) %>%
  rename(pop2016 = `2016`) %>%
  rename(pop2017 = `2017`) %>%
  rename(pop2018 = `2018`)
# for normalisation N = sum(!is.na(r_diff))
# Check later if this correct for the regression data

# Scotland is excluded as the methodology for Scottish Settlements and Localities is different:
# https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/population-estimates/special-area-population-estimates/settlements-and-localities/background-information
# pop %>% filter(str_detect(GEOGRAPHY_CODE, "S")) 
```

## IMD NOT USED

```{r}
# #imd.en.2019 <- read_csv(unzip("https://data.cdrc.ac.uk/system/files/English%20IMD%202019.zip" )
# temp <- tempfile()
# download.file("https://data.cdrc.ac.uk/system/files/English%20IMD%202019.zip",temp)
# imd.en.2019 <- read.table(unz(temp, "English"))
# unlink(temp)

# England 2010
#england.imd2010 <- read_csv(url("https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/15240/1871702.csv"))
# also saved as imd_england2010.csv

# England 2019
#england.imd2019 <- read_csv("https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/845345/File_7_-_All_IoD2019_Scores__Ranks__Deciles_and_Population_Denominators_3.csv")

# Wales 2011
#url <- "https://gov.wales/sites/default/files/statistics-and-research/2019-04/wimd-2011-individual-domain-scores-and-overall-index-scores-for-each-lsoa.xls"
#GET(url, write_disk(tf <- tempfile(fileext = ".xls")))
#wales.imd2011 <- read_excel(tf, 2L)
# also saved as wimd-2011-individual-domain-scores-and-overall-index-scores-for-each-lsoa.xls

# Wales 2019
#url <- "https://gov.wales/sites/default/files/statistics-and-research/2019-12/wimd-2019-index-and-domain-scores-by-small-area_0.ods"
#GET(url, write_disk(tf <- tempfile(fileext = ".xls")))
#wales.imd2019 <- read_ods(tf, 2L, skip = 2)
# also saved as wimd-2019-index-and-domain-scores-by-small-area_0.ods

```

## Jobseeker's Allowance with rates and proportions NOT USED

```{r}
# jobseeker <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_1_1.data.csv?geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&date=latestMINUS108,latestMINUS96,latestMINUS84,latestMINUS72,latestMINUS60,latestMINUS48,latestMINUS36,latestMINUS24,latestMINUS12&sex=7&item=1&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xbd6481e37da64ded2c7523fd7226460d3db46f0f")
# # saved as wide .csv in \ict_un_us_uk\data\uk\raw\job_seeker_allowance.csv
# 
# jobseeker <- read_json("https://www.nomisweb.co.uk/api/v01/dataset/NM_1_1.jsonstat.json?geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&date=latestMINUS108,latestMINUS96,latestMINUS84,latestMINUS72,latestMINUS60,latestMINUS48,latestMINUS36,latestMINUS24,latestMINUS12&sex=7&item=1&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xea3fd585c7b9709c1f247b9a304c1355a9b5e13a")
# 
# jobseeker %>% spread_all
# 
# jobseeker_ <- jobseeker %>% 
#   dplyr::select(DATE, GEOGRAPHY_CODE, GEOGRAPHY_NAME, OBS_VALUE) %>%
#   spread(DATE, OBS_VALUE) #%>% 
#   mutate(r2011 = dense_rank(desc(`2011`)))

```

## Population density

```{r}
pop.dens <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_143_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&rural_urban=0&cell=0,1&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xbd7005144df55f1693a38d07f3e255171c673600")
# saved also as pop_dens.csv

pop.dens <- pop.dens %>% 
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, CELL_NAME, OBS_VALUE) %>%
  spread(CELL_NAME, OBS_VALUE) %>% 
  mutate(pop.dens = `All usual residents` / `Area Hectares`) %>%
  dplyr::select(GEOGRAPHY_CODE, pop.dens)
```

## urban / rural classification

```{r}
# England and Wales 2011 urban/rural classification for OA
# http://geoportal.statistics.gov.uk/datasets/rural-urban-classification-2011-of-output-areas-in-england-and-wales
path.data <- paste(path, "/data/uk/raw/RUC11_OA11_EW.zip", sep = "")
ur.ew <- read_csv(unzip(path.data, "RUC11_OA11_EW.csv"))

# NOT USED
# Scotland
# The Scottish Government's Urban Rural Classification 2013/14, 2011 Output Area
# https://www.isdscotland.org/Products-and-Services/GPD-Support/Geography/Urban-Rural-Classification/
# classification info: https://www2.gov.scot/Topics/Statistics/About/Methodology/UrbanRuralClassification
# ur.sc <- read_csv("https://www.isdscotland.org/Products-and-Services/GPD-Support/Geography/Urban-Rural-Classification/_docs/2013_2014_urban_rural/oa2011_urban_rural_2013_2014.csv")

# lookup file for OA-to-BUA/BUASD
lookup.out.bua <- read_csv("https://opendata.arcgis.com/datasets/edaf7c4b5e6e401d987c30c4de6b63e6_0.csv")
# saved as Output_Area__2011__to_Built-up_Area_Sub-division_to_Built-up_Area_to_Local_Authority_District_to_Region__December_2011__Lookup_in_England_and_Wales.csv
dim(lookup.out.bua)
lookup.out.bua <- lookup.out.bua %>% dplyr::select(OA11CD, BUASD11CD, BUASD11NM, BUA11CD, BUA11NM)

ur.ew <- merge(ur.ew, lookup.out.bua, by = "OA11CD", all.x = T)
sapply(ur.ew, function(x) sum(is.na(x)))

# create new id which contains BUA or BUASD id
ur.ew$BUA_SD_CD <- ifelse(is.na(ur.ew$BUASD11CD), ur.ew$BUA11CD, ur.ew$BUASD11CD) #replace BUA sub divisions
ur.ew$BUA_SD_NM <- ifelse(is.na(ur.ew$BUASD11NM), ur.ew$BUA11NM, ur.ew$BUASD11NM) #replace BUA sub divisions
# with BUA for those BUA which don't have sub divisions
# sum(ur.ew$BUASD11NM==ur.ew$BUA11NM) #check

length(unique(ur.ew$BUA_SD_CD)) #6621, 6622 now 
length(unique(ur.ew$BUA_SD_NM)) #6621, 6622 now

ur.ew <- ur.ew %>% 
  filter(!is.na(BUA11CD) | !is.na(BUASD11CD) | !is.na(BUA_SD_CD)) %>%
  mutate(urban = gsub("([A-Za-z]+).*", "\\1", RUC11)) %>% # take the first word from split string
  dplyr::select(BUA_SD_CD, urban) %>%
  distinct() # this leads to 6621 unique obs., so every BUA_SD_CD has a unique value for urban
```

## Ethnicity in 2011

```{r}
british.pop <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_574_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&rural_urban=0&cell=0,1&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xadba1f3dfe1e84341e0d33f6d082210022a1418e")
# saved as british_pop.csv

british.pop <- british.pop %>% 
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, CELL_NAME, OBS_VALUE) %>%
  spread(CELL_NAME, OBS_VALUE) %>% 
  rename(all.pop = `All categories: Ethnic group`) %>%
  rename(british.pop = `English/Welsh/Scottish/Northern Irish/British`) %>%
  mutate(british.pop.share = british.pop / all.pop) %>%
  dplyr::select(GEOGRAPHY_CODE, british.pop.share)
```

## Working from home 2011

```{r}
work.home <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_153_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&rural_urban=0&cell=0,9&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xfa76d191fd681a934f45a8a0f35bad74f3fe0877")
# save as work_home.csv

work.home <- work.home %>% 
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, CELL_NAME, OBS_VALUE) %>%
  spread(CELL_NAME, OBS_VALUE) %>% 
  rename(working.pop = `All categories: Distance travelled to work`) %>%
  rename(work.home = `Work mainly at or from home`) %>%
  mutate(work.home.share = work.home / working.pop) %>%
  dplyr::select(GEOGRAPHY_CODE, work.home.share)
```

## Services 2011

```{r}
services <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_1020_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&c_indgpuk11=0...8&c_sex=0&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xe16647b1bfd6ad3e2bbea89fa2de3e97bbbab3a9")
# saved as industries.csv

services <- services %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, C_INDGPUK11_NAME, OBS_VALUE) %>% # C_INDGPUK11: this is the industry code
  spread(C_INDGPUK11_NAME, OBS_VALUE) %>% 
  mutate(services.share = (`G, I Distribution, hotels and restaurants` +
                             `H, J Transport and communication` +
                             `K, L, M, N Financial, Real Estate, Professional and Administrative activities` +
                             `O, P, Q Public administration, education and health` +
                             `R, S, T, U Other`) / `All categories: Industry`)  %>%
  dplyr::select(GEOGRAPHY_CODE, services.share)
```

## Unemployment 2011

```{r}
unemployment <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_1046_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&c_sex=0&c_age=0&c_ecopuk11=0,1,9&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0x9e220902be948949eb24ef622423687fba63356a")

unemployment <- unemployment %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, C_ECOPUK11_NAME, OBS_VALUE) %>% 
  spread(C_ECOPUK11_NAME, OBS_VALUE) %>% 
  mutate(unemployment.share = (`Economically active: Unemployed (including full-time students)` / 
                                 `Economically active: Total`)) %>%
  dplyr::select(GEOGRAPHY_CODE, unemployment.share)
```

## Universities

```{r}
# source: https://www.hesa.ac.uk/support/providers, keep HEI

url <- "https://www.hesa.ac.uk/support/providers"
webpage <- xml2::read_html(url)

uni.table <- rvest::html_table(webpage)[[1]] %>% 
  tibble::as_tibble(.name_repair = "unique") %>%               # repair the repeated columns
  filter(Region!= "Scotland" & Region!="Northern Ireland") %>% # drop Scotland and N. Ireland
  filter(`Provider type` == "HEI")

latlong <-  geocode_OSM(uni.table$Name) # 131 results

# wrong results based on the leaflet map below
# these universities are added in the uni.table.nomatch
wrong.location <- c("University of Plymouth", 
                    "The University of Southampton",
                    "The Open University",
                    "The University of Portsmouth",
                    "Goldsmiths College")

`%nin%` = Negate(`%in%`)

latlong <- latlong %>%
  filter(query %nin% wrong.location)

uni.table.nomatch <- tibble(
  Name = c("Conservatoire for Dance and Drama",
           "Hartpury University",
           "Imperial College of Science, Technology and Medicine",
           "Rose Bruford College of Theatre and Performance",
           "Royal Holloway and Bedford New College",
           "St George's, University of London",
           "The University of Huddersfield",
           "The University of Wales (central functions)",
           "University of London (Institutes and activities)",
           "University of Northumbria at Newcastle",
           "University of St Mark and St John",
           "University of the Arts, London",
           "University of Plymouth", 
           "The University of Southampton",
           "The Open University",
           "The University of Portsmouth",
           "Goldsmiths College"),
  postcode = c("WC1H 9JJ",
               "GL19 3BE",
               "SW7 2BU",
               "DA15 9DF",
               "TW20 0EX",
               "SW17 0RE",
               "HD1 3DH",
               "CF10 3NS",
               "WC1E 7HU",
               "NE1 8QH",
               "PL6 8BH",
               "WC1V 7EY",
               "PL4 8AA",
               "SO17 1BJ",
               "MK7 6BJ",
               "PO1 2UP",
               "SE14 6NW"))

latlong.nomatch <-  geocode_OSM(uni.table.nomatch$postcode) # 16 results
latlong.nomatch$query <- uni.table.nomatch$Name

uni.table <- rbind(latlong, latlong.nomatch)
uni.table <- uni.table %>%
  dplyr::select(query, lat, lon) %>%
  rename(university = query)

# convert uni points to spatial object
coords_uni <- cbind(uni.table$lon, uni.table$lat)
uni.bua <- SpatialPointsDataFrame(coords_uni, data = data.frame(uni.table))
proj4string(uni.bua) <- CRS("+init=epsg:4326") #define projection

# spatial join to bua_buasd
uni.sp <- over(uni.bua, bua_buasd[, "bua11cd"]) # not a spatial object
uni.table$bua11cd <- uni.sp$bua11cd

# Uni map. NOT TO RUN
# leaflet(bua_buasd) %>% 
#   addPolygons() %>% 
#   addCircles(lng=uni.table$lon, lat=uni.table$lat, popup=uni.table$university) %>%
#   addMarkers(lng=uni.table$lon, lat=uni.table$lat, popup=uni.table$university)

uni.freq <- uni.table %>%
  group_by(bua11cd) %>%
  summarise(uni.freq = n())

uni.freq <- merge(bua_buasd@data, uni.freq, by = "bua11cd", all.x = T)
uni.freq$uni.freq <- ifelse(is.na(uni.freq$uni.freq), 0, uni.freq$uni.freq)

uni.freq <- uni.freq %>%
  dplyr::select(bua11cd, uni.freq) %>%
  rename(GEOGRAPHY_CODE = bua11cd)

path.uni <- paste(path, "/data/uk/raw/uni_freq.csv", sep = "")
write_csv(uni.freq, path.uni)
```

## Merge objects

```{r}
data <- list(bb2017, bb2011, pop, pop.dens, ur.ew, british.pop, work.home,
services, unemployment, uni.freq)
sapply(data, function(x) dim(x))
sapply(data, function(x) names(x))

# rename bua11cd to GEOGRAPHY_CODE
names(data[[1]])[1] <- "GEOGRAPHY_CODE"
names(data[[1]])[2] <- "download2017"
names(data[[1]])[3] <- "upload2017"
names(data[[1]])[4] <- "n.tests2017"

names(data[[2]])[1] <- "GEOGRAPHY_CODE"
names(data[[2]])[2] <- "download2011"
names(data[[2]])[3] <- "upload2011"
names(data[[2]])[4] <- "n.tests2011"

names(data[[5]])[1] <- "GEOGRAPHY_CODE"

# merge with reduce
data <- data %>% 
  reduce(inner_join, by = "GEOGRAPHY_CODE")
sapply(data, function(x) sum(is.na(x)))

# NOT RUN
# export file for future reference / backup
# data.out.path <- paste0(path, "/data/data_inter/data_for_uk_regressions.csv")
# write.csv(data, data.out.path)
```

## descriptives

```{r}
ggplot(data = data, aes(x=download2011, y=r_diff_tr)) + geom_point() + geom_smooth(method = "lm")
ggplot(data = data, aes(x=download2017, y=r_diff_tr)) + geom_point() + geom_smooth(method = "lm")

# ggplot(data[(data$n.tests2017>50 & data$download2017<60000),], aes(x=download2017, y=r_diff_tr)) + geom_point() + geom_smooth(method = "lm")
# ggplot(data[(data$n.tests2011>50),], aes(x=download2011, y=r_diff_tr)) + geom_point() + geom_smooth(method = "lm")

f.n.tests2011 <- data %>%
  dplyr::select(n.tests2011) %>%
  group_by(ints = cut_width(n.tests2011, width = 20, boundary = 0)) %>%
  summarise(n = n())
head(f.n.tests2011)
```

## Regressions

F/or all the regressions, we include the aerial units with more than 20 tests in 2011.
This leads to a decrease in the number of observations include in the regressions from 6096 to 6096 - 2561 = 3535

```{r}

base <- lm(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
#nobs(base)

# for robust with library(jtools)
# summ(base, robust = "HC1")

# upload is not a significant predictor

base_glm <- glm(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share +
                          british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
#nobs(base_glm)

#base_betareg <- betareg(r_diff_tr ~ log(download2011) + log(pop2011) + log(pop2018) + I(n.tests2011/pop2011) + unemployment.share +
#                          british.pop.share +  pop.dens + work.home.share + services.share, data = data, 
#                          link = "logit", subset = n.tests2011>20)
#nobs(base_betareg)

# not converging, try quasipoisson
#base_poisson <- glm(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share +
#                          british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20, 
#                          family = "poisson")
#nobs(base_poisson)

# interactions
int_density <- lm(r_diff_tr ~ log(download2011)*pop.dens + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
#nobs(int_density)

int_pop <- lm(r_diff_tr ~ log(download2011)*log(pop2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
#nobs(int_pop)

# 2 SLS with IV
model_iv1 <- ivreg(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share 
                   | log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share + uni.freq
                     , data = data, subset = n.tests2011>30)
#summary(model_iv1, vcov = sandwich, diagnostics = TRUE)
#nobs(base)

model_iv2 <- ivreg(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share 
                   | log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share + uni.freq + n.tests2011
                     , data = data, subset = n.tests2011>30)
#summary(model_iv2, vcov = sandwich, diagnostics = TRUE) #yes
```

## Table of Descriptives and Regression tables

```{r}
# Table 9
data.for.decr <- data %>%
  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011,  
         unemployment.share, british.pop.share, pop.dens, 
         work.home.share, services.share, uni.freq) %>%
  mutate(n.tests2011perhab = n.tests2011/pop2011) %>%
  dplyr::select(-n.tests2011) %>%
  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011perhab,  # re-redring
         unemployment.share, british.pop.share, pop.dens, 
         work.home.share, services.share, uni.freq)

path.out <- paste(path, "/outputs/table9.htm", sep = "")

stargazer(as.data.frame(data.for.decr), out=path.out, type="html", summary = T,
          covariate.labels=c("normalized difference in ranks, 2011-14", "download speed, 2011", "population, 2011",  
                             "broadband tests per capita, 2011", "% of unemployment, 2011", "% of British population, 2011", 
                             "population density, 2011", "% of people working from home, 2011",
                             "employment in service, 2011 (%)", "Number of universities"))

#Table 10
#regressions table robust SE
rob.base        <- coeftest(base, function(x) vcovHC(x, type="HC0"))
rob.int_density <- coeftest(int_density, function(x) vcovHC(x, type="HC0"))
rob.int_pop <- coeftest(int_pop, function(x) vcovHC(x, type="HC0"))
rob.base_glm <- coeftest(base_glm, function(x) vcovHC(x, type="HC0"))

path.out <- paste(path, "/outputs/table10.htm", sep = "")

stargazer(base, int_density, int_pop, base_glm, type="html",
          se = list(rob.base[,"Std. Error"], rob.int_density[,"Std. Error"], rob.int_pop[,"Std. Error"], rob.base_glm[, "Std. Error"]), 
          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
          #dep.var.caption = "Normalized difference in ranks, 2011-18",
          column.labels = c("OLS", "GLM"),
          column.separate = c(3,1),
          covariate.labels=c("download speed, 2011 (log)", "population, 2011 (log)",  
                             "broadband tests per capita, 2011", "% of unemployment, 2011", "% of British population, 2011", 
                             "population density, 2011", "% of people working from home, 2011",
                             "employment in service, 2011 (%)", "download speed, 2011 (log) x pop. density, 2011", 
                             "download speed, 2011 (log) x population, 2011 (log)"), 
          out=path.out, single.row = FALSE, df = FALSE, omit.stat = c("rsq", "f"), 
          notes = "Robust Std. Error parenthesis")

# Table 11
# 2SLS robust SE table
#model_iv1, model_iv2, 
rob.model_iv1 <- coeftest(model_iv1, function(x) vcovHC(x, type="HC0"))
rob.model_iv2 <- coeftest(model_iv2, function(x) vcovHC(x, type="HC0"))
summ.fit1 <- summary(model_iv1, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
summ.fit2 <- summary(model_iv2, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

path.out <- paste(path, "/outputs/table11.htm", sep = "")

stargazer(model_iv1, model_iv2, type = "html", 
          se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]), 
          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
          covariate.labels=c("download speed, 2011 (log)", "population, 2011 (log)",  
                             "broadband tests per capita, 2011", "% of unemployment, 2011", "% of British population, 2011", 
                             "population density, 2011", "% of people working from home, 2011",
                             "employment in service, 2011 (%)"),
          out=path.out, single.row = FALSE, df = FALSE, 
          omit.stat = c("rsq", "f"), notes = c("Robust Std. Error in parenthesis", "IVs: (1) Number of universities",
                                               "IVs: (2) Number of universities, Number of broadband tests, 2011"),
          add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.fit1$diagnostics[1, "statistic"], 2), 
                             round(summ.fit2$diagnostics[1, "statistic"], 2)), 
                           c(rownames(summ.fit1$diagnostics)[2], 
                             round(summ.fit1$diagnostics[2, "statistic"], 2), 
                             round(summ.fit2$diagnostics[2, "statistic"], 2)), 
                           c("P-value", 
                             round(summ.fit1$diagnostics[2, "p-value"], 2), 
                             round(summ.fit2$diagnostics[2, "p-value"], 2)), 
                           c(rownames(summ.fit1$diagnostics)[3], 
                             round(summ.fit1$diagnostics[3, "statistic"], 2), 
                             round(summ.fit2$diagnostics[3, "statistic"], 2)),
                           c("P-value", 
                             round(summ.fit1$diagnostics[3, "p-value"], 2), 
                             round(summ.fit2$diagnostics[3, "p-value"], 2))
          ))
```

## Autocorrelation

The below indicates no Spatial Autocorrelation in r_diff_tr and in the residulas of the base regression.

```{r}
# auto correlation

# create a copy of the spatial object
bua_buasd_copy <- bua_buasd

# merge data used in the regressions with the spatial object 
bua_buasd_copy@data <- data.frame(bua_buasd_copy@data, data[match(bua_buasd_copy@data$bua11cd, data$GEOGRAPHY_CODE),])
bua_buasd_copy <- sp.na.omit(bua_buasd_copy)

# Moran's I
nearest_neigh = knearneigh(coordinates(bua_buasd_copy),k=6, RANN=F)
nearest_neigh = knn2nb(nearest_neigh)
spweights = nb2listw(nearest_neigh)

# Moran's I plot
moran.plot(data$r_diff_tr, spweights)

# Moran's I test
moran.test(data$r_diff_tr, spweights)

# subset bua_buasd based on the obs. included in base
inc.obs <- augment(base)[1]
data.inc.obs <-data %>% 
  slice(as.numeric(inc.obs$.rownames))

# merge data used in the base regression with the spatial object 
bua_buasd_copy <- bua_buasd
bua_buasd_copy@data <- data.frame(bua_buasd_copy@data, data[match(bua_buasd_copy@data$bua11cd, data.inc.obs$GEOGRAPHY_CODE),])
bua_buasd_copy <- sp.na.omit(bua_buasd_copy)

nearest_neigh = knearneigh(coordinates(bua_buasd_copy),k=6, RANN=F)
nearest_neigh = knn2nb(nearest_neigh)
spweights = nb2listw(nearest_neigh)

# Autocorrelation in residuals: NO AUTOCORRELATION
lm.morantest(base, spweights)
```

```{r}
path.out <- paste0(path, "/data/data_inter/UK.RData")
save.image(path.out)
```

### 2020-09-02 src/un.Rmd
all but .RData

---
title: "UN"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs")
  })
---

```{r}
library(rmarkdown)
library(data.table)
library(tidyverse)
library(readxl)
library(httr)
library(rprojroot)

library(ggplot2)
#library(geojsonio)
library(rgdal)
library(maptools)
library(sp)
library(sf)
library(rgeos)
#library(classInt)
#library(RColorBrewer)
library(knitr)
#library(viridis)
#library(hrbrthemes)
library(fs)
library(betareg)
library(broom)

library(AER)
library(MASS)
#library(foreign)
library(stargazer)
library(ivpack) 
library(leaflet)
library(rvest)
library(xml2)
library(tmaptools)
library(spdep)
library(spatialreg)

library(acs)
library(tidycensus)
library(rio)
library(wbstats)
library(WDI)
library(kableExtra)

# This is the project path
path <- find_rstudio_root_file()
```

## Zipf regressions

```{r}
# get urban pop data
url <- "https://population.un.org/wup/Download/Files/WUP2018-F22-Cities_Over_300K_Annual.xls"
GET(url, write_disk(tf <- tempfile(fileext = ".xls")))
urban.pop.raw <- read_excel(tf, 1L, range = "A17:CP1877", col_names = T)
# also saved as ./Data/UN/WUP2018-F22-Cities_Over_300K_Annual.xls

# get country codes
url <- "https://population.un.org/wup/Download/Files/WUP2018-F00-Locations.xlsx"
# also saved as ./Data/UN/WUP2018-F22-Cities_Over_300K_Annual.xls

codes <- rio::import(url)  # , startRow = 32)
codes <- codes[27:285,2:5]
codes <- codes %>%
  dplyr::select(-`...3`) %>%
  rename(`Country or area` = `...2`,
         `Country Code` = `United Nations`,
         ISO = `...5`) %>%
  remove_rownames() %>%
  dplyr::select(-`Country or area`) # no need of this at the end

# names
names <- urban.pop.raw %>% 
  dplyr::select(1:5 | starts_with("2")) %>%
  dplyr::select(-(25:41)) %>%
  rename_at(vars(-c(Index, 
                    `Country Code`, 
                    `Country or area`, 
                    `City Code`, 
                    `Urban Agglomeration`)),
            function(x) paste0("pop", x))

# # rankings OLD
# ranks <- urban.pop.raw %>% 
#   dplyr::select(`Country Code` | starts_with("2")) %>%
#   group_by(`Country Code`) %>%
#   summarise(across(everything(),dense_rank)) %>%
#   dplyr::select(-c(1, 21:37)) %>%
#   rename_at(vars(-`Country Code`),
#             function(x) paste0("rank", x))

dense_rank_d <- function(x){
  dense_rank(-x)
}

# rankings
ranks <- urban.pop.raw %>% 
  dplyr::select(`Country Code`, `City Code`, starts_with("2")) %>%
  group_by(`Country Code`) %>%
  mutate_at(3:38, dense_rank_d) %>%
  #mutate(dense_rank(-c(3:38))) %>%
  dplyr::select(-c(22:38)) %>%
  rename_at(vars(-`Country Code`, -`City Code`),
            function(x) paste0("rank", x))

# merge
#urban.pop <- bind_cols(names, ranks)
urban.pop <- left_join(names, ranks, by = "City Code") %>%
  dplyr::select(-`Country Code.y`) %>%
  rename(`Country Code` = `Country Code.x`)

#urban.pop$`Country Code...25` <- NULL
#urban.pop <- rename(urban.pop, `Country Code` = `Country Code...2`)
urban.pop <- merge(urban.pop, codes, by = "Country Code", all.x = T)
urban.pop <- urban.pop %>%
  relocate(ISO, .before = `City Code`) %>%
  dplyr::select(-Index) %>%
  group_by(ISO) %>%
  add_tally %>%
  rename(n.cities = n) %>%
  relocate(n.cities, .before = pop2000) %>%
  filter(n.cities > 5) # keep the countries with x or more urban agglomeration of 300k or above

# Zipf regressions
zipf <- NULL
years <- 2000:2018
countries <- unique(urban.pop$ISO)
for (c in countries){
  for (i in years){
  model.name <- paste(c, i, sep = ".")
  formula <- paste0("log(rank", i, " - 0.5) ~ log(pop", i, ")") # log transformation and Gabaix and Ibragimov (2011) correction
  model <- lm(formula, data = urban.pop[urban.pop$ISO==c,])
  b <- coef(summary(model))[2, 1]
  se <- coef(summary(model))[2, 2]
  df <- cbind(b = b, se = se, country = c, year = i)
  zipf <- rbind(zipf, df)
  # assign(model.name, model) # assign object (x) to a name (model.name)
}
}

# model$coefficients[2]
# se <- sqrt(diag(vcov(model)))
# coef(summary(model))[2, 2]

zipf <- as_tibble(zipf) %>% 
  mutate(b = as.numeric(b),
         se = as.numeric(se),
         year = as.numeric(year)) %>%
  dplyr::select(country, year, b, se)
```

## Data for second stage regressions - search World Bank API

```{r}

# This are just asearches in the world bank database

wb_search(pattern = "broadband")
wb_search(pattern = "internet")
wb_search(pattern = "mobile")
wb_search(pattern = "fixed")
wb_search(pattern = "total population")
wb_search(pattern = "population density")
wb_search(pattern = "GDP per capita")
wb_search(pattern = "GDP growth")
wb_search(pattern = "trade")
wb_search(pattern = "government final consumption expenditure")
wb_search(pattern = "value added")
wb_search(pattern = "Mortality rate")
wb_search(pattern = "Female labor force")
wb_search(pattern = "Secure internet servers")
wb_search(pattern = "households with television")
wb_search(pattern = "telecommunication sector")

# IT.NET.BBND.P3	    Broadband subscribers (per 1,000 people)
# IT.NET.USER.P3	    Internet users (per 1,000 people)	
# IT.CEL.SETS.P3      Mobile phone subscribers (per 1,000 people)	
# IT.MLT.MAIN.P3	    Telephone mainlines (per 1,000 people)	
# SP.POP.TOTL         Population, total	
# EN.POP.DNST	        Population density (people per sq. km of land area)
# NY.GDP.PCAP.PP.KD	  GDP per capita, PPP (constant 2017 international $)
# NY.GDP.MKTP.KD.ZG	  GDP growth (annual %)
# NE.TRD.GNFS.ZS	    Trade (% of GDP)
# NE.CON.GOVT.ZS	    General government final consumption expenditure (% of GDP)
# NV.AGR.TOTL.ZS	    Agriculture, forestry, and fishing, value added (% of GDP)
# HF.DYN.MORT	        Mortality rate, under-5 (per 1,000)	
# SL.TLF.TOTL.FE.ZS	  Labor force, female (% of total labor force)	
# IT.NET.SECR.P6	    Secure Internet servers (per 1 million people)
# IT.TVS.HOUS.ZS	    Households with television (%) **20 YEAR LAG IS NEEDED**
# SL.EMP.TRAN	        Number of people employed in transportation and telecommunication sector **20 YEAR LAG IS NEEDED**

# test <- wb_data(indicator = "IT.NET.BBND.P3", start_date = 2000, end_date = 2018)

wb_data("IT.NET.BBND.P3") # no, Broadband subscribers (per 1,000 people)
wb_data("IT.NET.BBND")    # Broadband subscribers
WDI("IT.NET.BBND.P3", country = "all") # until 2012
# localy saved in ./Data/UN/API_IT.NET.BBND.P2_DS2_en_csv_v2_1221909.zip

wb_data("IT.NET.USER.P3") # no, Internet users (per 1,000 people)	
wb_data("IT.NET.USER")    # no, Internet users (per 1,000 people)	
# http://api.worldbank.org/v2/country/all/indicator/IT.NET.USER.P3?per_page=4000
WDI("IT.NET.USER.P3", country = "all") # until 2012
# localy saved in ./Data/UN/API_IT.NET.USER.ZS_DS2_en_csv_v2_1307380.zip

wb_data("IT.CEL.SETS.P3") # no, Mobile phone subscribers (per 1,000 people)	
wb_data("IT.CEL.SETS")    # Mobile phone subscribers
WDI("IT.CEL.SETS.P3", country = "all") # until 2012
# localy saved in ./Data/UN/API_IT.CEL.SETS.P2_DS2_en_csv_v2_1217961.zip

wb_data("IT.MLT.MAIN.P3") # no, Telephone mainlines (per 1,000 people)	
wb_data("IT.MLT.MAIN")    # Telephone mainlines 
# localy saved in ./Data/UN/API_IT.MLT.MAIN.P2_DS2_en_csv_v2_1221907.zip

wb_data("SP.POP.TOTL")
wb_data("EN.POP.DNST")
wb_data("NY.GDP.PCAP.PP.KD")
wb_data("NY.GDP.MKTP.KD.ZG")
wb_data("NE.TRD.GNFS.ZS")
wb_data("NE.CON.GOVT.ZS")
wb_data("NV.AGR.TOTL.ZS")
wb_data("HF.DYN.MORT")
wb_data("SL.TLF.TOTL.FE.ZS")
wb_data("IT.NET.SECR.P6") # no, too many missing
wb_data("IT.TVS.HOUS.ZS") # no, Households with television (%) **20 YEAR LAG IS NEEDED**
WDI("IT.TVS.HOUS.ZS", country = "all") #no 
# localy saved in ./Data/UN/
wb_data("SL.EMP.TRAN")    #	no, Number of people employed in transportation and telecommunication sector **20 YEAR LAG IS NEEDED**
# localy saved in ./Data/UN/
WDI("SL.EMP.TRAN", country = "all") #no
```

## Data for second stage regressions

```{r}
# Getting the data for the second stage (after Zipf) regressions.
# This is where the data build starts.
# The below codes were obtained after the above searches.

rhs1 <- wb_data(c("SP.POP.TOTL",
                  "EN.POP.DNST",
                  "NY.GDP.PCAP.PP.KD",
                  "NY.GDP.MKTP.KD.ZG",
                  "NE.TRD.GNFS.ZS",
                  "NE.CON.GOVT.ZS",
                  "NV.AGR.TOTL.ZS",
                  #"HF.DYN.MORT", # too many missing data points
                  #"IT.NET.SECR.P6", # too many missing data points
                  "SL.TLF.TOTL.FE.ZS"),
                start_date = 2000,
                end_date = 2018)

rhs1 <- rhs1 %>% 
  rename(country_name = country,
         country = iso3c,
         year = date,
         pop.dens = EN.POP.DNST,
         # mort5 = HF.DYN.MORT,
         # sec.serv = IT.NET.SECR.P6,  
         gov.exp = NE.CON.GOVT.ZS,
         trade = NE.TRD.GNFS.ZS,
         agri = NV.AGR.TOTL.ZS,
         gdp.growth = NY.GDP.MKTP.KD.ZG, 
         gdppc = NY.GDP.PCAP.PP.KD,
         fem.lab = SL.TLF.TOTL.FE.ZS,
         pop = SP.POP.TOTL) %>%
  mutate(non.agri = 100-agri) %>%
  dplyr::select(-agri)

# internet users
net.path <- paste0(path, "/Data/UN/API_IT.NET.USER.ZS_DS2_en_csv_v2_1307380.zip")
con <- unz(net.path, "API_IT.NET.USER.ZS_DS2_en_csv_v2_1307380.csv")
net <- read_csv(con, skip = 4) %>%
  dplyr::select(-(5:44), -(64:65)) %>%
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,3:4)) %>%
  gather("year", "net100hab", -c(country)) %>%
  mutate(year = as.numeric(year))

# broadband users
bb.path <- paste0(path, "/Data/UN/API_IT.NET.BBND.P2_DS2_en_csv_v2_1221909.zip")
con <- unz(bb.path, "API_IT.NET.BBND.P2_DS2_en_csv_v2_1221909.csv")
bb <- read_csv(con, skip = 4) %>%
  dplyr::select(-(5:44), -(64:65)) %>%
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,3:4)) %>%
  gather("year", "bb100hab", -c(country)) %>%
  mutate(year = as.numeric(year))

# mobile phone users
mobile.path <- paste0(path, "/Data/UN/API_IT.CEL.SETS.P2_DS2_en_csv_v2_1217961.zip")
con <- unz(mobile.path, "API_IT.CEL.SETS.P2_DS2_en_csv_v2_1217961.csv")
mobile <- read_csv(con, skip = 4) %>%
  dplyr::select(-(5:44), -(64:65)) %>%
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,3:4)) %>%
  gather("year", "mobile100hab", -c(country)) %>%
  mutate(year = as.numeric(year))

# fixed telephony users
fixed.path <- paste0(path, "/Data/UN/API_IT.MLT.MAIN.P2_DS2_en_csv_v2_1221907.zip")
con <- unz(fixed.path, "API_IT.MLT.MAIN.P2_DS2_en_csv_v2_1221907.csv")
fixed <- read_csv(con, skip = 4) %>%
  dplyr::select(-(5:44), -(64:65)) %>%
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,3:4)) %>%
  gather("year", "fixed100hab", -c(country)) %>%
  mutate(year = as.numeric(year))

# tv 
# tv <- WDI("IT.TVS.HOUS.ZS", country = "all") %>%
#   rename(country = iso2c, country_name = country, tv.hous100 = IT.TVS.HOUS.ZS) %>% 
#   mutate(year = year + 10) # this is for the 10 year lag IV

tv.path <- paste0(path, "/Data/UN/itu_database2.xls")
tv <- read_excel(tv.path) %>%
  filter(`Series Name` == "Percentage of households with TV") %>%
  dplyr::select(-(5:35)) %>% #, -(64:65)
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,2,4)) %>%
  gather("year", "tv100hh", -c(country)) %>%
  mutate(tv100hh = as.numeric(tv100hh),
         year = as.numeric(year) + 10) %>% # this is for the 10 year lag IV 
  rename(tv100hh10L = tv100hh)

# employment in telecommunications
tele.emp.path <- paste0(path, "/Data/UN/tele_employment.csv")
tele.emp <- read_csv(tele.emp.path) %>%
  dplyr::select(-(5:34)) %>% #, -(64:65)
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,2,4)) %>%
  gather("year", "tele.emp", -c(country)) %>%
  mutate(tele.emp = as.numeric(tele.emp),
         year = as.numeric(year) + 10) %>% # this is for the 10 year lag IV 
  rename(tele.emp10L = tele.emp)

# TODO: GAPS
```

## Merge objects

```{r}
data <- list(zipf, rhs1, net, bb, mobile, fixed)# tele.emp and tv are not used in the regressions
sapply(data, function(x) dim(x))
sapply(data, function(x) names(x))
sapply(data, function(x) str(x))

# merge with reduce
data <- data %>% 
  reduce(inner_join, by = c("country", "year")) %>%
  relocate(iso2c, .after = country) %>%
  relocate(country_name, .after = country)
sapply(data, function(x) sum(is.na(x)))

# NOT RUN
# export file for future reference / backup
# data.out.path <- paste0(path, "/data/data_inter/data_for_un_regressions.csv")
# write.csv(data, data.out.path)
```

## descriptives

```{r}
ggplot(data = data, aes(x=bb100hab, y=b)) + geom_point() + geom_smooth(method = "lm")
```

## regressions

```{r}

# These are just tests

library(plm)
data.plm <- pdata.frame(data, index=c("country","year"), drop.index=TRUE, row.names=TRUE)

test.lm <- plm(b ~ pop.dens  + gov.exp + trade + non.agri + gdp.growth + gdppc + fem.lab + pop + net100hab + as.factor(year), 
               data = data, model = "within")
summary(test.lm)

test.2w <- plm(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop), 
               data = data, model = "within", effect = "twoways", weights = 1/(se^2))
summary(test.2w)

test.w.ols <- lm(b ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))
summary(test.w.ols)

test.2w.2sls <- ivreg(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(test.2w.2sls, vcov = sandwich, diagnostics = TRUE)

test.2w.2sls <- plm(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2), model = "fd", effect = "twoways")
summary(test.2w.2sls, vcov = sandwich, diagnostics = TRUE)

# + bb100hab + mobile100hab + fixed100hab 
# + mort5 + fem.lab +

```

## Table 4

```{r}
net.model <- lm(b ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))

bb.model <- lm(b ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))

fixed.model <- lm(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))

mobile.model <- lm(b ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))

# Table 4
net.rse <- coeftest(net.model, function(x) vcovHC(x, type="HC0"))
bb.rse <- coeftest(bb.model, function(x) vcovHC(x, type="HC0"))
fixed.rse <- coeftest(fixed.model, function(x) vcovHC(x, type="HC0"))
mobile.rse <- coeftest(mobile.model, function(x) vcovHC(x, type="HC0"))

path.out <- paste(path, "/outputs/table4.htm", sep = "")

stargazer(net.model, bb.model, fixed.model, mobile.model, type="html",
          se = list(net.rse[,"Std. Error"], bb.rse[,"Std. Error"], fixed.rse[,"Std. Error"], mobile.rse[, "Std. Error"]), 
          dep.var.labels=c("Zipf Coefficient 2000-18"),
          # covariate.labels=c("households w. internet, 2013 (%)", 
          #                    "population, 2013",  
          #                    "% of unemployment, 2013", 
          #                    "% of white population, 2013", 
          #                    "income, 2013 (log)",
          #                    "population density, 2013", 
          #                    "employment in service, 2011 (%)", 
          #                    "commute in minutes, 2013", 
          #                    "households w. internet, 2013 (%) x pop. density, 2013", 
          #                    "households w. internet, 2013 (%) x population, 2013"), 
          out=path.out, single.row = FALSE, df = FALSE, 
          omit = c("country", "year"),
          add.lines = list(c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          omit.stat = c("rsq", "f"), notes = "Robust Std. Error in parenthesis, corrected Zipf coefficient following Gabaix and Ibragimov (2011)")

```

## Table 5

```{r}
net.2sls <- ivreg(b ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(net.2sls, vcov = sandwich, diagnostics = TRUE)

# net.2sls.2 <- ivreg(b ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
#                       + as.factor(year) + as.factor(country) |
#                       log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
#                       + as.factor(year) + as.factor(country)  +
#                       fem.lab + tv100hh10L, # 
#                data = data, weights = 1/(se^2))
# summary(net.2sls.2, vcov = sandwich, diagnostics = TRUE)

bb.2sls <- ivreg(b ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(bb.2sls, vcov = sandwich, diagnostics = TRUE)

fixed.2sls <- ivreg(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(fixed.2sls, vcov = sandwich, diagnostics = TRUE)

mobile.2sls <- ivreg(b ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(mobile.2sls, vcov = sandwich, diagnostics = TRUE)

net.2sls.rse <- coeftest(net.2sls, function(x) vcovHC(x, type="HC0"))
summ.net.2sls <- summary(net.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

bb.2sls.rse <- coeftest(bb.2sls, function(x) vcovHC(x, type="HC0"))
summ.bb.2sls <- summary(bb.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

fixed.2sls.rse <- coeftest(fixed.2sls, function(x) vcovHC(x, type="HC0"))
summ.fixed.2sls <- summary(fixed.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

mobile.2sls.rse <- coeftest(mobile.2sls, function(x) vcovHC(x, type="HC0"))
summ.mobile.2sls <- summary(mobile.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

path.out <- paste(path, "/outputs/table5.htm", sep = "")

stargazer(net.2sls, bb.2sls, fixed.2sls, mobile.2sls, type = "html", 
          se = list(net.2sls.rse[,"Std. Error"], bb.2sls.rse[,"Std. Error"],
                    fixed.2sls.rse[,"Std. Error"], mobile.2sls.rse[,"Std. Error"]), 
          dep.var.labels=c("Normalized difference in ranks, 2013-18"),
          # covariate.labels=c("households w. internet, 2013 (%)", 
          #                    "population, 2013",  
          #                    "% of unemployment, 2013", 
          #                    "% of white population, 2013", 
          #                    "income, 2013 (log)",
          #                    "population density, 2013", 
          #                    "employment in service, 2011 (%)", 
          #                    "commute in minutes, 2013"), 
          out=path.out, single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"), notes = c("Robust Std. Error in parenthesis", 
                                               "IV: Female paticipation in labour force"),
          add.lines = list(c(rownames(summ.net.2sls$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.2sls$diagnostics[1, "statistic"], 2), 
                             round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)), 
                           c(rownames(summ.net.2sls$diagnostics)[2], 
                             round(summ.net.2sls$diagnostics[2, "statistic"], 2), 
                             round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)), 
                           c("P-value", 
                             round(summ.net.2sls$diagnostics[2, "p-value"], 2), 
                             round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)), 
                           c(rownames(summ.net.2sls$diagnostics)[3], 
                             round(summ.net.2sls$diagnostics[3, "statistic"], 2), 
                             round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           c("P-value", 
                             round(summ.net.2sls$diagnostics[3, "p-value"], 2), 
                             round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                             round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                             round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")
          ))

```

## Table 1

```{r}
zipf.table <- net.model$model %>%
  dplyr::select(`as.factor(country)`, `as.factor(year)`, b) %>%
  spread(`as.factor(year)`, b) %>%
  rename(`Zipf coef.` = `as.factor(country)`)

zipf.table <- zipf.table %>% kable(format = "html", digits = 3)

path.out <- paste(path, "/outputs/table1.htm", sep = "")
save_kable(zipf.table, path.out)
```

```{r}
path.out <- paste0(path, "/data/data_inter/UN.RData")
save.image(path.out)
```

### 2020-11-18 src/s1.Rmd
S1

---
title: "S1 Appendix"
author: ""
output: pdf_document
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs")
  })
---

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

# This .Rmd create teh S1 Appendix for the estimation of eq. 3
# using alternative specifications: Gini, Herfindahl-Hirschman
# and CV. 
# The output S1.pdf is located in /outputs

library(rmarkdown)
library(data.table)
library(tidyverse)
library(readxl)
library(httr)
library(rprojroot)

library(ggplot2)
#library(geojsonio)
library(rgdal)
library(maptools)
library(sp)
library(sf)
library(rgeos)
#library(classInt)
#library(RColorBrewer)
library(knitr)
#library(viridis)
#library(hrbrthemes)
library(fs)
library(betareg)
library(broom)

library(AER)
library(MASS)
#library(foreign)
library(stargazer)
library(ivpack) 
library(leaflet)
library(rvest)
library(xml2)
library(tmaptools)
library(spdep)
library(spatialreg)

library(acs)
library(tidycensus)
library(rio)
library(wbstats)
library(WDI)
library(kableExtra)

library(REAT)

# This is the project path
# path <- find_rstudio_root_file()

load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/UN.RData")
```

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

# Gini

net.gini.2sls <- ivreg(gini ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(net.gini.2sls, vcov = sandwich, diagnostics = TRUE)

bb.gini.2sls <- ivreg(gini ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(bb.gini.2sls, vcov = sandwich, diagnostics = TRUE)

fixed.gini.2sls <- ivreg(gini ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(fixed.gini.2sls, vcov = sandwich, diagnostics = TRUE)

mobile.gini.2sls <- ivreg(gini ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(mobile.gini.2sls, vcov = sandwich, diagnostics = TRUE)

net.gini.2sls.rse <- coeftest(net.gini.2sls, function(x) vcovHC(x, type="HC0"))
summ.net.gini.2sls <- summary(net.gini.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

bb.gini.2sls.rse <- coeftest(bb.gini.2sls, function(x) vcovHC(x, type="HC0"))
summ.bb.gini.2sls <- summary(bb.gini.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

fixed.gini.2sls.rse <- coeftest(fixed.gini.2sls, function(x) vcovHC(x, type="HC0"))
summ.fixed.gini.2sls <- summary(fixed.gini.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

mobile.gini.2sls.rse <- coeftest(mobile.gini.2sls, function(x) vcovHC(x, type="HC0"))
summ.mobile.gini.2sls <- summary(mobile.gini.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
```

```{r, echo=FALSE, warning=FALSE, results='asis'}
stargazer(net.gini.2sls, bb.gini.2sls, fixed.gini.2sls, mobile.gini.2sls, type = "latex",
          se = list(net.gini.2sls.rse[,"Std. Error"], bb.gini.2sls.rse[,"Std. Error"],
                    fixed.gini.2sls.rse[,"Std. Error"], mobile.gini.2sls.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (3), alt. specifications (i)\\label{2sls.gini.global}",
          font.size = "small",
          column.sep.width = "1pt",
          dep.var.labels=c("Gini coefficient 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(c(rownames(summ.net.gini.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.gini.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.bb.gini.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.gini.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.gini.2sls$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.net.gini.2sls$diagnostics)[2],
                             round(summ.net.gini.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.bb.gini.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.gini.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.gini.2sls$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.net.gini.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.bb.gini.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.gini.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.gini.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis",
                    "IV: Female participation in labour force"))

```

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

# herf

net.herf.2sls <- ivreg(herf ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(net.herf.2sls, vcov = sandwich, diagnostics = TRUE)

bb.herf.2sls <- ivreg(herf ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(bb.herf.2sls, vcov = sandwich, diagnostics = TRUE)

fixed.herf.2sls <- ivreg(herf ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(fixed.herf.2sls, vcov = sandwich, diagnostics = TRUE)

mobile.herf.2sls <- ivreg(herf ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(mobile.herf.2sls, vcov = sandwich, diagnostics = TRUE)

net.herf.2sls.rse <- coeftest(net.herf.2sls, function(x) vcovHC(x, type="HC0"))
summ.net.herf.2sls <- summary(net.herf.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

bb.herf.2sls.rse <- coeftest(bb.herf.2sls, function(x) vcovHC(x, type="HC0"))
summ.bb.herf.2sls <- summary(bb.herf.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

fixed.herf.2sls.rse <- coeftest(fixed.herf.2sls, function(x) vcovHC(x, type="HC0"))
summ.fixed.herf.2sls <- summary(fixed.herf.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

mobile.herf.2sls.rse <- coeftest(mobile.herf.2sls, function(x) vcovHC(x, type="HC0"))
summ.mobile.herf.2sls <- summary(mobile.herf.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
```

```{r, echo=FALSE, warning=FALSE, results='asis'}
stargazer(net.herf.2sls, bb.herf.2sls, fixed.herf.2sls, mobile.herf.2sls, type = "latex",
          se = list(net.herf.2sls.rse[,"Std. Error"], bb.herf.2sls.rse[,"Std. Error"],
                    fixed.herf.2sls.rse[,"Std. Error"], mobile.herf.2sls.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (3), alt. specifications (ii)\\label{2sls.herf.global}",
          font.size = "small",
          column.sep.width = "1pt",
          dep.var.labels=c("Herfindahl-Hirschman coefficient 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(c(rownames(summ.net.herf.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.herf.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.bb.herf.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.herf.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.herf.2sls$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.net.herf.2sls$diagnostics)[2],
                             round(summ.net.herf.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.bb.herf.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.herf.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.herf.2sls$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.net.herf.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.bb.herf.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.herf.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.herf.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis",
                    "IV: Female participation in labour force"))

```

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

# cv

net.cv.2sls <- ivreg(cv ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(net.cv.2sls, vcov = sandwich, diagnostics = TRUE)

bb.cv.2sls <- ivreg(cv ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(bb.cv.2sls, vcov = sandwich, diagnostics = TRUE)

fixed.cv.2sls <- ivreg(cv ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(fixed.cv.2sls, vcov = sandwich, diagnostics = TRUE)

mobile.cv.2sls <- ivreg(cv ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(mobile.cv.2sls, vcov = sandwich, diagnostics = TRUE)

net.cv.2sls.rse <- coeftest(net.cv.2sls, function(x) vcovHC(x, type="HC0"))
summ.net.cv.2sls <- summary(net.cv.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

bb.cv.2sls.rse <- coeftest(bb.cv.2sls, function(x) vcovHC(x, type="HC0"))
summ.bb.cv.2sls <- summary(bb.cv.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

fixed.cv.2sls.rse <- coeftest(fixed.cv.2sls, function(x) vcovHC(x, type="HC0"))
summ.fixed.cv.2sls <- summary(fixed.cv.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

mobile.cv.2sls.rse <- coeftest(mobile.cv.2sls, function(x) vcovHC(x, type="HC0"))
summ.mobile.cv.2sls <- summary(mobile.cv.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
```

```{r, echo=FALSE, warning=FALSE, results='asis'}
stargazer(net.cv.2sls, bb.cv.2sls, fixed.cv.2sls, mobile.cv.2sls, type = "latex",
          se = list(net.cv.2sls.rse[,"Std. Error"], bb.cv.2sls.rse[,"Std. Error"],
                    fixed.cv.2sls.rse[,"Std. Error"], mobile.cv.2sls.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (3), alt. specifications (iii)\\label{2sls.cv.global}",
          font.size = "small",
          column.sep.width = "1pt",
          dep.var.labels=c("Coefficient of variation 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(c(rownames(summ.net.cv.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.cv.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.bb.cv.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.cv.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.cv.2sls$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.net.cv.2sls$diagnostics)[2],
                             round(summ.net.cv.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.bb.cv.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.cv.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.cv.2sls$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.net.cv.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.bb.cv.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.cv.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.cv.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis",
                    "IV: Female participation in labour force"))

```

### 2020-11-18 src/un.Rmd
S1

@@
 
+library(REAT)
+
 # This is the project path
@@
 ```
+
+Add description
 
@@
 
+## Alternative measures of spatial dispertion
+
+```{r}
+gini <- urban.pop %>% 
+  rename(country = ISO) %>%
+  dplyr::select((contains("pop")) | country) %>%
+  group_by(country) %>% 
+  summarise(across(everything(), list(REAT::gini))) %>%
+  pivot_longer(!country, names_to = "year", values_to = "gini") %>%
+  mutate(year = gsub("pop", "", year)) %>% 
+  mutate(year = str_sub(year, end=-3)) %>%
+  mutate(year = as.numeric(year))
+
+herf <- urban.pop %>% 
+  rename(country = ISO) %>%
+  dplyr::select((contains("pop")) | country) %>%
+  group_by(country) %>% 
+  summarise(across(everything(), list(REAT::herf))) %>%
+  pivot_longer(!country, names_to = "year", values_to = "herf") %>%
+  mutate(year = gsub("pop", "", year)) %>% 
+  mutate(year = str_sub(year, end=-3)) %>%
+  mutate(year = as.numeric(year))
+
+cv <- urban.pop %>% 
+  rename(country = ISO) %>%
+  dplyr::select((contains("pop")) | country) %>%
+  group_by(country) %>% 
+  summarise(across(everything(), list(EnvStats::cv))) %>% # raster::cv / 100
+  pivot_longer(!country, names_to = "year", values_to = "cv") %>%
+  mutate(year = gsub("pop", "", year)) %>% 
+  mutate(year = str_sub(year, end=-3)) %>%
+  mutate(year = as.numeric(year))
+
+```
+
 ## Data for second stage regressions - search World Bank API
@@
 
-# This are just asearches in the world bank database
+# These are just searches in the world bank database
 
@@
 
-# TODO: GAPS
 ```
@@
 ```{r}
-data <- list(zipf, rhs1, net, bb, mobile, fixed)# tele.emp and tv are not used in the regressions
+data <- list(zipf, rhs1, net, bb, mobile, fixed,
+             gini, herf, cv)# tele.emp and tv are not used in the regressions
 sapply(data, function(x) dim(x))
@@
           #                    "commute in minutes, 2013"), 
-          out=path.out, single.row = FALSE, df = FALSE,
+          out=path.out, 
+          single.row = FALSE, df = FALSE,
           omit = c("country", "year"),

### 2020-11-19 paper/2_resubmission/Response_to_Reviewers.tex
paper on git

\documentclass[11pt, a4paper, answers]{exam}

\usepackage{shadow, fancybox, a4wide, bm, harvard, epsfig, theorem,  epic, verbatim, amsmath, amssymb,  microtype, fullpage, setspace, caption, makeidx, subfigure, marvosym, booktabs, dcolumn, mdwlist, lastpage, tikz, eurosym}
\usepackage[OT1]{fontenc}
\usepackage[utf8]{inputenx}
\renewcommand{\baselinestretch}{1.25}

%\hypersetup{%
%	colorlinks,urlcolor=blue,
%	pdfborder = {0 0 0}
%}

\usepackage{xcolor}
%\definecolor{SolutionColor}{rgb}{0.1,0.3,1}
%\renewcommand{\thequestion}{\alph{question} }
\renewcommand\questionlabel{\llap{\thequestion)}}

\renewcommand{\solutiontitle}{\noindent\textbf{Response: }}

%\pointsinrightmargin
%\boxedpoints
%\unframedsolutions
\shadedsolutions
\definecolor{SolutionColor}{rgb}{0.9,0.9,1}

\begin{document}

\noindent Dear Editor and Reviewers, \bigskip\bigskip

Please find attached a report covering our detailed responses to the editor and 
the reviewer comments. Overall, we feel the paper has been substantially improved 
as a result of the constructive criticism and suggestions made in the  review process,
and we would like to sincerely thank both reviewers as well as the editor for
that. We have structured this response by listing every
comment and immediately following it with a blue box that includes our
response.

%\today
\clearpage

\section*{Responses to Editor}

\begin{questions}
\question The second paragraph in the literature is too long. Consider breaking 
it into two or three paragraphs based on reclassification of the literature, for example, 
by substitution and complementarity effects, or by urban and rural areas.

\begin{solution}
\textbf{ET}
\end{solution}

\question For clarity, I would like to suggest you replace Zipf coefficient by Pareto 
exponent unless the coefficient equals 1, as you rightfully pointed it out in the 
paragraph below equation (1). This will help readers avoid some confusions.

\begin{solution}
\textbf{YMI}
\end{solution}

\question Please clarify in equation (3) whether you want to include time trend 
delta*t or year fixed effect delta subscript t. You used delta*t in the equation 
but used delta subscript t in line 234. Tables 4 and 5 say year fixed effects.

\begin{solution}
\textbf{ET}
\end{solution}

\question Using female labor force participation rate as IV for ICT usage is a bit 
exotic. Even females stay home as housewives, don’t they still need to use phones 
and internet at home? You did not provide any empirical evidence that female labor 
force participation indeed increases ICT usage. It would be more convincing if 
you can find and cite such supportive references.

\begin{solution}
\textbf{ET}
\end{solution}

\question Table 5: the impact on Zipf coefficient is very small: 0.001 or 0.002. 
How economically significant is this number in terms of city size distribution? 
Particularly, since a country uses both internet and fixed line phones and they 
have opposite, offsetting effects on Zipf coefficient, does this mean at the 
aggregate level there is no effect on the overall city size distribution? 
These are subtle questions and you may discuss a bit on this.

\begin{solution}
\textbf{YMI}
\end{solution}

\question Line 504: you mentioned using absolute number of broadband tests 
in 2011 as an additional IV, isn’t it a direct measure of ICT and similar 
to the main explanatory variable “broadband tests per capita 2011”?

\begin{solution}
\textbf{ET}
\end{solution}

\end{questions}

\section*{Responses to Reviewer 1}

\begin{questions}

\question There are two kinds of spatial dispersion: within a city or 
across different cities. They are conceptually different; there could be 
decreased dispersion within a city but at the same time increased 
dispersion across cities. Both types of spatial dispersion can be 
affected by the advancement of information technology. This study 
examines the “Zipf coefficient” of an urban system or a city’s rank 
in an urban system, thus it focuses on spatial dispersion across cities. 
Nothing is said about spatial dispersion within a city (e.g., whether 
ICT has resulted in more urban sprawl). The authors should make this 
clear upfront.

\begin{solution}
\textbf{YMI: maybe add a line or two in the intro where we talk about data?}
\end{solution}

\question After equation (1), it’ll be useful to add a brief discussion 
on why the Zipf coefficient measures spatial dispersion. Many 
readers should have seen the variance, Gini coefficient, or 
Herfindahl index used as a measure of dispersion. I am not 
sure how many of them has seen the Zipf coefficient, so it 
requires some explanation. I also wonder why those other 
measures are not used.

\begin{solution}
\textbf{YMI for the text, ET to check if we have estimates for 
alternative measures}
\end{solution}

\question In Table 2, the Zipf coefficient ranges between -2 and -1, 
but has a mean of -1 and a standard deviation of 0. That seems odd. 
Also, what is the “St. error for Zipf est.” in the second row?

\begin{solution}
\textbf{ET}
\end{solution}

\question On several occasions (e.g., p. 9 and 15), the authors 
seem to suggest that a valid IV should not affect (or be correlated 
with) the outcome variable. This is not entirely accurate. A valid 
IV must be correlated with the endogenous variable. Thus, if the 
endogenous variable indeed has a causal effect on the outcome variable, 
the IV must be correlated with the outcome variable. It would be 
correct to say that an IV would not affect the outcome variable 
directly and would only affect it indirectly (i.e., through the 
endogenous variable). However, no direct effect is different 
from a zero correlation between the IV and the outcome variable. 
Thus showing a low correlation between the IV and the outcome 
variable (p. 12, 15, and 18) is not informative about the validity of the IV.

\begin{solution}
\textbf{ET will remove correlations}
\end{solution}

\question In the global study, the use of an IV has reversed the 
sign of an OLS estimate and made two statistically insignificant 
OLS estimates become significant. In both case studies, the use 
of an IV has produced much larger estimates than the OLS 
estimates. If the authors really trust the IV estimates, it will be 
helpful to discuss what have biased the OLS estimates in the 
particular direction. For example, in both case studies, one 
might suspect that omitted variables could bias the OLS 
estimates upward. However, surprisingly the IV estimates 
turned out to be even larger than the OLS estimates. How do 
we understand this?

\begin{solution}
\textbf{ET will give first it a go then YMI }
\end{solution}

\question When using an instrumental variable, it will be 
helpful to offer some intuition why the IV is correlated with 
the endogenous variable. For example, why does female 
participation in labor force affect internet and phone usage? 
Why does the absolute number of broadband tests in 2011 
affect the endogenous variable (Internet speed)? These are 
not obvious, so some discussion is needed. Also, it might be 
helpful to show (or simply mention) some of the first-stage 
regression results to let readers know how the IV and the 
endogenous variable are correlated. For example, I would 
like to know whether and to what extent the absolute number 
of broadband tests in 2011 increased Internet speed. This 
information helps readers gauge whether the overall IV 
strategy makes sense.

\begin{solution}
\textbf{ET}
\end{solution}

\question In the two case studies, I am not sure that the 
normalization of the changed rank is helpful. I would prefer 
to use the unnormalized change in rank as the left-hand 
side variable for the OLS regression, because it makes the 
key coefficient easier to interpret. Normalization allows for 
the estimation of the quasibinomial GLM, but the GLM results 
are never discussed. It might make sense to use the unnormalized 
change in rank as the left-hand side variable, keep the OLS results 
only, and simply mention the GLM exercise in a footnote.

\begin{solution}
\textbf{ET}
\end{solution}

\end{questions}

\section*{Responses to Reviewer 2}

\begin{questions}

\question The authors used different instrumental variables for 
the multi-county, US, and UK analyses. Why is female labor 
participation not used as a consistent IV throughout the 
analyses? Is it because of different contexts or limited variation? 
The US and UK analyses also use similar yet different IV's. 
Perhaps the authors can further discuss their rationale in changing choices of IV's.

\begin{solution}
\textbf{ET}
\end{solution}

\question In interpreting the empirical results, the authors, 
to some extent, highlight the potential mechanism in explaining 
the positive relationship between internet adoption and urban 
structure. However, there is limited discussion on other interesting 
findings. For instance, why does fixed telephony have an opposite 
effect to internet and digital communication technologies?

\begin{solution}
\textbf{YMI, maybe an argument based on time/maturity of technology/link to the 
2008 paper?}
\end{solution}

\question Visualizing the patterns in Table 1 in the format of maps 
may make the information more reader friendly.

\begin{solution}
\textbf{ET}
\end{solution}

\question There are minor formatting issues and typos throughout 
the paper. For examples, "Government expenditure ( Trade (\% of GDP)" 
in Table 4 and "cites" in line 343.

\begin{solution}
\textbf{ET}
\end{solution}
\end{questions}

\clearpage
\end{document}

### 2020-11-19 paper/2_resubmission/tranos_ioannides.Rmd
paper on git

---
title: Ubiquitous Digital Technologies and Spatial Structure; an update
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliation: University of Bristol and The Alan Turing Institute
    corresponding: e.tranos@bristol.ac.uk
  - name: Yannis M. Ioannides
    email: Yannis.Ioannides@tufts.edu
    affiliation: Tufts University
address:
  - code: University of Bristol and The Alan Turing Institute
    address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
  - code: Tufts University
    address: Tufts University, Department of Economics, Medford, MA, USA
abstract: |
    This paper examines the impact of widespread adoption of information and communication technologies (ICT) on urban structure worldwide. Has it offset agglomeration benefits and led to more dispersed spatial structures, or has it strengthened urban externalities and thus resulted in more concentrated spatial structures? Theoretical and empirical studies on this question have produced contradictory findings. The present study recognizes that assumptions made earlier about the evolution of technological capabilities do not necessarily hold today. As cutting-edge digital technologies have matured considerably, a fresh look at this question is called for.

    The paper addresses this issue by means of several data sets using instrumental variable methods. One is the UN data on Urban Settlements with more than $300,000$ inhabitants. Estimation methods with these data show that increased adoption of ICT  has resulted in national urban systems that are less uniform in terms of city sizes and are characterized by higher population concentrations in larger cities, when concentration is proxied the Pareto (Zipf) coefficient for national city size distributions. Two, is disaggregated data for the urban systems of the US, defined as Micropolitan and Metropolitan Areas, and for the UK, defined as Built-up Areas in England and Wales, respectively. These data allow for the impacts to be studied for cities smaller than those included in the cross-country data. Increased internet usage improved a city’s ranking in the US urban system. Similarly, increased download speed improves a built-up area’s ranking in England and Wales.

author_summary: |
  TO ADD

bibliography: tranos_ioannides.bib
output: rticles::plos_article
csl: plos.csl
header-includes:
- \usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}

---

```{r echo=FALSE, message = FALSE, warning=FALSE}
library(stargazer)
library(tidyverse)
library(rmarkdown)
# library(data.table)
library(tidyverse)
# library(readxl)
# library(httr)
library(rprojroot)

library(ggplot2)
#library(geojsonio)
library(rgdal)
library(maptools)
library(sp)
library(sf)
library(rgeos)
#library(classInt)
#library(RColorBrewer)
library(knitr)
#library(viridis)
#library(hrbrthemes)
# library(fs)
library(betareg)
library(broom)

library(AER)
library(MASS)
#library(foreign)
library(ivpack)
# library(leaflet)
# library(rvest)
# library(xml2)
# library(tmaptools)
# library(spdep)
# library(spatialreg)

#library(acs)
#library(tidycensus)
#library(rio)
#library(wbstats)
#library(WDI)
library(kableExtra)
```

# Introduction {#sec:1}

Geographers, planners and urban economists spent effort in exploring the spatial footprint of the internet even at its early stages.
They theorized about the spatial impacts that rapid internet penetration might generate on individual cities and the national spatial structure.
The outcome of various such efforts was rather conjectural and even fanciful and not data-based.
Cases in point are celebrations of the emergence of telecottages [@toffler_third_1980], the rise of a borderless world [@ohmae_borderless_1995], the death of cities [@gilder_notitle_1995; @peter_f_drucker_capitalism_1998], and, more generally, the end of geography [@obrien_global_1992], the death of distance [@frances_cairncross_death_2001] and the emergence of a new flat world [@friedman_world_2005].
Today, $25$ years after the commercialization of the internet [@kende_digital_2003], we know that the above narratives overstated the potential of the internet and other digital communication technologies to supplement face-to-face interactions, diminish the cost of distance and, indeed weaken agglomeration economies.
The high and steadily increasing urbanization rates [@united_nations_world_2014] falsify such predictions.

The adoption rate and the pervasiveness of new internet-based information and communication technologies, ICT for short, such as online social media and mobile-telephony hosted internet, which have increased rapidly during the last 10-15 years in the developing world, too, raise questions about how exactly ICT might have affected agglomeration economies. Conflicting technology examples can be illustrated. On the one hand, despite the broader agreement that no digital technology can reach the content richness of face-to-face communications, empirical research from the management field suggests that current digital technologies can effectively facilitate the sharing of knowledge with low to medium tacitness and even support knowledge sharing of a high degree of tacitness [@panahi_towards_2013].
On the other hand, the very same technologies can further enhance what Storper and Venables [@storper_buzz_2004] termed as buzz, as the constant publication of our personal and professional updates and whereabouts enabled by ICT can directly facilitate deliberate as well as unplanned or even unintentional face-to-face meetings [@tranos2020social].

Urban economics suggests that a key source of agglomeration externalities is proximity, which  facilitates interaction and knowledge spillovers [@rosenthal_evidence_2004].
Hence, face-to-face interactions and the implied knowledge spillovers within cities flourish because of lower transportation and interaction costs [@storper_buzz_2004].
However, ICT have the capacity to directly affect this process by further reducing transportation cost^[Or, more broader, spatial transmission costs to follow McCann’s terminology [@mccann_globalization_2008]] [@mccann_globalization_2008].
In essence, the internet and digital communications promote further learning and matching, which are at the heart of the micro-foundations of agglomeration economies identified by Duranton and Puga [@duranton_micro-foundations_2004].
Web-based applications such as Massive Open Online Courses decrease the need for co-location of actors in order to participate in formal learning activities.
Furthermore, online social media such as LinkedIn^[LinkedIn has been identified as "the largest professional matchmaker site in the world" [@van_dijck_you_2013], p. 207.
See also [@gee2019more].]
can enhance the probability of matching and indeed the quality of matches even within cities. Therefore, it is natural to ask whether the widespread adoption of ICT could offset the benefits of  immediate physical proximity and result in more dispersed spatial structures, or further reinforce agglomeration  externalities and lead to more concentrated spatial structures.

This paper contributes to the above discussion by presenting empirical research on whether the internet and digital communications have affected spatial structure, as seen via the size distribution of cities.
Contrary to most of the previous empirical studies, which are reviewed in the next section, this paper returns first to an empirical setting similar to that of Ioannides et al. [@ioannides2008effect] in order to re-examine their results by employing updated data on internet use, associated with the enhanced  technological maturity of ICT, and using a number of different data sets.
In contrast to Tranos and Ioannides [@tranos2020ict], which revisits the question by seeking to replicate Ioannides et al. [@ioannides2008effect], but with the same but augmented data as that earlier study, the present paper probes the significance of different levels of aggregation for the relation between agglomeration externalities and digital communications by employing  different scales of spatial aggregation.
One is a global multi-country analysis using urban agglomeration data for many countries; a second is a more granular analysis for the US and the UK. For those two countries the paper  brings  novel data to bear on the question.
In addition, the paper also distinguishes the effects of a broader range of ICT on spatial structure, including internet and broadband internet as well as mobile and landline telephony adoption rates.

Interestingly, in contrast to Tranos and Ioannides [@tranos2020ict], most of the differences in the results -- from the global level analysis to the case studies –- support a complementarity argument.
Specifically, the paper examines econometrically whether spatial structures have been affected by the adoption rates of the different digital communication technologies described above.
The results are robust against potential endogeneity concerns, as one might claim that the take-up of these technologies could have been affected by spatial structures themselves. That is, individuals living in more dispersed spatial systems might have made greater use of such technologies in order to overcome the lower level of agglomeration externalities.
Notably, we find some evidence that such effects are stronger in smaller urban areas. Our findings can directly inform the urban policy agenda as they advocate in favor of including digital strategies in policies aiming to enhance agglomeration externalities and to improve the position of a city within its national urban hierarchy.

The impact of ICT adoption on urban spatial structure is of such paramount importance in understanding its impact on modern economies that replicating a study by extending and updating its data is a worthwhile undertaking.
By the same token, it behooves us to examine its robustness by means of alternative data sets.
Unlike the data used by Ioannides et al [@ioannides2008effect] and Tranos and Ioannides [@tranos2020ict], the data employed by the current study is arguable more consistent.
The structure of the paper is as follows:
[Section 2](#sec2) provides a brief literature review;
[Section 3](#sec3) describes the methods and the data we use;
[Section 4](#sec4) presents the results of the  multi-country analysis;
[Section 5](#sec5) narrows down to the two case studies, the US and UK;
and, [Section 6](#sec6) concludes.

# Literature review {#sec2}

Gaspar and Glaeser [@jess_gaspar_information_1998] were the first to model the effect of telecommunication improvements on the intensity of face-to-face interactions and city size.
Their results indicate that technological improvements in telecommunications may lead to increased demand for face-to-face interactions, which will then increase the importance of cities.
Their theoretical model allows for a complementary relation between agglomeration externalities and advances in telecommunications. However, their results rest on a critical assumption, namely that face-to-face interactions are superior to any technology-mediated interactions.
Were it not for this assumption then the opposite prediction would have followed. As indicated above and as the management literature suggests,  face-to-face interactions do not necessarily dominate digitally mediated ones.
Certain elements of knowledge sharing can also be achieved by via online interactions [@panahi_towards_2013; @hildrum_sharing_2009].
Indeed, to a certain extent, this argument is technology dependent.
E.g., current teleconferencing capabilities are much more advanced than the ones available in the late 1990s.
Hence, it behooves us to return empirically to the impact on agglomeration externalities from improved telecommunication technologies.

\color{blue}
Kolko [@kolko_death_2000] uses internet diffusion data and identifies a clear complementary link between internet usage and city size.
Interestingly, he identifies higher internet domain densities in remote cities which indicated a substitution effect of the internet for longer-distance non-electronic communications. His results are consistent for different measures of internet diffusion (internet domain density and internet take-up).
Sinai and Waldfogel [@todd_sinai_geography_2004] approach the same question from the consumers’ point of view and find support for complementarity between internet and cities.
They study the link between market size and locally targeted online content and find that more online local content in  larger markets.
However, they also obtain evidence for a substitution effect: they find that holding constant local online content constant, market size has a negative effect on individual connectivity.
Forman et al. [@chris_forman_how_2005] examined whether commercial internet adoption is higher in cities than in rural areas. While the former would indicate a complementarity between internet adoption and cities, the latter would reflect a complementary relation, according to which the internet is used as a means to offset costs and lack of opportunities related to peripheriality.
Their results indicate that despite internet adoption by firms with more than 100 employees being greater in smaller urban agglomerations, the adoption of more sophisticated internet-based applications is positively related with city size in $2000$.
Sohn, Kim, and Hewings [@jungyul_sohn_information_2003] compare how information technologies are related to urban spatial structure for Chicago and Seoul.
Although they find a clear complementary link for Chicago, this was not the case for Seoul, where information technologies contribute to a more dispersed spatial pattern.
Focusing on the municipalities in the province of Barcelona, Pons-Novell and Viladecans-Marsal [@pons-novell_cities_2006] find a complementary link between individual internet take-up and off-line commercial offerings, but their results cannot safely reject a substitution effect.
Craig, Hoang, and Kohlhase [@craig_does_2016] focus on internet take-up rates across US states during the period $2000-2011$.
Their analysis provides suggestive evidence of a complementary role that internet connectivity performs on urban living.
Ioannides et al. [@ioannides2008effect] use country-level data on city size distributions to examine the impact of fixed line telephony on urban structure $1980-2000$.
Using a panel dataset of spatial dispersion measures, they find robust evidence that an increase in the number of telephone lines per capita encourages the spatial dispersion of population in that they lead to a more concentrated distribution of city sizes.
In addition, by the end of their study period the internet has come into use, but their results with internet usage is more speculative but do show that it goes in the same direction.

\color{blue}
Focusing on rural areas, Partridge et al. [@mark_d_partridge_employment_2008] find no evidence that rural distance penalties in the US have substantially changed since $1970s$ indicating that technological changes including the internet and digital communications have not affected spatial structure. Also interesting are the findings of a more recent study by Kim and Orazem [@kim_broadband_2016] on the economic effects of broadband internet in rural areas. They identify a positive effect on new firm location decisions, with the effect being higher in rural areas with larger population and in rural areas adjacent to a metropolitan area. These results suggest a complementary relation between the internet and agglomeration economies.

\color{blue}
Moving beyond modelling the direct relationship between ICT and spatial structure,
Bekkerman and Gilpin [@bekkerman_high-speed_2013] focused on the role of locally 
based information resources using a dataset about the US libraries during the 
period $2000-2008$. Their results suggest that internet access increases the 
demand and the value of locally accessible information, and such complementarities 
are higher in larger metropolitan areas. Anenberg and Kung 
[@anenberg_information_2015] also identified a complementary relation between 
the internet and consumption variety in cities by focusing on food truck industry 
in the US.

\color{black}
Thus, clearly, previous research suggests that the declaration of the “death of distance” has  proven to be premature [@rietveld_transport_2004].
However, the exact impact of digital communications on spatial structure is still an open question.
As Leamer and Storper [@edward_e_leamer_economic_2001] indicate, the internet can affect both centripetal and centrifugal forces.
The only cross country study [@ioannides2008effect] supports a clear substitution effect, whereas most of the above studies have focused either on the US or on some specific cities.
Moreover, most of the above studies examined the complementarity/substitutability question over time periods when the internet and other digital communication technologies were still emerging.
For instance, internet penetration in the US in $2000$, which was the focus for quite a few of the above studies including Ioannides et al. [@ioannides2008effect], was just above $50$ percent, while in $2016$ it reached almost $90$ percent [@pew_internet_pew_2016].
At a global scale internet penetration increased from $7$ to $46$ percent during the same period [@internet_live_stats_internet_2017].
In addition, although email and instant messaging technologies were widespread in the developed world in early $2000s$, network externalities due to mobile internet and online social media were nowhere close to what we are familiar with today.
For instance, Facebook users increased from $1$ million in $2004$ to more than $1.5$ billion in $2015$^[[http://newsroom.fb.com/company-info/](http://newsroom.fb.com/company-info/)].
Hence, it might have been premature for the spatial economic effects of ICT adoption to have been materialized by the time that most of the above studies were conducted.

Most recently, Tranos and Ioannides [@tranos2020ict] return to the setting of Ioannides et al. [@ioannides2008effect], update the data employed and obtain results that confirm their original findings.
That is, the diffusion of fixed telephony has caused more dispersed urban structures worldwide, in other words, greater urban decentralization.
Similar causal effects are established for mobile telephony, which are novel in relation to Ioannides et al. [@ioannides2008effect], who lacked such data, and for the internet, which extends their earlier findings.
The robustness of their results is confirmed for such alternative measures of dispersion as the Gini coefficient, the Herfindahl index, and the coefficient of variation.
This is notable because several years of additional data were used that pertain to an era of rapid expansion of the internet and web-based technologies, more generally.

# Materials and Methods {#sec3}

The main aim of this paper is to return to the estimation of the causal impact of ICT adoption on the spatial dispersion of economic activities and consequently population.
It adopts an approach that is carried out at several spatial scales and uses different data than those used by Ioannides et al. [@ioannides2008effect] and Tranos and Ioannides [@tranos2020ict].

We start with a multi-country exercise, which includes both developed and developing countries.
[Section 3.1](#sec3.1) discusses the methods and the different data we use and [Section 4](#sec4) reports the results.
Because of limitations related with multi-country urban population data (see discussion below), we complement our analysis with two case studies for the US and the UK urban system for which we have access to much more granular data.
[Section 3.2](#sec3.2) discusses the methods and [Section 5](#sec5) reports the results.

## Multi-country analysis {#sec3.1}

The multi-country identification strategy is a two-step approach [@ioannides2008effect].
The first step of our methodology is to estimate the Zipf coefficient for a broad sample of countries over time.
Zipf coefficient is one of the most widely used measures of spatial dispersion with numerous applications in urban economics and economic geography ([see for example @black_urban_2003; @frenken_theoretical_2007; @giesen_zipfs_2011; @rauch_cities_2013; @ioannides_zipfs_2003; @ioannides_walled_2017; @nitsch_zipf_2005).
This an appropriate measure of dispersion because of the extreme heterogeneity of the city size distribution and the very good fits normally obtained with such estimations. Skipping details pertaining of the suitability of the approach, which may be found in Ioannides et al. [@ioannides2008effect] and Tranos and Ioannides [@tranos2020ict], we proceed with the estimation of the logarithmic form of:

\begin{align}
ln(rank_{i}) = ln(S_0) + \zeta ln(Size_i) + e_{i}. \label{eq1}
\end{align}

This leads to an estimate of $\zeta$, known as the Zipf coefficient, but more appropriately the Pareto exponent, when $\zeta$ is not estimated to be near $1$.
While much of the literature focuses on the Zipf coefficient being close to $1$, several estimations of city size distributions lead to estimates  of $\zeta$ that are not necessarily equal to $1$, or even near it, in which case we refer to $\zeta$ as the Pareto, or power law, exponent. Given that our aim here is to estimate $\zeta$ for a number of countries over time as a measure of dispersion, equation \ref{eq1} describes the rank of city $i$ in country $c$ in year $t$:

\begin{align}
ln(rank_{ict}) = ln(S_{0ct}) + \zeta_{ct} ln(Size_{ict}) + e_{ict}. \label{eq2}
\end{align}

The estimation of equation \ref{eq2} has typically been performed by Ordinary Least Squares (OLS).
Gabaix and Ioannides [@gabaix_evolution_2004] discuss the downward bias of estimates of equation \ref{eq2} using OLS on small samples.
Gabaix and Ibragimov [@gabaix_rank_2011] propose a practical remedy to correct this bias, which we do adopt in this paper:
instead of using the $log$ of $rank$ of a city $i$ in a country $c$ in year $t$, they propose to use the $log(rank-0.5)$, which has indeed been widely adopted.
Researchers working in this area must contend with definitional differences as well as differences in availability of different kinds of data sources.
Definitions of cities differ across countries, for political, administrative and legal reasons^[See [@ioannides2013neighborhoods], Ch. 8, for issues and pitfalls associated with different definitions.].
In contrast to the data drawn from Thomas Brinkhoff’s City Population project [@thomas_brinkhoff_city_2014]^[[http://www.citypopulation.de](http://www.citypopulation.de)], which were  used by Soo [@soo_zipfs_2005], Ioannides et al. [@ioannides2008effect] and Tranos and Ioannides [@tranos2020ict], the multi-country analysis in the present paper is based on the annual population data for urban agglomerations with $300,000$ inhabitants or more from the Department of Economic and Social Affairs of the United Nations [@united_nations_world_2014].
Despite some criticism about the consistency of the urban agglomeration definitions across different countries [@cohen_urban_2004], this is the only available source for yearly, multi-country population data for urban agglomerations [@montgomery_urban_2008; @chen_absolute_2007; @decker_global_2007].
The results based on these data are reported and discussed in [Section 4](#sec4).

Table \ref{zipf} presents the estimated $\zeta$ coefficients for the panel of countries that the second stage of the analysis focuses on using the UN urban agglomerations data.
It becomes evident that there is considerable variation in the estimates of $\zeta$ across the different countries in the sample.
Because of the empirically established heavy upper tail of data for cities and urban agglomerations, the $\zeta$ coefficient constitutes a convenient measure of dispersion.
The larger its absolute value, the thinner the upper tail; equivalently, the larger is the coefficient algebraically, the heavier the upper tail.
This key observation is basis for the second step of our methodology.

```{r, echo=FALSE, warning=FALSE, results='asis'}

load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/UN.RData")

zipf.table <- net.model$model %>%
  dplyr::select(`as.factor(country)`, `as.factor(year)`, b) %>%
  spread(`as.factor(year)`, b) %>%
  rename(`Zipf coef.` = `as.factor(country)`)

# stargazer(zipf.table, type = "latex", digits = 3, summary = FALSE,
#           header = FALSE,
#           no.space=TRUE,
#           title = "Zipf coefficients\\label{zipf}",
#           font.size = "small",
#           column.sep.width = "1pt",
#           float.env = "sidewaystable")

kable(zipf.table, format = "latex",
      caption = "Zipf coefficients\\label{zipf}",
      booktabs = T,
      digits = 2) %>%
  footnote(general = "Corrected as per Gabaix and Ibragimov (2011)") %>%
  kable_styling(latex_options = "scale_down",
                position = "left")
```

The second step of our methodology involves estimating the following empirical model [@ioannides2008effect]:

\color{blue}

\begin{align}
\zeta_{ct} = \theta_{c} + \delta_{t} + X_{ct} + e_{ct}. \label{eq3}
\end{align}

\color{black}
\noindent The model enable us to estimate the effects of a vector $X$ of explanatory variables which are included to account for the spatial structure of country $c$ in year $t$.
The main variables of interest here are internet and digital communications variables that include: internet users per $100$ inhabitants, broadband users per $100$ inhabitants, mobile phone users per $100$ inhabitants, and fixed phone users per $100$ inhabitants.
To address a potential omitted variable bias equation \ref{eq3} includes 
\color{blue}
country ($\theta_c$) and year ($\delta_t$) fixed effects;
\color{black}
$e_{ct}$ is the error term.
In addition, vector $X$ includes several control variables, whose descriptive statistics together with those for the other variables used to estimate \ref{eq3} are reported in Table \ref{desc.global}.

Referring to the control variables, total country population is an important measure of size, GDP per capita, and GDP growth are intimately related to urbanization and so are population density, and non-agricultural value-added as a share of GDP.
Trade, that is conventionally measured as exports plus imports as a share of GDP, is an important time varying measure of trade openness. Government expenditure as a share of GDP may be a proxy of public investment in some countries and government waste in others.

\color{blue}

As Table \ref{cor} indicates, although there some rather strong correlations among these variables, mobile phone penetration appears to have a distinct character from its fixed phone counterpart: their correlation coefficient is only $0.445$.
This probably highlights the different composition of the population or infrastructure development patterns in the developing world, where mobile telephony helped overcome the lack of fixed line infrastructure and mobile phone networks are also used as the main way to use the internet [@donner_research_2008; @hamilton_are_2003].

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 2

stargazer(as.data.frame(data[,c(5:18)]),
          type = "latex",
          digits = 1,
          header = FALSE,
          no.space=TRUE,
          title = "Descriptive statistics for the multi-country model\\label{desc.global}",
          font.size = "footnotesize",
          column.sep.width = "1pt",
          summary.stat = c("n", "min", "mean", "sd", "max"),
          covariate.labels=c("Zipf coefficient",
                             "Zipf inv. sq. error\\textsuperscript{a}",
                             "Population density",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Female labour force (\\%)",
                             "Population",
                             "Non agriculture value added (\\% GDP)",
                             "Internet users per 100 hab.",
                             "Broadband users per 100 hab.",
                             "Mobile phone users per 100 hab.",
                             "Fixed phone users per 100 hab."),
          notes = c("\\textsuperscript{a}This is the inverse squared standard error of the estimated Zipf, coefficient, which is used for weighting", 
                    "the observations for the estimation, of equation 3. See Section 4"))
```

\color{black}

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 3

corr.mat <- data %>%
  dplyr::select(net100hab, bb100hab, mobile100hab, fixed100hab)

corr.mat <- cor(corr.mat, use = "na.or.complete")
row.names(corr.mat) <- c("Internet users per 100 hab.",
                             "Broadband users per 100 hab.",
                             "Mobile phone users per 100 hab.",
                             "Fixed phone users per 100 hab.")
colnames(corr.mat) <- c("Internet",
                             "Broadband",
                             "Mobile",
                             "Fixed")

kable(corr.mat, format = "latex",
      caption = "Correlations between ICT variables\\label{cor}",
      booktabs = T,
      digits = 2) %>%
  kable_styling(latex_options = "scale_down",
                position = "left",
                font_size = 5)
```

The availability of a panel dataset for city sizes across countries enables us to use country fixed effects, which can address potential endogeneity issues related to unobserved country specific characteristics of city size distributions.
However, such a strategy does not address potential simultaneity issues.
Simply put, internet penetration might be affected by spatial structure, as reflected in Zipf coefficients, or both internet penetration and spatial structure might be jointly determined by a third variable.
E.g., if a country already has a dispersed spatial structure, internet is particularly suitable in facilitating communication.
Potential endogeneity in our specification will prevent us from being able to determine the causal impact of internet and digital communication technologies usage on spatial structure, which is the main aim of this paper. In order to address this problem, we will adopt an instrumented variable strategy.
Table \ref{desc.global} also includes the descriptive statistics for the instrumental variable we are using and ($female\:labour\:force$), will be discussed in [Section 4](#sec4).

## 3.2	Case study approach {#sec3.2}

The above global level analysis is followed by two case studies.
They allow us to examine the potential effects of the internet on two mature urban systems in greater detail and depth and without the exogenously imposed threshold of the $300,000$ habitants that the global level analysis affords us.
We focus on the US and the UK, for which we are able to utilize more granular internet-related data (see [Section 5](#sec5) for the data description).
Given that we are dealing with specific countries and not with a panel of countries we cannot employ an identification strategy similar to that of the global level analysis discussed in the previous section.
Therefore, we propose a one-step approach involving estimation of the following empirical cross-sectional model:

\begin{align}
Diff\:in\:ranks_{i} = a + \beta ICT_{i} + B C_{i} + e_{i}.\label{eq4}
\end{align}

In order to capture the micro-dynamics of the urban systems in the two case studies, we follow Batty [@batty_rank_2006] and Havlin [@havlin_distance_1995] and focus on the difference in ranks for individual cities during the study period ^[Changes in ranking of individuals has also been used elsewhere in economics because of its robustness properties. C.f. Chetty and Hendren [@chetty2018impacts].].
Contrary to their approach, we are interested in the real difference in rankings rather than in absolute differences, in order to capture whether or a city improves its ranking within the urban system during the study period.
We then test whether our internet-related variable has an effect on the change in ranking.
Hence, we define the left-hand side (LHS) variable of equation \ref{eq4} as follows:

\begin{align}
Diff\:in\:ranks_{i} = r_{i(t-1)} - r_{it},\label{eq5}
\end{align}

\noindent where $r_{it}$ is the population rank of city $i$ in year $t$. A negative (positive) value for the $Diff\:in\:ranks_{i}$ variable indicates that city’s $i$ position in the urban hierarchy of the country worsens (improves) in relative terms, also due  to the population changes of the other cities of the urban system.
Notably, this variable does not only consider the population change of a specific city, but it also considers the overall urban system dynamics by focusing on the rank and not on population per se.
Given that the data we use and the definitions of cities vary between the UK and the US we are going to discuss these data in the relevant sections.
What we highlight here is that the estimation of equation \ref{eq4}, just like equation \ref{eq3}, might suffer from endogeneity and therefore an instrumental variables strategy is employed in order to address this issue.

# Digital technologies and spatial structure: a global view {#sec4}

This section presents the estimation results of \ref{eq3}.
The LHS variable is $\zeta$, the Zipf coefficient, as estimated according to the Gabaix and Ibragimov [@gabaix_rank_2011] correction.
The main explanatory variables of interest are provided at the top of Table \ref{ols.global}; namely internet, broadband, mobile and fixed telephony per $100$ habitants,  expressed in natural logarithms.
The main variables of interest are introduced successively on their own in the regressions reported in Table \ref{ols.global}.
All regressions include country fixed effects to control for unobserved heterogeneity and a time trend.
In addition, the observations are weighted with the inverse squared standard error of the estimated Zipf coefficient (see Table \ref{desc.global}) to address potential noise that is carried over from the first part of our identification strategy.
In regards to the interpretation of the estimated coefficients, given that the Zipf coefficient has entered the regression not as an absolute value, but instead as a real number a positive coefficient for a RHS variable indicates an impact towards the decrease of the spatial dispersion of population. In other words, a positive coefficient indicates an effect towards less uniform city sizes that is more dispersion of city sizes.
The latter is indicative of enhancement of agglomeration economies because of the expansion of digital technologies.

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 4

net.rse <- coeftest(net.model, function(x) vcovHC(x, type="HC0"))
bb.rse <- coeftest(bb.model, function(x) vcovHC(x, type="HC0"))
fixed.rse <- coeftest(fixed.model, function(x) vcovHC(x, type="HC0"))
mobile.rse <- coeftest(mobile.model, function(x) vcovHC(x, type="HC0"))

stargazer(net.model, bb.model, fixed.model, mobile.model, type="latex",
          header = FALSE,
          no.space=TRUE,
          title = "OLS estimation of equation (3)\\label{ols.global}",
          font.size = "small",
          column.sep.width = "1pt",
          se = list(net.rse[,"Std. Error"], bb.rse[,"Std. Error"], fixed.rse[,"Std. Error"], mobile.rse[, "Std. Error"]),
          dep.var.labels=c("Zipf Coefficient 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          add.lines = list(c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          omit.stat = c("rsq", "f"),
          notes = "Robust Std. Errors in parenthesis") #, corrected Zipf coefficient following Gabaix and Ibragimov (2011)")
```

Table \ref{ols.global} and \ref{2sls.global} report the estimation results based on the UN urban agglomeration data.
Starting with the former, we note that an agglomerative effect is only detected for the internet users and while a marginally opposing result can be seen for broadband users as indicated by the significant coefficients in columns (1) and (2).
For the telephony variables the estimation of \ref{eq3} did not yield statistically significant coefficients.
Before discussing further these results and the effects of the other control variables, we need to highlight that the main challenge of estimating equation \ref{eq3} is the potential endogenous nature of the share of internet users which might prevent us from being able to infer a truly causal effect.
Endogeneity might be an issue here as spatial structure, which is proxied by the Zipf coefficient, might be affected by another source, which also affects internet penetration.
For instance, economic development might affect the concentration of population in large cities and at the same time enable more people to go online.
If we do not address this issue, the coefficient for the main variable of interest will capture potential effects that internet penetration has on spatial structure, but also potential reverse causality effects that spatial structure might generate on internet penetration.
To overcome this potential problem, Table \ref{2sls.global} reports estimates of equation \ref{eq3} using two-stage least squares (2SLS) with instrumental variables (IVs).
The latter are variables which are correlated with our endogenous variables, but do not influence current spatial structure.
Such an approach will enable us to estimate the causal effect -- if any -- of the internet and digital communications on spatial structure.
At a first stage, our endogenous variables are regressed against the IVs.
Then, the predicted values of the endogenous variables based on the IVs and the other control variables are used instead of the endogenous variable to estimate eq. \ref{eq3}.
A significant effect will verify the causal impact of the internet and digital communication usage on spatial structure.

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 5

net.2sls.rse <- coeftest(net.2sls, function(x) vcovHC(x, type="HC0"))
summ.net.2sls <- summary(net.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

bb.2sls.rse <- coeftest(bb.2sls, function(x) vcovHC(x, type="HC0"))
summ.bb.2sls <- summary(bb.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

fixed.2sls.rse <- coeftest(fixed.2sls, function(x) vcovHC(x, type="HC0"))
summ.fixed.2sls <- summary(fixed.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

mobile.2sls.rse <- coeftest(mobile.2sls, function(x) vcovHC(x, type="HC0"))
summ.mobile.2sls <- summary(mobile.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

stargazer(net.2sls, bb.2sls, fixed.2sls, mobile.2sls, type = "latex",
          se = list(net.2sls.rse[,"Std. Error"], bb.2sls.rse[,"Std. Error"],
                    fixed.2sls.rse[,"Std. Error"], mobile.2sls.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (3)\\label{2sls.global}",
          font.size = "small",
          column.sep.width = "1pt",
          dep.var.labels=c("Zipf Coefficient 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(c(rownames(summ.net.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.net.2sls$diagnostics)[2],
                             round(summ.net.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.net.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis",
                    "IV: Female participation in labour force"))
```

The main challenge for such an exercise is to find a valid IV.
In our case, the challenge is even bigger as we need to find a variable which is not only fit for purpose, but it also contain as few missing values as possible in order to retain the same number of observations as the OLS models.
This is particularly difficult for a diverse sample of countries as the one we are dealing with here.
We propose here the percentage of female participation in the labor force, which is directly related to digital infrastructure, but not to spatial structure (see Table \ref{desc.global} for the descriptive statistics).
\color{blue}
The literature provides empirical and theoretical evidence that ICT are correlated with increase in female participation in labour markets mostly because they enable teleworking [@dettling2017; @arntz2019working; @watson2018digitalisation; @Lechman2020; @Valberg2020]
\color{black}
At the same time, female labor participation is not related with spatial structure especially given the diverse sample of countries.
<!-- Firstly, the correlation coefficient between the IV and the Zipf coefficient is very low ($-0.14$). -->
Although one might think that increased female labor participation may lead to relocation of households to large cities and therefore affect the spatial structure of cites, such a process varies a lot within our dataset.
Even if this might be true for developed countries with mature urban systems for which a location within a large urban center might provide opportunities for both male and female workers to find jobs, our dataset also includes countries, $41$ per cent of the GDP of which is attributed to primary activities (see Table \ref{desc.global}).
So, female labor participation might also be related to economic activities located outside large urban centers.
In terms of the relevant tests presented in Table \ref{2sls.global} the weak identification first stage F-test exceeds by far all the Stock-Yogo critical values for all, but the mobile telephony regression.
Hence, we are confident to interpret the effects of internet, broadband internet and fixed telephony usage.

The estimations presented in Table \ref{2sls.global} indicate that increases in internet and broadband internet usage have resulted to a decrease of the spatial dispersion of population for the time period and for the panel of countries included in our data.
In other words, increases in internet and digital communications resulted to national urban systems which are less uniform in terms of city sizes and are characterized by higher population concentrations in larger cities.
The exact opposite appears to happen for fixed telephony, which our results indicate that has led to an increase of spatial dispersion of population.
Importantly, our identification strategy enables us to address potential reverse causality issues and treat the results of Table \ref{2sls.global} as causal. Hence, the main finding of our multi-country, global analysis based on data for urban agglomerations of at least 300,000 inhabitants is that internet and broadband internet usage appear to act in favor of agglomeration economies and result to urban systems with more dominant cities on the top of the urban hierarchy.
On the contrary, fixed telephony usage acted in favour of centrifugal forces.
Our identification strategy does not enable us to make similar interpretations for mobile telephony usage.

Regarding the control variables, only a few of them have significant effects on spatial structure probably because the fixed effects estimation masks the between-country variation.
The effects of these control variables are in agreement with previous research (Ioannides et al. 2008).
Namely, GDP per capita has a negative effect which indicates that wealthier countries tend to have more balanced urban systems, while the opposite seems to happen for countries experiencing a growing economy.

In total, the results indicate that our measures of internet and telephony penetration have further enhanced agglomeration forces, at least for large urban agglomerations, while the opposite happens for fixed telephony.
These results are robust against endogeneity issues, but are limited to urban agglomeration included in our data.
Indeed, for the estimation of the Zipf coefficient we only included urban agglomeration of $300,000$ inhabitants or more due to data availability.
Hence, the above estimations cannot verify whether such an effect is also valid for smaller cities.
Interestingly, the above results regarding the role of fixed telephony are in agreement with previous results from [@ioannides2008effect] [@tranos2020ict] which used a smaller panel of countries but without constraints regarding the city size.
However, our results do not support the findings of [@tranos2020ict] regarding internet usage.
In order to overcome the city size limitation, the next section presents two case studies, for which we have obtained much more granular data and therefore we are in a position to test the effect of the internet on the tail of the urban population distribution for these countries.

# The impact of ICT on the US and the UK spatial structures {#sec5}

This section focuses on the US and the UK urban systems, for the cities of which equation \ref{eq4} from [Section 3.2](#sec3.2) is estimated separately.

## Internet and the US Spatial Structure: Evidence from the US Micropolitan and Metropolitan Statistical Areas, 2013-2018 {#sec5.1}

We pursue further our investigation of the impact of internet adoption by using a previously unutilized, to the best of our knowledge, for this purpose data source.
That is, for the first time in $2013$, data on internet use was made available via the American Community Survey and is provided at the metropolitan area (urban areas comprised of one or more adjacent counties or county equivalents that have at least one urban core area of at least $50,000$ population, plus adjacent territory that has a high degree of social and economic integration with the core as measured by commuting ties) and at the micropolitan area (defined like metropolitan areas except that are comprised of an urban core of at least $10,000$, but less than $50,000$ population) level of aggregation [@file_computer_2014].
These functional definitions of a city represent in essence labor markets and according to Table \ref{desc.usa} the observable minimum size of city population used for this analysis is just above $62,000$ habitants.

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 6
rm(list=ls())
load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/USA.RData")

stargazer(as.data.frame(data.for.decr),
          type="latex",
          summary = T,
          summary.stat = c("n", "min", "mean", "sd", "max"),
          header = FALSE,
          no.space=TRUE,
          title = "Descriptive statistics for the USA model\\label{desc.usa}",
          font.size = "small",
          single.row = FALSE,
          column.sep.width = "1pt",
          covariate.labels=c("normalized difference in ranks, 2013-18",
                             "households w. internet 2013, \\%",
                             "population 2013",
                             "\\% of unemployment 2013",
                             "\\% of white population 2013",
                             "income 2013",
                             "population density 2013",
                             "employment in service 2011, \\%",
                             "commute in minutes 2013",
                             "pop. above 25 w. Bachelor's \\\\
                             degree 2005, \\%",
                             "commute in minutes, 2005"))
```

The LHS variable we use here is the outcome of equation \ref{eq5} in a normalized form^[$$Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.$$]
, so as that variable is bounded between $0$ and $1$ and with a mean of $0.5$.
This transformation enables us to estimate \ref{eq4} not only with OLS, but also with a quasibinomial GLM estimator given that the original form of our LHS variable does not vary continuously, but is instead defined as a difference between two count variables, which may also assume negative values.
The results of the different estimations remain qualitatively the same regardless of the form of the LHS variable we use.
Table \ref{ols.usa} reports the results of the estimation of equation \ref{eq4} using OLS.
The sign and significance of the main variable of interest -- percentage of population with computer and broadband connection -- verifies our global level results.
That is an increase in internet usage improves the position of a city in the US urban system.
The results are consistent among the different estimators that are OLS and GLM.
In addition, Table \ref{ols.usa} also reports the estimates of interaction effects between the share of population with broadband connection with population and population density.
Although the interaction terms are not significant, the main effect remains qualitatively unchanged.

```{r, echo=FALSE, warning=FALSE, results='asis'}

# Table 7
rob.base        <- coeftest(base, function(x) vcovHC(x, type="HC0"))
rob.int_density <- coeftest(int_density, function(x) vcovHC(x, type="HC0"))
rob.int_pop <- coeftest(int_pop, function(x) vcovHC(x, type="HC0"))
rob.base_glm <- coeftest(base_glm, function(x) vcovHC(x, type="HC0"))

stargazer(base, int_density, int_pop, base_glm, type="latex",
          se = list(rob.base[,"Std. Error"], rob.int_density[,"Std. Error"], rob.int_pop[,"Std. Error"], rob.base_glm[, "Std. Error"]),
          summary = F,
          header = FALSE,
          no.space=TRUE,
          title = "OLS estimation of equation (4) for USA\\label{ols.usa}",
          font.size = "small",
          column.sep.width = "1pt",
          column.labels   = c("OLS", "GLM"),
          column.separate = c(3, 1),
          model.names = FALSE,
          dep.var.labels=c("Normalized difference in ranks 2013-18"),
          covariate.labels=c("\\% of households w. internet 2013",
                             "population 2013 (log)",
                             "\\% of unemployment 2013",
                             "\\% of white population 2013",
                             "income 2013 (log)",
                             "population density 2013",
                             "\\% of employment in service 2011",
                             "commute in minutes 2013",
                             "\\% of households w. internet, \\\\
                             2013 x pop. density, 2013",
                             "\\% of households w. internet, \\\\
                             2013 x population, 2013"),
          single.row = FALSE,
          df = FALSE,
          omit.stat = c("rsq", "f"), notes = "Robust Std. Errors in parenthesis")
```

Although not directly comparable, the above estimations are in accordance with our global model. In addition, the potential presence of endogeneity might be a problem here as well.
Therefore, Table \ref{2sls.usa} reports 2SLS estimations.
In addition to the strategy employed for the cross-country analysis -- that is the inclusion of one instrument for which we have strong reasons to believe is uncorrelated with the error term -- data availability allows to employ a second IV in order to estimate the Sargan over-identification restrictions test (Column 2).
The main instrument we propose here is Bachelor’s degrees per inhabitant in $2005$.
Even if the quality of human capital can affect the population growth of a city $10$ years later, our LHS variable adopts a systemic understanding of the US urban system as it measures the relative position of a city within the overall urban system instead of its population growth.
In other words, a city might experience population growth between two periods, but if other cities have also experienced population growth, this might not affect its relative position within the urban system.
Hence, we do not expect that this IV affects the LHS variable.
<!-- In addition, the correlation coefficient between these two variables is still very low ($r=0.24$). -->
Moreover, when we add the second IV, which in this case is the commute time in minutes also in $2005$, we fail to reject the null hypothesis of the Sargan test, something which also adds on the validity of our strategy.
Furthermore, our results do not suffer from weak identification according to the relevant tests in Table \ref{2sls.usa}.

```{r, echo=FALSE, warning=FALSE, results='asis'}

# Table 8
rob.model_iv1 <- coeftest(model_iv1, function(x) vcovHC(x, type="HC0"))
rob.model_iv2 <- coeftest(model_iv2, function(x) vcovHC(x, type="HC0"))
summ.fit1 <- summary(model_iv1, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
summ.fit2 <- summary(model_iv2, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

stargazer(model_iv1, model_iv2, type = "latex",
          se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]),
          summary = F,
          header = FALSE,
          no.space=TRUE,
          column.separate = c(1,1,1,1),
          title = "2SLS estimation of equation (4) for USA\\label{2sls.usa}",
          font.size = "small",
          column.sep.width = "-15pt",
          align = TRUE,
          dep.var.labels=c("Normalized difference in ranks, 2013-18"),
          covariate.labels=c("\\% of households w. internet 2013",
                             "population 2013 (log)",
                             "\\% of unemployment 2013",
                             "\\% of white population 2013",
                             "income 2013 (log)",
                             "population density 2013",
                             "\\% of employment in service 2011",
                             "commute in minutes 2013"),
          single.row = FALSE, df = FALSE,
          omit.stat = c("rsq", "f"), notes = c("Robust Std. Errors in parenthesis",
                                               "IVs: (1) Bachelors degree per hab. in 2005",
                                               "IVs: (2) Bachelors degree per hab. in 2005, Commute, minutes, 2005"),
          add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.fit1$diagnostics[1, "statistic"], 2),
                             round(summ.fit2$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.fit1$diagnostics)[2],
                             round(summ.fit1$diagnostics[2, "statistic"], 2),
                             round(summ.fit2$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[2, "p-value"], 2),
                             round(summ.fit2$diagnostics[2, "p-value"], 2)),
                           c(rownames(summ.fit1$diagnostics)[3],
                             round(summ.fit1$diagnostics[3, "statistic"], 2),
                             round(summ.fit2$diagnostics[3, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[3, "p-value"], 2),
                             round(summ.fit2$diagnostics[3, "p-value"], 2))
          ))

```

In total, the estimates of Table \ref{2sls.usa} enable us to identify a causal effect of the share of households with internet connection on the position of a city in the urban system.
Specifically, an increase in internet penetration led to an improvement of the position of a city in the US urban hierarchy for the cities and the time period included in our analysis.
Everything else being equal, using the normalization formula for the $difference\:in\:ranks_{i}$, if a city whose position in the US urban system did not change during the period $2013-2018$ (that is $Normalized\:difference\:in\:ranks_{i} = 0.5$) had experienced an increase of 10 per cent points of internet penetration, this increase would have improved its relative position by $9$ places.
As in the global model reported in [Section 4](#sec4), internet penetration appears to work in favor of agglomeration externalities.
In regards to the control variables, we can identify a negative effect of income, which is consistent with the effect of GDP per capita in the global model presented in Table \ref{2sls.global}.
During the study period, unemployment rates and the share of white population negatively affected the ranking of micropolitan and metropolitan areas in the US as reflected in the relevant regressions.

## Internet and the UK Spatial Structure: Evidence from the Built-up Areas in England and Wales, 2011-2018 {#sec5.2}

The next step in our analysis is to estimate equation \ref{eq4} for cities in England and Wales for which we were able to access internet speed micro-data.
More specifically, we obtained individual speed internet tests from [broadbandspeedchecker.co.uk](broadbandspeedchecker.co.uk).
This website enables individuals to directly measure their upload and download internet speed.
The results of the tests as well as the geo-location of the users are recorded by the website operator and were provided to us in a fully anonymized manner. More discussion about the nature and the validity of this data can be found in the work of Riddlesden and Singleton [@riddlesden_broadband_2014].
The point nature of these individual level data enables us to aggregate them up to any urban level that we are interested in.
Given that all the above analyses focused on functional definitions of cities, we adopt here a morphological definition of a city or a town for the UK, which enables us to test the effect of internet on smaller areas irrespective of being part of a wider urban agglomeration.
Therefore we aggregate the internet speed data at the level of Built-up Areas (BUA) for England and Wales.
This is a 'bricks and mortar' approach which refers to land which is "irreversibly urban in character" including villages, towns or cities.
Some key characteristics of these areas include: minimum size of $20$ hectares;
areas with less than $200$ meters between them are linked to a single built-up area;
larger built-up areas are separated to smaller sub-divisions of built-up areas [@ons2013; see also @johnston_changing_2014; @johnston2016ethnic for other research applications of BUA].

In order to obtain information about the tail of the urban size distribution we include, wherever available, the sub-divisions of BUA.
As Table \ref{desc.uk} illustrates, our approach results to $6,163$ observations which include built-up areas and sub-divisions of built-up areas for which we have internet speed data.
The lowest population of a built-up area in our data is just above $100$ inhabitants in $2011$.
Figure \ref{figure1} illustrates the built-up areas we use for the South-East of England and the Greater London Area.

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 9
rm(list=ls())
load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/UK.RData")

data.for.decr <- data %>%
  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011,  
         unemployment.share, british.pop.share, pop.dens,
         work.home.share, services.share, uni.freq) %>%
  mutate(n.tests2011perhab = n.tests2011/pop2011) %>%
  dplyr::select(-n.tests2011) %>%
  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011perhab,  # re-redring
         unemployment.share, british.pop.share, pop.dens,
         work.home.share, services.share, uni.freq)

stargazer(as.data.frame(data.for.decr),
          type="latex",
          summary = T,
          summary.stat = c("n", "min", "mean", "sd", "max"),
          header = FALSE,
          no.space=TRUE,
          title = "Descriptive statistics for the UK model\\label{desc.uk}",
          font.size = "small",
          single.row = FALSE,
          covariate.labels=c("normalized difference in ranks, 2011-18",
                             "download speed, 2011",
                             "population, 2011",
                             "broadband tests per capita, 2011",
                             "\\% of unemployment, 2011",
                             "\\% of British population, 2011",
                             "population density, 2011",
                             "% of people working from home, 2011",
                             "employment in service, 2011 (\\%)",
                             "Number of universities"))

```

```{r, echo=FALSE, out.width="80%", fig.cap="\\label{figure1}Built-up areas in London and the South East of England"}
knitr::include_graphics("Figure1.jpg")
```

In terms of our identification strategy, we follow the same approach as we did for the US case.
Table \ref{ols.uk} presents the OLS and GLM estimation of equation \ref{eq4} while Table \ref{2sls.uk} presents the 2SLS estimation.
Starting from Table \ref{ols.uk} we see a significant and positive effect of the average download speed on the relative position of a BUA in the urban hierarchy in England and Wales.
This is consistent with the previous results from the US and also with the global model.
Of course, the same endogeneity issues might be present here too and we are addressing that below.
In terms of control variables, we include a measure of broadband tests per inhabitant in order to control for potential differences in the take-up of this service.
In addition, we include a number of socio-economic variables that we believe can affect the relative position of a city in the urban hierarchy.
Unemployment rate has a negative and significant effect as in the US case.
The same applies for the percentage of British population, which indicates the importance of migration in urban growth.
Population density and the percentage of people working from home also negatively affected the relative position of BUA during $2011-2018$.
Finally, Table \ref{ols.uk} includes two interaction terms (columns $2$ and $3$) between download speed and population and population density.
The negative and significant sign for both interaction coefficients indicates that the effect of download speed decreases as the size of the BUA or its density increases.
Contrary to the US case, this is indicative of a larger digital connectivity effect for smaller BUA.

```{r, echo=FALSE, warning=FALSE, results='asis'}

#Table 10
#regressions table robust SE
rob.base        <- coeftest(base, function(x) vcovHC(x, type="HC0"))
rob.int_density <- coeftest(int_density, function(x) vcovHC(x, type="HC0"))
rob.int_pop <- coeftest(int_pop, function(x) vcovHC(x, type="HC0"))
rob.base_glm <- coeftest(base_glm, function(x) vcovHC(x, type="HC0"))

stargazer(base, int_density, int_pop, base_glm,
          se = list(rob.base[,"Std. Error"], rob.int_density[,"Std. Error"], rob.int_pop[,"Std. Error"], rob.base_glm[, "Std. Error"]),
          type="latex",
          summary = F,
          header = FALSE,
          no.space=TRUE,
          title = "OLS estimation of equation (4) for the UK\\label{ols.uk}",
          font.size = "small",
          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
          #dep.var.caption = "Normalized difference in ranks, 2011-18",
          column.labels = c("OLS", "GLM"),
          column.separate = c(3,1),
          model.names = FALSE,
          covariate.labels=c("download speed, 2011 (log)",
                             "population, 2011 (log)",  
                             "broadband tests per capita, 2011",
                             "\\% of unemployment, 2011",
                             "\\% of British population, 2011",
                             "population density, 2011",
                             "\\% of people working from home, 2011",
                             "employment in service, 2011 (\\%)",
                             "download speed, 2011 (log) \\\\
                             x pop. density, 2011",
                             "download speed, 2011 (log) \\\\
                             x population, 2011 (log)"),
          single.row = FALSE,
          df = FALSE,
          omit.stat = c("rsq", "f"),
          notes = "Robust Std. Errors")
```

In regards to IV strategy, we use here the number of universities as the main instrument.
Universities in England and Wales have been established long time ago and therefore we do not expect them to affect the change in the relative position of built-up areas in $2011$.
<!-- The correlation coefficient with the LHS variables is smaller than $0.05$. -->
The addition of the second instrument (absolute number of broadband tests in 2011) in column $2$ results to a Sargan test, the null hypothesis of which cannot be rejected.
Moreover, the estimations reported in Table 11 do not suffer from weak instruments as the relevant test is above the rule of thumb value of $10$.
The coefficients derived from OLS (Table \ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are of the same magnitude and always positive and significant, something which advocates towards the effect of broadband speed in enhancing agglomeration forces.
A back of the envelope calculation indicates that, everything else being equal, $1$ per cent increase in download speed for a BUA, the relative position of which did not change during the period $2011-2014$, would have increased its position by $0.3$ places.
In absolute numbers, $1$ per cent increase for the average BUA would mean a marginal change from $13.8$ Mbps to $13.9$ Mbps.

```{r, echo=FALSE, warning=FALSE, results='asis'}

# Table 11
# 2SLS robust SE table
#model_iv1, model_iv2,
rob.model_iv1 <- coeftest(model_iv1, function(x) vcovHC(x, type="HC0"))
rob.model_iv2 <- coeftest(model_iv2, function(x) vcovHC(x, type="HC0"))
summ.fit1 <- summary(model_iv1, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
summ.fit2 <- summary(model_iv2, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

stargazer(model_iv1, model_iv2,
          type="latex",
          summary = F,
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (4) for the UK\\label{2sls.uk}",
          font.size = "small",
          se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]),
          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
          covariate.labels=c("download speed, 2011 (log)",
                             "population, 2011 (log)",  
                             "broadband tests per capita, 2011",
                             "\\% of unemployment, 2011",
                             "\\% of British population, 2011",
                             "population density, 2011",
                             "\\% of people working from home, 2011",
                             "employment in service, 2011 (\\%)"),
          single.row = FALSE,
          df = FALSE,
          omit.stat = c("rsq", "f"), notes = c("Robust Std. Errors", "IVs: (1) N.r of universities",
                                               "IVs: (2) N. of universities, N. of broadband tests, 2011"),
          add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.fit1$diagnostics[1, "statistic"], 2),
                             round(summ.fit2$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.fit1$diagnostics)[2],
                             round(summ.fit1$diagnostics[2, "statistic"], 2),
                             round(summ.fit2$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[2, "p-value"], 2),
                             round(summ.fit2$diagnostics[2, "p-value"], 2)),
                           c(rownames(summ.fit1$diagnostics)[3],
                             round(summ.fit1$diagnostics[3, "statistic"], 2),
                             round(summ.fit2$diagnostics[3, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[3, "p-value"], 2),
                             round(summ.fit2$diagnostics[3, "p-value"], 2))
          ))

```

In total, the results of the UK case study are in accordance with the previous results of the global, multi-country analysis and of the US urban system. Interestingly, the positive effect of internet is still apparent when the analysis adopts a morphological definition of cities and when the emphasis is not on internet penetration, but instead on the quality of internet infrastructure as reflected in download broadband speed.
What is also interesting for the UK case study is that this effect appears to be higher for small and less dense BUA.

# Discussion and conclusions {#sec6}

This paper reports empirical findings on a question at the heart of urban economics and economic geography:
has  the proliferation of information and communication technologies offset the benefits of agglomeration economies and resulted in more dispersed spatial population structures, or has it further reinforced such urban externalities and led to more concentrated spatial structures?
Previous studies have led to contradictory results regarding whether ICT adoption and urban agglomeration externalities are complementary or substitutable.
As Leamer and Storper [@edward_e_leamer_economic_2001] stress, the internet can affect both centripetal and centrifugal forces.
The present study revisits Ioannides et al. [@ioannides2008effect] with a completely open mind in view of the availability of several years of additional data on internet penetration across the world.
It examines the robustness of the findings of Tranos and Ioannides [@tranos2020ict] by employing alternative data sets.
In general, quite a few earlier studies have either been based on assumptions about technological capabilities that might no longer hold today, or use data that do not fully capture  the widespread adoption and maturity of communication technologies that has taken place  since the data coverage of those earlier studies.

We report estimation results using international country-level data on urban agglomerations with more than $300,000$ inhabitants from the UN Urban Settlements data and test the effect of ICT adoption on the Zipf, or Pareto, coefficients for national city size distributions, as measures of dispersion for heavy-tailed data. We report results which are robust to endogeneity concerns.
Then, in order to examine such effects for cities that are smaller than those included in our international data, we focus on the US and the UK urban systems.
Specifically, we test the effect of internet usage and internet speed on the changes over time in rankings of Micropolitan  and Metropolitan Areas in the US, and of Built Up Areas in the UK.
The results favor a complementary relation between the internet and agglomeration externalities.
While the cross-country estimates indicate that increase adoption rates of such technologies has resulted in less dispersed urban spatial structures, the two case studies reveal that internet adoption and internet speed improved the relative population rank of a city within its urban system.
Interestingly, the latter results indicate that such effects might be even stronger for smaller and less dense urban areas at least in the UK.
On the contrary, our estimates regarding the effect of fixed telephony advocate towards the reverse effect: that is fixed telephony acting against of agglomeration forces.

While our results are in agreement with Ioannides et al. [@ioannides2008effect] regarding fixed telephony, they contradict Tranos and Ioannides [@tranos2020ict]
The latter cannot of course be ignored.
We believe strongly that the differences lie in the different definitions of urban areas.
To this date, and in spite of numerous efforts by teams worldwide, there has not been a universally accepted definition, which could moreover be implemented consistently by governments worldwide^[
The increasingly popular reliance on high-resolution, pioneered by Rozenfeld et al. [@rozenfeld2011area],  and satellite-based lights data, most recently used by Duben and Krause [@duben2019population], for measuring city sizes, are very exciting developments, but they have not yet been accompanied by complementary information on ICT adoption at a comparably granular level.
A promising new development is the EU’s Global Human Settlements Initiative and the associated spatial urban data sets; see Pesaresi et al. [@pesaresi2019ghs].
].
Barring the availability of such data, scholars must be careful in drawing conclusions.
Arguably, the measurement of the impact of ICT at the level of agglomerations, on one hand, and on changes in rankings, on the other, reported by this paper constitute a contribution to the literature.

We believe that our results, apart from their theoretical value, have the capacity to inform urban policy.
The ability of the internet and digital communications to further enhance agglomeration economies can be used as a tool to support urban growth.
In addition, the indication that such effects might be stronger for smaller and less dense urban areas, at least in England and Wales, might be helpful to further orient digital strategies towards such locations.
Of course, improvements in internet speed is not a trivial policy instrument as it involves numerous complexities.
Apart from infrastructure installation costs and engineering challenges, governance issues regarding the ownership of such digital networks as well as the provision of state subsidies create obstacles for the inclusion of such strategies in the urban growth agenda.
To further inform such policies, more research at granular scales is needed in order to shed lights on the micro-mechanisms behind such urban processes.

# References {#references .unnumbered}

### 2020-11-19 paper/2_resubmission/tranos_ioannides.tex
paper on git

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

\usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{University of Bristol and The Alan Turing Institute,Tufts University}
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ubiquitous Digital Technologies and Spatial Structure; an update} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Emmanouil Tranos\textsuperscript{\getAff{University of Bristol and The Alan Turing Institute}}\textsuperscript{*},
Yannis M. Ioannides\textsuperscript{\getAff{Tufts University}}\\
\bigskip
\textbf{\getAff{University of Bristol and The Alan Turing Institute}}University of Bristol, School of Geographical Sciences, Bristol, UK; The
Alan Turing Institute, London, UK\\
\textbf{\getAff{Tufts University}}Tufts University, Department of Economics, Medford, MA, USA\\
\bigskip
* Corresponding author: e.tranos@bristol.ac.uk\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
This paper examines the impact of widespread adoption of information and
communication technologies (ICT) on urban structure worldwide. Has it
offset agglomeration benefits and led to more dispersed spatial
structures, or has it strengthened urban externalities and thus resulted
in more concentrated spatial structures? Theoretical and empirical
studies on this question have produced contradictory findings. The
present study recognizes that assumptions made earlier about the
evolution of technological capabilities do not necessarily hold today.
As cutting-edge digital technologies have matured considerably, a fresh
look at this question is called for.

The paper addresses this issue by means of several data sets using
instrumental variable methods. One is the UN data on Urban Settlements
with more than \(300,000\) inhabitants. Estimation methods with these
data show that increased adoption of ICT has resulted in national urban
systems that are less uniform in terms of city sizes and are
characterized by higher population concentrations in larger cities, when
concentration is proxied the Pareto (Zipf) coefficient for national city
size distributions. Two, is disaggregated data for the urban systems of
the US, defined as Micropolitan and Metropolitan Areas, and for the UK,
defined as Built-up Areas in England and Wales, respectively. These data
allow for the impacts to be studied for cities smaller than those
included in the cross-country data. Increased internet usage improved a
city's ranking in the US urban system. Similarly, increased download
speed improves a built-up area's ranking in England and Wales.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
TO ADD

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\hypertarget{sec:1}{%
\section{Introduction}\label{sec:1}}

Geographers, planners and urban economists spent effort in exploring the
spatial footprint of the internet even at its early stages. They
theorized about the spatial impacts that rapid internet penetration
might generate on individual cities and the national spatial structure.
The outcome of various such efforts was rather conjectural and even
fanciful and not data-based. Cases in point are celebrations of the
emergence of telecottages {[}1{]}, the rise of a borderless world
{[}2{]}, the death of cities {[}3,4{]}, and, more generally, the end of
geography {[}5{]}, the death of distance {[}6{]} and the emergence of a
new flat world {[}7{]}. Today, \(25\) years after the commercialization
of the internet {[}8{]}, we know that the above narratives overstated
the potential of the internet and other digital communication
technologies to supplement face-to-face interactions, diminish the cost
of distance and, indeed weaken agglomeration economies. The high and
steadily increasing urbanization rates {[}9{]} falsify such predictions.

The adoption rate and the pervasiveness of new internet-based
information and communication technologies, ICT for short, such as
online social media and mobile-telephony hosted internet, which have
increased rapidly during the last 10-15 years in the developing world,
too, raise questions about how exactly ICT might have affected
agglomeration economies. Conflicting technology examples can be
illustrated. On the one hand, despite the broader agreement that no
digital technology can reach the content richness of face-to-face
communications, empirical research from the management field suggests
that current digital technologies can effectively facilitate the sharing
of knowledge with low to medium tacitness and even support knowledge
sharing of a high degree of tacitness {[}10{]}. On the other hand, the
very same technologies can further enhance what Storper and Venables
{[}11{]} termed as buzz, as the constant publication of our personal and
professional updates and whereabouts enabled by ICT can directly
facilitate deliberate as well as unplanned or even unintentional
face-to-face meetings {[}12{]}.

Urban economics suggests that a key source of agglomeration
externalities is proximity, which facilitates interaction and knowledge
spillovers {[}13{]}. Hence, face-to-face interactions and the implied
knowledge spillovers within cities flourish because of lower
transportation and interaction costs {[}11{]}. However, ICT have the
capacity to directly affect this process by further reducing
transportation cost\footnote{Or, more broader, spatial transmission
  costs to follow McCann's terminology {[}14{]}} {[}14{]}. In essence,
the internet and digital communications promote further learning and
matching, which are at the heart of the micro-foundations of
agglomeration economies identified by Duranton and Puga {[}15{]}.
Web-based applications such as Massive Open Online Courses decrease the
need for co-location of actors in order to participate in formal
learning activities. Furthermore, online social media such as
LinkedIn\footnote{LinkedIn has been identified as ``the largest
  professional matchmaker site in the world'' {[}16{]}, p.~207. See also
  {[}17{]}.} can enhance the probability of matching and indeed the
quality of matches even within cities. Therefore, it is natural to ask
whether the widespread adoption of ICT could offset the benefits of
immediate physical proximity and result in more dispersed spatial
structures, or further reinforce agglomeration externalities and lead to
more concentrated spatial structures.

This paper contributes to the above discussion by presenting empirical
research on whether the internet and digital communications have
affected spatial structure, as seen via the size distribution of cities.
Contrary to most of the previous empirical studies, which are reviewed
in the next section, this paper returns first to an empirical setting
similar to that of Ioannides et al.~{[}18{]} in order to re-examine
their results by employing updated data on internet use, associated with
the enhanced technological maturity of ICT, and using a number of
different data sets. In contrast to Tranos and Ioannides {[}19{]}, which
revisits the question by seeking to replicate Ioannides et al.~{[}18{]},
but with the same but augmented data as that earlier study, the present
paper probes the significance of different levels of aggregation for the
relation between agglomeration externalities and digital communications
by employing different scales of spatial aggregation. One is a global
multi-country analysis using urban agglomeration data for many
countries; a second is a more granular analysis for the US and the UK.
For those two countries the paper brings novel data to bear on the
question. In addition, the paper also distinguishes the effects of a
broader range of ICT on spatial structure, including internet and
broadband internet as well as mobile and landline telephony adoption
rates.

Interestingly, in contrast to Tranos and Ioannides {[}19{]}, most of the
differences in the results -- from the global level analysis to the case
studies --- support a complementarity argument. Specifically, the paper
examines econometrically whether spatial structures have been affected
by the adoption rates of the different digital communication
technologies described above. The results are robust against potential
endogeneity concerns, as one might claim that the take-up of these
technologies could have been affected by spatial structures themselves.
That is, individuals living in more dispersed spatial systems might have
made greater use of such technologies in order to overcome the lower
level of agglomeration externalities. Notably, we find some evidence
that such effects are stronger in smaller urban areas. Our findings can
directly inform the urban policy agenda as they advocate in favor of
including digital strategies in policies aiming to enhance agglomeration
externalities and to improve the position of a city within its national
urban hierarchy.

The impact of ICT adoption on urban spatial structure is of such
paramount importance in understanding its impact on modern economies
that replicating a study by extending and updating its data is a
worthwhile undertaking. By the same token, it behooves us to examine its
robustness by means of alternative data sets. Unlike the data used by
Ioannides et al {[}18{]} and Tranos and Ioannides {[}19{]}, the data
employed by the current study is arguable more consistent. The structure
of the paper is as follows: \protect\hyperlink{sec2}{Section 2} provides
a brief literature review; \protect\hyperlink{sec3}{Section 3} describes
the methods and the data we use; \protect\hyperlink{sec4}{Section 4}
presents the results of the multi-country analysis;
\protect\hyperlink{sec5}{Section 5} narrows down to the two case
studies, the US and UK; and, \protect\hyperlink{sec6}{Section 6}
concludes.

\hypertarget{sec2}{%
\section{Literature review}\label{sec2}}

Gaspar and Glaeser {[}20{]} were the first to model the effect of
telecommunication improvements on the intensity of face-to-face
interactions and city size. Their results indicate that technological
improvements in telecommunications may lead to increased demand for
face-to-face interactions, which will then increase the importance of
cities. Their theoretical model allows for a complementary relation
between agglomeration externalities and advances in telecommunications.
However, their results rest on a critical assumption, namely that
face-to-face interactions are superior to any technology-mediated
interactions. Were it not for this assumption then the opposite
prediction would have followed. As indicated above and as the management
literature suggests, face-to-face interactions do not necessarily
dominate digitally mediated ones. Certain elements of knowledge sharing
can also be achieved by via online interactions {[}10,21{]}. Indeed, to
a certain extent, this argument is technology dependent. E.g., current
teleconferencing capabilities are much more advanced than the ones
available in the late 1990s. Hence, it behooves us to return empirically
to the impact on agglomeration externalities from improved
telecommunication technologies.

\color{blue}

Kolko {[}22{]} uses internet diffusion data and identifies a clear
complementary link between internet usage and city size. Interestingly,
he identifies higher internet domain densities in remote cities which
indicated a substitution effect of the internet for longer-distance
non-electronic communications. His results are consistent for different
measures of internet diffusion (internet domain density and internet
take-up). Sinai and Waldfogel {[}23{]} approach the same question from
the consumers' point of view and find support for complementarity
between internet and cities. They study the link between market size and
locally targeted online content and find that more online local content
in larger markets. However, they also obtain evidence for a substitution
effect: they find that holding constant local online content constant,
market size has a negative effect on individual connectivity. Forman et
al.~{[}24{]} examined whether commercial internet adoption is higher in
cities than in rural areas. While the former would indicate a
complementarity between internet adoption and cities, the latter would
reflect a complementary relation, according to which the internet is
used as a means to offset costs and lack of opportunities related to
peripheriality. Their results indicate that despite internet adoption by
firms with more than 100 employees being greater in smaller urban
agglomerations, the adoption of more sophisticated internet-based
applications is positively related with city size in \(2000\). Sohn,
Kim, and Hewings {[}25{]} compare how information technologies are
related to urban spatial structure for Chicago and Seoul. Although they
find a clear complementary link for Chicago, this was not the case for
Seoul, where information technologies contribute to a more dispersed
spatial pattern. Focusing on the municipalities in the province of
Barcelona, Pons-Novell and Viladecans-Marsal {[}26{]} find a
complementary link between individual internet take-up and off-line
commercial offerings, but their results cannot safely reject a
substitution effect. Craig, Hoang, and Kohlhase {[}27{]} focus on
internet take-up rates across US states during the period \(2000-2011\).
Their analysis provides suggestive evidence of a complementary role that
internet connectivity performs on urban living. Ioannides et
al.~{[}18{]} use country-level data on city size distributions to
examine the impact of fixed line telephony on urban structure
\(1980-2000\). Using a panel dataset of spatial dispersion measures,
they find robust evidence that an increase in the number of telephone
lines per capita encourages the spatial dispersion of population in that
they lead to a more concentrated distribution of city sizes. In
addition, by the end of their study period the internet has come into
use, but their results with internet usage is more speculative but do
show that it goes in the same direction.

\color{blue}

Focusing on rural areas, Partridge et al.~{[}28{]} find no evidence that
rural distance penalties in the US have substantially changed since
\(1970s\) indicating that technological changes including the internet
and digital communications have not affected spatial structure. Also
interesting are the findings of a more recent study by Kim and Orazem
{[}29{]} on the economic effects of broadband internet in rural areas.
They identify a positive effect on new firm location decisions, with the
effect being higher in rural areas with larger population and in rural
areas adjacent to a metropolitan area. These results suggest a
complementary relation between the internet and agglomeration economies.

\color{blue}

Moving beyond modelling the direct relationship between ICT and spatial
structure, Bekkerman and Gilpin {[}30{]} focused on the role of locally
based information resources using a dataset about the US libraries
during the period \(2000-2008\). Their results suggest that internet
access increases the demand and the value of locally accessible
information, and such complementarities are higher in larger
metropolitan areas. Anenberg and Kung {[}31{]} also identified a
complementary relation between the internet and consumption variety in
cities by focusing on food truck industry in the US.

\color{black}

Thus, clearly, previous research suggests that the declaration of the
``death of distance'' has proven to be premature {[}32{]}. However, the
exact impact of digital communications on spatial structure is still an
open question. As Leamer and Storper {[}33{]} indicate, the internet can
affect both centripetal and centrifugal forces. The only cross country
study {[}18{]} supports a clear substitution effect, whereas most of the
above studies have focused either on the US or on some specific cities.
Moreover, most of the above studies examined the
complementarity/substitutability question over time periods when the
internet and other digital communication technologies were still
emerging. For instance, internet penetration in the US in \(2000\),
which was the focus for quite a few of the above studies including
Ioannides et al.~{[}18{]}, was just above \(50\) percent, while in
\(2016\) it reached almost \(90\) percent {[}34{]}. At a global scale
internet penetration increased from \(7\) to \(46\) percent during the
same period {[}35{]}. In addition, although email and instant messaging
technologies were widespread in the developed world in early \(2000s\),
network externalities due to mobile internet and online social media
were nowhere close to what we are familiar with today. For instance,
Facebook users increased from \(1\) million in \(2004\) to more than
\(1.5\) billion in \(2015\)\footnote{\url{http://newsroom.fb.com/company-info/}}.
Hence, it might have been premature for the spatial economic effects of
ICT adoption to have been materialized by the time that most of the
above studies were conducted.

Most recently, Tranos and Ioannides {[}19{]} return to the setting of
Ioannides et al.~{[}18{]}, update the data employed and obtain results
that confirm their original findings. That is, the diffusion of fixed
telephony has caused more dispersed urban structures worldwide, in other
words, greater urban decentralization. Similar causal effects are
established for mobile telephony, which are novel in relation to
Ioannides et al.~{[}18{]}, who lacked such data, and for the internet,
which extends their earlier findings. The robustness of their results is
confirmed for such alternative measures of dispersion as the Gini
coefficient, the Herfindahl index, and the coefficient of variation.
This is notable because several years of additional data were used that
pertain to an era of rapid expansion of the internet and web-based
technologies, more generally.

\hypertarget{sec3}{%
\section{Materials and Methods}\label{sec3}}

The main aim of this paper is to return to the estimation of the causal
impact of ICT adoption on the spatial dispersion of economic activities
and consequently population. It adopts an approach that is carried out
at several spatial scales and uses different data than those used by
Ioannides et al.~{[}18{]} and Tranos and Ioannides {[}19{]}.

We start with a multi-country exercise, which includes both developed
and developing countries. \protect\hyperlink{sec3.1}{Section 3.1}
discusses the methods and the different data we use and
\protect\hyperlink{sec4}{Section 4} reports the results. Because of
limitations related with multi-country urban population data (see
discussion below), we complement our analysis with two case studies for
the US and the UK urban system for which we have access to much more
granular data. \protect\hyperlink{sec3.2}{Section 3.2} discusses the
methods and \protect\hyperlink{sec5}{Section 5} reports the results.

\hypertarget{sec3.1}{%
\subsection{Multi-country analysis}\label{sec3.1}}

The multi-country identification strategy is a two-step approach
{[}18{]}. The first step of our methodology is to estimate the Zipf
coefficient for a broad sample of countries over time. Zipf coefficient
is one of the most widely used measures of spatial dispersion with
numerous applications in urban economics and economic geography ({[}see
for example {[}36{]}; {[}37{]}; {[}38{]}; {[}39{]}; {[}40{]}; {[}41{]};
{[}42{]}). This an appropriate measure of dispersion because of the
extreme heterogeneity of the city size distribution and the very good
fits normally obtained with such estimations. Skipping details
pertaining of the suitability of the approach, which may be found in
Ioannides et al.~{[}18{]} and Tranos and Ioannides {[}19{]}, we proceed
with the estimation of the logarithmic form of:

\begin{align}
ln(rank_{i}) = ln(S_0) + \zeta ln(Size_i) + e_{i}. \label{eq1}
\end{align}

This leads to an estimate of \(\zeta\), known as the Zipf coefficient,
but more appropriately the Pareto exponent, when \(\zeta\) is not
estimated to be near \(1\). While much of the literature focuses on the
Zipf coefficient being close to \(1\), several estimations of city size
distributions lead to estimates of \(\zeta\) that are not necessarily
equal to \(1\), or even near it, in which case we refer to \(\zeta\) as
the Pareto, or power law, exponent. Given that our aim here is to
estimate \(\zeta\) for a number of countries over time as a measure of
dispersion, equation \ref{eq1} describes the rank of city \(i\) in
country \(c\) in year \(t\):

\begin{align}
ln(rank_{ict}) = ln(S_{0ct}) + \zeta_{ct} ln(Size_{ict}) + e_{ict}. \label{eq2}
\end{align}

The estimation of equation \ref{eq2} has typically been performed by
Ordinary Least Squares (OLS). Gabaix and Ioannides {[}43{]} discuss the
downward bias of estimates of equation \ref{eq2} using OLS on small
samples. Gabaix and Ibragimov {[}44{]} propose a practical remedy to
correct this bias, which we do adopt in this paper: instead of using the
\(log\) of \(rank\) of a city \(i\) in a country \(c\) in year \(t\),
they propose to use the \(log(rank-0.5)\), which has indeed been widely
adopted. Researchers working in this area must contend with definitional
differences as well as differences in availability of different kinds of
data sources. Definitions of cities differ across countries, for
political, administrative and legal reasons\footnote{See {[}45{]}, Ch.
  8, for issues and pitfalls associated with different definitions.}. In
contrast to the data drawn from Thomas Brinkhoff's City Population
project {[}46{]}\footnote{\url{http://www.citypopulation.de}}, which
were used by Soo {[}47{]}, Ioannides et al.~{[}18{]} and Tranos and
Ioannides {[}19{]}, the multi-country analysis in the present paper is
based on the annual population data for urban agglomerations with
\(300,000\) inhabitants or more from the Department of Economic and
Social Affairs of the United Nations {[}9{]}. Despite some criticism
about the consistency of the urban agglomeration definitions across
different countries {[}48{]}, this is the only available source for
yearly, multi-country population data for urban agglomerations
{[}49--51{]}. The results based on these data are reported and discussed
in \protect\hyperlink{sec4}{Section 4}.

Table \ref{zipf} presents the estimated \(\zeta\) coefficients for the
panel of countries that the second stage of the analysis focuses on
using the UN urban agglomerations data. It becomes evident that there is
considerable variation in the estimates of \(\zeta\) across the
different countries in the sample. Because of the empirically
established heavy upper tail of data for cities and urban
agglomerations, the \(\zeta\) coefficient constitutes a convenient
measure of dispersion. The larger its absolute value, the thinner the
upper tail; equivalently, the larger is the coefficient algebraically,
the heavier the upper tail. This key observation is basis for the second
step of our methodology.

\begin{table}

\caption{\label{tab:unnamed-chunk-2}Zipf coefficients\label{zipf}}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrrr}
\toprule
Zipf coef. & 2000 & 2001 & 2002 & 2003 & 2004 & 2005 & 2006 & 2007 & 2008 & 2009 & 2010 & 2011 & 2012 & 2013 & 2014 & 2015 & 2016 & 2017 & 2018\\
\midrule
AGO & NA & -0.94 & -0.94 & -0.94 & -0.94 & -0.94 & -0.93 & -0.93 & -0.92 & -0.92 & -0.92 & -0.91 & -0.90 & -0.90 & -0.89 & -0.90 & -0.90 & -0.91 & NA\\
ARG & -0.97 & -0.97 & -0.97 & -0.97 & -0.97 & -0.97 & -0.97 & -0.97 & -0.97 & -0.97 & -0.97 & -0.97 & -0.97 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & NA\\
AUS & -0.82 & -0.82 & NA & NA & NA & -0.83 & -0.83 & -0.82 & -0.82 & -0.82 & -0.82 & -0.82 & -0.81 & -0.81 & -0.81 & -0.81 & -0.81 & -0.80 & NA\\
BGD & -0.65 & -0.65 & -0.66 & -0.67 & -0.67 & -0.68 & -0.69 & -0.69 & -0.70 & -0.70 & -0.71 & -0.72 & -0.72 & -0.72 & -0.73 & -0.73 & -0.73 & -0.73 & NA\\
BLR & -1.26 & -1.26 & -1.26 & NA & NA & NA & -1.24 & -1.23 & -1.23 & -1.22 & -1.22 & -1.23 & -1.23 & -1.23 & -1.23 & -1.23 & -1.23 & -1.23 & -1.23\\
\addlinespace
BRA & -0.93 & -0.93 & -0.93 & -0.93 & -0.93 & -0.93 & -0.93 & -0.93 & -0.93 & -0.93 & -0.94 & -0.94 & -0.94 & -0.94 & -0.94 & -0.94 & -0.94 & -0.93 & -0.93\\
CAN & -1.06 & -1.05 & -1.05 & -1.05 & -1.05 & -1.04 & -1.04 & -1.04 & -1.04 & -1.03 & -1.03 & -1.03 & -1.03 & -1.02 & -1.02 & -1.02 & -1.02 & NA & NA\\
CHL & -0.73 & -0.73 & -0.73 & -0.74 & -0.74 & -0.75 & -0.75 & -0.76 & -0.76 & -0.76 & -0.77 & -0.77 & -0.77 & -0.78 & -0.78 & -0.78 & -0.78 & -0.79 & NA\\
CHN & -1.15 & -1.16 & -1.16 & -1.17 & -1.17 & -1.18 & -1.18 & -1.18 & -1.19 & -1.19 & -1.19 & -1.19 & -1.19 & -1.20 & -1.20 & -1.20 & -1.20 & -1.21 & NA\\
CMR & -0.76 & -0.76 & -0.76 & -0.76 & -0.76 & -0.77 & -0.77 & -0.77 & -0.77 & -0.76 & -0.76 & -0.76 & -0.76 & -0.76 & -0.77 & -0.77 & -0.77 & -0.77 & NA\\
\addlinespace
COD & -0.94 & -0.94 & -0.95 & -0.95 & -0.96 & -0.96 & -0.97 & -0.97 & -0.98 & -0.98 & -0.99 & -0.99 & -0.99 & -0.99 & -0.99 & -0.99 & -1.00 & -1.00 & NA\\
COL & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98 & -0.98\\
DEU & -1.53 & -1.53 & -1.53 & -1.53 & -1.53 & -1.52 & -1.52 & -1.52 & -1.51 & -1.51 & -1.51 & -1.51 & -1.50 & -1.50 & -1.50 & -1.50 & -1.50 & -1.50 & -1.50\\
DZA & -1.10 & -1.11 & -1.12 & -1.13 & -1.14 & -1.15 & -1.15 & -1.16 & -1.16 & -1.17 & -1.18 & -1.18 & -1.18 & -1.19 & -1.19 & -1.19 & -1.19 & -1.20 & -1.21\\
EGY & -0.73 & -0.73 & -0.73 & -0.73 & -0.73 & -0.73 & -0.73 & -0.73 & -0.72 & -0.72 & -0.72 & -0.72 & -0.72 & -0.72 & -0.72 & -0.72 & -0.72 & -0.71 & -0.71\\
\addlinespace
ESP & -0.94 & -0.94 & -0.93 & -0.93 & -0.93 & -0.93 & -0.93 & -0.92 & -0.92 & -0.92 & -0.92 & -0.92 & -0.91 & -0.91 & -0.91 & -0.91 & -0.91 & -0.90 & -0.90\\
FRA & -1.15 & -1.15 & -1.15 & -1.15 & -1.14 & -1.14 & -1.14 & -1.14 & -1.14 & -1.14 & -1.14 & -1.14 & -1.14 & -1.14 & -1.13 & -1.13 & -1.13 & -1.13 & -1.13\\
GBR & -1.19 & -1.19 & -1.19 & -1.19 & -1.19 & -1.19 & -1.19 & -1.19 & -1.18 & -1.18 & -1.18 & -1.18 & -1.18 & -1.17 & -1.17 & -1.17 & -1.17 & -1.17 & -1.16\\
IDN & -1.11 & -1.11 & -1.12 & -1.12 & -1.12 & -1.13 & -1.13 & -1.13 & -1.13 & -1.13 & -1.13 & -1.13 & -1.13 & -1.13 & -1.13 & -1.13 & -1.12 & -1.12 & -1.12\\
IND & -1.05 & -1.06 & -1.06 & -1.06 & -1.07 & -1.07 & -1.08 & -1.08 & -1.08 & -1.09 & -1.09 & -1.09 & -1.09 & -1.09 & -1.09 & -1.09 & -1.09 & -1.09 & NA\\
\addlinespace
IRN & -1.03 & -1.05 & -1.07 & -1.09 & -1.11 & -1.13 & -1.14 & -1.15 & -1.16 & -1.16 & -1.17 & -1.17 & -1.18 & -1.18 & -1.18 & -1.19 & -1.19 & -1.19 & NA\\
IRQ & NA & -1.06 & -1.07 & -1.07 & -1.08 & -1.07 & -1.07 & -1.07 & -1.05 & -1.03 & -1.06 & -1.10 & -1.14 & -1.17 & -1.20 & -1.24 & -1.23 & -1.23 & -1.23\\
ITA & -1.38 & -1.39 & -1.39 & -1.39 & -1.40 & -1.40 & -1.40 & -1.41 & -1.41 & -1.41 & -1.42 & -1.42 & -1.42 & -1.43 & -1.43 & -1.43 & -1.44 & -1.44 & -1.44\\
JOR & -0.94 & -0.97 & -1.00 & -1.04 & -1.07 & -1.09 & -1.11 & -1.12 & -1.13 & -1.15 & -1.16 & -1.17 & -1.19 & -1.20 & -1.22 & -1.23 & -1.25 & -1.26 & NA\\
JPN & -0.77 & -0.77 & -0.76 & -0.76 & -0.76 & -0.76 & -0.76 & -0.75 & -0.75 & -0.75 & -0.75 & -0.75 & -0.75 & -0.75 & -0.75 & -0.75 & -0.75 & -0.75 & -0.74\\
\addlinespace
KAZ & -1.93 & -1.91 & -1.88 & -1.85 & -1.83 & -1.82 & -1.81 & -1.79 & -1.78 & -1.76 & -1.72 & -1.68 & -1.63 & -1.59 & -1.55 & -1.51 & -1.47 & -1.43 & -1.39\\
KEN & -0.75 & -0.76 & -0.77 & -0.78 & -0.79 & -0.80 & -0.80 & -0.81 & -0.81 & -0.82 & -0.82 & -0.81 & -0.81 & -0.81 & -0.81 & -0.81 & -0.81 & -0.81 & NA\\
KOR & -1.04 & -1.05 & -1.06 & -1.08 & -1.09 & -1.10 & -1.10 & -1.11 & -1.11 & -1.11 & -1.12 & -1.12 & -1.12 & -1.12 & -1.12 & -1.13 & -1.13 & -1.13 & -1.13\\
MAR & -1.23 & -1.23 & -1.23 & -1.23 & -1.23 & -1.23 & -1.23 & -1.23 & -1.24 & -1.24 & -1.24 & -1.24 & -1.24 & -1.24 & -1.24 & -1.24 & -1.24 & -1.24 & -1.24\\
MEX & -1.17 & -1.18 & -1.18 & -1.18 & -1.18 & -1.18 & -1.18 & -1.18 & -1.18 & -1.18 & -1.19 & -1.19 & -1.19 & -1.19 & -1.19 & -1.19 & -1.19 & -1.19 & -1.19\\
\addlinespace
MOZ & -1.15 & -1.16 & -1.17 & -1.18 & -1.18 & -1.19 & -1.20 & -1.21 & -1.23 & -1.24 & -1.26 & -1.26 & -1.27 & -1.28 & -1.31 & -1.34 & -1.36 & -1.38 & NA\\
MYS & -1.04 & -1.03 & -1.03 & -1.03 & -1.03 & -1.02 & -1.02 & -1.02 & -1.01 & -1.01 & -1.01 & -1.00 & -1.00 & -1.00 & -0.99 & -0.99 & -0.99 & -0.98 & -0.98\\
NGA & -1.13 & -1.16 & -1.17 & -1.18 & -1.19 & -1.20 & -1.21 & -1.22 & -1.22 & -1.23 & -1.23 & -1.25 & -1.25 & -1.26 & -1.26 & -1.27 & -1.27 & -1.27 & NA\\
PAK & NA & -0.90 & -0.90 & -0.90 & -0.90 & -0.90 & -0.90 & -0.90 & -0.90 & -0.90 & -0.90 & -0.90 & -0.89 & -0.89 & -0.89 & -0.89 & -0.89 & -0.89 & NA\\
PER & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.84 & -0.84\\
\addlinespace
PHL & -1.13 & -1.15 & -1.16 & -1.17 & -1.18 & -1.18 & -1.21 & -1.22 & -1.22 & -1.23 & -1.23 & -1.24 & -1.24 & -1.25 & -1.25 & -1.25 & -1.26 & -1.26 & NA\\
POL & -1.85 & -1.85 & -1.85 & -1.85 & -1.84 & -1.84 & -1.83 & -1.83 & -1.83 & -1.82 & -1.82 & -1.82 & -1.81 & -1.80 & -1.79 & -1.78 & -1.78 & -1.77 & -1.76\\
RUS & -1.46 & -1.46 & -1.46 & -1.46 & -1.47 & -1.47 & -1.47 & -1.47 & -1.47 & -1.47 & -1.46 & -1.47 & -1.47 & -1.47 & -1.47 & -1.47 & -1.46 & -1.46 & -1.46\\
SAU & -0.97 & -0.97 & -0.98 & -0.98 & -0.99 & -0.99 & -0.99 & -0.99 & -1.00 & -1.00 & -1.00 & -1.00 & -1.00 & -1.01 & -1.01 & -1.01 & -1.01 & -1.01 & -1.01\\
THA & -1.14 & -1.16 & -1.18 & -1.20 & -1.21 & -1.23 & -1.24 & -1.26 & -1.27 & -1.29 & -1.30 & -1.30 & -1.31 & -1.31 & -1.32 & -1.32 & -1.33 & -1.33 & -1.33\\
\addlinespace
TUR & -1.07 & -1.07 & -1.06 & -1.06 & -1.06 & -1.06 & -1.06 & -1.05 & -1.05 & -1.05 & -1.04 & -1.04 & -1.04 & -1.04 & -1.04 & -1.04 & -1.05 & -1.05 & -1.05\\
TZA & -0.86 & -0.86 & -0.86 & -0.86 & -0.87 & -0.87 & -0.87 & -0.87 & -0.88 & -0.88 & -0.88 & -0.88 & -0.88 & -0.88 & -0.88 & -0.87 & -0.87 & -0.87 & NA\\
UKR & -1.53 & -1.53 & -1.53 & -1.53 & -1.52 & -1.52 & -1.52 & -1.52 & -1.52 & -1.52 & -1.51 & -1.51 & -1.51 & -1.51 & -1.51 & -1.51 & -1.51 & -1.51 & -1.51\\
USA & -0.97 & -0.98 & -0.99 & -1.00 & -1.01 & -1.01 & -1.02 & -1.03 & -1.03 & -1.04 & -1.05 & -1.05 & -1.06 & -1.06 & -1.07 & -1.07 & -1.08 & -1.08 & NA\\
VNM & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.85 & -0.84 & -0.84 & -0.84 & -0.84 & -0.83 & -0.83 & -0.83 & -0.82 & -0.82\\
\addlinespace
ZAF & -0.80 & -0.81 & -0.82 & -0.83 & -0.84 & -0.85 & -0.86 & -0.87 & -0.88 & -0.88 & -0.89 & -0.90 & -0.91 & -0.92 & -0.92 & -0.93 & -0.94 & -0.94 & NA\\
\bottomrule
\multicolumn{20}{l}{\textit{Note: }}\\
\multicolumn{20}{l}{Corrected as per Gabaix and Ibragimov (2011)}\\
\end{tabular}}
\end{table}

The second step of our methodology involves estimating the following
empirical model {[}18{]}:

\color{blue}

\begin{align}
\zeta_{ct} = \theta_{c} + \delta_{t} + X_{ct} + e_{ct}. \label{eq3}
\end{align}

\color{black}

\noindent The model enable us to estimate the effects of a vector \(X\)
of explanatory variables which are included to account for the spatial
structure of country \(c\) in year \(t\). The main variables of interest
here are internet and digital communications variables that include:
internet users per \(100\) inhabitants, broadband users per \(100\)
inhabitants, mobile phone users per \(100\) inhabitants, and fixed phone
users per \(100\) inhabitants. To address a potential omitted variable
bias equation \ref{eq3} includes \color{blue} country (\(\theta_c\)) and
year (\(\delta_t\)) fixed effects; \color{black} \(e_{ct}\) is the error
term. In addition, vector \(X\) includes several control variables,
whose descriptive statistics together with those for the other variables
used to estimate \ref{eq3} are reported in Table \ref{desc.global}.

Referring to the control variables, total country population is an
important measure of size, GDP per capita, and GDP growth are intimately
related to urbanization and so are population density, and
non-agricultural value-added as a share of GDP. Trade, that is
conventionally measured as exports plus imports as a share of GDP, is an
important time varying measure of trade openness. Government expenditure
as a share of GDP may be a proxy of public investment in some countries
and government waste in others.

\color{blue}

As Table \ref{cor} indicates, although there some rather strong
correlations among these variables, mobile phone penetration appears to
have a distinct character from its fixed phone counterpart: their
correlation coefficient is only \(0.445\). This probably highlights the
different composition of the population or infrastructure development
patterns in the developing world, where mobile telephony helped overcome
the lack of fixed line infrastructure and mobile phone networks are also
used as the main way to use the internet {[}52,53{]}.

\begin{table}[!htbp] \centering 
  \caption{Descriptive statistics for the multi-country model\label{desc.global}} 
  \label{} 
\footnotesize 
\begin{tabular}{@{\extracolsep{1pt}}lccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
Zipf coefficient & 950 & $-$1.9 & $-$1.1 & 0.2 & $-$0.7 \\ 
Zipf inv. sq. error\textsuperscript{a} & 950 & 0.01 & 0.1 & 0.05 & 0.2 \\ 
Population density & 931 & 2.5 & 126.2 & 182.1 & 1,239.6 \\ 
Government expenditure (\% GDP) & 913 & 1.0 & 14.7 & 4.9 & 30.0 \\ 
Trade (\% GDP) & 914 & 19.1 & 65.3 & 33.9 & 220.4 \\ 
GDP growth & 934 & $-$33.1 & 4.0 & 4.2 & 54.2 \\ 
GDP per capita (log) & 892 & 630.7 & 18,049.2 & 15,432.9 & 61,391.4 \\ 
Female labour force (\%) & 950 & 7.9 & 37.8 & 11.9 & 55.2 \\ 
Population & 950 & 5,122,493 & 116,073,037.0 & 244,323,372.0 & 1,392,730,000 \\ 
Non agriculture value added (\% GDP) & 930 & 58.8 & 90.3 & 8.2 & 99.4 \\ 
Internet users per 100 hab. & 916 & 0.01 & 32.5 & 28.2 & 96.0 \\ 
Broadband users per 100 hab. & 824 & 0.0 & 8.8 & 11.2 & 44.8 \\ 
Mobile phone users per 100 hab. & 949 & 0.0 & 72.9 & 46.4 & 191.0 \\ 
Fixed phone users per 100 hab. & 946 & 0.0 & 19.8 & 18.6 & 68.4 \\ 
\hline \\[-1.8ex] 
\multicolumn{6}{l}{\textsuperscript{a}This is the inverse squared standard error of the estimated Zipf, coefficient, which is used for weighting} \\ 
\multicolumn{6}{l}{the observations for the estimation, of equation 3. See Section 4} \\ 
\end{tabular} 
\end{table}

\color{black}

\begin{table}

\caption{\label{tab:unnamed-chunk-4}Correlations between ICT variables\label{cor}}
\resizebox{\linewidth}{!}{
\fontsize{5}{7}\selectfont
\begin{tabular}[t]{lrrrr}
\toprule
  & Internet & Broadband & Mobile & Fixed\\
\midrule
Internet users per 100 hab. & 1.00 & 0.87 & 0.70 & 0.66\\
Broadband users per 100 hab. & 0.87 & 1.00 & 0.53 & 0.71\\
Mobile phone users per 100 hab. & 0.70 & 0.53 & 1.00 & 0.28\\
Fixed phone users per 100 hab. & 0.66 & 0.71 & 0.28 & 1.00\\
\bottomrule
\end{tabular}}
\end{table}

The availability of a panel dataset for city sizes across countries
enables us to use country fixed effects, which can address potential
endogeneity issues related to unobserved country specific
characteristics of city size distributions. However, such a strategy
does not address potential simultaneity issues. Simply put, internet
penetration might be affected by spatial structure, as reflected in Zipf
coefficients, or both internet penetration and spatial structure might
be jointly determined by a third variable. E.g., if a country already
has a dispersed spatial structure, internet is particularly suitable in
facilitating communication. Potential endogeneity in our specification
will prevent us from being able to determine the causal impact of
internet and digital communication technologies usage on spatial
structure, which is the main aim of this paper. In order to address this
problem, we will adopt an instrumented variable strategy. Table
\ref{desc.global} also includes the descriptive statistics for the
instrumental variable we are using and (\(female\:labour\:force\)), will
be discussed in \protect\hyperlink{sec4}{Section 4}.

\hypertarget{sec3.2}{%
\subsection{3.2 Case study approach}\label{sec3.2}}

The above global level analysis is followed by two case studies. They
allow us to examine the potential effects of the internet on two mature
urban systems in greater detail and depth and without the exogenously
imposed threshold of the \(300,000\) habitants that the global level
analysis affords us. We focus on the US and the UK, for which we are
able to utilize more granular internet-related data (see
\protect\hyperlink{sec5}{Section 5} for the data description). Given
that we are dealing with specific countries and not with a panel of
countries we cannot employ an identification strategy similar to that of
the global level analysis discussed in the previous section. Therefore,
we propose a one-step approach involving estimation of the following
empirical cross-sectional model:

\begin{align}
Diff\:in\:ranks_{i} = a + \beta ICT_{i} + B C_{i} + e_{i}.\label{eq4}
\end{align}

In order to capture the micro-dynamics of the urban systems in the two
case studies, we follow Batty {[}54{]} and Havlin {[}55{]} and focus on
the difference in ranks for individual cities during the study
period.\footnote{Changes in ranking of individuals has also been used
  elsewhere in economics because of its robustness properties. C.f.
  Chetty and Hendren {[}56{]}.} Contrary to their approach, we are
interested in the real difference in rankings rather than in absolute
differences, in order to capture whether or a city improves its ranking
within the urban system during the study period. We then test whether
our internet-related variable has an effect on the change in ranking.
Hence, we define the left-hand side (LHS) variable of equation \ref{eq4}
as follows:

\begin{align}
Diff\:in\:ranks_{i} = r_{i(t-1)} - r_{it},\label{eq5}
\end{align}

\noindent where \(r_{it}\) is the population rank of city \(i\) in year
\(t\). A negative (positive) value for the \(Diff\:in\:ranks_{i}\)
variable indicates that city's \(i\) position in the urban hierarchy of
the country worsens (improves) in relative terms, also due to the
population changes of the other cities of the urban system. Notably,
this variable does not only consider the population change of a specific
city, but it also considers the overall urban system dynamics by
focusing on the rank and not on population per se. Given that the data
we use and the definitions of cities vary between the UK and the US we
are going to discuss these data in the relevant sections. What we
highlight here is that the estimation of equation \ref{eq4}, just like
equation \ref{eq3}, might suffer from endogeneity and therefore an
instrumental variables strategy is employed in order to address this
issue.

\hypertarget{sec4}{%
\section{Digital technologies and spatial structure: a global
view}\label{sec4}}

This section presents the estimation results of \ref{eq3}. The LHS
variable is \(\zeta\), the Zipf coefficient, as estimated according to
the Gabaix and Ibragimov {[}44{]} correction. The main explanatory
variables of interest are provided at the top of Table \ref{ols.global};
namely internet, broadband, mobile and fixed telephony per \(100\)
habitants, expressed in natural logarithms. The main variables of
interest are introduced successively on their own in the regressions
reported in Table \ref{ols.global}. All regressions include country
fixed effects to control for unobserved heterogeneity and a time trend.
In addition, the observations are weighted with the inverse squared
standard error of the estimated Zipf coefficient (see Table
\ref{desc.global}) to address potential noise that is carried over from
the first part of our identification strategy. In regards to the
interpretation of the estimated coefficients, given that the Zipf
coefficient has entered the regression not as an absolute value, but
instead as a real number a positive coefficient for a RHS variable
indicates an impact towards the decrease of the spatial dispersion of
population. In other words, a positive coefficient indicates an effect
towards less uniform city sizes that is more dispersion of city sizes.
The latter is indicative of enhancement of agglomeration economies
because of the expansion of digital technologies.

\begin{table}[!htbp] \centering 
  \caption{OLS estimation of equation (3)\label{ols.global}} 
  \label{} 
\small 
\begin{tabular}{@{\extracolsep{1pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{Zipf Coefficient 2000-18} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 Internet users per 100 hab. (log) & 0.001$^{***}$ &  &  &  \\ 
  & (0.0001) &  &  &  \\ 
  Broadband users per 100 hab. (log) &  & $-$0.001$^{**}$ &  &  \\ 
  &  & (0.0002) &  &  \\ 
  Mobile phone users per 100 hab. (log) &  &  & 0.0003 &  \\ 
  &  &  & (0.0003) &  \\ 
  Fixed phone users per 100 hab. (log) &  &  &  & $-$0.00003 \\ 
  &  &  &  & (0.0001) \\ 
  Population density (log) & $-$1.378 & $-$1.060 & $-$1.505 & $-$1.892 \\ 
  & (1.870) & (1.985) & (1.896) & (2.007) \\ 
  Government expenditure (% GDP) & 0.0002 & 0.003$^{***}$ & 0.001 & 0.001 \\ 
  & (0.001) & (0.001) & (0.001) & (0.001) \\ 
  Trade (\% of GDP) & 0.0002$^{***}$ & $-$0.0001 & 0.0001 & 0.0001 \\ 
  & (0.0001) & (0.0001) & (0.0001) & (0.0001) \\ 
  Non agriculture value added (\% GDP) & 0.0001 & 0.001 & $-$0.0001 & $-$0.0002 \\ 
  & (0.001) & (0.001) & (0.001) & (0.001) \\ 
  GDP growth & 0.001$^{***}$ & 0.001$^{***}$ & 0.001$^{**}$ & 0.001$^{**}$ \\ 
  & (0.0003) & (0.0004) & (0.0004) & (0.0004) \\ 
  GDP per capita (log) & $-$0.020$^{***}$ & $-$0.023$^{***}$ & $-$0.019$^{***}$ & $-$0.016$^{**}$ \\ 
  & (0.007) & (0.006) & (0.007) & (0.008) \\ 
  Population (log) & 1.300 & 0.955 & 1.390 & 1.785 \\ 
  & (1.873) & (1.985) & (1.899) & (2.011) \\ 
  Constant & $-$18.795 & $-$13.969 & $-$19.948 & $-$25.516 \\ 
  & (26.314) & (27.886) & (26.670) & (28.261) \\ 
 \hline \\[-1.8ex] 
Country fixed effects & Yes & Yes & Yes & Yes \\ 
Yearly fixed effects & Yes & Yes & Yes & Yes \\ 
Observations & 844 & 757 & 865 & 867 \\ 
Adjusted R$^{2}$ & 0.987 & 0.988 & 0.986 & 0.986 \\ 
Residual Std. Error & 0.516 & 0.497 & 0.540 & 0.540 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{4}{r}{Robust Std. Errors in parenthesis} \\ 
\end{tabular} 
\end{table}

Table \ref{ols.global} and \ref{2sls.global} report the estimation
results based on the UN urban agglomeration data. Starting with the
former, we note that an agglomerative effect is only detected for the
internet users and while a marginally opposing result can be seen for
broadband users as indicated by the significant coefficients in columns
(1) and (2). For the telephony variables the estimation of \ref{eq3} did
not yield statistically significant coefficients. Before discussing
further these results and the effects of the other control variables, we
need to highlight that the main challenge of estimating equation
\ref{eq3} is the potential endogenous nature of the share of internet
users which might prevent us from being able to infer a truly causal
effect. Endogeneity might be an issue here as spatial structure, which
is proxied by the Zipf coefficient, might be affected by another source,
which also affects internet penetration. For instance, economic
development might affect the concentration of population in large cities
and at the same time enable more people to go online. If we do not
address this issue, the coefficient for the main variable of interest
will capture potential effects that internet penetration has on spatial
structure, but also potential reverse causality effects that spatial
structure might generate on internet penetration. To overcome this
potential problem, Table \ref{2sls.global} reports estimates of equation
\ref{eq3} using two-stage least squares (2SLS) with instrumental
variables (IVs). The latter are variables which are correlated with our
endogenous variables, but do not influence current spatial structure.
Such an approach will enable us to estimate the causal effect -- if any
-- of the internet and digital communications on spatial structure. At a
first stage, our endogenous variables are regressed against the IVs.
Then, the predicted values of the endogenous variables based on the IVs
and the other control variables are used instead of the endogenous
variable to estimate eq. \ref{eq3}. A significant effect will verify the
causal impact of the internet and digital communication usage on spatial
structure.

\begin{table}[!htbp] \centering 
  \caption{2SLS estimation of equation (3)\label{2sls.global}} 
  \label{} 
\small 
\begin{tabular}{@{\extracolsep{1pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{Zipf Coefficient 2000-18} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 Internet users per 100 hab. (log) & 0.001$^{***}$ &  &  &  \\ 
  & (0.0004) &  &  &  \\ 
  Broadband users per 100 hab. (log) &  & 0.002$^{**}$ &  &  \\ 
  &  & (0.001) &  &  \\ 
  Mobile phone users per 100 hab. (log) &  &  & $-$0.007$^{*}$ &  \\ 
  &  &  & (0.004) &  \\ 
  Fixed phone users per 100 hab. (log) &  &  &  & $-$0.001$^{***}$ \\ 
  &  &  &  & (0.0003) \\ 
  Population density (log) & $-$0.827 & $-$1.736 & $-$12.765 & $-$0.828 \\ 
  & (1.841) & (2.622) & (8.326) & (2.262) \\ 
  Government expenditure (\% GDP) & $-$0.0003 & 0.0002 & $-$0.002 & $-$0.0003 \\ 
  & (0.001) & (0.002) & (0.002) & (0.001) \\ 
  Trade (\% of GDP) & 0.0004$^{***}$ & 0.0004$^{*}$ & 0.001 & $-$0.0003 \\ 
  & (0.0001) & (0.0002) & (0.0005) & (0.0002) \\ 
  Non agriculture value added (\% GDP) & 0.0003 & $-$0.0003 & $-$0.001 & $-$0.001 \\ 
  & (0.001) & (0.001) & (0.001) & (0.001) \\ 
  GDP growth & 0.001$^{***}$ & 0.002$^{***}$ & 0.001$^{*}$ & 0.0001 \\ 
  & (0.0004) & (0.001) & (0.001) & (0.0004) \\ 
  GDP per capita (log) & $-$0.023$^{***}$ & 0.003 & 0.034 & $-$0.0003 \\ 
  & (0.007) & (0.012) & (0.038) & (0.011) \\ 
  Population (log) & 0.776 & 1.867 & 12.794 & 0.740 \\ 
  & (1.841) & (2.646) & (8.410) & (2.267) \\ 
  Constant & $-$11.513 & $-$27.468 & $-$180.823 & $-$10.915 \\ 
  & (25.852) & (37.246) & (118.577) & (31.852) \\ 
 \hline \\[-1.8ex] 
Weak instruments & 44.77 & 22.23 & 3.51 & 27.24 \\ 
Wu-Hausman & 3 & 15.96 & 12.26 & 12 \\ 
P-value & 0.08 & 0 & 0 & 0 \\ 
Country fixed effects & Yes & Yes & Yes & Yes \\ 
Yearly fixed effects & Yes & Yes & Yes & Yes \\ 
Observations & 844 & 757 & 865 & 867 \\ 
Adjusted R$^{2}$ & 0.986 & 0.984 & 0.944 & 0.979 \\ 
Residual Std. Error & 0.534 & 0.579 & 1.078 & 0.657 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{4}{r}{Robust Std. Errors in parenthesis} \\ 
 & \multicolumn{4}{r}{IV: Female participation in labour force} \\ 
\end{tabular} 
\end{table}

The main challenge for such an exercise is to find a valid IV. In our
case, the challenge is even bigger as we need to find a variable which
is not only fit for purpose, but it also contain as few missing values
as possible in order to retain the same number of observations as the
OLS models. This is particularly difficult for a diverse sample of
countries as the one we are dealing with here. We propose here the
percentage of female participation in the labor force, which is directly
related to digital infrastructure, but not to spatial structure (see
Table \ref{desc.global} for the descriptive statistics). \color{blue}
The literature provides empirical and theoretical evidence that ICT are
correlated with increase in female participation in labour markets
mostly because they enable teleworking {[}57--61{]} \color{black} At the
same time, female labor participation is not related with spatial
structure especially given the diverse sample of countries. Although one
might think that increased female labor participation may lead to
relocation of households to large cities and therefore affect the
spatial structure of cites, such a process varies a lot within our
dataset. Even if this might be true for developed countries with mature
urban systems for which a location within a large urban center might
provide opportunities for both male and female workers to find jobs, our
dataset also includes countries, \(41\) per cent of the GDP of which is
attributed to primary activities (see Table \ref{desc.global}). So,
female labor participation might also be related to economic activities
located outside large urban centers. In terms of the relevant tests
presented in Table \ref{2sls.global} the weak identification first stage
F-test exceeds by far all the Stock-Yogo critical values for all, but
the mobile telephony regression. Hence, we are confident to interpret
the effects of internet, broadband internet and fixed telephony usage.

The estimations presented in Table \ref{2sls.global} indicate that
increases in internet and broadband internet usage have resulted to a
decrease of the spatial dispersion of population for the time period and
for the panel of countries included in our data. In other words,
increases in internet and digital communications resulted to national
urban systems which are less uniform in terms of city sizes and are
characterized by higher population concentrations in larger cities. The
exact opposite appears to happen for fixed telephony, which our results
indicate that has led to an increase of spatial dispersion of
population. Importantly, our identification strategy enables us to
address potential reverse causality issues and treat the results of
Table \ref{2sls.global} as causal. Hence, the main finding of our
multi-country, global analysis based on data for urban agglomerations of
at least 300,000 inhabitants is that internet and broadband internet
usage appear to act in favor of agglomeration economies and result to
urban systems with more dominant cities on the top of the urban
hierarchy. On the contrary, fixed telephony usage acted in favour of
centrifugal forces. Our identification strategy does not enable us to
make similar interpretations for mobile telephony usage.

Regarding the control variables, only a few of them have significant
effects on spatial structure probably because the fixed effects
estimation masks the between-country variation. The effects of these
control variables are in agreement with previous research (Ioannides et
al.~2008). Namely, GDP per capita has a negative effect which indicates
that wealthier countries tend to have more balanced urban systems, while
the opposite seems to happen for countries experiencing a growing
economy.

In total, the results indicate that our measures of internet and
telephony penetration have further enhanced agglomeration forces, at
least for large urban agglomerations, while the opposite happens for
fixed telephony. These results are robust against endogeneity issues,
but are limited to urban agglomeration included in our data. Indeed, for
the estimation of the Zipf coefficient we only included urban
agglomeration of \(300,000\) inhabitants or more due to data
availability. Hence, the above estimations cannot verify whether such an
effect is also valid for smaller cities. Interestingly, the above
results regarding the role of fixed telephony are in agreement with
previous results from {[}18{]} {[}19{]} which used a smaller panel of
countries but without constraints regarding the city size. However, our
results do not support the findings of {[}19{]} regarding internet
usage. In order to overcome the city size limitation, the next section
presents two case studies, for which we have obtained much more granular
data and therefore we are in a position to test the effect of the
internet on the tail of the urban population distribution for these
countries.

\hypertarget{sec5}{%
\section{The impact of ICT on the US and the UK spatial
structures}\label{sec5}}

This section focuses on the US and the UK urban systems, for the cities
of which equation \ref{eq4} from \protect\hyperlink{sec3.2}{Section 3.2}
is estimated separately.

\hypertarget{sec5.1}{%
\subsection{Internet and the US Spatial Structure: Evidence from the US
Micropolitan and Metropolitan Statistical Areas,
2013-2018}\label{sec5.1}}

We pursue further our investigation of the impact of internet adoption
by using a previously unutilized, to the best of our knowledge, for this
purpose data source. That is, for the first time in \(2013\), data on
internet use was made available via the American Community Survey and is
provided at the metropolitan area (urban areas comprised of one or more
adjacent counties or county equivalents that have at least one urban
core area of at least \(50,000\) population, plus adjacent territory
that has a high degree of social and economic integration with the core
as measured by commuting ties) and at the micropolitan area (defined
like metropolitan areas except that are comprised of an urban core of at
least \(10,000\), but less than \(50,000\) population) level of
aggregation {[}62{]}. These functional definitions of a city represent
in essence labor markets and according to Table \ref{desc.usa} the
observable minimum size of city population used for this analysis is
just above \(62,000\) habitants.

\begin{table}[!htbp] \centering 
  \caption{Descriptive statistics for the USA model\label{desc.usa}} 
  \label{} 
\small 
\begin{tabular}{@{\extracolsep{1pt}}lccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
normalized difference in ranks, 2013-18 & 461 & 0.480 & 0.500 & 0.008 & 0.546 \\ 
households w. internet 2013, \% & 461 & 0.307 & 0.716 & 0.073 & 0.871 \\ 
population 2013 & 461 & 62,282 & 568,372.100 & 1,370,757.000 & 19,949,502 \\ 
\% of unemployment 2013 & 461 & 2.300 & 8.530 & 2.591 & 19.900 \\ 
\% of white population 2013 & 443 & 0.173 & 0.806 & 0.129 & 0.963 \\ 
income 2013 & 461 & 15,455 & 24,336.230 & 3,802.734 & 41,498 \\ 
population density 2013 & 461 & 0.028 & 0.918 & 1.036 & 9.288 \\ 
employment in service 2011, \% & 461 & 49.000 & 78.996 & 6.277 & 92.300 \\ 
commute in minutes 2013 & 461 & 15 & 22.428 & 3.280 & 38 \\ 
pop. above 25 w. Bachelor's \\
                             degree 2005, \% & 461 & 5.600 & 14.832 & 4.727 & 33.900 \\ 
commute in minutes, 2005 & 461 & 14.000 & 21.739 & 3.436 & 40.700 \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

The LHS variable we use here is the outcome of equation \ref{eq5} in a
normalized form\footnote{\[Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.\]}
, so as that variable is bounded between \(0\) and \(1\) and with a mean
of \(0.5\). This transformation enables us to estimate \ref{eq4} not
only with OLS, but also with a quasibinomial GLM estimator given that
the original form of our LHS variable does not vary continuously, but is
instead defined as a difference between two count variables, which may
also assume negative values. The results of the different estimations
remain qualitatively the same regardless of the form of the LHS variable
we use. Table \ref{ols.usa} reports the results of the estimation of
equation \ref{eq4} using OLS. The sign and significance of the main
variable of interest -- percentage of population with computer and
broadband connection -- verifies our global level results. That is an
increase in internet usage improves the position of a city in the US
urban system. The results are consistent among the different estimators
that are OLS and GLM. In addition, Table \ref{ols.usa} also reports the
estimates of interaction effects between the share of population with
broadband connection with population and population density. Although
the interaction terms are not significant, the main effect remains
qualitatively unchanged.

\begin{table}[!htbp] \centering 
  \caption{OLS estimation of equation (4) for USA\label{ols.usa}} 
  \label{} 
\small 
\begin{tabular}{@{\extracolsep{1pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{Normalized difference in ranks 2013-18} \\ 
 & \multicolumn{3}{c}{OLS} & GLM \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 \% of households w. internet 2013 & 0.050$^{***}$ & 0.050$^{***}$ & 0.111$^{*}$ & 0.201$^{***}$ \\ 
  & (0.009) & (0.010) & (0.064) & (0.036) \\ 
  population 2013 (log) & $-$0.0002 & $-$0.0002 & 0.003 & $-$0.001 \\ 
  & (0.0004) & (0.0004) & (0.004) & (0.002) \\ 
  \% of unemployment 2013 & $-$0.0004$^{*}$ & $-$0.0003$^{*}$ & $-$0.0004$^{*}$ & $-$0.001$^{*}$ \\ 
  & (0.0002) & (0.0002) & (0.0002) & (0.001) \\ 
  \% of white population 2013 & 0.0001 & 0.0001 & $-$0.0002 & 0.0003 \\ 
  & (0.003) & (0.003) & (0.003) & (0.013) \\ 
  income 2013 (log) & $-$0.011$^{***}$ & $-$0.011$^{***}$ & $-$0.011$^{***}$ & $-$0.045$^{***}$ \\ 
  & (0.004) & (0.004) & (0.004) & (0.015) \\ 
  population density 2013 & $-$0.001$^{***}$ & $-$0.001 & $-$0.001$^{***}$ & $-$0.004$^{***}$ \\ 
  & (0.0003) & (0.004) & (0.0003) & (0.001) \\ 
  \% of employment in service 2011 & $-$0.00002 & $-$0.00002 & $-$0.00001 & $-$0.0001 \\ 
  & (0.0001) & (0.0001) & (0.0001) & (0.0003) \\ 
  commute in minutes 2013 & 0.0003 & 0.0003 & 0.0003 & 0.001 \\ 
  & (0.0002) & (0.0002) & (0.0002) & (0.001) \\ 
  \% of households w. internet, \\
                             2013 x pop. density, 2013 &  & 0.001 &  &  \\ 
  &  & (0.006) &  &  \\ 
  \% of households w. internet, \\
                             2013 x population, 2013 &  &  & $-$0.005 &  \\ 
  &  &  & (0.005) &  \\ 
  Constant & 0.581$^{***}$ & 0.581$^{***}$ & 0.532$^{***}$ & 0.323$^{**}$ \\ 
  & (0.035) & (0.036) & (0.063) & (0.142) \\ 
 \hline \\[-1.8ex] 
Observations & 443 & 443 & 443 & 443 \\ 
Adjusted R$^{2}$ & 0.122 & 0.120 & 0.122 &  \\ 
Residual Std. Error & 0.007 & 0.007 & 0.007 &  \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{4}{r}{Robust Std. Errors in parenthesis} \\ 
\end{tabular} 
\end{table}

Although not directly comparable, the above estimations are in
accordance with our global model. In addition, the potential presence of
endogeneity might be a problem here as well. Therefore, Table
\ref{2sls.usa} reports 2SLS estimations. In addition to the strategy
employed for the cross-country analysis -- that is the inclusion of one
instrument for which we have strong reasons to believe is uncorrelated
with the error term -- data availability allows to employ a second IV in
order to estimate the Sargan over-identification restrictions test
(Column 2). The main instrument we propose here is Bachelor's degrees
per inhabitant in \(2005\). Even if the quality of human capital can
affect the population growth of a city \(10\) years later, our LHS
variable adopts a systemic understanding of the US urban system as it
measures the relative position of a city within the overall urban system
instead of its population growth. In other words, a city might
experience population growth between two periods, but if other cities
have also experienced population growth, this might not affect its
relative position within the urban system. Hence, we do not expect that
this IV affects the LHS variable. Moreover, when we add the second IV,
which in this case is the commute time in minutes also in \(2005\), we
fail to reject the null hypothesis of the Sargan test, something which
also adds on the validity of our strategy. Furthermore, our results do
not suffer from weak identification according to the relevant tests in
Table \ref{2sls.usa}.

\begin{table}[!htbp] \centering 
  \caption{2SLS estimation of equation (4) for USA\label{2sls.usa}} 
  \label{} 
\small 
\begin{tabular}{@{\extracolsep{-15pt}}lD{.}{.}{-3} D{.}{.}{-3} } 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
\cline{2-3} 
\\[-1.8ex] & \multicolumn{2}{c}{Normalized difference in ranks, 2013-18} \\ 
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)}\\ 
\hline \\[-1.8ex] 
 \% of households w. internet 2013 & 0.091^{***} & 0.084^{***} \\ 
  & (0.020) & (0.021) \\ 
  population 2013 (log) & -0.001 & -0.001 \\ 
  & (0.0005) & (0.0005) \\ 
  \% of unemployment 2013 & -0.0004^{**} & -0.0004^{**} \\ 
  & (0.0002) & (0.0002) \\ 
  \% of white population 2013 & -0.006 & -0.005 \\ 
  & (0.005) & (0.005) \\ 
  income 2013 (log) & -0.024^{***} & -0.022^{***} \\ 
  & (0.006) & (0.006) \\ 
  population density 2013 & -0.001^{***} & -0.001^{***} \\ 
  & (0.0004) & (0.0004) \\ 
  \% of employment in service 2011 & -0.0001 & -0.0001 \\ 
  & (0.0001) & (0.0001) \\ 
  commute in minutes 2013 & 0.0004^{*} & 0.0003^{*} \\ 
  & (0.0002) & (0.0002) \\ 
  Constant & 0.691^{***} & 0.673^{***} \\ 
  & (0.059) & (0.058) \\ 
 \hline \\[-1.8ex] 
Weak instruments & 114.85 & 62.05 \\ 
Wu-Hausman & 5.23 & 3.58 \\ 
P-value & 0.02 & 0.06 \\ 
Sargan &  & 5.65 \\ 
P-value &  & 0.02 \\ 
Observations & \multicolumn{1}{c}{443} & \multicolumn{1}{c}{443} \\ 
Adjusted R$^{2}$ & \multicolumn{1}{c}{0.061} & \multicolumn{1}{c}{0.080} \\ 
Residual Std. Error & \multicolumn{1}{c}{0.008} & \multicolumn{1}{c}{0.008} \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{2}{r}{Robust Std. Errors in parenthesis} \\ 
 & \multicolumn{2}{r}{IVs: (1) Bachelors degree per hab. in 2005} \\ 
 & \multicolumn{2}{r}{IVs: (2) Bachelors degree per hab. in 2005, Commute, minutes, 2005} \\ 
\end{tabular} 
\end{table}

In total, the estimates of Table \ref{2sls.usa} enable us to identify a
causal effect of the share of households with internet connection on the
position of a city in the urban system. Specifically, an increase in
internet penetration led to an improvement of the position of a city in
the US urban hierarchy for the cities and the time period included in
our analysis. Everything else being equal, using the normalization
formula for the \(difference\:in\:ranks_{i}\), if a city whose position
in the US urban system did not change during the period \(2013-2018\)
(that is \(Normalized\:difference\:in\:ranks_{i} = 0.5\)) had
experienced an increase of 10 per cent points of internet penetration,
this increase would have improved its relative position by \(9\) places.
As in the global model reported in \protect\hyperlink{sec4}{Section 4},
internet penetration appears to work in favor of agglomeration
externalities. In regards to the control variables, we can identify a
negative effect of income, which is consistent with the effect of GDP
per capita in the global model presented in Table \ref{2sls.global}.
During the study period, unemployment rates and the share of white
population negatively affected the ranking of micropolitan and
metropolitan areas in the US as reflected in the relevant regressions.

\hypertarget{sec5.2}{%
\subsection{Internet and the UK Spatial Structure: Evidence from the
Built-up Areas in England and Wales, 2011-2018}\label{sec5.2}}

The next step in our analysis is to estimate equation \ref{eq4} for
cities in England and Wales for which we were able to access internet
speed micro-data. More specifically, we obtained individual speed
internet tests from \url{broadbandspeedchecker.co.uk}. This website
enables individuals to directly measure their upload and download
internet speed. The results of the tests as well as the geo-location of
the users are recorded by the website operator and were provided to us
in a fully anonymized manner. More discussion about the nature and the
validity of this data can be found in the work of Riddlesden and
Singleton {[}63{]}. The point nature of these individual level data
enables us to aggregate them up to any urban level that we are
interested in. Given that all the above analyses focused on functional
definitions of cities, we adopt here a morphological definition of a
city or a town for the UK, which enables us to test the effect of
internet on smaller areas irrespective of being part of a wider urban
agglomeration. Therefore we aggregate the internet speed data at the
level of Built-up Areas (BUA) for England and Wales. This is a `bricks
and mortar' approach which refers to land which is ``irreversibly urban
in character'' including villages, towns or cities. Some key
characteristics of these areas include: minimum size of \(20\) hectares;
areas with less than \(200\) meters between them are linked to a single
built-up area; larger built-up areas are separated to smaller
sub-divisions of built-up areas {[}64--66{]}.

In order to obtain information about the tail of the urban size
distribution we include, wherever available, the sub-divisions of BUA.
As Table \ref{desc.uk} illustrates, our approach results to \(6,163\)
observations which include built-up areas and sub-divisions of built-up
areas for which we have internet speed data. The lowest population of a
built-up area in our data is just above \(100\) inhabitants in \(2011\).
Figure \ref{figure1} illustrates the built-up areas we use for the
South-East of England and the Greater London Area.

\begin{table}[!htbp] \centering 
  \caption{Descriptive statistics for the UK model\label{desc.uk}} 
  \label{} 
\small 
\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
normalized difference in ranks, 2011-18 & 6,163 & 0.455 & 0.498 & 0.005 & 0.593 \\ 
download speed, 2011 & 6,163 & 520 & 3,745.519 & 2,523.129 & 50,280 \\ 
population, 2011 & 6,163 & 101 & 8,629.648 & 34,895.950 & 1,087,558 \\ 
broadband tests per capita, 2011 & 6,163 & 0.0001 & 0.034 & 0.062 & 2.220 \\ 
\% of unemployment, 2011 & 6,163 & 0.000 & 0.048 & 0.024 & 0.279 \\ 
\% of British population, 2011 & 6,163 & 0.150 & 0.944 & 0.063 & 1.000 \\ 
population density, 2011 & 6,163 & 0.578 & 24.973 & 11.607 & 141.744 \\ 
% of people working from home, 2011 & 6,163 & 0.024 & 0.155 & 0.063 & 0.564 \\ 
employment in service, 2011 (\%) & 6,163 & 0.511 & 0.776 & 0.054 & 0.971 \\ 
Number of universities & 6,163 & 0 & 0.021 & 0.250 & 11 \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

\begin{figure}
\includegraphics[width=0.8\linewidth]{Figure1} \caption{\label{figure1}Built-up areas in London and the South East of England}\label{fig:unnamed-chunk-11}
\end{figure}

In terms of our identification strategy, we follow the same approach as
we did for the US case. Table \ref{ols.uk} presents the OLS and GLM
estimation of equation \ref{eq4} while Table \ref{2sls.uk} presents the
2SLS estimation. Starting from Table \ref{ols.uk} we see a significant
and positive effect of the average download speed on the relative
position of a BUA in the urban hierarchy in England and Wales. This is
consistent with the previous results from the US and also with the
global model. Of course, the same endogeneity issues might be present
here too and we are addressing that below. In terms of control
variables, we include a measure of broadband tests per inhabitant in
order to control for potential differences in the take-up of this
service. In addition, we include a number of socio-economic variables
that we believe can affect the relative position of a city in the urban
hierarchy. Unemployment rate has a negative and significant effect as in
the US case. The same applies for the percentage of British population,
which indicates the importance of migration in urban growth. Population
density and the percentage of people working from home also negatively
affected the relative position of BUA during \(2011-2018\). Finally,
Table \ref{ols.uk} includes two interaction terms (columns \(2\) and
\(3\)) between download speed and population and population density. The
negative and significant sign for both interaction coefficients
indicates that the effect of download speed decreases as the size of the
BUA or its density increases. Contrary to the US case, this is
indicative of a larger digital connectivity effect for smaller BUA.

\begin{table}[!htbp] \centering 
  \caption{OLS estimation of equation (4) for the UK\label{ols.uk}} 
  \label{} 
\small 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{Normalized difference in ranks, 2011-18} \\ 
 & \multicolumn{3}{c}{OLS} & GLM \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 download speed, 2011 (log) & 0.0004$^{***}$ & 0.001$^{***}$ & 0.002$^{**}$ & 0.0004$^{***}$ \\ 
  & (0.0002) & (0.0004) & (0.001) & (0.0002) \\ 
  population, 2011 (log) & 0.0004$^{***}$ & 0.0004$^{***}$ & 0.002$^{***}$ & 0.0004$^{***}$ \\ 
  & (0.0001) & (0.0001) & (0.001) & (0.0001) \\ 
  broadband tests per capita, 2011 & $-$0.0003 & $-$0.0002 & $-$0.0002 & $-$0.0003 \\ 
  & (0.001) & (0.001) & (0.001) & (0.001) \\ 
  \% of unemployment, 2011 & $-$0.011$^{***}$ & $-$0.011$^{***}$ & $-$0.011$^{***}$ & $-$0.011$^{***}$ \\ 
  & (0.004) & (0.004) & (0.004) & (0.004) \\ 
  \% of British population, 2011 & $-$0.003$^{**}$ & $-$0.003$^{**}$ & $-$0.003$^{**}$ & $-$0.003$^{**}$ \\ 
  & (0.001) & (0.001) & (0.001) & (0.001) \\ 
  population density, 2011 & 0.00000 & 0.0002$^{*}$ & 0.00000 & 0.00000 \\ 
  & (0.00001) & (0.0001) & (0.00001) & (0.00001) \\ 
  \% of people working from home, 2011 & $-$0.006$^{**}$ & $-$0.005$^{*}$ & $-$0.006$^{*}$ & $-$0.006$^{**}$ \\ 
  & (0.003) & (0.003) & (0.003) & (0.003) \\ 
  employment in service, 2011 (\%) & $-$0.005$^{**}$ & $-$0.005$^{**}$ & $-$0.005$^{**}$ & $-$0.005$^{**}$ \\ 
  & (0.002) & (0.002) & (0.002) & (0.002) \\ 
  download speed, 2011 (log) \\
                             x pop. density, 2011 &  & $-$0.00003$^{*}$ &  &  \\ 
  &  & (0.00001) &  &  \\ 
  download speed, 2011 (log) \\
                             x population, 2011 (log) &  &  & $-$0.0002$^{**}$ &  \\ 
  &  &  & (0.0001) &  \\ 
  Constant & 0.499$^{***}$ & 0.494$^{***}$ & 0.490$^{***}$ & 0.499$^{***}$ \\ 
  & (0.003) & (0.004) & (0.006) & (0.003) \\ 
 \hline \\[-1.8ex] 
Observations & 3,546 & 3,546 & 3,546 & 3,546 \\ 
Adjusted R$^{2}$ & 0.049 & 0.050 & 0.049 &  \\ 
Log Likelihood &  &  &  & 14,294.670 \\ 
Akaike Inf. Crit. &  &  &  & $-$28,571.330 \\ 
Residual Std. Error & 0.004 & 0.004 & 0.004 &  \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{4}{r}{Robust Std. Errors} \\ 
\end{tabular} 
\end{table}

In regards to IV strategy, we use here the number of universities as the
main instrument. Universities in England and Wales have been established
long time ago and therefore we do not expect them to affect the change
in the relative position of built-up areas in \(2011\). The addition of
the second instrument (absolute number of broadband tests in 2011) in
column \(2\) results to a Sargan test, the null hypothesis of which
cannot be rejected. Moreover, the estimations reported in Table 11 do
not suffer from weak instruments as the relevant test is above the rule
of thumb value of \(10\). The coefficients derived from OLS (Table
\ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are of the same magnitude
and always positive and significant, something which advocates towards
the effect of broadband speed in enhancing agglomeration forces. A back
of the envelope calculation indicates that, everything else being equal,
\(1\) per cent increase in download speed for a BUA, the relative
position of which did not change during the period \(2011-2014\), would
have increased its position by \(0.3\) places. In absolute numbers,
\(1\) per cent increase for the average BUA would mean a marginal change
from \(13.8\) Mbps to \(13.9\) Mbps.

\begin{table}[!htbp] \centering 
  \caption{2SLS estimation of equation (4) for the UK\label{2sls.uk}} 
  \label{} 
\small 
\begin{tabular}{@{\extracolsep{5pt}}lcc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
\cline{2-3} 
\\[-1.8ex] & \multicolumn{2}{c}{Normalized difference in ranks, 2011-18} \\ 
\\[-1.8ex] & (1) & (2)\\ 
\hline \\[-1.8ex] 
 download speed, 2011 (log) & 0.003$^{***}$ & 0.004$^{***}$ \\ 
  & (0.001) & (0.0005) \\ 
  population, 2011 (log) & $-$0.0001 & $-$0.0002$^{*}$ \\ 
  & (0.0002) & (0.0001) \\ 
  broadband tests per capita, 2011 & $-$0.001 & $-$0.0004 \\ 
  & (0.001) & (0.001) \\ 
  \% of unemployment, 2011 & $-$0.007 & $-$0.007 \\ 
  & (0.005) & (0.004) \\ 
  \% of British population, 2011 & $-$0.003$^{**}$ & $-$0.003$^{**}$ \\ 
  & (0.001) & (0.001) \\ 
  population density, 2011 & 0.00000 & $-$0.00000 \\ 
  & (0.00001) & (0.00001) \\ 
  \% of people working from home, 2011 & $-$0.007$^{**}$ & $-$0.007$^{**}$ \\ 
  & (0.003) & (0.003) \\ 
  employment in service, 2011 (\%) & $-$0.004$^{*}$ & $-$0.004$^{*}$ \\ 
  & (0.002) & (0.002) \\ 
  Constant & 0.481$^{***}$ & 0.478$^{***}$ \\ 
  & (0.007) & (0.004) \\ 
 \hline \\[-1.8ex] 
Weak instruments & 7.32 & 42.23 \\ 
Wu-Hausman & 6.15 & 40.92 \\ 
P-value & 0.01 & 0 \\ 
Sargan &  & 0.03 \\ 
P-value &  & 0.86 \\ 
Observations & 3,032 & 3,032 \\ 
Adjusted R$^{2}$ & $-$0.058 & $-$0.097 \\ 
Residual Std. Error & 0.004 & 0.004 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{2}{r}{Robust Std. Errors} \\ 
 & \multicolumn{2}{r}{IVs: (1) N.r of universities} \\ 
 & \multicolumn{2}{r}{IVs: (2) N. of universities, N. of broadband tests, 2011} \\ 
\end{tabular} 
\end{table}

In total, the results of the UK case study are in accordance with the
previous results of the global, multi-country analysis and of the US
urban system. Interestingly, the positive effect of internet is still
apparent when the analysis adopts a morphological definition of cities
and when the emphasis is not on internet penetration, but instead on the
quality of internet infrastructure as reflected in download broadband
speed. What is also interesting for the UK case study is that this
effect appears to be higher for small and less dense BUA.

\hypertarget{sec6}{%
\section{Discussion and conclusions}\label{sec6}}

This paper reports empirical findings on a question at the heart of
urban economics and economic geography: has the proliferation of
information and communication technologies offset the benefits of
agglomeration economies and resulted in more dispersed spatial
population structures, or has it further reinforced such urban
externalities and led to more concentrated spatial structures? Previous
studies have led to contradictory results regarding whether ICT adoption
and urban agglomeration externalities are complementary or
substitutable. As Leamer and Storper {[}33{]} stress, the internet can
affect both centripetal and centrifugal forces. The present study
revisits Ioannides et al.~{[}18{]} with a completely open mind in view
of the availability of several years of additional data on internet
penetration across the world. It examines the robustness of the findings
of Tranos and Ioannides {[}19{]} by employing alternative data sets. In
general, quite a few earlier studies have either been based on
assumptions about technological capabilities that might no longer hold
today, or use data that do not fully capture the widespread adoption and
maturity of communication technologies that has taken place since the
data coverage of those earlier studies.

We report estimation results using international country-level data on
urban agglomerations with more than \(300,000\) inhabitants from the UN
Urban Settlements data and test the effect of ICT adoption on the Zipf,
or Pareto, coefficients for national city size distributions, as
measures of dispersion for heavy-tailed data. We report results which
are robust to endogeneity concerns. Then, in order to examine such
effects for cities that are smaller than those included in our
international data, we focus on the US and the UK urban systems.
Specifically, we test the effect of internet usage and internet speed on
the changes over time in rankings of Micropolitan and Metropolitan Areas
in the US, and of Built Up Areas in the UK. The results favor a
complementary relation between the internet and agglomeration
externalities. While the cross-country estimates indicate that increase
adoption rates of such technologies has resulted in less dispersed urban
spatial structures, the two case studies reveal that internet adoption
and internet speed improved the relative population rank of a city
within its urban system. Interestingly, the latter results indicate that
such effects might be even stronger for smaller and less dense urban
areas at least in the UK. On the contrary, our estimates regarding the
effect of fixed telephony advocate towards the reverse effect: that is
fixed telephony acting against of agglomeration forces.

While our results are in agreement with Ioannides et al.~{[}18{]}
regarding fixed telephony, they contradict Tranos and Ioannides {[}19{]}
The latter cannot of course be ignored. We believe strongly that the
differences lie in the different definitions of urban areas. To this
date, and in spite of numerous efforts by teams worldwide, there has not
been a universally accepted definition, which could moreover be
implemented consistently by governments worldwide\footnote{The
  increasingly popular reliance on high-resolution, pioneered by
  Rozenfeld et al.~{[}67{]}, and satellite-based lights data, most
  recently used by Duben and Krause {[}68{]}, for measuring city sizes,
  are very exciting developments, but they have not yet been accompanied
  by complementary information on ICT adoption at a comparably granular
  level. A promising new development is the EU's Global Human
  Settlements Initiative and the associated spatial urban data sets; see
  Pesaresi et al.~{[}69{]}.}. Barring the availability of such data,
scholars must be careful in drawing conclusions. Arguably, the
measurement of the impact of ICT at the level of agglomerations, on one
hand, and on changes in rankings, on the other, reported by this paper
constitute a contribution to the literature.

We believe that our results, apart from their theoretical value, have
the capacity to inform urban policy. The ability of the internet and
digital communications to further enhance agglomeration economies can be
used as a tool to support urban growth. In addition, the indication that
such effects might be stronger for smaller and less dense urban areas,
at least in England and Wales, might be helpful to further orient
digital strategies towards such locations. Of course, improvements in
internet speed is not a trivial policy instrument as it involves
numerous complexities. Apart from infrastructure installation costs and
engineering challenges, governance issues regarding the ownership of
such digital networks as well as the provision of state subsidies create
obstacles for the inclusion of such strategies in the urban growth
agenda. To further inform such policies, more research at granular
scales is needed in order to shed lights on the micro-mechanisms behind
such urban processes.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-toffler_third_1980}{}%
1. Toffler A. Third way. New York: William Morrow; 1980.

\leavevmode\hypertarget{ref-ohmae_borderless_1995}{}%
2. Ohmae K. The Borderless World: Power and Strategy in an
Interdependent Economy. New York: Harper Business; 1995.

\leavevmode\hypertarget{ref-gilder_notitle_1995}{}%
3. Gilder G. Forbes ASAP. 1995; 56.

\leavevmode\hypertarget{ref-peter_f_drucker_capitalism_1998}{}%
4. Drucker PF. From Capitalism to Knowledge Society. In: Neef D, editor.
The knowledge economy. Woburn, MA: Butterworth-Heinemann; 1998. pp.
15--34.

\leavevmode\hypertarget{ref-obrien_global_1992}{}%
5. O'Brien R. Global Financial Integration: The End of Geography.
London: Pinter; 1992.

\leavevmode\hypertarget{ref-frances_cairncross_death_2001}{}%
6. Cairncross F. The death of distance 2.0. London: Texere Publishing
Limited; 2001.

\leavevmode\hypertarget{ref-friedman_world_2005}{}%
7. Friedman TL. The world is flat. New York: Farrar, Straus; Giroux;
2005.

\leavevmode\hypertarget{ref-kende_digital_2003}{}%
8. Kende M. The digital handshake: Connecting Internet backbones.
CommLaw Conspectus. 2003;11: 45.

\leavevmode\hypertarget{ref-united_nations_world_2014}{}%
9. Nations U. World Urbanization Prospects: The 2014 Revision.
Department of Economic; Social Affairs, Population Division, UN; 2014.

\leavevmode\hypertarget{ref-panahi_towards_2013}{}%
10. Panahi S, Watson J, Partridge H. Towards tacit knowledge sharing
over social web tools. Journal of Knowledge Management. 2013;17:
379--397.

\leavevmode\hypertarget{ref-storper_buzz_2004}{}%
11. Storper M, Venables AJ. Buzz: Face-to-face contact and the urban
economy. Journal of economic geography. 2004;4: 351--370.

\leavevmode\hypertarget{ref-tranos2020social}{}%
12. Tranos E. Social network sites and knowledge transfer: An urban
perspective. Journal of Planning Literature. SAGE Publications Sage CA:
Los Angeles, CA; 2020; 0885412220921526.

\leavevmode\hypertarget{ref-rosenthal_evidence_2004}{}%
13. Rosenthal SS, Strange WC. Evidence on the nature and sources of
agglomeration economies. Handbook of regional and urban economics.
2004;4: 2119--2171.

\leavevmode\hypertarget{ref-mccann_globalization_2008}{}%
14. McCann P. Globalization and economic geography: The world is curved,
not flat. Cambridge Journal of Regions, Economy and Society. 2008;1:
351--370.

\leavevmode\hypertarget{ref-duranton_micro-foundations_2004}{}%
15. Duranton G, Puga D. Micro-foundations of urban agglomeration
economies. Handbook of regional and urban economics. 2004;4: 2063--2117.

\leavevmode\hypertarget{ref-van_dijck_you_2013}{}%
16. Van Dijck J. ``You have one identity'': Performing the self on
Facebook and LinkedIn. Media, Culture \& Society. 2013;35: 199--215.

\leavevmode\hypertarget{ref-gee2019more}{}%
17. Gee LK. The more you know: Information effects on job application
rates in a large field experiment. Management Science. INFORMS; 2019;65:
2077--2094.

\leavevmode\hypertarget{ref-ioannides2008effect}{}%
18. Ioannides YM, Overman HG, Rossi-Hansberg E, Schmidheiny K. The
effect of information and communication technologies on urban structure.
Economic Policy. Oxford University Press; 2008;23: 202--242.

\leavevmode\hypertarget{ref-tranos2020ict}{}%
19. Tranos E, Ioannides YM. ICT and cities revisited. Telematics and
Informatics. Elsevier; 2020; 101439.

\leavevmode\hypertarget{ref-jess_gaspar_information_1998}{}%
20. Gaspar J, Glaeser EL. Information Technology and the Future of
Cities. Journal of Urban Economics. 1998;43: 136--156.

\leavevmode\hypertarget{ref-hildrum_sharing_2009}{}%
21. Hildrum JM. Sharing Tacit Knowledge Online: A Case Study of
e-Learning in Cisco's Network of System Integrator Partner Firms.
Industry and Innovation. 2009;16: 197--218.

\leavevmode\hypertarget{ref-kolko_death_2000}{}%
22. Kolko J. The death of cities? The death of distance? Evidence from
the geography of commercial Internet usage. In: Vogelsang I, Compaine
BM, editors. The internet upheaval: Raising questions, seeking answers
in communications policy. Cambridge, MA; London, UK: The MIT Press;
2000. pp. 73--98.

\leavevmode\hypertarget{ref-todd_sinai_geography_2004}{}%
23. Sinai T, Waldfogel J. Geography and the Internet: Is the Internet a
substitute or a complement for cities? Journal of Urban Economics.
2004;56: 1--24.

\leavevmode\hypertarget{ref-chris_forman_how_2005}{}%
24. Forman C, Goldfarb A, Greenstein S. How did location affect adoption
of the commercial Internet? Global village vs. Urban leadership. Journal
of Urban Economics. 2005;58: 389--420.

\leavevmode\hypertarget{ref-jungyul_sohn_information_2003}{}%
25. Sohn J, Kim TJ, Hewings GJD. Information technology and urban
spatial structure: A comparative analysis of the Chicago and Seoul
regions. Annals of Regional Science. 2003;37: 447--462.

\leavevmode\hypertarget{ref-pons-novell_cities_2006}{}%
26. Pons-Novell J, Viladecans-Marsal E. Cities and the Internet: The end
of distance? Journal of Urban Technology. 2006;13: 109--132.

\leavevmode\hypertarget{ref-craig_does_2016}{}%
27. Craig SG, Hoang EC, Kohlhase JE. Does closeness in virtual space
complement urban space? Socio-Economic Planning Sciences. 2016;

\leavevmode\hypertarget{ref-mark_d_partridge_employment_2008}{}%
28. Partridge MD, Rickman DS, Ali K, Olfert MR. Employment Growth in the
American Urban Hierarchy: Long Live Distance. The BE Journal of
Macroeconomics. 2008;8: Article 10.

\leavevmode\hypertarget{ref-kim_broadband_2016}{}%
29. Kim Y, Orazem PF. Broadband Internet and New Firm Location Decisions
in Rural Areas. American Journal of Agricultural Economics. 2016;99:
285--302.

\leavevmode\hypertarget{ref-bekkerman_high-speed_2013}{}%
30. Bekkerman A, Gilpin G. High-speed Internet growth and the demand for
locally accessible information content. Journal of Urban Economics.
2013;77: 1--10.

\leavevmode\hypertarget{ref-anenberg_information_2015}{}%
31. Anenberg E, Kung E. Information technology and product variety in
the city: The case of food trucks. Journal of Urban Economics. 2015;90:
60--78.

\leavevmode\hypertarget{ref-rietveld_transport_2004}{}%
32. Rietveld P, Vickerman RW. Transport in regional science: The ``death
of distance'' is premature. Papers in Regional Science. 2004;83:
229--248.

\leavevmode\hypertarget{ref-edward_e_leamer_economic_2001}{}%
33. Leamer EE, Storper M. The Economic Geography of the Internet Age.
Journal of International Business Studies. 2001;32: 641--65.

\leavevmode\hypertarget{ref-pew_internet_pew_2016}{}%
34. Internet P. Pew Internet \& American Life Project. Pew Internet;
2016.

\leavevmode\hypertarget{ref-internet_live_stats_internet_2017}{}%
35. Stats IL. Internet Users. 2017.

\leavevmode\hypertarget{ref-black_urban_2003}{}%
36. Black D, Henderson V. Urban evolution in the USA. Journal of
economic geography. 2003;3: 343--372.

\leavevmode\hypertarget{ref-frenken_theoretical_2007}{}%
37. Frenken K, Boschma RA. A theoretical framework for evolutionary
economic geography: Industrial dynamics and urban growth as a branching
process. Journal of economic geography. 2007;7: 635--649.

\leavevmode\hypertarget{ref-giesen_zipfs_2011}{}%
38. Giesen K, Südekum J. Zipf's law for cities in the regions and the
country. Journal of economic geography. 2011;11: 667--686.

\leavevmode\hypertarget{ref-rauch_cities_2013}{}%
39. Rauch F. Cities as spatial clusters. Journal of economic geography.
2013; lbt034.

\leavevmode\hypertarget{ref-ioannides_zipfs_2003}{}%
40. Ioannides YM, Overman HG. Zipf's law for cities: An empirical
examination. Regional science and urban economics. 2003;33: 127--137.

\leavevmode\hypertarget{ref-ioannides_walled_2017}{}%
41. Ioannides YM, Zhang J. Walled cities in late imperial China. Journal
of Urban Economics. 2017;97: 71--88.

\leavevmode\hypertarget{ref-nitsch_zipf_2005}{}%
42. Nitsch V. Zipf zipped. Journal of Urban Economics. 2005;57: 86--100.

\leavevmode\hypertarget{ref-gabaix_evolution_2004}{}%
43. Gabaix X, Ioannides YM. The evolution of city size distributions.
Handbook of regional and urban economics. 2004;4: 2341--2378.

\leavevmode\hypertarget{ref-gabaix_rank_2011}{}%
44. Gabaix X, Ibragimov R. Rank− 1/2: A simple way to improve the OLS
estimation of tail exponents. Journal of Business \& Economic
Statistics. 2011;29: 24--39.

\leavevmode\hypertarget{ref-ioannides2013neighborhoods}{}%
45. Ioannides Y. From neighborhoods to nations: The economics of social
interactions. Princeton University Press; 2013.

\leavevmode\hypertarget{ref-thomas_brinkhoff_city_2014}{}%
46. Brinkhoff T. City Population. 2014.

\leavevmode\hypertarget{ref-soo_zipfs_2005}{}%
47. Soo KT. Zipf's Law for cities: A cross-country investigation.
Regional science and urban economics. 2005;35: 239--263.
doi:\href{https://doi.org/http://dx.doi.org/10.1016/j.regsciurbeco.2004.04.004}{http://dx.doi.org/10.1016/j.regsciurbeco.2004.04.004}

\leavevmode\hypertarget{ref-cohen_urban_2004}{}%
48. Cohen B. Urban growth in developing countries: A review of current
trends and a caution regarding existing forecasts. World development.
2004;32: 23--51.

\leavevmode\hypertarget{ref-montgomery_urban_2008}{}%
49. Montgomery MR. The urban transformation of the developing world.
science. 2008;319: 761--764.

\leavevmode\hypertarget{ref-chen_absolute_2007}{}%
50. Chen S, Ravallion M. Absolute poverty measures for the developing
world, 1981--2004. Proceedings of the National Academy of Sciences.
2007;104: 16757--16762.

\leavevmode\hypertarget{ref-decker_global_2007}{}%
51. Decker EH, Kerkhoff AJ, Moses ME. Global patterns of city size
distributions and their fundamental drivers. PLoS One. 2007;2: e934.

\leavevmode\hypertarget{ref-donner_research_2008}{}%
52. Donner J. Research approaches to mobile use in the developing world:
A review of the literature. The information society. 2008;24: 140--159.

\leavevmode\hypertarget{ref-hamilton_are_2003}{}%
53. Hamilton J. Are main lines and mobile phones substitutes or
complements? Evidence from Africa. Telecommunications Policy. 2003;27:
109--133.

\leavevmode\hypertarget{ref-batty_rank_2006}{}%
54. Batty M. Rank clocks. Nature. 2006;444: 592--596.

\leavevmode\hypertarget{ref-havlin_distance_1995}{}%
55. Havlin S. The distance between Zipf plots. Physica A: Statistical
Mechanics and its Applications. 1995;216: 148--150.

\leavevmode\hypertarget{ref-chetty2018impacts}{}%
56. Chetty R, Hendren N. The impacts of neighborhoods on
intergenerational mobility i: Childhood exposure effects. The Quarterly
Journal of Economics. Oxford University Press; 2018;133: 1107--1162.

\leavevmode\hypertarget{ref-dettling2017}{}%
57. Dettling LJ. Broadband in the labor market: The impact of
residential high-speed internet on married women's labor force
participation. ILR Review. 2017;70: 451--482.
doi:\href{https://doi.org/10.1177/0019793916644721}{10.1177/0019793916644721}

\leavevmode\hypertarget{ref-arntz2019working}{}%
58. Arntz M, Sarra BY, Berlingieri F. Working from home: Heterogeneous
effects on hours worked and wages. ZEW-Centre for European Economic
Research Discussion Paper. 2019;

\leavevmode\hypertarget{ref-watson2018digitalisation}{}%
59. Watson T, Corliss M, Le M, others. Digitalisation and women's
workforce participation in the indo-pacific. Australian Journal of
Labour Economics. Centre for Labour Market Research; 2018;21: 45.

\leavevmode\hypertarget{ref-Lechman2020}{}%
60. Lechman E, Popowska M. Enhancing women's engagement in economic
activities through information and communication technology deployment:
Evidence from central--eastern european countries. Gender, Technology
and Development. Routledge; 2020;24: 314--340.
doi:\href{https://doi.org/10.1080/09718524.2020.1824568}{10.1080/09718524.2020.1824568}

\leavevmode\hypertarget{ref-Valberg2020}{}%
61. Valberg S. ICT, gender, and the labor market: A cross-country
analysis. In: Maiti D, Castellacci F, Melchior A, editors.
Digitalisation and development: Issues for india and beyond. Singapore:
Springer Singapore; 2020. pp. 375--405.
doi:\href{https://doi.org/10.1007/978-981-13-9996-1_15}{10.1007/978-981-13-9996-1\_15}

\leavevmode\hypertarget{ref-file_computer_2014}{}%
62. File T, Ryan C. Computer and Internet use in the United States:
2013. American Community Survey Reports. 2014;

\leavevmode\hypertarget{ref-riddlesden_broadband_2014}{}%
63. Riddlesden D, Singleton AD. Broadband speed equity: A new digital
divide? Applied Geography. 2014;52: 25--33.

\leavevmode\hypertarget{ref-ons2013}{}%
64. ONS. 2011 Built-up Areas - Methodology and Guidance. Office for
National Statistics; 2013 pp. 1--15.

\leavevmode\hypertarget{ref-johnston_changing_2014}{}%
65. Johnston R, Poulsen M, Forrest J. The changing ethnic composition of
urban neighbourhoods in England and Wales, 2001-2011: Creating nations
of strangers? Geography. 2014;99: 67.

\leavevmode\hypertarget{ref-johnston2016ethnic}{}%
66. Johnston R, Poulsen M, Forrest J. Ethnic residential patterns in
urban england and wales, 2001--2011: A system-wide analysis. Tijdschrift
voor economische en sociale geografie. Wiley Online Library; 2016;107:
1--15.

\leavevmode\hypertarget{ref-rozenfeld2011area}{}%
67. Rozenfeld HD, Rybski D, Gabaix X, Makse HA. The area and population
of cities: New insights from a different perspective on cities. American
Economic Review. 2011;101: 2205--25.

\leavevmode\hypertarget{ref-duben2019population}{}%
68. Duben C, Krause M, others. Population, light, and the size
distribution of cities. 2019.

\leavevmode\hypertarget{ref-pesaresi2019ghs}{}%
69. Pesaresi M, Florczyk A, Schiavina M, Melchiorri M, Maffenini L. GHS
settlement grid, updated and refined regio model 2014 in application to
ghs-built r2018a and ghs-pop r2019a, multitemporal (1975-1990-2000-2015)
r2019a. European Commission, Joint Research Centre (JRC). 2019;

\nolinenumbers

\end{document}

### 2020-11-19 src/s1.Rmd
paper on git

@@
 knit: (function(inputFile, encoding) {
-    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs")
+    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/2_resubmission")
   })
@@
 # and CV. 
-# The output S1.pdf is located in /outputs
+# The output S1.pdf is located in /paper/2_resubmission
 

### 2020-12-01 paper/2_resubmission/Tranos Ioannides Revision November 19 2020.tex
responses

--- 2020-11-19 paper/2_resubmission/tranos_ioannides.tex
@@
 
+The present paper does not address the issue that adoption of ICT may have different effects on urban dispersion on a national scale, that is how far are major areas, that is, cities, from one another, versus dispersion within major urban areas, that is urban sprawl.
+Such an inquiry would require detailed data about patterns how urban sprawl  may hamper or promote economic interaction, given advancements in information technology.
+We think that data from the Covid-19 pandemic would likely throw light at this issue, as it is clear that not all economic activities might continue in an unencumbered manner; see Dingel and Neiman (2020).
+
+Jonathan Dingel and Brett Neiman, "How Many Jobs Can Be Done at Home?", Becker Friedman Institute, University of Chicago, April 2020. https://tinyurl.com/y9nbmkmz
+
 \hypertarget{sec3}{%
@@
 The multi-country identification strategy is a two-step approach
-{[}18{]}. The first step of our methodology is to estimate the Zipf
-coefficient for a broad sample of countries over time. Zipf coefficient
+{[}18{]}. The first step of our methodology is to estimate the Pareto exponent
+coefficient for a broad sample of countries over time. \textbf{The Pareto} coefficient
 is one of the most widely used measures of spatial dispersion with
@@
 
-This leads to an estimate of \(\zeta\), known as the Zipf coefficient,
-but more appropriately the Pareto exponent, when \(\zeta\) is not
-estimated to be near \(1\). While much of the literature focuses on the
-Zipf coefficient being close to \(1\), several estimations of city size
+This leads to an estimate of \(\zeta\), known as the the Zipf coefficient,
+but more appropriately the Pareto exponent, a terminology we will adhere to for the remained of the text, even  when \(\zeta\) is 
+estimated to be near \(1\). The term Zipf coefficient should be reserved fo the special case of \(\zeta=1\), While much of the literature focuses on the
+Pareto exponent being close to \(1\), several estimations of city size
 distributions lead to estimates of \(\zeta\) that are not necessarily
@@
 
-\caption{\label{tab:unnamed-chunk-2}Zipf coefficients\label{zipf}}
+\caption{\label{tab:unnamed-chunk-2}Pareto Exponent\label{zipf}}
 \resizebox{\linewidth}{!}{
@@
 \toprule
-Zipf coef. & 2000 & 2001 & 2002 & 2003 & 2004 & 2005 & 2006 & 2007 & 2008 & 2009 & 2010 & 2011 & 2012 & 2013 & 2014 & 2015 & 2016 & 2017 & 2018\\
+Pareto exp. & 2000 & 2001 & 2002 & 2003 & 2004 & 2005 & 2006 & 2007 & 2008 & 2009 & 2010 & 2011 & 2012 & 2013 & 2014 & 2015 & 2016 & 2017 & 2018\\
 \midrule
@@
 \hline \\[-1.8ex]
-Zipf coefficient & 950 & $-$1.9 & $-$1.1 & 0.2 & $-$0.7 \\ 
-Zipf inv. sq. error\textsuperscript{a} & 950 & 0.01 & 0.1 & 0.05 & 0.2 \\ 
+Pareto exponent & 950 & $-$1.9 & $-$1.1 & 0.2 & $-$0.7 \\
+Pareto inv. sq. error\textsuperscript{a} & 950 & 0.01 & 0.1 & 0.05 & 0.2 \\
 Population density & 931 & 2.5 & 126.2 & 182.1 & 1,239.6 \\
@@
 \hline \\[-1.8ex]
-\multicolumn{6}{l}{\textsuperscript{a}This is the inverse squared standard error of the estimated Zipf, coefficient, which is used for weighting} \\ 
+\multicolumn{6}{l}{\textsuperscript{a}This is the inverse squared standard error of the estimated Pareto exponent, which is used for weighting} \\
 \multicolumn{6}{l}{the observations for the estimation, of equation 3. See Section 4} \\
@@
 does not address potential simultaneity issues. Simply put, internet
-penetration might be affected by spatial structure, as reflected in Zipf
-coefficients, or both internet penetration and spatial structure might
+penetration might be affected by spatial structure, as reflected in the Pareto 
+exponents, or both internet penetration and spatial structure might
 be jointly determined by a third variable. E.g., if a country already
@@
 This section presents the estimation results of \ref{eq3}. The LHS
-variable is \(\zeta\), the Zipf coefficient, as estimated according to
+variable is \(\zeta\), the Pareto exponent, as estimated according to
 the Gabaix and Ibragimov {[}44{]} correction. The main explanatory
@@
 In addition, the observations are weighted with the inverse squared
-standard error of the estimated Zipf coefficient (see Table
+standard error of the estimated  the Pareto exponent (see Table
 \ref{desc.global}) to address potential noise that is carried over from
 the first part of our identification strategy. In regards to the
-interpretation of the estimated coefficients, given that the Zipf
-coefficient has entered the regression not as an absolute value, but
+interpretation of the estimated coefficients, given that the Pareto exponent
+ has entered the regression not as an absolute value, but
 instead as a real number a positive coefficient for a RHS variable
@@
 \cline{2-5}
-\\[-1.8ex] & \multicolumn{4}{c}{Zipf Coefficient 2000-18} \\ 
+\\[-1.8ex] & \multicolumn{4}{c}{ Pareto Exponent 2000-18} \\
 \\[-1.8ex] & (1) & (2) & (3) & (4)\\
@@
 effect. Endogeneity might be an issue here as spatial structure, which
-is proxied by the Zipf coefficient, might be affected by another source,
+is proxied by the Pareto exponent, might be affected by another source,
 which also affects internet penetration. For instance, economic
@@
 \cline{2-5}
-\\[-1.8ex] & \multicolumn{4}{c}{Zipf Coefficient 2000-18} \\ 
+\\[-1.8ex] & \multicolumn{4}{c}{ Pareto exponent, 2000-18} \\
 \\[-1.8ex] & (1) & (2) & (3) & (4)\\
@@
 but are limited to urban agglomeration included in our data. Indeed, for
-the estimation of the Zipf coefficient we only included urban
+the estimation of  the Pareto exponent we only included urban
 agglomeration of \(300,000\) inhabitants or more due to data
@@
 results do not support the findings of {[}19{]} regarding internet
-usage. In order to overcome the city size limitation, the next section
+usage.
+
+It is indeed puzzling that the results on Table \ref{2sls.global} that the effects on urban concentration for internet users and broadband users, on the one hand, and mobile and fixed telephony users are in opposite directions.
+We wish to broach this subject in the context of some related literature, namely Calabrese \emph{et al.} (2011)
+
+Interplay between Telecommunications and Face-to-Face Interactions: A Study Using Mobile Phone Data
+
+Francesco Calabrese ,Zbigniew Smoreda,Vincent D. Blondel,Carlo Ratti
+
+Published: July 13, 2011https://doi.org/10.1371/journal.pone.0020814
+
+and
+
+B\"{u}chel and Ehrlich
+
+Cities and the Structure of Social Interactions:
+Evidence from Mobile Phone Data
+
+http://staff.vwi.unibe.ch/vonehrlich/Papers/BuchelEhrlich012019.pdf
+
+The latter supports the notion that ICT favors agglomeration in the sense that in spite of their convenience modern technologies do no make up for the need of proximity in  social interactions.
+The former supports the notion people who communicate more frequently are likely to be near one another.
+ICT, as viewed through such granular studies of individuals' communications does not make up for physical distance; individuals nearer one another communicate more.
+
+In order to overcome the city size limitation, the next section
 presents two case studies, for which we have obtained much more granular
@@
 urban agglomerations with more than \(300,000\) inhabitants from the UN
-Urban Settlements data and test the effect of ICT adoption on the Zipf,
-or Pareto, coefficients for national city size distributions, as
+Urban Settlements data and test the effect of ICT adoption on the Pareto exponents for national city size distributions, as
 measures of dispersion for heavy-tailed data. We report results which
@@
 \leavevmode\hypertarget{ref-gabaix_rank_2011}{}%
-44. Gabaix X, Ibragimov R. Rank− 1/2: A simple way to improve the OLS
+44. Gabaix X, Ibragimov R. Rank- 1/2: A simple way to improve the OLS
 estimation of tail exponents. Journal of Business \& Economic

### 2020-12-01 paper/2_resubmission/tranos_ioannides.Rmd
responses

@@
 
+\color{blue}
+The present paper does not address the issue that adoption of ICT may have
+different effects on urban dispersion on a national scale, that is how far are
+major areas, that is, cities, from one another, versus dispersion within major
+urban areas, that is urban sprawl. Such an inquiry would require detailed data
+about patterns of how urban sprawl may hamper or promote economic interaction,
+given advancements in information technology.
+We think that data from the Covid-19 pandemic would likely throw light at this
+issue, as it is clear that not all economic activities might continue in an
+unencumbered manner; see Dingel and Neiman [@dingel2020many].
+\color{black}
+
 # Materials and Methods {#sec3}
@@
 The multi-country identification strategy is a two-step approach [@ioannides2008effect].
-The first step of our methodology is to estimate the Zipf coefficient for a broad sample of countries over time.
-Zipf coefficient is one of the most widely used measures of spatial dispersion with numerous applications in urban economics and economic geography ([see for example @black_urban_2003; @frenken_theoretical_2007; @giesen_zipfs_2011; @rauch_cities_2013; @ioannides_zipfs_2003; @ioannides_walled_2017; @nitsch_zipf_2005).
-This an appropriate measure of dispersion because of the extreme heterogeneity of the city size distribution and the very good fits normally obtained with such estimations.
+The first step of our methodology is to estimate the
+\color{blue}
+Pareto exponent coefficient
+\color{black}
+for a broad sample of countries over time.
+\color{blue}
+The Pareto coefficient
+\color{black}
+is one of the most widely used measures of spatial dispersion with numerous applications in urban economics and economic geography ([see for example @black_urban_2003; @frenken_theoretical_2007; @giesen_zipfs_2011; @rauch_cities_2013; @ioannides_zipfs_2003; @ioannides_walled_2017; @nitsch_zipf_2005).
+This an appropriate measure of dispersion because of the extreme heterogeneity of the city size distribution and the very good fits normally obtained with such estimations^[
+\color{blue}
+We also replicate our analysis using other measures of dispersion in Appendix S1. Our results remain consistent.].
+\color{black}
 Skipping details pertaining of the suitability of the approach, which may be found in Ioannides et al.
@@
 
-This leads to an estimate of $\zeta$, known as the Zipf coefficient, but more appropriately the Pareto exponent, when $\zeta$ is not estimated to be near $1$.
-While much of the literature focuses on the Zipf coefficient being close to $1$, several estimations of city size distributions lead to estimates  of $\zeta$ that are not necessarily equal to $1$, or even near it, in which case we refer to $\zeta$ as the Pareto, or power law, exponent.
-Given that our aim here is to estimate $\zeta$ for a number of countries over time as a measure of dispersion, equation \ref{eq1} describes the rank of city $i$ in country $c$ in year $t$:
+\color{blue}
+This leads to an estimate of $\zeta$, known as the Zipf coefficient,
+but more appropriately the Pareto exponent, a terminology we will adhere to for
+the remained of the text, even  when $\zeta$ is estimated to be near $1$.
+The term Zipf coefficient should be reserved for the special case of
+$\zeta=1$.
+While much of the literature focuses on the Pareto exponent being
+close to $1$, several estimations of city size distributions lead to estimates
+of $\zeta$ that are not necessarily equal to $1$, or even near it, in which
+case we refer to $\zeta$ as the Pareto, or power law, exponent.
+Given that our aim here is to estimate $\zeta$ for a number of countries over
+time as a measure of dispersion, equation \ref{eq1} describes the rank of city
+$i$ in country $c$ in year $t$:
+\color{black}
 
@@
   spread(`as.factor(year)`, b) %>%
-  rename(`Zipf coef.` = `as.factor(country)`)
+  rename(`Countries` = `as.factor(country)`)
 
@@
 kable(zipf.table, format = "latex",
-      caption = "Zipf coefficients\\label{zipf}",
+      caption = "Pareto exponent\\label{zipf}",
       booktabs = T,
@@
           summary.stat = c("n", "min", "mean", "sd", "max"),
-          covariate.labels=c("Zipf coefficient",
-                             "Zipf inv. sq. error\\textsuperscript{a}",
+          covariate.labels=c("Pareto exponent",
+                             "Pareto exp. inv. sq. error\\textsuperscript{a}",
                              "Population density",
@@
                              "Fixed phone users per 100 hab."),
-          notes = c("\\textsuperscript{a}This is the inverse squared standard error of the estimated Zipf, coefficient, which is used for weighting", 
-                    "the observations for the estimation, of equation 3. See Section 4"))
+          notes = c("\\textsuperscript{a}This is the inverse squared standard error of the estimated Pareto exponent, which is used for weighting",
+                    "the observations for the estimation of equation 3. See Section 4"))
 ```
@@
 However, such a strategy does not address potential simultaneity issues.
-Simply put, internet penetration might be affected by spatial structure, as reflected in Zipf coefficients, or both internet penetration and spatial structure might be jointly determined by a third variable.
+Simply put, internet penetration might be affected by spatial structure, as reflected in Pareto exponents, or both internet penetration and spatial structure might be jointly determined by a third variable.
 E.g., if a country already has a dispersed spatial structure, internet is particularly suitable in facilitating communication.
@@
 This section presents the estimation results of \ref{eq3}.
-The LHS variable is $\zeta$, the Zipf coefficient, as estimated according to the Gabaix and Ibragimov [@gabaix_rank_2011] correction.
+The LHS variable is $\zeta$, the
+\color{blue}
+Pareto exponent,
+\color{black}
+as estimated according to the Gabaix and Ibragimov [@gabaix_rank_2011] correction.
 The main explanatory variables of interest are provided at the top of Table \ref{ols.global}; namely internet, broadband, mobile and fixed telephony per $100$ habitants, expressed in natural logarithms.
@@
 All regressions include country fixed effects to control for unobserved heterogeneity and a time trend.
-In addition, the observations are weighted with the inverse squared standard error of the estimated Zipf coefficient (see Table \ref{desc.global}) to address potential noise that is carried over from the first part of our identification strategy.
-In regards to the interpretation of the estimated coefficients, given that the Zipf coefficient has entered the regression not as an absolute value, but instead as a real number a positive coefficient for a RHS variable indicates an impact towards the decrease of the spatial dispersion of population.
+In addition, the observations are weighted with the inverse squared standard error of the estimated
+\color{blue}
+Pareto exponent
+\color{black}
+(see Table \ref{desc.global}) to address potential noise that is carried over from the first part of our identification strategy.
+In regards to the interpretation of the estimated coefficients, given that the
+\color{blue}
+Pareto exponent
+\color{black}
+has entered the regression not as an absolute value, but instead as a real number a positive coefficient for a RHS variable indicates an impact towards the decrease of the spatial dispersion of population.
 In other words, a positive coefficient indicates an effect towards less uniform city sizes that is more dispersion of city sizes.
@@
           se = list(net.rse[,"Std. Error"], bb.rse[,"Std. Error"], fixed.rse[,"Std. Error"], mobile.rse[, "Std. Error"]),
-          dep.var.labels=c("Zipf Coefficient 2000-18"),
+          dep.var.labels=c("Pareto exponent 2000-18"),
           column.separate = c(1,1,1,1),
@@
           omit.stat = c("rsq", "f"),
-          notes = "Robust Std. Errors in parenthesis") #, corrected Zipf coefficient following Gabaix and Ibragimov (2011)")
+          notes = "Robust std. errors in parenthesis") #, corrected Zipf coefficient following Gabaix and Ibragimov (2011)")
 ```
@@
 Before discussing further these results and the effects of the other control variables, we need to highlight that the main challenge of estimating equation \ref{eq3} is the potential endogenous nature of the share of internet users which might prevent us from being able to infer a truly causal effect.
-Endogeneity might be an issue here as spatial structure, which is proxied by the Zipf coefficient, might be affected by another source, which also affects internet penetration.
+Endogeneity might be an issue here as spatial structure, which is proxied by the
+\color{blue}
+Pareto exponent,
+\color{black}
+might be affected by another source, which also affects internet penetration.
 For instance, economic development might affect the concentration of population in large cities and at the same time enable more people to go online.
@@
           column.sep.width = "1pt",
-          dep.var.labels=c("Zipf Coefficient 2000-18"),
+          dep.var.labels=c("Pareto exponent 2000-18"),
           column.separate = c(1,1,1,1),
@@
                            c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
-          notes = c("Robust Std. Errors in parenthesis",
-                    "IV: Female participation in labour force"))
+          notes = c("Robust std. errors in parenthesis",
+                    "IV: Female participation in labour force",
+                    "First stage regressions can be found in S2"))
 ```
@@
 \color{blue}
-The literature provides empirical and theoretical evidence that ICT are correlated with increase in female participation in labour markets mostly because they enable teleworking [@dettling2017; @arntz2019working; @watson2018digitalisation; @Lechman2020; @Valberg2020]
+The literature provides empirical and theoretical evidence that ICT are correlated with female participation in labour markets mostly because they enable teleworking [@dettling2017; @arntz2019working; @watson2018digitalisation; @Lechman2020; @Valberg2020]
 \color{black}
@@
 <!-- Firstly, the correlation coefficient between the IV and the Zipf coefficient is very low ($-0.14$). -->
-Although one might think that increased female labor participation may lead to relocation of households to large cities and therefore affect the spatial structure of cites, such a process varies a lot within our dataset.
+Although one might think that increased female labor participation may lead to relocation of households to large cities and therefore affect the spatial structure of cities, such a process varies a lot within our dataset.
 Even if this might be true for developed countries with mature urban systems for which a location within a large urban center might provide opportunities for both male and female workers to find jobs, our dataset also includes countries, $41$ per cent of the GDP of which is attributed to primary activities (see Table \ref{desc.global}).
@@
 The exact opposite appears to happen for fixed telephony, which our results indicate that has led to an increase of spatial dispersion of population.
+\color{blue}
+We expect that the IV estimates capture the local average treatment effect
+[@angrist1995identification] of these labour markets, which are characterised by
+more active engage with the internet, and, therefore, experience a larger effect
+on the Pareto exponents.
+\color{black}
 Importantly, our identification strategy enables us to address potential reverse causality issues and treat the results of Table \ref{2sls.global} as causal.
@@
 These results are robust against endogeneity issues, but are limited to urban agglomeration included in our data.
-Indeed, for the estimation of the Zipf coefficient we only included urban agglomeration of $300,000$ inhabitants or more due to data availability.
+Indeed, for the estimation of the
+\color{blue}
+Pareto exponent
+\color{black}
+we only included urban agglomeration of $300,000$ inhabitants or more due to data availability.
 Hence, the above estimations cannot verify whether such an effect is also valid for smaller cities.
@@
 However, our results do not support the findings of [@tranos2020ict] regarding internet usage.
+
+\color{blue}
+It is indeed puzzling that the results on Table \ref{2sls.global} that the
+effects on urban concentration for internet users and broadband users, on the
+one hand, and mobile and fixed telephony users are in opposite directions.
+We wish to broach this subject in the context of some related literature,
+namely [@calabrese2011interplay] and [@buchel2020cities].
+The latter supports the notion that ICT favors agglomeration in the sense that
+in spite of their convenience modern technologies do no make up for the need of
+proximity in  social interactions.
+The former supports the notion people who communicate more frequently are likely
+to be near one another.
+ICT, as viewed through such granular studies of individuals' communications
+does not make up for physical distance; individuals nearer one another
+communicate more.  
+\color{black}
+
 In order to overcome the city size limitation, the next section presents two case studies, for which we have obtained much more granular data and therefore we are in a position to test the effect of the internet on the tail of the urban population distribution for these countries.
@@
           column.sep.width = "1pt",
-          covariate.labels=c("normalized difference in ranks, 2013-18",
+          covariate.labels=c("difference in ranks, 2013-18",
                              "households w. internet 2013, \\%",
@@
 
-The LHS variable we use here is the outcome of equation \ref{eq5} in a normalized form^[$$Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.$$]
-, so as that variable is bounded between $0$ and $1$ and with a mean of $0.5$.
-This transformation enables us to estimate \ref{eq4} not only with OLS, but also with a quasibinomial GLM estimator given that the original form of our LHS variable does not vary continuously, but is instead defined as a difference between two count variables, which may also assume negative values.
+The LHS variable we use here is the outcome of equation \ref{eq5}.
+\color{blue}
+We use this variable to estimate \ref{eq4} with OLS.
+Additionally, we employ a normalized form of the dependent variable, so as that
+variable is bounded between $0$ and $1$ and with a mean of  $0.5$^[$$Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.$$].
+This transformation enables us to estimate \ref{eq4}
+with a quasibinomial GLM estimator given that the original form of our LHS
+variable does not vary continuously, but is instead defined as a difference between
+two count variables, which may also assume negative values.
+\color{blue}
 The results of the different estimations remain qualitatively the same regardless of the form of the LHS variable we use.
@@
           no.space=TRUE,
-          title = "OLS estimation of equation (4) for USA\\label{ols.usa}",
+          title = "OLS and GLM estimations of equation (4) for USA\\label{ols.usa}",
           font.size = "small",
           column.sep.width = "1pt",
-          column.labels   = c("OLS", "GLM"),
-          column.separate = c(3, 1),
-          model.names = FALSE,
-          dep.var.labels=c("Normalized difference in ranks 2013-18"),
+          #column.labels   = c("OLS", "GLM"),
+          column.separate = 4, #c(3, 1),
+          model.names = F,
+          dep.var.labels.include = F,
+          column.labels = "Difference in ranks, 2013-18",
+          #dep.var.labels=c("Difference in ranks, 2013-18"),
           covariate.labels=c("\\% of households w. internet 2013",
@@
           df = FALSE,
-          omit.stat = c("rsq", "f"), notes = "Robust Std. Errors in parenthesis")
-```
-
-Although not directly comparable, the above estimations are in accordance with our global model. In addition, the potential presence of endogeneity might be a problem here as well.
+          omit.stat = c("rsq", "f"),
+          notes = c("(1)-(3) is based on OLS, (4) on GLM",
+                    "For the GLM the Normalized diff. in ranks is used",
+                    "Robust std. errors in parenthesis"))
+```
+
+Although not directly comparable, the above estimations are in accordance with our global model. Moreover, the potential presence of endogeneity might be a problem here as well.
 Therefore, Table \ref{2sls.usa} reports 2SLS estimations.
 In addition to the strategy employed for the cross-country analysis -- that is the inclusion of one instrument for which we have strong reasons to believe is uncorrelated with the error term -- data availability allows to employ a second IV in order to estimate the Sargan over-identification restrictions test (Column 2).
-The main instrument we propose here is Bachelor’s degrees per inhabitant in $2005$.
+The main instrument we propose here is Bachelor’s degrees per inhabitant in $2005$^[
+\color{blue}
+Due to differences in scale, time and the endogenous variables, different IVs
+have been unutilized for the different sections of the paper.
+\color{black}].
 Even if the quality of human capital can affect the population growth of a city $10$ years later, our LHS variable adopts a systemic understanding of the US urban system as it measures the relative position of a city within the overall urban system instead of its population growth.
@@
           no.space=TRUE,
-          column.separate = c(1,1,1,1),
+          #column.separate = c(1,1,1,1),
           title = "2SLS estimation of equation (4) for USA\\label{2sls.usa}",
@@
           align = TRUE,
-          dep.var.labels=c("Normalized difference in ranks, 2013-18"),
+          dep.var.labels=c("Difference in ranks, 2013-18"),
           covariate.labels=c("\\% of households w. internet 2013",
@@
                                                "IVs: (1) Bachelors degree per hab. in 2005",
-                                               "IVs: (2) Bachelors degree per hab. in 2005, Commute, minutes, 2005"),
+                                               "IVs: (2) Bachelors degree per hab. in 2005, Commute, minutes, 2005",
+                                               "First stage regressions can be found in S2"),
           add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
@@
 Specifically, an increase in internet penetration led to an improvement of the position of a city in the US urban hierarchy for the cities and the time period included in our analysis.
-Everything else being equal, using the normalization formula for the $difference\:in\:ranks_{i}$, if a city whose position in the US urban system did not change during the period $2013-2018$ (that is $Normalized\:difference\:in\:ranks_{i} = 0.5$) had experienced an increase of 10 per cent points of internet penetration, this increase would have improved its relative position by $9$ places.
+\color{blue}
+Everything else being equal, if a city had experienced an increase of 10 per cent
+of internet penetration, this increase would have improved its relative position by $9$ places.
+\color{black}
 As in the global model reported in [Section 4](#sec4), internet penetration appears to work in favor of agglomeration externalities.
@@
 data.for.decr <- data %>%
-  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011,  
+  dplyr::select(r_diff, download2011, pop2011, n.tests2011,  
          unemployment.share, british.pop.share, pop.dens,
@@
   dplyr::select(-n.tests2011) %>%
-  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011perhab,  # re-redring
+  dplyr::select(r_diff, download2011, pop2011, n.tests2011perhab,  # re-redring
          unemployment.share, british.pop.share, pop.dens,
@@
           single.row = FALSE,
-          covariate.labels=c("normalized difference in ranks, 2011-18",
+          covariate.labels=c("difference in ranks, 2011-18",
                              "download speed, 2011",
@@
           no.space=TRUE,
-          title = "OLS estimation of equation (4) for the UK\\label{ols.uk}",
+          title = "OLS and GLM estimations of equation (4) for the UK\\label{ols.uk}",
           font.size = "small",
-          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
+          column.sep.width = "1pt",
+          #dep.var.labels=c("Difference in ranks, 2011-18"),
           #dep.var.caption = "Normalized difference in ranks, 2011-18",
-          column.labels = c("OLS", "GLM"),
-          column.separate = c(3,1),
-          model.names = FALSE,
+          #column.labels = c("OLS", "GLM"),
+          column.separate = 4, #c(3,1),
+          model.names = F,
+          dep.var.labels.include = F,
+          column.labels = "Difference in ranks, 2013-18",
           covariate.labels=c("download speed, 2011 (log)",
@@
           omit.stat = c("rsq", "f"),
-          notes = "Robust Std. Errors")
-```
-
-In regards to IV strategy, we use here the number of universities as the main instrument.
+          notes = c("(1)-(3) is based on OLS, (4) on GLM",
+                    "For the GLM the Normalized diff. in ranks is used",
+                    "Robust std. errors in parenthesis"))
+```
+
+In regards to IV strategy, we use here the *number of universities* as the main instrument.
 Universities in England and Wales have been established long time ago and therefore we do not expect them to affect the change in the relative position of built-up areas in $2011$.
+\color{blue}
+The presence of universities and large student populations tend to be correlated
+with the quality of internet infrastructure provision [@GRUBESIC201950].
+\color{black}
 <!-- The correlation coefficient with the LHS variables is smaller than $0.05$. -->
@@
 Moreover, the estimations reported in Table 11 do not suffer from weak instruments as the relevant test is above the rule of thumb value of $10$.
-The coefficients derived from OLS (Table \ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are of the same magnitude and always positive and significant, something which advocates towards the effect of broadband speed in enhancing agglomeration forces.
-A back of the envelope calculation indicates that, everything else being equal, $1$ per cent increase in download speed for a BUA, the relative position of which did not change during the period $2011-2014$, would have increased its position by $0.3$ places.
-In absolute numbers, $1$ per cent increase for the average BUA would mean a marginal change from $13.8$ Mbps to $13.9$ Mbps.
+\color{blue}
+The coefficients derived from OLS (Table \ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are always positive and significant, something which advocates towards the effect of broadband speed in enhancing agglomeration forces.
+Interestingly, the magnitude of the 2SLS coefficient (column 1 in Table \ref{2sls.uk})
+is almost 8 times higher than the OLS coefficient (column 1 in Table \ref{ols.uk}).
+We attribute the OLS underestimation, which is more or less consistent in all
+our estimations, to the difficulty in directly observing active engagement with
+ICT. Our endogenous variable captures the quality of the internet infrastructure
+as the experienced internet speed.
+Similarly to the endogenous variables of the global and the US model (internet
+and telephony take-up), what these variables do not capture is what individuals
+actually do with ICT [see @tranos2020digital for such an application]. For example, if we assume a bi-modal internet usage for
+either entertainment or work purposes, one would expect  that the latter would
+have a larger effect on spatial structure than the former. We believe that our
+endogenous variables capture mostly the former. Therefore, the IV estimations,
+especially for UK case, reveal the local average treatment effect
+[@angrist1995identification] by capturing the effect of these users/households
+who actively engage with ICT (e.g. work on ICT-related sectors) and, therefore, have a higher effect on spatial structure.
+A back of the envelope calculation indicates that, everything else being equal,
+$10$ per cent increase in download speed for a BUA,
+would have increased its position by $4$ places.
+\color{black}
 
@@
           se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]),
-          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
+          dep.var.labels=c("Difference in ranks, 2011-18"),
           covariate.labels=c("download speed, 2011 (log)",
@@
           df = FALSE,
-          omit.stat = c("rsq", "f"), notes = c("Robust Std. Errors", "IVs: (1) N.r of universities",
-                                               "IVs: (2) N. of universities, N. of broadband tests, 2011"),
+          omit.stat = c("rsq", "f"), notes = c("Robust std. errors in parenthesis", "IVs: (1) N.r of universities",
+                                               "IVs: (2) N. of universities, N. of broadband tests, 2011",
+                                               "First stage regressions can be found in S2"),
           add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
@@
 
-We report estimation results using international country-level data on urban agglomerations with more than $300,000$ inhabitants from the UN Urban Settlements data and test the effect of ICT adoption on the Zipf, or Pareto, coefficients for national city size distributions, as measures of dispersion for heavy-tailed data.
+We report estimation results using international country-level data on urban agglomerations with more than $300,000$ inhabitants from the UN Urban Settlements data and test the effect of ICT adoption on the Pareto exponents for national city size distributions, as measures of dispersion for heavy-tailed data.
 We report results which are robust to endogeneity concerns.

### 2020-12-01 src/s2.Rmd
responses

---
title: "S2 Appendix"
author: ""
output: pdf_document
header-includes:
- \usepackage{multirow}
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/2_resubmission")
  })
---

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(rmarkdown)
library(tidyverse)
library(car)
library(lmtest)
library(plm)
library(stargazer)
library(knitr)

un <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_un_regressions.csv")
us <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_usa_regressions.csv")
uk <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_uk_regressions.csv")
```

# First stage regressions for the 2SLS estimations

## Global model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

net.2sls.s1 <- lm(net100hab ~  fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

bb.2sls.s1 <- lm(bb100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

fixed.2sls.s1 <- lm(fixed100hab ~ fem.lab + log(pop.dens) + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

mobile.2sls.s1 <- lm(mobile100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

net.2sls.s1.rse <- coeftest(net.2sls.s1, function(x) vcovHC(x, type="HC0"))

bb.2sls.s1.rse <- coeftest(bb.2sls.s1, function(x) vcovHC(x, type="HC0"))

fixed.2sls.s1.rse <- coeftest(fixed.2sls.s1, function(x) vcovHC(x, type="HC0"))

mobile.2sls.s1.rse <- coeftest(mobile.2sls.s1, function(x) vcovHC(x, type="HC0"))

stargazer(net.2sls.s1, bb.2sls.s1, fixed.2sls.s1, mobile.2sls.s1, type = "latex",
          se = list(net.2sls.s1.rse[,"Std. Error"], bb.2sls.s1.rse[,"Std. Error"],
                    fixed.2sls.s1.rse[,"Std. Error"], mobile.2sls.s1.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "First stage regressions for Table 5 \\label{2sls.global.s1}",
          font.size = "footnotesize",
          column.sep.width = "1pt",
          # dep.var.labels=c("Internet users per 100 hab. (log)",
          #                 "Broadband users per 100 hab. (log)",
          #                 "Mobile phone users per 100 hab. (log)",
          #                 "Fixed phone users per 100 hab. (log)"),
          column.labels=c('Internet users',
                           "Broadband users",
                           "Mobile phone users",
                           "Fixed phone users"),
          dep.var.labels.include = FALSE,
          #column.separate = c(1,1,1,1),
          covariate.labels=c("Female labour force (\\%)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(
            # c(rownames(summ.net.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
            #                  round(summ.net.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)),
            #                c(rownames(summ.net.2sls$diagnostics)[2],
            #                  round(summ.net.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)),
            #                c("P-value",
            #                  round(summ.net.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis"))

net.test <- car::linearHypothesis(net.2sls.s1, c("fem.lab=0"))
bb.test <- car::linearHypothesis(bb.2sls.s1, c("fem.lab=0"))
mob.test <- car::linearHypothesis(fixed.2sls.s1, c("fem.lab=0"))
fixed.test <- car::linearHypothesis(mobile.2sls.s1, c("fem.lab=0"))

tests <- data.frame("F" = c(net.test$F[2], bb.test$F[2], mob.test$F[2], fixed.test$F[2]),
                    "p" = c(net.test$`Pr(>F)`[2], bb.test$`Pr(>F)`[2], mob.test$`Pr(>F)`[2], fixed.test$`Pr(>F)`[2]),
                    row.names = c("internet users", "broadband users", "mobile phone users", "fixed telephony users"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IV (female labour force (\\%)) from first stage regressions for Table 5 \\label{2sls.global.s1.f}",
          header = FALSE
)
```

\newpage

## US model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

model_iv1.s1 <- lm(hh.net.share ~ bachelors2005 + 
                     log(pop2013) + un13 + white2013share + 
                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
                   data = us)

model_iv2.s1 <- lm(hh.net.share ~ bachelors2005 + commuting.minutes2005 + 
                     log(pop2013) + un13 + white2013share + 
                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
                   data = us)

rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0")) # rob = rse, naming inconsistency
rob.model_iv2.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))

stargazer(model_iv1.s1, model_iv2.s1, 
          type = "latex",
          title = "First stage regressions for Table 8 \\label{2sls.us.s1}",
          header = FALSE,
          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]), 
          dep.var.labels=c("\\% of households w. internet 2013"),
          covariate.labels=c("bachelors degree per hab. 2005",
                             "commute in minutes 2005",
                             "population 2013 (log)",
                             "\\% of unemployment 2013",
                             "\\% of white population 2013",
                             "income 2013 (log)",
                             "population density 2013",
                             "\\% of employment in service 2011",
                             "commute in minutes 2012"),
          single.row = FALSE,
          df = FALSE, 
          omit.stat = c("rsq", "f"), 
          notes = c("Robust Std. Error in parenthesis"))

model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("bachelors2005=0"))
model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("bachelors2005=0","commuting.minutes2005=0"))

tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
                    row.names = c("bachelors degree per hab., 2005", "the above + commute in minutes, 2005"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
          header = FALSE
)
```

\newpage

## UK model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

model_iv1.s1 <- lm(log(download2011) ~ uni.freq +
                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share
                     , data = uk, subset = n.tests2011>30)

model_iv2.s1 <- lm(log(download2011) ~ uni.freq + n.tests2011 +
                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share
                     , data = uk, subset = n.tests2011>30)

rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))
rob.model_iv2.s1 <- coeftest(model_iv2.s1, function(x) vcovHC(x, type="HC0"))

stargazer(model_iv1.s1, model_iv2.s1,
          type="latex",
          summary = F,
          header = FALSE,
          no.space=TRUE,
          title = "First stage regressions for Table 11 \\label{2sls.us.s1}",
          font.size = "small",
          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]),
          dep.var.labels=c("download speed, 2011 (log)"),
          covariate.labels=c("N. of universities",
                             "N. of broadband tests, 2011",
                             "population, 2011 (log)",  
                             "broadband tests per capita, 2011",
                             "\\% of unemployment, 2011",
                             "\\% of British population, 2011",
                             "population density, 2011",
                             "\\% of people working from home, 2011",
                             "employment in service, 2011 (\\%)"),
          single.row = FALSE,
          df = FALSE,
          omit.stat = c("rsq", "f"), 
          notes = c("Robust Std. Errors"))

# N. of universities",
# N. of universities, N. of broadband tests, 2011"),

model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("uni.freq=0"))
model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("uni.freq=0", "n.tests2011=0"))

tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
                    row.names = c("N. of universities", "the above + N. of broadband tests, 2011"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
          header = FALSE
)

```

### 2020-12-01 src/uk.Rmd
responses

@@
 
-base <- lm(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
+base <- lm(r_diff ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                      british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
@@
 # interactions
-int_density <- lm(r_diff_tr ~ log(download2011)*pop.dens + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
+int_density <- lm(r_diff ~ log(download2011)*pop.dens + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                      british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
@@
 
-int_pop <- lm(r_diff_tr ~ log(download2011)*log(pop2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
+int_pop <- lm(r_diff ~ log(download2011)*log(pop2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                      british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
@@
 # 2 SLS with IV
-model_iv1 <- ivreg(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
+model_iv1 <- ivreg(r_diff ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                      british.pop.share +  pop.dens + work.home.share + services.share 
@@
 
-model_iv2 <- ivreg(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
+model_iv2 <- ivreg(r_diff ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                      british.pop.share +  pop.dens + work.home.share + services.share 
@@
 data.for.decr <- data %>%
-  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011,  
+  dplyr::select(r_diff, download2011, pop2011, n.tests2011,  
          unemployment.share, british.pop.share, pop.dens, 
@@
   dplyr::select(-n.tests2011) %>%
-  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011perhab,  # re-redring
+  dplyr::select(r_diff, download2011, pop2011, n.tests2011perhab,  # re-redring
          unemployment.share, british.pop.share, pop.dens, 
@@
 
-path.out <- paste(path, "/outputs/table9.htm", sep = "")
+#path.out <- paste(path, "/outputs/table9.htm", sep = "")
 
-stargazer(as.data.frame(data.for.decr), out=path.out, type="html", summary = T,
-          covariate.labels=c("normalized difference in ranks, 2011-14", "download speed, 2011", "population, 2011",  
+stargazer(as.data.frame(data.for.decr), 
+          #out=path.out, 
+          type="text", summary = T,
+          covariate.labels=c("difference in ranks, 2011-14", "download speed, 2011", "population, 2011",  
                              "broadband tests per capita, 2011", "% of unemployment, 2011", "% of British population, 2011", 
@@
 
-path.out <- paste(path, "/outputs/table10.htm", sep = "")
+# path.out <- paste(path, "/outputs/table10.htm", sep = "")
 
-stargazer(base, int_density, int_pop, base_glm, type="html",
+stargazer(base, int_density, int_pop, base_glm, 
+          type="text",
           se = list(rob.base[,"Std. Error"], rob.int_density[,"Std. Error"], rob.int_pop[,"Std. Error"], rob.base_glm[, "Std. Error"]), 
-          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
+          dep.var.labels=c("Difference in ranks, 2011-18"),
           #dep.var.caption = "Normalized difference in ranks, 2011-18",
@@
                              "download speed, 2011 (log) x population, 2011 (log)"), 
-          out=path.out, single.row = FALSE, df = FALSE, omit.stat = c("rsq", "f"), 
-          notes = "Robust Std. Error parenthesis")
+          #out=path.out, 
+          single.row = FALSE, df = FALSE, omit.stat = c("rsq", "f"), 
+          notes = c("Robust Std. Error parenthesis",
+                    "For the GLM the Normalized diff. in ranks is used"))
 
@@
 
-path.out <- paste(path, "/outputs/table11.htm", sep = "")
+# path.out <- paste(path, "/outputs/table11.htm", sep = "")
 
-stargazer(model_iv1, model_iv2, type = "html", 
+stargazer(model_iv1, model_iv2, 
+          type = "text", 
           se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]), 
-          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
+          dep.var.labels=c("Difference in ranks, 2011-18"),
           covariate.labels=c("download speed, 2011 (log)", "population, 2011 (log)",  
@@
                              "employment in service, 2011 (%)"),
-          out=path.out, single.row = FALSE, df = FALSE, 
+          #out=path.out, 
+          single.row = FALSE, df = FALSE, 
           omit.stat = c("rsq", "f"), notes = c("Robust Std. Error in parenthesis", "IVs: (1) Number of universities",

### 2020-12-01 src/help.R

@@ -1,0 +1,159 @@
+a <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_uk_regressions.csv")
+glimpse(a)
+
+cor(I(a$n.tests2011/a$pop2011), a$download2011)
+
+cor((a$n.tests2011), a$download2011)
+
+cor(I(a$uni.freq/a$pop2011), a$download2011)
+
+model_iv1_s1 <- lm(log(download2011) ~ log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
+                     british.pop.share +  pop.dens + work.home.share + services.share + n.tests2011, data = a, subset = n.tests2011>30)
+summary(model_iv1_s1)
+
+
+us <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_usa_regressions.csv")
+
+summary(model_iv1 <- lm(hh.net.share ~ log(pop2013) + un13 + white2013share +
+                             log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013 +
+                             bachelors2005, data = us))
+
+t.test(model_iv1$coefficients[9], 0)
+car::linearHypothesis(model_iv1, c("bachelors2005=0"))
+
+
+
+# counter cumulatve
+#pop_i = popd*r_i^{-a}
+
+rank <- c(1:1000)
+pop_0 <- 30000000
+# a <- 1
+# a <- 1.095888
+
+# zipf plot
+pop.zipf <- pop_0*rank^(-1)
+pop.zipf <- as.data.frame(pop.zipf)
+
+pop.mean <- pop_0*rank^(-1.095888)
+pop.mean <- as.data.frame(pop.mean)
+
+pop.test <- pop_0*rank^(-1.095888 + .0003)
+pop.test <- as.data.frame(pop.test)
+
+pop <- cbind(pop.mean, pop.zipf, pop.test)
+pop$rank <- c(1:1000)
+
+pop.long <- pivot_longer(pop, !rank, names_to = "distr", values_to = "pop")
+
+ggplot(data=pop.long,
[…]

### 2020-12-02 paper/2_resubmission/tranos_ioannides.Rmd
minor changes

@@
 Kolko [@kolko_death_2000] uses internet diffusion data and identifies a clear complementary link between internet usage and city size.
-Interestingly, he identifies higher internet domain densities in remote cities which indicated a substitution effect of the internet for longer-distance non-electronic communications.
+Interestingly, he identifies higher internet domain densities in remote cities which indicates a substitution effect of the internet for longer-distance non-electronic communications.
 His results are consistent for different measures of internet diffusion (internet domain density and internet take-up).
 Sinai and Waldfogel [@todd_sinai_geography_2004] approach the same question from the consumers’ point of view and find support for complementarity between internet and cities.
-They study the link between market size and locally targeted online content and find that more online local content in  larger markets.
-However, they also obtain evidence for a substitution effect: they find that holding constant local online content constant, market size has a negative effect on individual connectivity.
+They study the link between market size and locally targeted online content and find more online local content in  larger markets.
+However, they also obtain evidence for a substitution effect: holding local online content constant, market size has a negative effect on individual connectivity.
 Forman et al.
-[@chris_forman_how_2005] examined whether commercial internet adoption is higher in cities than in rural areas.
-While the former would indicate a complementarity between internet adoption and cities, the latter would reflect a complementary relation, according to which the internet is used as a means to offset costs and lack of opportunities related to peripheriality.
+[@chris_forman_how_2005] examines whether commercial internet adoption is higher in cities than in rural areas.
+While the former would indicate a complementarity between internet adoption and cities, the latter would reflect a substitutability relation, according to which the internet is used as a means to offset costs and lack of opportunities related to peripheriality.
 Their results indicate that despite internet adoption by firms with more than 100 employees being greater in smaller urban agglomerations, the adoption of more sophisticated internet-based applications is positively related with city size in $2000$.
 Sohn, Kim, and Hewings [@jungyul_sohn_information_2003] compare how information technologies are related to urban spatial structure for Chicago and Seoul.
-Although they find a clear complementary link for Chicago, this was not the case for Seoul, where information technologies contribute to a more dispersed spatial pattern.
+Although they find a clear complementary link for Chicago, this is not the case for Seoul, where information technologies contribute to a more dispersed spatial pattern.
 Focusing on the municipalities in the province of Barcelona, Pons-Novell and Viladecans-Marsal [@pons-novell_cities_2006] find a complementary link between individual internet take-up and off-line commercial offerings, but their results cannot safely reject a substitution effect.
@@
 different effects on urban dispersion on a national scale, that is how far are
-major areas, that is, cities, from one another, versus dispersion within major
+major urban areas, that is, cities, from one another, versus dispersion within major
 urban areas, that is urban sprawl. Such an inquiry would require detailed data
@@
 We think that data from the Covid-19 pandemic would likely throw light at this
-issue, as it is clear that not all economic activities might continue in an
-unencumbered manner; see Dingel and Neiman [@dingel2020many].
+issue in the future, as it is clear that not all economic activities might
+continue in an unencumbered manner; see Dingel and Neiman [@dingel2020many].
 \color{black}
@@
 \color{blue}
-Pareto exponent coefficient
+Pareto exponent
 \color{black}
@@
 \color{blue}
-The Pareto coefficient
+The Pareto exponent
 \color{black}
@@
 [@angrist1995identification] of these labour markets, which are characterised by
-more active engage with the internet, and, therefore, experience a larger effect
+more active engagement with the internet, and, therefore, experience a larger effect
 on the Pareto exponents.
@@
 ICT, as viewed through such granular studies of individuals' communications
-does not make up for physical distance; individuals nearer one another
+do not make up for physical distance; individuals nearer one another
 communicate more.  
@@
 The sign and significance of the main variable of interest -- percentage of population with computer and broadband connection -- verifies our global level results.
-That is an increase in internet usage improves the position of a city in the US urban system.
-The results are consistent among the different estimators that are OLS and GLM.
+That is, an increase in internet usage improves the position of a city in the US urban system.
+The results are consistent among the different estimators that is OLS and GLM.
 In addition, Table \ref{ols.usa} also reports the estimates of interaction effects between the share of population with broadband connection with population and population density.
@@
 Therefore, Table \ref{2sls.usa} reports 2SLS estimations.
-In addition to the strategy employed for the cross-country analysis -- that is the inclusion of one instrument for which we have strong reasons to believe is uncorrelated with the error term -- data availability allows to employ a second IV in order to estimate the Sargan over-identification restrictions test (Column 2).
+In addition to the strategy employed for the cross-country analysis -- that is the inclusion of one instrument for which we have strong reasons to believe is uncorrelated with the error term -- data availability allows us to employ a second IV in order to estimate the Sargan over-identification restrictions test (Column 2).
 The main instrument we propose here is Bachelor’s degrees per inhabitant in $2005$^[
@@
 \color{black}].
-Even if the quality of human capital can affect the population growth of a city $10$ years later, our LHS variable adopts a systemic understanding of the US urban system as it measures the relative position of a city within the overall urban system instead of its population growth.
+Even if the quality of human capital affect the population growth of a city $10$ years later, our LHS variable adopts a systemic understanding of the US urban system as it measures the relative position of a city within the overall urban system instead of its population growth.
 In other words, a city might experience population growth between two periods, but if other cities have also experienced population growth, this might not affect its relative position within the urban system.
@@
 <!-- In addition, the correlation coefficient between these two variables is still very low ($r=0.24$). -->
-Moreover, when we add the second IV, which in this case is the commute time in minutes also in $2005$, we fail to reject the null hypothesis of the Sargan test, something which also adds on the validity of our strategy.
+Moreover, when we add the second IV, which in this case is the commute time in minutes also in $2005$, we fail to reject the null hypothesis of the Sargan test, something which also adds the validity of our strategy.
 Furthermore, our results do not suffer from weak identification according to the relevant tests in Table \ref{2sls.usa}.
@@
 \color{blue}
-The coefficients derived from OLS (Table \ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are always positive and significant, something which advocates towards the effect of broadband speed in enhancing agglomeration forces.
+The coefficients derived from OLS (Table \ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are always positive and significant, something which advocates in favour of the effect of broadband speed in enhancing agglomeration forces.
 Interestingly, the magnitude of the 2SLS coefficient (column 1 in Table \ref{2sls.uk})

### 2020-12-02 paper/2_resubmission/tranos_ioannides.tex
minor changes

@@
 he identifies higher internet domain densities in remote cities which
-indicated a substitution effect of the internet for longer-distance
+indicates a substitution effect of the internet for longer-distance
 non-electronic communications. His results are consistent for different
@@
 between internet and cities. They study the link between market size and
-locally targeted online content and find that more online local content
-in larger markets. However, they also obtain evidence for a substitution
-effect: they find that holding constant local online content constant,
-market size has a negative effect on individual connectivity. Forman et
-al.~{[}24{]} examined whether commercial internet adoption is higher in
-cities than in rural areas. While the former would indicate a
-complementarity between internet adoption and cities, the latter would
-reflect a complementary relation, according to which the internet is
-used as a means to offset costs and lack of opportunities related to
+locally targeted online content and find more online local content in
+larger markets. However, they also obtain evidence for a substitution
+effect: holding local online content constant, market size has a
+negative effect on individual connectivity. Forman et al.~{[}24{]}
+examines whether commercial internet adoption is higher in cities than
+in rural areas. While the former would indicate a complementarity
+between internet adoption and cities, the latter would reflect a
+substitutability relation, according to which the internet is used as a
+means to offset costs and lack of opportunities related to
 peripheriality. Their results indicate that despite internet adoption by
@@
 related to urban spatial structure for Chicago and Seoul. Although they
-find a clear complementary link for Chicago, this was not the case for
+find a clear complementary link for Chicago, this is not the case for
 Seoul, where information technologies contribute to a more dispersed
@@
 
+\color{blue}
+
+The present paper does not address the issue that adoption of ICT may
+have different effects on urban dispersion on a national scale, that is
+how far are major urban areas, that is, cities, from one another, versus
+dispersion within major urban areas, that is urban sprawl. Such an
+inquiry would require detailed data about patterns of how urban sprawl
+may hamper or promote economic interaction, given advancements in
+information technology. We think that data from the Covid-19 pandemic
+would likely throw light at this issue in the future, as it is clear
+that not all economic activities might continue in an unencumbered
+manner; see Dingel and Neiman {[}36{]}. \color{black}
+
 \hypertarget{sec3}{%
@@
 The multi-country identification strategy is a two-step approach
-{[}18{]}. The first step of our methodology is to estimate the Zipf
-coefficient for a broad sample of countries over time. Zipf coefficient
-is one of the most widely used measures of spatial dispersion with
-numerous applications in urban economics and economic geography ({[}see
-for example {[}36{]}; {[}37{]}; {[}38{]}; {[}39{]}; {[}40{]}; {[}41{]};
-{[}42{]}). This an appropriate measure of dispersion because of the
+{[}18{]}. The first step of our methodology is to estimate the
+\color{blue} Pareto exponent \color{black} for a broad sample of
+countries over time. \color{blue} The Pareto exponent \color{black} is
+one of the most widely used measures of spatial dispersion with numerous
+applications in urban economics and economic geography ({[}see for
+example {[}37{]}; {[}38{]}; {[}39{]}; {[}40{]}; {[}41{]}; {[}42{]};
+{[}43{]}). This an appropriate measure of dispersion because of the
 extreme heterogeneity of the city size distribution and the very good
-fits normally obtained with such estimations. Skipping details
-pertaining of the suitability of the approach, which may be found in
-Ioannides et al.~{[}18{]} and Tranos and Ioannides {[}19{]}, we proceed
-with the estimation of the logarithmic form of:
+fits normally obtained with such estimations\footnote{\color{blue} We
+  also replicate our analysis using other measures of dispersion in
+  Appendix S1. Our results remain consistent.}. \color{black} Skipping
+details pertaining of the suitability of the approach, which may be
+found in Ioannides et al.~{[}18{]} and Tranos and Ioannides {[}19{]}, we
+proceed with the estimation of the logarithmic form of:
 
@@
 
+\color{blue}
+
 This leads to an estimate of \(\zeta\), known as the Zipf coefficient,
-but more appropriately the Pareto exponent, when \(\zeta\) is not
-estimated to be near \(1\). While much of the literature focuses on the
-Zipf coefficient being close to \(1\), several estimations of city size
+but more appropriately the Pareto exponent, a terminology we will adhere
+to for the remained of the text, even when \(\zeta\) is estimated to be
+near \(1\). The term Zipf coefficient should be reserved for the special
+case of \(\zeta=1\). While much of the literature focuses on the Pareto
+exponent being close to \(1\), several estimations of city size
 distributions lead to estimates of \(\zeta\) that are not necessarily
@@
 dispersion, equation \ref{eq1} describes the rank of city \(i\) in
-country \(c\) in year \(t\):
+country \(c\) in year \(t\): \color{black}
 
@@
 The estimation of equation \ref{eq2} has typically been performed by
-Ordinary Least Squares (OLS). Gabaix and Ioannides {[}43{]} discuss the
+Ordinary Least Squares (OLS). Gabaix and Ioannides {[}44{]} discuss the
 downward bias of estimates of equation \ref{eq2} using OLS on small
-samples. Gabaix and Ibragimov {[}44{]} propose a practical remedy to
+samples. Gabaix and Ibragimov {[}45{]} propose a practical remedy to
 correct this bias, which we do adopt in this paper: instead of using the
@@
 data sources. Definitions of cities differ across countries, for
-political, administrative and legal reasons\footnote{See {[}45{]}, Ch.
+political, administrative and legal reasons\footnote{See {[}46{]}, Ch.
   8, for issues and pitfalls associated with different definitions.}. In
 contrast to the data drawn from Thomas Brinkhoff's City Population
-project {[}46{]}\footnote{\url{http://www.citypopulation.de}}, which
-were used by Soo {[}47{]}, Ioannides et al.~{[}18{]} and Tranos and
+project {[}47{]}\footnote{\url{http://www.citypopulation.de}}, which
+were used by Soo {[}48{]}, Ioannides et al.~{[}18{]} and Tranos and
 Ioannides {[}19{]}, the multi-country analysis in the present paper is
@@
 about the consistency of the urban agglomeration definitions across
-different countries {[}48{]}, this is the only available source for
+different countries {[}49{]}, this is the only available source for
 yearly, multi-country population data for urban agglomerations
-{[}49--51{]}. The results based on these data are reported and discussed
+{[}50--52{]}. The results based on these data are reported and discussed
 in \protect\hyperlink{sec4}{Section 4}.
@@
 
-\caption{\label{tab:unnamed-chunk-2}Zipf coefficients\label{zipf}}
+\caption{\label{tab:unnamed-chunk-2}Pareto exponent\label{zipf}}
 \resizebox{\linewidth}{!}{
@@
 \toprule
-Zipf coef. & 2000 & 2001 & 2002 & 2003 & 2004 & 2005 & 2006 & 2007 & 2008 & 2009 & 2010 & 2011 & 2012 & 2013 & 2014 & 2015 & 2016 & 2017 & 2018\\
+Countries & 2000 & 2001 & 2002 & 2003 & 2004 & 2005 & 2006 & 2007 & 2008 & 2009 & 2010 & 2011 & 2012 & 2013 & 2014 & 2015 & 2016 & 2017 & 2018\\
 \midrule
@@
 the lack of fixed line infrastructure and mobile phone networks are also
-used as the main way to use the internet {[}52,53{]}.
+used as the main way to use the internet {[}53,54{]}.
 
@@
 \hline \\[-1.8ex] 
-Zipf coefficient & 950 & $-$1.9 & $-$1.1 & 0.2 & $-$0.7 \\ 
-Zipf inv. sq. error\textsuperscript{a} & 950 & 0.01 & 0.1 & 0.05 & 0.2 \\ 
+Pareto exponent & 950 & $-$1.9 & $-$1.1 & 0.2 & $-$0.7 \\ 
+Pareto exp. inv. sq. error\textsuperscript{a} & 950 & 0.01 & 0.1 & 0.05 & 0.2 \\ 
 Population density & 931 & 2.5 & 126.2 & 182.1 & 1,239.6 \\ 
@@
 \hline \\[-1.8ex] 
-\multicolumn{6}{l}{\textsuperscript{a}This is the inverse squared standard error of the estimated Zipf, coefficient, which is used for weighting} \\ 
-\multicolumn{6}{l}{the observations for the estimation, of equation 3. See Section 4} \\ 
+\multicolumn{6}{l}{\textsuperscript{a}This is the inverse squared standard error of the estimated Pareto exponent, which is used for weighting} \\ 
+\multicolumn{6}{l}{the observations for the estimation of equation 3. See Section 4} \\ 
 \end{tabular} 
@@
 does not address potential simultaneity issues. Simply put, internet
-penetration might be affected by spatial structure, as reflected in Zipf
-coefficients, or both internet penetration and spatial structure might
-be jointly determined by a third variable. E.g., if a country already
-has a dispersed spatial structure, internet is particularly suitable in
-facilitating communication. Potential endogeneity in our specification
-will prevent us from being able to determine the causal impact of
-internet and digital communication technologies usage on spatial
-structure, which is the main aim of this paper. In order to address this
-problem, we will adopt an instrumented variable strategy. Table
-\ref{desc.global} also includes the descriptive statistics for the
+penetration might be affected by spatial structure, as reflected in
+Pareto exponents, or both internet penetration and spatial structure
+might be jointly determined by a third variable. E.g., if a country
+already has a dispersed spatial structure, internet is particularly
+suitable in facilitating communication. Potential endogeneity in our
+specification will prevent us from being able to determine the causal
+impact of internet and digital communication technologies usage on
+spatial structure, which is the main aim of this paper. In order to
+address this problem, we will adopt an instrumented variable strategy.
+Table \ref{desc.global} also includes the descriptive statistics for the
 instrumental variable we are using and (\(female\:labour\:force\)), will
@@
 In order to capture the micro-dynamics of the urban systems in the two
-case studies, we follow Batty {[}54{]} and Havlin {[}55{]} and focus on
+case studies, we follow Batty {[}55{]} and Havlin {[}56{]} and focus on
 the difference in ranks for individual cities during the study
@@
   elsewhere in economics because of its robustness properties. C.f.
-  Chetty and Hendren {[}56{]}.} Contrary to their approach, we are
+  Chetty and Hendren {[}57{]}.} Contrary to their approach, we are
 interested in the real difference in rankings rather than in absolute
@@
 This section presents the estimation results of \ref{eq3}. The LHS
-variable is \(\zeta\), the Zipf coefficient, as estimated according to
-the Gabaix and Ibragimov {[}44{]} correction. The main explanatory
-variables of interest are provided at the top of Table \ref{ols.global};
-namely internet, broadband, mobile and fixed telephony per \(100\)
-habitants, expressed in natural logarithms. The main variables of
-interest are introduced successively on their own in the regressions
-reported in Table \ref{ols.global}. All regressions include country
-fixed effects to control for unobserved heterogeneity and a time trend.
-In addition, the observations are weighted with the inverse squared
-standard error of the estimated Zipf coefficient (see Table
-\ref{desc.global}) to address potential noise that is carried over from
-the first part of our identification strategy. In regards to the
-interpretation of the estimated coefficients, given that the Zipf
-coefficient has entered the regression not as an absolute value, but
+variable is \(\zeta\), the \color{blue} Pareto exponent, \color{black}
+as estimated according to the Gabaix and Ibragimov {[}45{]} correction.
+The main explanatory variables of interest are provided at the top of
+Table \ref{ols.global}; namely internet, broadband, mobile and fixed
+telephony per \(100\) habitants, expressed in natural logarithms. The
+main variables of interest are introduced successively on their own in
+the regressions reported in Table \ref{ols.global}. All regressions
+include country fixed effects to control for unobserved heterogeneity
+and a time trend. In addition, the observations are weighted with the
+inverse squared standard error of the estimated \color{blue} Pareto
+exponent \color{black} (see Table \ref{desc.global}) to address
+potential noise that is carried over from the first part of our
+identification strategy. In regards to the interpretation of the
+estimated coefficients, given that the \color{blue} Pareto exponent
+\color{black} has entered the regression not as an absolute value, but
 instead as a real number a positive coefficient for a RHS variable
@@
 \cline{2-5} 
-\\[-1.8ex] & \multicolumn{4}{c}{Zipf Coefficient 2000-18} \\ 
+\\[-1.8ex] & \multicolumn{4}{c}{Pareto exponent 2000-18} \\ 
 \\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
@@
 \textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
- & \multicolumn{4}{r}{Robust Std. Errors in parenthesis} \\ 
+ & \multicolumn{4}{r}{Robust std. errors in parenthesis} \\ 
 \end{tabular} 
@@
 effect. Endogeneity might be an issue here as spatial structure, which
-is proxied by the Zipf coefficient, might be affected by another source,
-which also affects internet penetration. For instance, economic
-development might affect the concentration of population in large cities
-and at the same time enable more people to go online. If we do not
-address this issue, the coefficient for the main variable of interest
-will capture potential effects that internet penetration has on spatial
-structure, but also potential reverse causality effects that spatial
-structure might generate on internet penetration. To overcome this
-potential problem, Table \ref{2sls.global} reports estimates of equation
-\ref{eq3} using two-stage least squares (2SLS) with instrumental
-variables (IVs). The latter are variables which are correlated with our
-endogenous variables, but do not influence current spatial structure.
-Such an approach will enable us to estimate the causal effect -- if any
--- of the internet and digital communications on spatial structure. At a
-first stage, our endogenous variables are regressed against the IVs.
-Then, the predicted values of the endogenous variables based on the IVs
-and the other control variables are used instead of the endogenous
-variable to estimate eq. \ref{eq3}. A significant effect will verify the
-causal impact of the internet and digital communication usage on spatial
-structure.
+is proxied by the \color{blue} Pareto exponent, \color{black} might be
+affected by another source, which also affects internet penetration. For
+instance, economic development might affect the concentration of
+population in large cities and at the same time enable more people to go
+online. If we do not address this issue, the coefficient for the main
+variable of interest will capture potential effects that internet
+penetration has on spatial structure, but also potential reverse
+causality effects that spatial structure might generate on internet
+penetration. To overcome this potential problem, Table \ref{2sls.global}
+reports estimates of equation \ref{eq3} using two-stage least squares
+(2SLS) with instrumental variables (IVs). The latter are variables which
+are correlated with our endogenous variables, but do not influence
+current spatial structure. Such an approach will enable us to estimate
+the causal effect -- if any -- of the internet and digital
+communications on spatial structure. At a first stage, our endogenous
+variables are regressed against the IVs. Then, the predicted values of
+the endogenous variables based on the IVs and the other control
+variables are used instead of the endogenous variable to estimate eq.
+\ref{eq3}. A significant effect will verify the causal impact of the
+internet and digital communication usage on spatial structure.
 
@@
 \textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
- & \multicolumn{4}{r}{Robust Std. Errors in parenthesis} \\ 
+ & \multicolumn{4}{r}{Robust std. errors in parenthesis} \\ 
  & \multicolumn{4}{r}{IV: Female participation in labour force} \\ 
+ & \multicolumn{4}{r}{First stage regressions can be found in S2} \\ 
 \end{tabular} 
@@
 The literature provides empirical and theoretical evidence that ICT are
-correlated with increase in female participation in labour markets
-mostly because they enable teleworking {[}57--61{]} \color{black} At the
-same time, female labor participation is not related with spatial
-structure especially given the diverse sample of countries. Although one
-might think that increased female labor participation may lead to
-relocation of households to large cities and therefore affect the
-spatial structure of cites, such a process varies a lot within our
-dataset. Even if this might be true for developed countries with mature
-urban systems for which a location within a large urban center might
-provide opportunities for both male and female workers to find jobs, our
-dataset also includes countries, \(41\) per cent of the GDP of which is
-attributed to primary activities (see Table \ref{desc.global}). So,
-female labor participation might also be related to economic activities
-located outside large urban centers. In terms of the relevant tests
-presented in Table \ref{2sls.global} the weak identification first stage
-F-test exceeds by far all the Stock-Yogo critical values for all, but
-the mobile telephony regression. Hence, we are confident to interpret
-the effects of internet, broadband internet and fixed telephony usage.
+correlated with female participation in labour markets mostly because
+they enable teleworking {[}58--62{]} \color{black} At the same time,
+female labor participation is not related with spatial structure
+especially given the diverse sample of countries. Although one might
+think that increased female labor participation may lead to relocation
+of households to large cities and therefore affect the spatial structure
+of cities, such a process varies a lot within our dataset. Even if this
+might be true for developed countries with mature urban systems for
+which a location within a large urban center might provide opportunities
+for both male and female workers to find jobs, our dataset also includes
+countries, \(41\) per cent of the GDP of which is attributed to primary
+activities (see Table \ref{desc.global}). So, female labor participation
+might also be related to economic activities located outside large urban
+centers. In terms of the relevant tests presented in Table
+\ref{2sls.global} the weak identification first stage F-test exceeds by
+far all the Stock-Yogo critical values for all, but the mobile telephony
+regression. Hence, we are confident to interpret the effects of
+internet, broadband internet and fixed telephony usage.
 
@@
 indicate that has led to an increase of spatial dispersion of
-population. Importantly, our identification strategy enables us to
+population. \color{blue} We expect that the IV estimates capture the
+local average treatment effect {[}63{]} of these labour markets, which
+are characterised by more active engagement with the internet, and,
+therefore, experience a larger effect on the Pareto exponents.
+\color{black} Importantly, our identification strategy enables us to
 address potential reverse causality issues and treat the results of
@@
 but are limited to urban agglomeration included in our data. Indeed, for
-the estimation of the Zipf coefficient we only included urban
-agglomeration of \(300,000\) inhabitants or more due to data
-availability. Hence, the above estimations cannot verify whether such an
-effect is also valid for smaller cities. Interestingly, the above
-results regarding the role of fixed telephony are in agreement with
-previous results from {[}18{]} {[}19{]} which used a smaller panel of
-countries but without constraints regarding the city size. However, our
-results do not support the findings of {[}19{]} regarding internet
-usage. In order to overcome the city size limitation, the next section
-presents two case studies, for which we have obtained much more granular
-data and therefore we are in a position to test the effect of the
-internet on the tail of the urban population distribution for these
-countries.
+the estimation of the \color{blue} Pareto exponent \color{black} we only
+included urban agglomeration of \(300,000\) inhabitants or more due to
+data availability. Hence, the above estimations cannot verify whether
+such an effect is also valid for smaller cities. Interestingly, the
+above results regarding the role of fixed telephony are in agreement
+with previous results from {[}18{]} {[}19{]} which used a smaller panel
+of countries but without constraints regarding the city size. However,
+our results do not support the findings of {[}19{]} regarding internet
+usage.
+
+\color{blue}
+
+It is indeed puzzling that the results on Table \ref{2sls.global} that
+the effects on urban concentration for internet users and broadband
+users, on the one hand, and mobile and fixed telephony users are in
+opposite directions. We wish to broach this subject in the context of
+some related literature, namely {[}64{]} and {[}65{]}. The latter
+supports the notion that ICT favors agglomeration in the sense that in
+spite of their convenience modern technologies do no make up for the
+need of proximity in social interactions. The former supports the notion
+people who communicate more frequently are likely to be near one
+another. ICT, as viewed through such granular studies of individuals'
+communications do not make up for physical distance; individuals nearer
+one another communicate more.\\
+\color{black}
+
+In order to overcome the city size limitation, the next section presents
+two case studies, for which we have obtained much more granular data and
+therefore we are in a position to test the effect of the internet on the
+tail of the urban population distribution for these countries.
 
@@
 least \(10,000\), but less than \(50,000\) population) level of
-aggregation {[}62{]}. These functional definitions of a city represent
+aggregation {[}66{]}. These functional definitions of a city represent
 in essence labor markets and according to Table \ref{desc.usa} the
@@
 \hline \\[-1.8ex] 
-normalized difference in ranks, 2013-18 & 461 & 0.480 & 0.500 & 0.008 & 0.546 \\ 
+difference in ranks, 2013-18 & 461 & $-$21 & 0.028 & 8.171 & 47 \\ 
 households w. internet 2013, \% & 461 & 0.307 & 0.716 & 0.073 & 0.871 \\ 
@@
 
-The LHS variable we use here is the outcome of equation \ref{eq5} in a
-normalized form\footnote{\[Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.\]}
-, so as that variable is bounded between \(0\) and \(1\) and with a mean
-of \(0.5\). This transformation enables us to estimate \ref{eq4} not
-only with OLS, but also with a quasibinomial GLM estimator given that
-the original form of our LHS variable does not vary continuously, but is
-instead defined as a difference between two count variables, which may
-also assume negative values. The results of the different estimations
-remain qualitatively the same regardless of the form of the LHS variable
-we use. Table \ref{ols.usa} reports the results of the estimation of
+The LHS variable we use here is the outcome of equation \ref{eq5}.
+\color{blue} We use this variable to estimate \ref{eq4} with OLS.
+Additionally, we employ a normalized form of the dependent variable, so
+as that variable is bounded between \(0\) and \(1\) and with a mean of
+\(0.5\)\footnote{\[Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.\]}.
+This transformation enables us to estimate \ref{eq4} with a
+quasibinomial GLM estimator given that the original form of our LHS
+variable does not vary continuously, but is instead defined as a
+difference between two count variables, which may also assume negative
+values. \color{blue} The results of the different estimations remain
+qualitatively the same regardless of the form of the LHS variable we
+use. Table \ref{ols.usa} reports the results of the estimation of
 equation \ref{eq4} using OLS. The sign and significance of the main
 variable of interest -- percentage of population with computer and
-broadband connection -- verifies our global level results. That is an
+broadband connection -- verifies our global level results. That is, an
 increase in internet usage improves the position of a city in the US
 urban system. The results are consistent among the different estimators
-that are OLS and GLM. In addition, Table \ref{ols.usa} also reports the
+that is OLS and GLM. In addition, Table \ref{ols.usa} also reports the
 estimates of interaction effects between the share of population with
@@
 \begin{table}[!htbp] \centering 
-  \caption{OLS estimation of equation (4) for USA\label{ols.usa}} 
+  \caption{OLS and GLM estimations of equation (4) for USA\label{ols.usa}} 
   \label{} 
@@
 \cline{2-5} 
-\\[-1.8ex] & \multicolumn{4}{c}{Normalized difference in ranks 2013-18} \\ 
- & \multicolumn{3}{c}{OLS} & GLM \\ 
+ & \multicolumn{4}{c}{Difference in ranks, 2013-18} \\ 
 \\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
 \hline \\[-1.8ex] 
- \% of households w. internet 2013 & 0.050$^{***}$ & 0.050$^{***}$ & 0.111$^{*}$ & 0.201$^{***}$ \\ 
-  & (0.009) & (0.010) & (0.064) & (0.036) \\ 
-  population 2013 (log) & $-$0.0002 & $-$0.0002 & 0.003 & $-$0.001 \\ 
-  & (0.0004) & (0.0004) & (0.004) & (0.002) \\ 
-  \% of unemployment 2013 & $-$0.0004$^{*}$ & $-$0.0003$^{*}$ & $-$0.0004$^{*}$ & $-$0.001$^{*}$ \\ 
-  & (0.0002) & (0.0002) & (0.0002) & (0.001) \\ 
-  \% of white population 2013 & 0.0001 & 0.0001 & $-$0.0002 & 0.0003 \\ 
-  & (0.003) & (0.003) & (0.003) & (0.013) \\ 
-  income 2013 (log) & $-$0.011$^{***}$ & $-$0.011$^{***}$ & $-$0.011$^{***}$ & $-$0.045$^{***}$ \\ 
-  & (0.004) & (0.004) & (0.004) & (0.015) \\ 
-  population density 2013 & $-$0.001$^{***}$ & $-$0.001 & $-$0.001$^{***}$ & $-$0.004$^{***}$ \\ 
-  & (0.0003) & (0.004) & (0.0003) & (0.001) \\ 
-  \% of employment in service 2011 & $-$0.00002 & $-$0.00002 & $-$0.00001 & $-$0.0001 \\ 
-  & (0.0001) & (0.0001) & (0.0001) & (0.0003) \\ 
-  commute in minutes 2013 & 0.0003 & 0.0003 & 0.0003 & 0.001 \\ 
-  & (0.0002) & (0.0002) & (0.0002) & (0.001) \\ 
+ \% of households w. internet 2013 & 51.540$^{***}$ & 51.212$^{***}$ & 114.247$^{*}$ & 0.201$^{***}$ \\ 
+  & (9.226) & (10.686) & (65.765) & (0.036) \\ 
+  population 2013 (log) & $-$0.232 & $-$0.223 & 3.586 & $-$0.001 \\ 
+  & (0.407) & (0.429) & (3.764) & (0.002) \\ 
+  \% of unemployment 2013 & $-$0.360$^{*}$ & $-$0.359$^{*}$ & $-$0.367$^{*}$ & $-$0.001$^{*}$ \\ 
+  & (0.196) & (0.197) & (0.197) & (0.001) \\ 
+  \% of white population 2013 & 0.085 & 0.137 & $-$0.237 & 0.0003 \\ 
+  & (3.409) & (3.445) & (3.412) & (0.013) \\ 
+  income 2013 (log) & $-$11.663$^{***}$ & $-$11.695$^{***}$ & $-$11.210$^{***}$ & $-$0.045$^{***}$ \\ 
+  & (3.812) & (3.829) & (3.847) & (0.015) \\ 
+  population density 2013 & $-$1.056$^{***}$ & $-$1.464 & $-$0.987$^{***}$ & $-$0.004$^{***}$ \\ 
+  & (0.328) & (4.568) & (0.328) & (0.001) \\ 
+  \% of employment in service 2011 & $-$0.016 & $-$0.016 & $-$0.015 & $-$0.0001 \\ 
+  & (0.065) & (0.065) & (0.066) & (0.0003) \\ 
+  commute in minutes 2013 & 0.261 & 0.259 & 0.286 & 0.001 \\ 
+  & (0.174) & (0.175) & (0.177) & (0.001) \\ 
   \% of households w. internet, \\
-                             2013 x pop. density, 2013 &  & 0.001 &  &  \\ 
-  &  & (0.006) &  &  \\ 
+                             2013 x pop. density, 2013 &  & 0.528 &  &  \\ 
+  &  & (5.851) &  &  \\ 
   \% of households w. internet, \\
-                             2013 x population, 2013 &  &  & $-$0.005 &  \\ 
-  &  &  & (0.005) &  \\ 
-  Constant & 0.581$^{***}$ & 0.581$^{***}$ & 0.532$^{***}$ & 0.323$^{**}$ \\ 
-  & (0.035) & (0.036) & (0.063) & (0.142) \\ 
+                             2013 x population, 2013 &  &  & $-$5.236 &  \\ 
+  &  &  & (5.161) &  \\ 
+  Constant & 83.035$^{**}$ & 83.513$^{**}$ & 32.486 & 0.323$^{**}$ \\ 
+  & (36.479) & (36.997) & (64.658) & (0.142) \\ 
  \hline \\[-1.8ex] 
@@
 Adjusted R$^{2}$ & 0.122 & 0.120 & 0.122 &  \\ 
-Residual Std. Error & 0.007 & 0.007 & 0.007 &  \\ 
+Residual Std. Error & 7.561 & 7.569 & 7.562 &  \\ 
 \hline 
@@
 \textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
- & \multicolumn{4}{r}{Robust Std. Errors in parenthesis} \\ 
+ & \multicolumn{4}{r}{(1)-(3) is based on OLS, (4) on GLM} \\ 
+ & \multicolumn{4}{r}{For the GLM the Normalized diff. in ranks is used} \\ 
+ & \multicolumn{4}{r}{Robust std. errors in parenthesis} \\ 
 \end{tabular} 
@@
 Although not directly comparable, the above estimations are in
-accordance with our global model. In addition, the potential presence of
+accordance with our global model. Moreover, the potential presence of
 endogeneity might be a problem here as well. Therefore, Table
@@
 instrument for which we have strong reasons to believe is uncorrelated
-with the error term -- data availability allows to employ a second IV in
-order to estimate the Sargan over-identification restrictions test
+with the error term -- data availability allows us to employ a second IV
+in order to estimate the Sargan over-identification restrictions test
 (Column 2). The main instrument we propose here is Bachelor's degrees
-per inhabitant in \(2005\). Even if the quality of human capital can
-affect the population growth of a city \(10\) years later, our LHS
-variable adopts a systemic understanding of the US urban system as it
-measures the relative position of a city within the overall urban system
-instead of its population growth. In other words, a city might
-experience population growth between two periods, but if other cities
-have also experienced population growth, this might not affect its
-relative position within the urban system. Hence, we do not expect that
-this IV affects the LHS variable. Moreover, when we add the second IV,
-which in this case is the commute time in minutes also in \(2005\), we
-fail to reject the null hypothesis of the Sargan test, something which
-also adds on the validity of our strategy. Furthermore, our results do
-not suffer from weak identification according to the relevant tests in
-Table \ref{2sls.usa}.
+per inhabitant in \(2005\)\footnote{\color{blue} Due to differences in
+  scale, time and the endogenous variables, different IVs have been
+  unutilized for the different sections of the paper. \color{black}}.
+Even if the quality of human capital affect the population growth of a
+city \(10\) years later, our LHS variable adopts a systemic
+understanding of the US urban system as it measures the relative
+position of a city within the overall urban system instead of its
+population growth. In other words, a city might experience population
+growth between two periods, but if other cities have also experienced
+population growth, this might not affect its relative position within
+the urban system. Hence, we do not expect that this IV affects the LHS
+variable. Moreover, when we add the second IV, which in this case is the
+commute time in minutes also in \(2005\), we fail to reject the null
+hypothesis of the Sargan test, something which also adds the validity of
+our strategy. Furthermore, our results do not suffer from weak
+identification according to the relevant tests in Table \ref{2sls.usa}.
 
@@
 \cline{2-3} 
-\\[-1.8ex] & \multicolumn{2}{c}{Normalized difference in ranks, 2013-18} \\ 
+\\[-1.8ex] & \multicolumn{2}{c}{Difference in ranks, 2013-18} \\ 
 \\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)}\\ 
 \hline \\[-1.8ex] 
- \% of households w. internet 2013 & 0.091^{***} & 0.084^{***} \\ 
-  & (0.020) & (0.021) \\ 
-  population 2013 (log) & -0.001 & -0.001 \\ 
-  & (0.0005) & (0.0005) \\ 
-  \% of unemployment 2013 & -0.0004^{**} & -0.0004^{**} \\ 
-  & (0.0002) & (0.0002) \\ 
-  \% of white population 2013 & -0.006 & -0.005 \\ 
-  & (0.005) & (0.005) \\ 
-  income 2013 (log) & -0.024^{***} & -0.022^{***} \\ 
-  & (0.006) & (0.006) \\ 
-  population density 2013 & -0.001^{***} & -0.001^{***} \\ 
-  & (0.0004) & (0.0004) \\ 
-  \% of employment in service 2011 & -0.0001 & -0.0001 \\ 
-  & (0.0001) & (0.0001) \\ 
-  commute in minutes 2013 & 0.0004^{*} & 0.0003^{*} \\ 
-  & (0.0002) & (0.0002) \\ 
-  Constant & 0.691^{***} & 0.673^{***} \\ 
-  & (0.059) & (0.058) \\ 
+ \% of households w. internet 2013 & 93.908^{***} & 86.712^{***} \\ 
+  & (21.028) & (21.085) \\ 
+  population 2013 (log) & -0.624 & -0.558 \\ 
+  & (0.489) & (0.489) \\ 
+  \% of unemployment 2013 & -0.429^{**} & -0.417^{**} \\ 
+  & (0.206) & (0.202) \\ 
+  \% of white population 2013 & -5.973 & -4.944 \\ 
+  & (4.748) & (4.629) \\ 
+  income 2013 (log) & -24.364^{***} & -22.207^{***} \\ 
+  & (6.541) & (6.437) \\ 
+  population density 2013 & -1.110^{***} & -1.101^{***} \\ 
+  & (0.379) & (0.368) \\ 
+  \% of employment in service 2011 & -0.117 & -0.100 \\ 
+  & (0.084) & (0.082) \\ 
+  commute in minutes 2013 & 0.367^{*} & 0.349^{*} \\ 
+  & (0.192) & (0.188) \\ 
+  Constant & 196.828^{***} & 177.500^{***} \\ 
+  & (60.591) & (59.487) \\ 
  \hline \\[-1.8ex] 
@@
 Adjusted R$^{2}$ & \multicolumn{1}{c}{0.061} & \multicolumn{1}{c}{0.080} \\ 
-Residual Std. Error & \multicolumn{1}{c}{0.008} & \multicolumn{1}{c}{0.008} \\ 
+Residual Std. Error & \multicolumn{1}{c}{7.819} & \multicolumn{1}{c}{7.740} \\ 
 \hline 
@@
  & \multicolumn{2}{r}{IVs: (2) Bachelors degree per hab. in 2005, Commute, minutes, 2005} \\ 
+ & \multicolumn{2}{r}{First stage regressions can be found in S2} \\ 
 \end{tabular} 
@@
 the US urban hierarchy for the cities and the time period included in
-our analysis. Everything else being equal, using the normalization
-formula for the \(difference\:in\:ranks_{i}\), if a city whose position
-in the US urban system did not change during the period \(2013-2018\)
-(that is \(Normalized\:difference\:in\:ranks_{i} = 0.5\)) had
-experienced an increase of 10 per cent points of internet penetration,
-this increase would have improved its relative position by \(9\) places.
-As in the global model reported in \protect\hyperlink{sec4}{Section 4},
-internet penetration appears to work in favor of agglomeration
-externalities. In regards to the control variables, we can identify a
-negative effect of income, which is consistent with the effect of GDP
-per capita in the global model presented in Table \ref{2sls.global}.
-During the study period, unemployment rates and the share of white
-population negatively affected the ranking of micropolitan and
-metropolitan areas in the US as reflected in the relevant regressions.
+our analysis. \color{blue} Everything else being equal, if a city had
+experienced an increase of 10 per cent of internet penetration, this
+increase would have improved its relative position by \(9\) places.
+\color{black} As in the global model reported in
+\protect\hyperlink{sec4}{Section 4}, internet penetration appears to
+work in favor of agglomeration externalities. In regards to the control
+variables, we can identify a negative effect of income, which is
+consistent with the effect of GDP per capita in the global model
+presented in Table \ref{2sls.global}. During the study period,
+unemployment rates and the share of white population negatively affected
+the ranking of micropolitan and metropolitan areas in the US as
+reflected in the relevant regressions.
 
@@
 validity of this data can be found in the work of Riddlesden and
-Singleton {[}63{]}. The point nature of these individual level data
+Singleton {[}67{]}. The point nature of these individual level data
 enables us to aggregate them up to any urban level that we are
@@
 built-up area; larger built-up areas are separated to smaller
-sub-divisions of built-up areas {[}64--66{]}.
+sub-divisions of built-up areas {[}68--70{]}.
 
@@
 \hline \\[-1.8ex] 
-normalized difference in ranks, 2011-18 & 6,163 & 0.455 & 0.498 & 0.005 & 0.593 \\ 
+difference in ranks, 2011-18 & 6,163 & $-$602 & $-$21.749 & 70.668 & 1,235 \\ 
 download speed, 2011 & 6,163 & 520 & 3,745.519 & 2,523.129 & 50,280 \\ 
@@
 employment in service, 2011 (\%) & 6,163 & 0.511 & 0.776 & 0.054 & 0.971 \\ 
-Number of universities & 6,163 & 0 & 0.021 & 0.250 & 11 \\ 
+Number of universities & 6,163 & 0 & 0.021 & 0.245 & 11 \\ 
 \hline \\[-1.8ex] 
@@
 \begin{table}[!htbp] \centering 
-  \caption{OLS estimation of equation (4) for the UK\label{ols.uk}} 
+  \caption{OLS and GLM estimations of equation (4) for the UK\label{ols.uk}} 
   \label{} 
 \small 
-\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
+\begin{tabular}{@{\extracolsep{1pt}}lcccc} 
 \\[-1.8ex]\hline 
@@
 \cline{2-5} 
-\\[-1.8ex] & \multicolumn{4}{c}{Normalized difference in ranks, 2011-18} \\ 
- & \multicolumn{3}{c}{OLS} & GLM \\ 
+ & \multicolumn{4}{c}{Difference in ranks, 2013-18} \\ 
 \\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
 \hline \\[-1.8ex] 
- download speed, 2011 (log) & 0.0004$^{***}$ & 0.001$^{***}$ & 0.002$^{**}$ & 0.0004$^{***}$ \\ 
-  & (0.0002) & (0.0004) & (0.001) & (0.0002) \\ 
-  population, 2011 (log) & 0.0004$^{***}$ & 0.0004$^{***}$ & 0.002$^{***}$ & 0.0004$^{***}$ \\ 
-  & (0.0001) & (0.0001) & (0.001) & (0.0001) \\ 
-  broadband tests per capita, 2011 & $-$0.0003 & $-$0.0002 & $-$0.0002 & $-$0.0003 \\ 
-  & (0.001) & (0.001) & (0.001) & (0.001) \\ 
-  \% of unemployment, 2011 & $-$0.011$^{***}$ & $-$0.011$^{***}$ & $-$0.011$^{***}$ & $-$0.011$^{***}$ \\ 
-  & (0.004) & (0.004) & (0.004) & (0.004) \\ 
-  \% of British population, 2011 & $-$0.003$^{**}$ & $-$0.003$^{**}$ & $-$0.003$^{**}$ & $-$0.003$^{**}$ \\ 
-  & (0.001) & (0.001) & (0.001) & (0.001) \\ 
-  population density, 2011 & 0.00000 & 0.0002$^{*}$ & 0.00000 & 0.00000 \\ 
-  & (0.00001) & (0.0001) & (0.00001) & (0.00001) \\ 
-  \% of people working from home, 2011 & $-$0.006$^{**}$ & $-$0.005$^{*}$ & $-$0.006$^{*}$ & $-$0.006$^{**}$ \\ 
-  & (0.003) & (0.003) & (0.003) & (0.003) \\ 
-  employment in service, 2011 (\%) & $-$0.005$^{**}$ & $-$0.005$^{**}$ & $-$0.005$^{**}$ & $-$0.005$^{**}$ \\ 
-  & (0.002) & (0.002) & (0.002) & (0.002) \\ 
+ download speed, 2011 (log) & 5.484$^{***}$ & 14.807$^{***}$ & 21.973$^{**}$ & 0.0004$^{***}$ \\ 
+  & (2.029) & (5.348) & (8.791) & (0.0002) \\ 
+  population, 2011 (log) & 5.720$^{***}$ & 5.808$^{***}$ & 23.875$^{***}$ & 0.0004$^{***}$ \\ 
+  & (1.052) & (1.056) & (8.964) & (0.0001) \\ 
+  broadband tests per capita, 2011 & $-$3.991 & $-$2.247 & $-$2.017 & $-$0.0003 \\ 
+  & (10.382) & (10.556) & (10.471) & (0.001) \\ 
+  \% of unemployment, 2011 & $-$142.544$^{***}$ & $-$140.509$^{***}$ & $-$142.753$^{***}$ & $-$0.011$^{***}$ \\ 
+  & (53.532) & (53.629) & (53.493) & (0.004) \\ 
+  \% of British population, 2011 & $-$34.756$^{**}$ & $-$42.859$^{**}$ & $-$39.608$^{**}$ & $-$0.003$^{**}$ \\ 
+  & (16.009) & (16.687) & (16.466) & (0.001) \\ 
+  population density, 2011 & 0.008 & 2.793$^{*}$ & 0.011 & 0.00000 \\ 
+  & (0.150) & (1.571) & (0.150) & (0.00001) \\ 
+  \% of people working from home, 2011 & $-$80.250$^{**}$ & $-$71.360$^{*}$ & $-$74.454$^{*}$ & $-$0.006$^{**}$ \\ 
+  & (38.392) & (39.203) & (38.381) & (0.003) \\ 
+  employment in service, 2011 (\%) & $-$61.945$^{**}$ & $-$62.921$^{**}$ & $-$64.734$^{**}$ & $-$0.005$^{**}$ \\ 
+  & (28.073) & (28.066) & (28.098) & (0.002) \\ 
   download speed, 2011 (log) \\
-                             x pop. density, 2011 &  & $-$0.00003$^{*}$ &  &  \\ 
-  &  & (0.00001) &  &  \\ 
+                             x pop. density, 2011 &  & $-$0.338$^{*}$ &  &  \\ 
+  &  & (0.183) &  &  \\ 
   download speed, 2011 (log) \\
-                             x population, 2011 (log) &  &  & $-$0.0002$^{**}$ &  \\ 
-  &  &  & (0.0001) &  \\ 
-  Constant & 0.499$^{***}$ & 0.494$^{***}$ & 0.490$^{***}$ & 0.499$^{***}$ \\ 
-  & (0.003) & (0.004) & (0.006) & (0.003) \\ 
+                             x population, 2011 (log) &  &  & $-$2.175$^{**}$ &  \\ 
+  &  &  & (1.080) &  \\ 
+  Constant & $-$6.790 & $-$76.679 & $-$138.062$^{*}$ & 0.499$^{***}$ \\ 
+  & (34.074) & (48.125) & (73.822) & (0.003) \\ 
  \hline \\[-1.8ex] 
@@
 Akaike Inf. Crit. &  &  &  & $-$28,571.330 \\ 
-Residual Std. Error & 0.004 & 0.004 & 0.004 &  \\ 
+Residual Std. Error & 56.940 & 56.907 & 56.929 &  \\ 
 \hline 
@@
 \textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
- & \multicolumn{4}{r}{Robust Std. Errors} \\ 
+ & \multicolumn{4}{r}{(1)-(3) is based on OLS, (4) on GLM} \\ 
+ & \multicolumn{4}{r}{For the GLM the Normalized diff. in ranks is used} \\ 
+ & \multicolumn{4}{r}{Robust std. errors in parenthesis} \\ 
 \end{tabular} 
@@
 
-In regards to IV strategy, we use here the number of universities as the
-main instrument. Universities in England and Wales have been established
-long time ago and therefore we do not expect them to affect the change
-in the relative position of built-up areas in \(2011\). The addition of
-the second instrument (absolute number of broadband tests in 2011) in
-column \(2\) results to a Sargan test, the null hypothesis of which
-cannot be rejected. Moreover, the estimations reported in Table 11 do
-not suffer from weak instruments as the relevant test is above the rule
-of thumb value of \(10\). The coefficients derived from OLS (Table
-\ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are of the same magnitude
-and always positive and significant, something which advocates towards
-the effect of broadband speed in enhancing agglomeration forces. A back
-of the envelope calculation indicates that, everything else being equal,
-\(1\) per cent increase in download speed for a BUA, the relative
-position of which did not change during the period \(2011-2014\), would
-have increased its position by \(0.3\) places. In absolute numbers,
-\(1\) per cent increase for the average BUA would mean a marginal change
-from \(13.8\) Mbps to \(13.9\) Mbps.
+In regards to IV strategy, we use here the \emph{number of universities}
+as the main instrument. Universities in England and Wales have been
+established long time ago and therefore we do not expect them to affect
+the change in the relative position of built-up areas in \(2011\).
+\color{blue} The presence of universities and large student populations
+tend to be correlated with the quality of internet infrastructure
+provision {[}71{]}. \color{black} The addition of the second instrument
+(absolute number of broadband tests in 2011) in column \(2\) results to
+a Sargan test, the null hypothesis of which cannot be rejected.
+Moreover, the estimations reported in Table 11 do not suffer from weak
+instruments as the relevant test is above the rule of thumb value of
+\(10\). \color{blue} The coefficients derived from OLS (Table
+\ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are always positive and
+significant, something which advocates in favour of the effect of
+broadband speed in enhancing agglomeration forces. Interestingly, the
+magnitude of the 2SLS coefficient (column 1 in Table \ref{2sls.uk}) is
+almost 8 times higher than the OLS coefficient (column 1 in Table
+\ref{ols.uk}). We attribute the OLS underestimation, which is more or
+less consistent in all our estimations, to the difficulty in directly
+observing active engagement with ICT. Our endogenous variable captures
+the quality of the internet infrastructure as the experienced internet
+speed. Similarly to the endogenous variables of the global and the US
+model (internet and telephony take-up), what these variables do not
+capture is what individuals actually do with ICT {[}72{]}. For example,
+if we assume a bi-modal internet usage for either entertainment or work
+purposes, one would expect that the latter would have a larger effect on
+spatial structure than the former. We believe that our endogenous
+variables capture mostly the former. Therefore, the IV estimations,
+especially for UK case, reveal the local average treatment effect
+{[}63{]} by capturing the effect of these users/households who actively
+engage with ICT (e.g.~work on ICT-related sectors) and, therefore, have
+a higher effect on spatial structure. A back of the envelope calculation
+indicates that, everything else being equal, \(10\) per cent increase in
+download speed for a BUA, would have increased its position by \(4\)
+places. \color{black}
 
@@
 \cline{2-3} 
-\\[-1.8ex] & \multicolumn{2}{c}{Normalized difference in ranks, 2011-18} \\ 
+\\[-1.8ex] & \multicolumn{2}{c}{Difference in ranks, 2011-18} \\ 
 \\[-1.8ex] & (1) & (2)\\ 
 \hline \\[-1.8ex] 
- download speed, 2011 (log) & 0.003$^{***}$ & 0.004$^{***}$ \\ 
-  & (0.001) & (0.0005) \\ 
-  population, 2011 (log) & $-$0.0001 & $-$0.0002$^{*}$ \\ 
-  & (0.0002) & (0.0001) \\ 
-  broadband tests per capita, 2011 & $-$0.001 & $-$0.0004 \\ 
-  & (0.001) & (0.001) \\ 
-  \% of unemployment, 2011 & $-$0.007 & $-$0.007 \\ 
-  & (0.005) & (0.004) \\ 
-  \% of British population, 2011 & $-$0.003$^{**}$ & $-$0.003$^{**}$ \\ 
-  & (0.001) & (0.001) \\ 
-  population density, 2011 & 0.00000 & $-$0.00000 \\ 
-  & (0.00001) & (0.00001) \\ 
-  \% of people working from home, 2011 & $-$0.007$^{**}$ & $-$0.007$^{**}$ \\ 
-  & (0.003) & (0.003) \\ 
-  employment in service, 2011 (\%) & $-$0.004$^{*}$ & $-$0.004$^{*}$ \\ 
-  & (0.002) & (0.002) \\ 
-  Constant & 0.481$^{***}$ & 0.478$^{***}$ \\ 
-  & (0.007) & (0.004) \\ 
- \hline \\[-1.8ex] 
-Weak instruments & 7.32 & 42.23 \\ 
-Wu-Hausman & 6.15 & 40.92 \\ 
-P-value & 0.01 & 0 \\ 
-Sargan &  & 0.03 \\ 
-P-value &  & 0.86 \\ 
+ download speed, 2011 (log) & 43.058$^{***}$ & 47.226$^{***}$ \\ 
+  & (11.727) & (6.298) \\ 
+  population, 2011 (log) & $-$1.735 & $-$2.451$^{*}$ \\ 
+  & (2.327) & (1.389) \\ 
+  broadband tests per capita, 2011 & $-$6.658 & $-$5.763 \\ 
+  & (12.159) & (12.263) \\ 
+  \% of unemployment, 2011 & $-$93.185 & $-$89.366 \\ 
+  & (60.397) & (59.332) \\ 
+  \% of British population, 2011 & $-$36.963$^{**}$ & $-$38.269$^{**}$ \\ 
+  & (17.921) & (17.631) \\ 
+  population density, 2011 & $-$0.003 & $-$0.010 \\ 
+  & (0.144) & (0.147) \\ 
+  \% of people working from home, 2011 & $-$92.525$^{**}$ & $-$93.787$^{**}$ \\ 
+  & (40.412) & (40.729) \\ 
+  employment in service, 2011 (\%) & $-$47.395$^{*}$ & $-$48.132$^{*}$ \\ 
+  & (28.089) & (28.448) \\ 
+  Constant & $-$260.358$^{***}$ & $-$286.509$^{***}$ \\ 
+  & (81.193) & (51.613) \\ 
+ \hline \\[-1.8ex] 
+Weak instruments & 6.98 & 42.03 \\ 
+Wu-Hausman & 8.08 & 40.8 \\ 
+P-value & 0 & 0 \\ 
+Sargan &  & 0.02 \\ 
+P-value &  & 0.9 \\ 
 Observations & 3,032 & 3,032 \\ 
-Adjusted R$^{2}$ & $-$0.058 & $-$0.097 \\ 
-Residual Std. Error & 0.004 & 0.004 \\ 
+Adjusted R$^{2}$ & $-$0.069 & $-$0.096 \\ 
+Residual Std. Error & 57.001 & 57.721 \\ 
 \hline 
@@
 \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
- & \multicolumn{2}{r}{Robust Std. Errors} \\ 
+ & \multicolumn{2}{r}{Robust std. errors in parenthesis} \\ 
  & \multicolumn{2}{r}{IVs: (1) N.r of universities} \\ 
  & \multicolumn{2}{r}{IVs: (2) N. of universities, N. of broadband tests, 2011} \\ 
+ & \multicolumn{2}{r}{First stage regressions can be found in S2} \\ 
 \end{tabular} 
@@
 urban agglomerations with more than \(300,000\) inhabitants from the UN
-Urban Settlements data and test the effect of ICT adoption on the Zipf,
-or Pareto, coefficients for national city size distributions, as
-measures of dispersion for heavy-tailed data. We report results which
-are robust to endogeneity concerns. Then, in order to examine such
-effects for cities that are smaller than those included in our
-international data, we focus on the US and the UK urban systems.
-Specifically, we test the effect of internet usage and internet speed on
-the changes over time in rankings of Micropolitan and Metropolitan Areas
-in the US, and of Built Up Areas in the UK. The results favor a
-complementary relation between the internet and agglomeration
-externalities. While the cross-country estimates indicate that increase
-adoption rates of such technologies has resulted in less dispersed urban
-spatial structures, the two case studies reveal that internet adoption
-and internet speed improved the relative population rank of a city
-within its urban system. Interestingly, the latter results indicate that
-such effects might be even stronger for smaller and less dense urban
-areas at least in the UK. On the contrary, our estimates regarding the
-effect of fixed telephony advocate towards the reverse effect: that is
-fixed telephony acting against of agglomeration forces.
+Urban Settlements data and test the effect of ICT adoption on the Pareto
+exponents for national city size distributions, as measures of
+dispersion for heavy-tailed data. We report results which are robust to
+endogeneity concerns. Then, in order to examine such effects for cities
+that are smaller than those included in our international data, we focus
+on the US and the UK urban systems. Specifically, we test the effect of
+internet usage and internet speed on the changes over time in rankings
+of Micropolitan and Metropolitan Areas in the US, and of Built Up Areas
+in the UK. The results favor a complementary relation between the
+internet and agglomeration externalities. While the cross-country
+estimates indicate that increase adoption rates of such technologies has
+resulted in less dispersed urban spatial structures, the two case
+studies reveal that internet adoption and internet speed improved the
+relative population rank of a city within its urban system.
+Interestingly, the latter results indicate that such effects might be
+even stronger for smaller and less dense urban areas at least in the UK.
+On the contrary, our estimates regarding the effect of fixed telephony
+advocate towards the reverse effect: that is fixed telephony acting
+against of agglomeration forces.
 
@@
   increasingly popular reliance on high-resolution, pioneered by
-  Rozenfeld et al.~{[}67{]}, and satellite-based lights data, most
-  recently used by Duben and Krause {[}68{]}, for measuring city sizes,
+  Rozenfeld et al.~{[}73{]}, and satellite-based lights data, most
+  recently used by Duben and Krause {[}74{]}, for measuring city sizes,
   are very exciting developments, but they have not yet been accompanied
@@
   Settlements Initiative and the associated spatial urban data sets; see
-  Pesaresi et al.~{[}69{]}.}. Barring the availability of such data,
+  Pesaresi et al.~{[}75{]}.}. Barring the availability of such data,
 scholars must be careful in drawing conclusions. Arguably, the
@@
 
+\leavevmode\hypertarget{ref-dingel2020many}{}%
+36. Dingel JI, Neiman B. How many jobs can be done at home? National
+Bureau of Economic Research; 2020.
+doi:\href{https://doi.org/DOI\%2010.3386/w26948}{DOI 10.3386/w26948}
+
 \leavevmode\hypertarget{ref-black_urban_2003}{}%
-36. Black D, Henderson V. Urban evolution in the USA. Journal of
+37. Black D, Henderson V. Urban evolution in the USA. Journal of
 economic geography. 2003;3: 343--372.
@@
 \leavevmode\hypertarget{ref-frenken_theoretical_2007}{}%
-37. Frenken K, Boschma RA. A theoretical framework for evolutionary
+38. Frenken K, Boschma RA. A theoretical framework for evolutionary
 economic geography: Industrial dynamics and urban growth as a branching
@@
 \leavevmode\hypertarget{ref-giesen_zipfs_2011}{}%
-38. Giesen K, Südekum J. Zipf's law for cities in the regions and the
+39. Giesen K, Südekum J. Zipf's law for cities in the regions and the
 country. Journal of economic geography. 2011;11: 667--686.
@@
 \leavevmode\hypertarget{ref-rauch_cities_2013}{}%
-39. Rauch F. Cities as spatial clusters. Journal of economic geography.
+40. Rauch F. Cities as spatial clusters. Journal of economic geography.
 2013; lbt034.
@@
 \leavevmode\hypertarget{ref-ioannides_zipfs_2003}{}%
-40. Ioannides YM, Overman HG. Zipf's law for cities: An empirical
+41. Ioannides YM, Overman HG. Zipf's law for cities: An empirical
 examination. Regional science and urban economics. 2003;33: 127--137.
@@
 \leavevmode\hypertarget{ref-ioannides_walled_2017}{}%
-41. Ioannides YM, Zhang J. Walled cities in late imperial China. Journal
+42. Ioannides YM, Zhang J. Walled cities in late imperial China. Journal
 of Urban Economics. 2017;97: 71--88.
@@
 \leavevmode\hypertarget{ref-nitsch_zipf_2005}{}%
-42. Nitsch V. Zipf zipped. Journal of Urban Economics. 2005;57: 86--100.
+43. Nitsch V. Zipf zipped. Journal of Urban Economics. 2005;57: 86--100.
 
 \leavevmode\hypertarget{ref-gabaix_evolution_2004}{}%
-43. Gabaix X, Ioannides YM. The evolution of city size distributions.
+44. Gabaix X, Ioannides YM. The evolution of city size distributions.
 Handbook of regional and urban economics. 2004;4: 2341--2378.
@@
 \leavevmode\hypertarget{ref-gabaix_rank_2011}{}%
-44. Gabaix X, Ibragimov R. Rank− 1/2: A simple way to improve the OLS
+45. Gabaix X, Ibragimov R. Rank− 1/2: A simple way to improve the OLS
 estimation of tail exponents. Journal of Business \& Economic
@@
 \leavevmode\hypertarget{ref-ioannides2013neighborhoods}{}%
-45. Ioannides Y. From neighborhoods to nations: The economics of social
+46. Ioannides Y. From neighborhoods to nations: The economics of social
 interactions. Princeton University Press; 2013.
@@
 \leavevmode\hypertarget{ref-thomas_brinkhoff_city_2014}{}%
-46. Brinkhoff T. City Population. 2014.
+47. Brinkhoff T. City Population. 2014.
 
 \leavevmode\hypertarget{ref-soo_zipfs_2005}{}%
-47. Soo KT. Zipf's Law for cities: A cross-country investigation.
+48. Soo KT. Zipf's Law for cities: A cross-country investigation.
 Regional science and urban economics. 2005;35: 239--263.
@@
 \leavevmode\hypertarget{ref-cohen_urban_2004}{}%
-48. Cohen B. Urban growth in developing countries: A review of current
+49. Cohen B. Urban growth in developing countries: A review of current
 trends and a caution regarding existing forecasts. World development.
@@
 \leavevmode\hypertarget{ref-montgomery_urban_2008}{}%
-49. Montgomery MR. The urban transformation of the developing world.
+50. Montgomery MR. The urban transformation of the developing world.
 science. 2008;319: 761--764.
@@
 \leavevmode\hypertarget{ref-chen_absolute_2007}{}%
-50. Chen S, Ravallion M. Absolute poverty measures for the developing
+51. Chen S, Ravallion M. Absolute poverty measures for the developing
 world, 1981--2004. Proceedings of the National Academy of Sciences.
@@
 \leavevmode\hypertarget{ref-decker_global_2007}{}%
-51. Decker EH, Kerkhoff AJ, Moses ME. Global patterns of city size
+52. Decker EH, Kerkhoff AJ, Moses ME. Global patterns of city size
 distributions and their fundamental drivers. PLoS One. 2007;2: e934.
@@
 \leavevmode\hypertarget{ref-donner_research_2008}{}%
-52. Donner J. Research approaches to mobile use in the developing world:
+53. Donner J. Research approaches to mobile use in the developing world:
 A review of the literature. The information society. 2008;24: 140--159.
@@
 \leavevmode\hypertarget{ref-hamilton_are_2003}{}%
-53. Hamilton J. Are main lines and mobile phones substitutes or
+54. Hamilton J. Are main lines and mobile phones substitutes or
 complements? Evidence from Africa. Telecommunications Policy. 2003;27:
@@
 \leavevmode\hypertarget{ref-batty_rank_2006}{}%
-54. Batty M. Rank clocks. Nature. 2006;444: 592--596.
+55. Batty M. Rank clocks. Nature. 2006;444: 592--596.
 
 \leavevmode\hypertarget{ref-havlin_distance_1995}{}%
-55. Havlin S. The distance between Zipf plots. Physica A: Statistical
+56. Havlin S. The distance between Zipf plots. Physica A: Statistical
 Mechanics and its Applications. 1995;216: 148--150.
@@
 \leavevmode\hypertarget{ref-chetty2018impacts}{}%
-56. Chetty R, Hendren N. The impacts of neighborhoods on
+57. Chetty R, Hendren N. The impacts of neighborhoods on
 intergenerational mobility i: Childhood exposure effects. The Quarterly
@@
 \leavevmode\hypertarget{ref-dettling2017}{}%
-57. Dettling LJ. Broadband in the labor market: The impact of
+58. Dettling LJ. Broadband in the labor market: The impact of
 residential high-speed internet on married women's labor force
@@
 \leavevmode\hypertarget{ref-arntz2019working}{}%
-58. Arntz M, Sarra BY, Berlingieri F. Working from home: Heterogeneous
+59. Arntz M, Sarra BY, Berlingieri F. Working from home: Heterogeneous
 effects on hours worked and wages. ZEW-Centre for European Economic
@@
 \leavevmode\hypertarget{ref-watson2018digitalisation}{}%
-59. Watson T, Corliss M, Le M, others. Digitalisation and women's
+60. Watson T, Corliss M, Le M, others. Digitalisation and women's
 workforce participation in the indo-pacific. Australian Journal of
@@
 \leavevmode\hypertarget{ref-Lechman2020}{}%
-60. Lechman E, Popowska M. Enhancing women's engagement in economic
+61. Lechman E, Popowska M. Enhancing women's engagement in economic
 activities through information and communication technology deployment:
@@
 \leavevmode\hypertarget{ref-Valberg2020}{}%
-61. Valberg S. ICT, gender, and the labor market: A cross-country
+62. Valberg S. ICT, gender, and the labor market: A cross-country
 analysis. In: Maiti D, Castellacci F, Melchior A, editors.
@@
 
+\leavevmode\hypertarget{ref-angrist1995identification}{}%
+63. Angrist JD, Imbens GW. Identification and estimation of local
+average treatment effects. National Bureau of Economic Research; 1995.
+
+\leavevmode\hypertarget{ref-calabrese2011interplay}{}%
+64. Calabrese F, Smoreda Z, Blondel VD, Ratti C. Interplay between
+telecommunications and face-to-face interactions: A study using mobile
+phone data. PloS one. Public Library of Science; 2011;6: e20814.
+
+\leavevmode\hypertarget{ref-buchel2020cities}{}%
+65. Büchel K, Ehrlich MV. Cities and the structure of social
+interactions: Evidence from mobile phone data. Journal of Urban
+Economics. Elsevier; 2020;119: 103276.
+
 \leavevmode\hypertarget{ref-file_computer_2014}{}%
-62. File T, Ryan C. Computer and Internet use in the United States:
+66. File T, Ryan C. Computer and Internet use in the United States:
 2013. American Community Survey Reports. 2014;
@@
 \leavevmode\hypertarget{ref-riddlesden_broadband_2014}{}%
-63. Riddlesden D, Singleton AD. Broadband speed equity: A new digital
+67. Riddlesden D, Singleton AD. Broadband speed equity: A new digital
 divide? Applied Geography. 2014;52: 25--33.
@@
 \leavevmode\hypertarget{ref-ons2013}{}%
-64. ONS. 2011 Built-up Areas - Methodology and Guidance. Office for
+68. ONS. 2011 Built-up Areas - Methodology and Guidance. Office for
 National Statistics; 2013 pp. 1--15.
@@
 \leavevmode\hypertarget{ref-johnston_changing_2014}{}%
-65. Johnston R, Poulsen M, Forrest J. The changing ethnic composition of
+69. Johnston R, Poulsen M, Forrest J. The changing ethnic composition of
 urban neighbourhoods in England and Wales, 2001-2011: Creating nations
@@
 \leavevmode\hypertarget{ref-johnston2016ethnic}{}%
-66. Johnston R, Poulsen M, Forrest J. Ethnic residential patterns in
+70. Johnston R, Poulsen M, Forrest J. Ethnic residential patterns in
 urban england and wales, 2001--2011: A system-wide analysis. Tijdschrift
@@
 
+\leavevmode\hypertarget{ref-GRUBESIC201950}{}%
+71. Grubesic TH, Helderop E, Alizadeh T. Closing information
+asymmetries: A scale agnostic approach for exploring equity implications
+of broadband provision. Telecommunications Policy. 2019;43: 50--66.
+doi:\href{https://doi.org/https://doi.org/10.1016/j.telpol.2018.04.002}{https://doi.org/10.1016/j.telpol.2018.04.002}
+
+\leavevmode\hypertarget{ref-tranos2020digital}{}%
+72. Tranos E, Kitsos T, Ortega-Argilés R. Digital economy in the uk:
+Regional productivity effects of early adoption. Regional Studies.
+Routledge; 2020; 1--15.
+
 \leavevmode\hypertarget{ref-rozenfeld2011area}{}%
-67. Rozenfeld HD, Rybski D, Gabaix X, Makse HA. The area and population
+73. Rozenfeld HD, Rybski D, Gabaix X, Makse HA. The area and population
 of cities: New insights from a different perspective on cities. American
@@
 \leavevmode\hypertarget{ref-duben2019population}{}%
-68. Duben C, Krause M, others. Population, light, and the size
+74. Duben C, Krause M, others. Population, light, and the size
 distribution of cities. 2019.
@@
 \leavevmode\hypertarget{ref-pesaresi2019ghs}{}%
-69. Pesaresi M, Florczyk A, Schiavina M, Melchiorri M, Maffenini L. GHS
+75. Pesaresi M, Florczyk A, Schiavina M, Melchiorri M, Maffenini L. GHS
 settlement grid, updated and refined regio model 2014 in application to

### 2020-12-03 paper/2_resubmission/tranos_ioannides_unmarked.Rmd
unmarked

--- 2020-12-02 paper/2_resubmission/tranos_ioannides.Rmd
@@
 
-\color{blue}
+<!-- <!-- \color{blue} --> -->
 Kolko [@kolko_death_2000] uses internet diffusion data and identifies a clear complementary link between internet usage and city size.
@@
 
-\color{blue}
+<!-- <!-- \color{blue} --> -->
 Focusing on rural areas, Partridge et al.
@@
 
-\color{blue}
+<!-- <!-- \color{blue} --> -->
 Moving beyond modelling the direct relationship between ICT and spatial structure,
@@
 
-\color{blue}
+<!-- \color{blue} -->
 The present paper does not address the issue that adoption of ICT may have
@@
 The first step of our methodology is to estimate the
-\color{blue}
+<!-- \color{blue} -->
 Pareto exponent
@@
 for a broad sample of countries over time.
-\color{blue}
+<!-- \color{blue} -->
 The Pareto exponent
@@
 This an appropriate measure of dispersion because of the extreme heterogeneity of the city size distribution and the very good fits normally obtained with such estimations^[
-\color{blue}
+<!-- \color{blue} -->
 We also replicate our analysis using other measures of dispersion in Appendix S1. Our results remain consistent.].
@@
 
-\color{blue}
+<!-- \color{blue} -->
 This leads to an estimate of $\zeta$, known as the Zipf coefficient,
@@
 
-\color{blue}
+<!-- \color{blue} -->
 
@@
 To address a potential omitted variable bias equation \ref{eq3} includes
-\color{blue}
+<!-- \color{blue} -->
 country ($\theta_c$) and year ($\delta_t$) fixed effects;
@@
 The LHS variable is $\zeta$, the
-\color{blue}
+<!-- \color{blue} -->
 Pareto exponent,
@@
 In addition, the observations are weighted with the inverse squared standard error of the estimated
-\color{blue}
+<!-- \color{blue} -->
 Pareto exponent
@@
 In regards to the interpretation of the estimated coefficients, given that the
-\color{blue}
+<!-- \color{blue} -->
 Pareto exponent
@@
 Endogeneity might be an issue here as spatial structure, which is proxied by the
-\color{blue}
+<!-- \color{blue} -->
 Pareto exponent,
@@
 We propose here the percentage of female participation in the labor force, which is directly related to digital infrastructure, but not to spatial structure (see Table \ref{desc.global} for the descriptive statistics).
-\color{blue}
+<!-- \color{blue} -->
 The literature provides empirical and theoretical evidence that ICT are correlated with female participation in labour markets mostly because they enable teleworking [@dettling2017; @arntz2019working; @watson2018digitalisation; @Lechman2020; @Valberg2020]
@@
 The exact opposite appears to happen for fixed telephony, which our results indicate that has led to an increase of spatial dispersion of population.
-\color{blue}
+<!-- \color{blue} -->
 We expect that the IV estimates capture the local average treatment effect
@@
 Indeed, for the estimation of the
-\color{blue}
+<!-- \color{blue} -->
 Pareto exponent
@@
 
-\color{blue}
+<!-- \color{blue} -->
 It is indeed puzzling that the results on Table \ref{2sls.global} that the
@@
 The LHS variable we use here is the outcome of equation \ref{eq5}.
-\color{blue}
+<!-- \color{blue} -->
 We use this variable to estimate \ref{eq4} with OLS.
@@
 two count variables, which may also assume negative values.
-\color{blue}
+<!-- \color{blue} -->
 The results of the different estimations remain qualitatively the same regardless of the form of the LHS variable we use.
@@
 The main instrument we propose here is Bachelor’s degrees per inhabitant in $2005$^[
-\color{blue}
+<!-- \color{blue} -->
 Due to differences in scale, time and the endogenous variables, different IVs
@@
 Specifically, an increase in internet penetration led to an improvement of the position of a city in the US urban hierarchy for the cities and the time period included in our analysis.
-\color{blue}
+<!-- \color{blue} -->
 Everything else being equal, if a city had experienced an increase of 10 per cent
@@
 Universities in England and Wales have been established long time ago and therefore we do not expect them to affect the change in the relative position of built-up areas in $2011$.
-\color{blue}
+<!-- \color{blue} -->
 The presence of universities and large student populations tend to be correlated
@@
 Moreover, the estimations reported in Table 11 do not suffer from weak instruments as the relevant test is above the rule of thumb value of $10$.
-\color{blue}
+<!-- \color{blue} -->
 The coefficients derived from OLS (Table \ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are always positive and significant, something which advocates in favour of the effect of broadband speed in enhancing agglomeration forces.

### 2020-12-03 paper/2_resubmission/tranos_ioannides_unmarked.tex
unmarked

--- 2020-12-02 paper/2_resubmission/tranos_ioannides.tex
@@
 
-\color{blue}
-
-Kolko {[}22{]} uses internet diffusion data and identifies a clear
-complementary link between internet usage and city size. Interestingly,
-he identifies higher internet domain densities in remote cities which
-indicates a substitution effect of the internet for longer-distance
-non-electronic communications. His results are consistent for different
-measures of internet diffusion (internet domain density and internet
-take-up). Sinai and Waldfogel {[}23{]} approach the same question from
-the consumers' point of view and find support for complementarity
-between internet and cities. They study the link between market size and
-locally targeted online content and find more online local content in
-larger markets. However, they also obtain evidence for a substitution
-effect: holding local online content constant, market size has a
-negative effect on individual connectivity. Forman et al.~{[}24{]}
-examines whether commercial internet adoption is higher in cities than
-in rural areas. While the former would indicate a complementarity
-between internet adoption and cities, the latter would reflect a
-substitutability relation, according to which the internet is used as a
-means to offset costs and lack of opportunities related to
+--\textgreater{} Kolko {[}22{]} uses internet diffusion data and
+identifies a clear complementary link between internet usage and city
+size. Interestingly, he identifies higher internet domain densities in
+remote cities which indicates a substitution effect of the internet for
+longer-distance non-electronic communications. His results are
+consistent for different measures of internet diffusion (internet domain
+density and internet take-up). Sinai and Waldfogel {[}23{]} approach the
+same question from the consumers' point of view and find support for
+complementarity between internet and cities. They study the link between
+market size and locally targeted online content and find more online
+local content in larger markets. However, they also obtain evidence for
+a substitution effect: holding local online content constant, market
+size has a negative effect on individual connectivity. Forman et
+al.~{[}24{]} examines whether commercial internet adoption is higher in
+cities than in rural areas. While the former would indicate a
+complementarity between internet adoption and cities, the latter would
+reflect a substitutability relation, according to which the internet is
+used as a means to offset costs and lack of opportunities related to
 peripheriality. Their results indicate that despite internet adoption by
@@
 
-\color{blue}
-
-Focusing on rural areas, Partridge et al.~{[}28{]} find no evidence that
-rural distance penalties in the US have substantially changed since
-\(1970s\) indicating that technological changes including the internet
-and digital communications have not affected spatial structure. Also
-interesting are the findings of a more recent study by Kim and Orazem
-{[}29{]} on the economic effects of broadband internet in rural areas.
-They identify a positive effect on new firm location decisions, with the
-effect being higher in rural areas with larger population and in rural
-areas adjacent to a metropolitan area. These results suggest a
-complementary relation between the internet and agglomeration economies.
-
-\color{blue}
-
-Moving beyond modelling the direct relationship between ICT and spatial
-structure, Bekkerman and Gilpin {[}30{]} focused on the role of locally
-based information resources using a dataset about the US libraries
-during the period \(2000-2008\). Their results suggest that internet
-access increases the demand and the value of locally accessible
+--\textgreater{} Focusing on rural areas, Partridge et al.~{[}28{]} find
+no evidence that rural distance penalties in the US have substantially
+changed since \(1970s\) indicating that technological changes including
+the internet and digital communications have not affected spatial
+structure. Also interesting are the findings of a more recent study by
+Kim and Orazem {[}29{]} on the economic effects of broadband internet in
+rural areas. They identify a positive effect on new firm location
+decisions, with the effect being higher in rural areas with larger
+population and in rural areas adjacent to a metropolitan area. These
+results suggest a complementary relation between the internet and
+agglomeration economies.
+
+--\textgreater{} Moving beyond modelling the direct relationship between
+ICT and spatial structure, Bekkerman and Gilpin {[}30{]} focused on the
+role of locally based information resources using a dataset about the US
+libraries during the period \(2000-2008\). Their results suggest that
+internet access increases the demand and the value of locally accessible
 information, and such complementarities are higher in larger
@@
 
-\color{blue}
-
 The present paper does not address the issue that adoption of ICT may
@@
 The multi-country identification strategy is a two-step approach
-{[}18{]}. The first step of our methodology is to estimate the
-\color{blue} Pareto exponent \color{black} for a broad sample of
-countries over time. \color{blue} The Pareto exponent \color{black} is
-one of the most widely used measures of spatial dispersion with numerous
-applications in urban economics and economic geography ({[}see for
-example {[}37{]}; {[}38{]}; {[}39{]}; {[}40{]}; {[}41{]}; {[}42{]};
-{[}43{]}). This an appropriate measure of dispersion because of the
-extreme heterogeneity of the city size distribution and the very good
-fits normally obtained with such estimations\footnote{\color{blue} We
-  also replicate our analysis using other measures of dispersion in
-  Appendix S1. Our results remain consistent.}. \color{black} Skipping
-details pertaining of the suitability of the approach, which may be
-found in Ioannides et al.~{[}18{]} and Tranos and Ioannides {[}19{]}, we
-proceed with the estimation of the logarithmic form of:
+{[}18{]}. The first step of our methodology is to estimate the Pareto
+exponent \color{black} for a broad sample of countries over time. The
+Pareto exponent \color{black} is one of the most widely used measures of
+spatial dispersion with numerous applications in urban economics and
+economic geography ({[}see for example {[}37{]}; {[}38{]}; {[}39{]};
+{[}40{]}; {[}41{]}; {[}42{]}; {[}43{]}). This an appropriate measure of
+dispersion because of the extreme heterogeneity of the city size
+distribution and the very good fits normally obtained with such
+estimations\footnote{We also replicate our analysis using other measures
+  of dispersion in Appendix S1. Our results remain consistent.}.
+\color{black} Skipping details pertaining of the suitability of the
+approach, which may be found in Ioannides et al.~{[}18{]} and Tranos and
+Ioannides {[}19{]}, we proceed with the estimation of the logarithmic
+form of:
 
@@
 \end{align}
-
-\color{blue}
 
@@
 
-\color{blue}
-
 \begin{align}
@@
 users per \(100\) inhabitants. To address a potential omitted variable
-bias equation \ref{eq3} includes \color{blue} country (\(\theta_c\)) and
-year (\(\delta_t\)) fixed effects; \color{black} \(e_{ct}\) is the error
+bias equation \ref{eq3} includes country (\(\theta_c\)) and year
+(\(\delta_t\)) fixed effects; \color{black} \(e_{ct}\) is the error
 term. In addition, vector \(X\) includes several control variables,
@@
 and government waste in others.
-
-\color{blue}
 
@@
 This section presents the estimation results of \ref{eq3}. The LHS
-variable is \(\zeta\), the \color{blue} Pareto exponent, \color{black}
-as estimated according to the Gabaix and Ibragimov {[}45{]} correction.
-The main explanatory variables of interest are provided at the top of
-Table \ref{ols.global}; namely internet, broadband, mobile and fixed
-telephony per \(100\) habitants, expressed in natural logarithms. The
-main variables of interest are introduced successively on their own in
-the regressions reported in Table \ref{ols.global}. All regressions
-include country fixed effects to control for unobserved heterogeneity
-and a time trend. In addition, the observations are weighted with the
-inverse squared standard error of the estimated \color{blue} Pareto
-exponent \color{black} (see Table \ref{desc.global}) to address
-potential noise that is carried over from the first part of our
-identification strategy. In regards to the interpretation of the
-estimated coefficients, given that the \color{blue} Pareto exponent
-\color{black} has entered the regression not as an absolute value, but
-instead as a real number a positive coefficient for a RHS variable
-indicates an impact towards the decrease of the spatial dispersion of
-population. In other words, a positive coefficient indicates an effect
-towards less uniform city sizes that is more dispersion of city sizes.
-The latter is indicative of enhancement of agglomeration economies
-because of the expansion of digital technologies.
+variable is \(\zeta\), the Pareto exponent, \color{black} as estimated
+according to the Gabaix and Ibragimov {[}45{]} correction. The main
+explanatory variables of interest are provided at the top of Table
+\ref{ols.global}; namely internet, broadband, mobile and fixed telephony
+per \(100\) habitants, expressed in natural logarithms. The main
+variables of interest are introduced successively on their own in the
+regressions reported in Table \ref{ols.global}. All regressions include
+country fixed effects to control for unobserved heterogeneity and a time
+trend. In addition, the observations are weighted with the inverse
+squared standard error of the estimated Pareto exponent \color{black}
+(see Table \ref{desc.global}) to address potential noise that is carried
+over from the first part of our identification strategy. In regards to
+the interpretation of the estimated coefficients, given that the Pareto
+exponent \color{black} has entered the regression not as an absolute
+value, but instead as a real number a positive coefficient for a RHS
+variable indicates an impact towards the decrease of the spatial
+dispersion of population. In other words, a positive coefficient
+indicates an effect towards less uniform city sizes that is more
+dispersion of city sizes. The latter is indicative of enhancement of
+agglomeration economies because of the expansion of digital
+technologies.
 
@@
 effect. Endogeneity might be an issue here as spatial structure, which
-is proxied by the \color{blue} Pareto exponent, \color{black} might be
-affected by another source, which also affects internet penetration. For
-instance, economic development might affect the concentration of
-population in large cities and at the same time enable more people to go
-online. If we do not address this issue, the coefficient for the main
-variable of interest will capture potential effects that internet
-penetration has on spatial structure, but also potential reverse
-causality effects that spatial structure might generate on internet
-penetration. To overcome this potential problem, Table \ref{2sls.global}
-reports estimates of equation \ref{eq3} using two-stage least squares
-(2SLS) with instrumental variables (IVs). The latter are variables which
-are correlated with our endogenous variables, but do not influence
-current spatial structure. Such an approach will enable us to estimate
-the causal effect -- if any -- of the internet and digital
-communications on spatial structure. At a first stage, our endogenous
-variables are regressed against the IVs. Then, the predicted values of
-the endogenous variables based on the IVs and the other control
-variables are used instead of the endogenous variable to estimate eq.
-\ref{eq3}. A significant effect will verify the causal impact of the
-internet and digital communication usage on spatial structure.
+is proxied by the Pareto exponent, \color{black} might be affected by
+another source, which also affects internet penetration. For instance,
+economic development might affect the concentration of population in
+large cities and at the same time enable more people to go online. If we
+do not address this issue, the coefficient for the main variable of
+interest will capture potential effects that internet penetration has on
+spatial structure, but also potential reverse causality effects that
+spatial structure might generate on internet penetration. To overcome
+this potential problem, Table \ref{2sls.global} reports estimates of
+equation \ref{eq3} using two-stage least squares (2SLS) with
+instrumental variables (IVs). The latter are variables which are
+correlated with our endogenous variables, but do not influence current
+spatial structure. Such an approach will enable us to estimate the
+causal effect -- if any -- of the internet and digital communications on
+spatial structure. At a first stage, our endogenous variables are
+regressed against the IVs. Then, the predicted values of the endogenous
+variables based on the IVs and the other control variables are used
+instead of the endogenous variable to estimate eq. \ref{eq3}. A
+significant effect will verify the causal impact of the internet and
+digital communication usage on spatial structure.
 
@@
 related to digital infrastructure, but not to spatial structure (see
-Table \ref{desc.global} for the descriptive statistics). \color{blue}
-The literature provides empirical and theoretical evidence that ICT are
-correlated with female participation in labour markets mostly because
-they enable teleworking {[}58--62{]} \color{black} At the same time,
-female labor participation is not related with spatial structure
-especially given the diverse sample of countries. Although one might
-think that increased female labor participation may lead to relocation
-of households to large cities and therefore affect the spatial structure
-of cities, such a process varies a lot within our dataset. Even if this
-might be true for developed countries with mature urban systems for
-which a location within a large urban center might provide opportunities
-for both male and female workers to find jobs, our dataset also includes
-countries, \(41\) per cent of the GDP of which is attributed to primary
-activities (see Table \ref{desc.global}). So, female labor participation
-might also be related to economic activities located outside large urban
-centers. In terms of the relevant tests presented in Table
-\ref{2sls.global} the weak identification first stage F-test exceeds by
-far all the Stock-Yogo critical values for all, but the mobile telephony
-regression. Hence, we are confident to interpret the effects of
-internet, broadband internet and fixed telephony usage.
+Table \ref{desc.global} for the descriptive statistics). The literature
+provides empirical and theoretical evidence that ICT are correlated with
+female participation in labour markets mostly because they enable
+teleworking {[}58--62{]} \color{black} At the same time, female labor
+participation is not related with spatial structure especially given the
+diverse sample of countries. Although one might think that increased
+female labor participation may lead to relocation of households to large
+cities and therefore affect the spatial structure of cities, such a
+process varies a lot within our dataset. Even if this might be true for
+developed countries with mature urban systems for which a location
+within a large urban center might provide opportunities for both male
+and female workers to find jobs, our dataset also includes countries,
+\(41\) per cent of the GDP of which is attributed to primary activities
+(see Table \ref{desc.global}). So, female labor participation might also
+be related to economic activities located outside large urban centers.
+In terms of the relevant tests presented in Table \ref{2sls.global} the
+weak identification first stage F-test exceeds by far all the Stock-Yogo
+critical values for all, but the mobile telephony regression. Hence, we
+are confident to interpret the effects of internet, broadband internet
+and fixed telephony usage.
 
@@
 indicate that has led to an increase of spatial dispersion of
-population. \color{blue} We expect that the IV estimates capture the
-local average treatment effect {[}63{]} of these labour markets, which
-are characterised by more active engagement with the internet, and,
+population. We expect that the IV estimates capture the local average
+treatment effect {[}63{]} of these labour markets, which are
+characterised by more active engagement with the internet, and,
 therefore, experience a larger effect on the Pareto exponents.
@@
 but are limited to urban agglomeration included in our data. Indeed, for
-the estimation of the \color{blue} Pareto exponent \color{black} we only
-included urban agglomeration of \(300,000\) inhabitants or more due to
-data availability. Hence, the above estimations cannot verify whether
-such an effect is also valid for smaller cities. Interestingly, the
-above results regarding the role of fixed telephony are in agreement
-with previous results from {[}18{]} {[}19{]} which used a smaller panel
-of countries but without constraints regarding the city size. However,
-our results do not support the findings of {[}19{]} regarding internet
+the estimation of the Pareto exponent \color{black} we only included
+urban agglomeration of \(300,000\) inhabitants or more due to data
+availability. Hence, the above estimations cannot verify whether such an
+effect is also valid for smaller cities. Interestingly, the above
+results regarding the role of fixed telephony are in agreement with
+previous results from {[}18{]} {[}19{]} which used a smaller panel of
+countries but without constraints regarding the city size. However, our
+results do not support the findings of {[}19{]} regarding internet
 usage.
-
-\color{blue}
 
@@
 
-The LHS variable we use here is the outcome of equation \ref{eq5}.
-\color{blue} We use this variable to estimate \ref{eq4} with OLS.
-Additionally, we employ a normalized form of the dependent variable, so
-as that variable is bounded between \(0\) and \(1\) and with a mean of
-\(0.5\)\footnote{\[Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.\]}.
+The LHS variable we use here is the outcome of equation \ref{eq5}. We
+use this variable to estimate \ref{eq4} with OLS. Additionally, we
+employ a normalized form of the dependent variable, so as that variable
+is bounded between \(0\) and \(1\) and with a mean of \(0.5\)\footnote{\[Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.\]}.
 This transformation enables us to estimate \ref{eq4} with a
@@
 difference between two count variables, which may also assume negative
-values. \color{blue} The results of the different estimations remain
-qualitatively the same regardless of the form of the LHS variable we
-use. Table \ref{ols.usa} reports the results of the estimation of
-equation \ref{eq4} using OLS. The sign and significance of the main
-variable of interest -- percentage of population with computer and
-broadband connection -- verifies our global level results. That is, an
-increase in internet usage improves the position of a city in the US
-urban system. The results are consistent among the different estimators
-that is OLS and GLM. In addition, Table \ref{ols.usa} also reports the
-estimates of interaction effects between the share of population with
-broadband connection with population and population density. Although
-the interaction terms are not significant, the main effect remains
+values. The results of the different estimations remain qualitatively
+the same regardless of the form of the LHS variable we use. Table
+\ref{ols.usa} reports the results of the estimation of equation
+\ref{eq4} using OLS. The sign and significance of the main variable of
+interest -- percentage of population with computer and broadband
+connection -- verifies our global level results. That is, an increase in
+internet usage improves the position of a city in the US urban system.
+The results are consistent among the different estimators that is OLS
+and GLM. In addition, Table \ref{ols.usa} also reports the estimates of
+interaction effects between the share of population with broadband
+connection with population and population density. Although the
+interaction terms are not significant, the main effect remains
 qualitatively unchanged.
@@
 (Column 2). The main instrument we propose here is Bachelor's degrees
-per inhabitant in \(2005\)\footnote{\color{blue} Due to differences in
-  scale, time and the endogenous variables, different IVs have been
-  unutilized for the different sections of the paper. \color{black}}.
-Even if the quality of human capital affect the population growth of a
-city \(10\) years later, our LHS variable adopts a systemic
-understanding of the US urban system as it measures the relative
-position of a city within the overall urban system instead of its
-population growth. In other words, a city might experience population
-growth between two periods, but if other cities have also experienced
-population growth, this might not affect its relative position within
-the urban system. Hence, we do not expect that this IV affects the LHS
-variable. Moreover, when we add the second IV, which in this case is the
-commute time in minutes also in \(2005\), we fail to reject the null
-hypothesis of the Sargan test, something which also adds the validity of
-our strategy. Furthermore, our results do not suffer from weak
-identification according to the relevant tests in Table \ref{2sls.usa}.
+per inhabitant in \(2005\)\footnote{Due to differences in scale, time
+  and the endogenous variables, different IVs have been unutilized for
+  the different sections of the paper. \color{black}}. Even if the
+quality of human capital affect the population growth of a city \(10\)
+years later, our LHS variable adopts a systemic understanding of the US
+urban system as it measures the relative position of a city within the
+overall urban system instead of its population growth. In other words, a
+city might experience population growth between two periods, but if
+other cities have also experienced population growth, this might not
+affect its relative position within the urban system. Hence, we do not
+expect that this IV affects the LHS variable. Moreover, when we add the
+second IV, which in this case is the commute time in minutes also in
+\(2005\), we fail to reject the null hypothesis of the Sargan test,
+something which also adds the validity of our strategy. Furthermore, our
+results do not suffer from weak identification according to the relevant
+tests in Table \ref{2sls.usa}.
 
@@
 the US urban hierarchy for the cities and the time period included in
-our analysis. \color{blue} Everything else being equal, if a city had
-experienced an increase of 10 per cent of internet penetration, this
-increase would have improved its relative position by \(9\) places.
-\color{black} As in the global model reported in
-\protect\hyperlink{sec4}{Section 4}, internet penetration appears to
-work in favor of agglomeration externalities. In regards to the control
-variables, we can identify a negative effect of income, which is
-consistent with the effect of GDP per capita in the global model
-presented in Table \ref{2sls.global}. During the study period,
-unemployment rates and the share of white population negatively affected
-the ranking of micropolitan and metropolitan areas in the US as
-reflected in the relevant regressions.
+our analysis. Everything else being equal, if a city had experienced an
+increase of 10 per cent of internet penetration, this increase would
+have improved its relative position by \(9\) places. \color{black} As in
+the global model reported in \protect\hyperlink{sec4}{Section 4},
+internet penetration appears to work in favor of agglomeration
+externalities. In regards to the control variables, we can identify a
+negative effect of income, which is consistent with the effect of GDP
+per capita in the global model presented in Table \ref{2sls.global}.
+During the study period, unemployment rates and the share of white
+population negatively affected the ranking of micropolitan and
+metropolitan areas in the US as reflected in the relevant regressions.
 
@@
 established long time ago and therefore we do not expect them to affect
-the change in the relative position of built-up areas in \(2011\).
-\color{blue} The presence of universities and large student populations
-tend to be correlated with the quality of internet infrastructure
-provision {[}71{]}. \color{black} The addition of the second instrument
-(absolute number of broadband tests in 2011) in column \(2\) results to
-a Sargan test, the null hypothesis of which cannot be rejected.
-Moreover, the estimations reported in Table 11 do not suffer from weak
-instruments as the relevant test is above the rule of thumb value of
-\(10\). \color{blue} The coefficients derived from OLS (Table
-\ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are always positive and
-significant, something which advocates in favour of the effect of
-broadband speed in enhancing agglomeration forces. Interestingly, the
-magnitude of the 2SLS coefficient (column 1 in Table \ref{2sls.uk}) is
-almost 8 times higher than the OLS coefficient (column 1 in Table
-\ref{ols.uk}). We attribute the OLS underestimation, which is more or
-less consistent in all our estimations, to the difficulty in directly
-observing active engagement with ICT. Our endogenous variable captures
-the quality of the internet infrastructure as the experienced internet
-speed. Similarly to the endogenous variables of the global and the US
-model (internet and telephony take-up), what these variables do not
-capture is what individuals actually do with ICT {[}72{]}. For example,
-if we assume a bi-modal internet usage for either entertainment or work
-purposes, one would expect that the latter would have a larger effect on
-spatial structure than the former. We believe that our endogenous
-variables capture mostly the former. Therefore, the IV estimations,
-especially for UK case, reveal the local average treatment effect
-{[}63{]} by capturing the effect of these users/households who actively
-engage with ICT (e.g.~work on ICT-related sectors) and, therefore, have
-a higher effect on spatial structure. A back of the envelope calculation
-indicates that, everything else being equal, \(10\) per cent increase in
-download speed for a BUA, would have increased its position by \(4\)
-places. \color{black}
+the change in the relative position of built-up areas in \(2011\). The
+presence of universities and large student populations tend to be
+correlated with the quality of internet infrastructure provision
+{[}71{]}. \color{black} The addition of the second instrument (absolute
+number of broadband tests in 2011) in column \(2\) results to a Sargan
+test, the null hypothesis of which cannot be rejected. Moreover, the
+estimations reported in Table 11 do not suffer from weak instruments as
+the relevant test is above the rule of thumb value of \(10\). The
+coefficients derived from OLS (Table \ref{ols.uk}) and 2SLS (Table
+\ref{2sls.uk}) are always positive and significant, something which
+advocates in favour of the effect of broadband speed in enhancing
+agglomeration forces. Interestingly, the magnitude of the 2SLS
+coefficient (column 1 in Table \ref{2sls.uk}) is almost 8 times higher
+than the OLS coefficient (column 1 in Table \ref{ols.uk}). We attribute
+the OLS underestimation, which is more or less consistent in all our
+estimations, to the difficulty in directly observing active engagement
+with ICT. Our endogenous variable captures the quality of the internet
+infrastructure as the experienced internet speed. Similarly to the
+endogenous variables of the global and the US model (internet and
+telephony take-up), what these variables do not capture is what
+individuals actually do with ICT {[}72{]}. For example, if we assume a
+bi-modal internet usage for either entertainment or work purposes, one
+would expect that the latter would have a larger effect on spatial
+structure than the former. We believe that our endogenous variables
+capture mostly the former. Therefore, the IV estimations, especially for
+UK case, reveal the local average treatment effect {[}63{]} by capturing
+the effect of these users/households who actively engage with ICT
+(e.g.~work on ICT-related sectors) and, therefore, have a higher effect
+on spatial structure. A back of the envelope calculation indicates that,
+everything else being equal, \(10\) per cent increase in download speed
+for a BUA, would have increased its position by \(4\) places.
+\color{black}
 

### 2021-02-01 paper/2_final_submission/tranos_ioannides.Rmd
visualisations

--- 2020-12-03 paper/2_resubmission/tranos_ioannides_unmarked.Rmd
@@
 library(rmarkdown)
-# library(data.table)
 library(tidyverse)
-# library(readxl)
-# library(httr)
 library(rprojroot)
@@
 library(ggplot2)
-#library(geojsonio)
 library(rgdal)
@@
 library(MASS)
-#library(foreign)
 library(ivpack)
-# library(leaflet)
-# library(rvest)
-# library(xml2)
-# library(tmaptools)
-# library(spdep)
-# library(spatialreg)
-
-#library(acs)
-#library(tidycensus)
-#library(rio)
-#library(wbstats)
-#library(WDI)
 library(kableExtra)
+
+path <- find_rstudio_root_file()
 ```
@@
 Table \ref{zipf} presents the estimated $\zeta$ coefficients for the panel of countries that the second stage of the analysis focuses on using the UN urban agglomerations data. 
+More detailed and interactive visualisations can be found in Appendix S2.
 It becomes evident that there is considerable variation in the estimates of $\zeta$ across the different countries in the sample.
@@
 
-```{r, echo=FALSE, warning=FALSE, results='asis'}
-
-load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/UN.RData")
-
+```{r, echo=FALSE, warning=FALSE, out.width = "85%"}
+
+# load data
+path.data <- paste0(path, "/data/data_inter/UN.RData")
+load(path.data)
+
+# a zipf table for the appendix, which includes the Zipf coef. per year and country
 zipf.table <- net.model$model %>%
@@
 
-# stargazer(zipf.table, type = "latex", digits = 3, summary = FALSE,
-#           header = FALSE,
-#           no.space=TRUE,
-#           title = "Zipf coefficients\\label{zipf}",
-#           font.size = "small",
-#           column.sep.width = "1pt",
-#           float.env = "sidewaystable")
-
 kable(zipf.table, format = "latex",
-      caption = "Pareto exponent\\label{zipf}",
+      caption = "Pareto exponents\\label{zipf}",
       booktabs = T,
@@
                     "IV: Female participation in labour force",
-                    "First stage regressions can be found in S2"))
+                    "First stage regressions can be found in S3"))
 ```
@@
 rm(list=ls())
-load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/USA.RData")
+
+# load data
+path <- find_rstudio_root_file()
+path.data <- paste0(path, "/data/data_inter/USA.RData")
+load(path.data)
 
@@
                                                "IVs: (2) Bachelors degree per hab. in 2005, Commute, minutes, 2005",
-                                               "First stage regressions can be found in S2"),
+                                               "First stage regressions can be found in S3"),
           add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
@@
 rm(list=ls())
-load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/UK.RData")
+
+path <- find_rstudio_root_file()
+path.data <- paste0(path, "/data/data_inter/UK.RData")
+load(path.data)
 
@@
                                                "IVs: (2) N. of universities, N. of broadband tests, 2011",
-                                               "First stage regressions can be found in S2"),
+                                               "First stage regressions can be found in S3"),
           add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns

### 2021-02-01 paper/2_final_submission/tranos_ioannides.tex
visualisations

--- 2020-12-03 paper/2_resubmission/tranos_ioannides_unmarked.tex
@@
 panel of countries that the second stage of the analysis focuses on
-using the UN urban agglomerations data. It becomes evident that there is
-considerable variation in the estimates of \(\zeta\) across the
+using the UN urban agglomerations data. More detailed and interactive
+visualisations can be found in Appendix S2. It becomes evident that
+there is considerable variation in the estimates of \(\zeta\) across the
 different countries in the sample. Because of the empirically
@@
 
-\caption{\label{tab:unnamed-chunk-2}Pareto exponent\label{zipf}}
+\caption{\label{tab:unnamed-chunk-2}Pareto exponents\label{zipf}}
 \resizebox{\linewidth}{!}{
@@
 \bottomrule
-\multicolumn{20}{l}{\textit{Note: }}\\
-\multicolumn{20}{l}{Corrected as per Gabaix and Ibragimov (2011)}\\
+\multicolumn{20}{l}{\rule{0pt}{1em}\textit{Note: }}\\
+\multicolumn{20}{l}{\rule{0pt}{1em}Corrected as per Gabaix and Ibragimov (2011)}\\
 \end{tabular}}
@@
  & \multicolumn{4}{r}{IV: Female participation in labour force} \\ 
- & \multicolumn{4}{r}{First stage regressions can be found in S2} \\ 
+ & \multicolumn{4}{r}{First stage regressions can be found in S3} \\ 
 \end{tabular} 
@@
  & \multicolumn{2}{r}{IVs: (2) Bachelors degree per hab. in 2005, Commute, minutes, 2005} \\ 
- & \multicolumn{2}{r}{First stage regressions can be found in S2} \\ 
+ & \multicolumn{2}{r}{First stage regressions can be found in S3} \\ 
 \end{tabular} 
@@
  & \multicolumn{2}{r}{IVs: (2) N. of universities, N. of broadband tests, 2011} \\ 
- & \multicolumn{2}{r}{First stage regressions can be found in S2} \\ 
+ & \multicolumn{2}{r}{First stage regressions can be found in S3} \\ 
 \end{tabular} 

### 2021-02-01 src/s1.Rmd
visualisations

@@
 knit: (function(inputFile, encoding) {
-    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/2_resubmission")
+    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/2_final_submission")
   })
@@
 # This is the project path
-# path <- find_rstudio_root_file()
+path <- find_rstudio_root_file()
 
-load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/UN.RData")
+# load data
+path <- find_rstudio_root_file()
+path.un <- paste0(path, "/data/data_inter/UN.RData")
+
+load(path.un)
 ```

### 2021-02-01 src/s2.Rmd
visualisations

@@
 author: ""
-output: pdf_document
+output: html_document
 header-includes:
@@
 library(tidyverse)
-library(car)
-library(lmtest)
-library(plm)
-library(stargazer)
-library(knitr)
+library(plotly)
+library(rnaturalearth)
+library(rnaturalearthdata)
+library(gridExtra)
+library(cowplot)
+library(rprojroot)
 
-un <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_un_regressions.csv")
-us <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_usa_regressions.csv")
-uk <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_uk_regressions.csv")
+path <- find_rstudio_root_file()
 ```
 
-# First stage regressions for the 2SLS estimations
+# Pareto exponents
 
-## Global model
+The below maps plot the Pareto exponents for all the countries included in our data 
+set for the years 2000, 2005, 2010 and 2010. The quantiles of the yearly exponent 
+distribution are used for the maps with the first quantile always representing 
+the countries with the of the smallest Pareto exponent.
 
-```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}
+```{r, echo=FALSE, warning=FALSE, fig.width = 8}
 
-net.2sls.s1 <- lm(net100hab ~  fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
-                      + as.factor(year) + as.factor(country), 
-               data = un, weights = 1/(se^2))
+# load data
+path.data <- paste0(path, "/data/data_inter/UN.RData")
+load(path.data)
 
-bb.2sls.s1 <- lm(bb100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
-                      + as.factor(year) + as.factor(country), 
-               data = un, weights = 1/(se^2))
+# a zipf table for the appendix, which includes the Zipf coef. per year and country
+zipf.table <- net.model$model %>%
+  dplyr::select(`as.factor(country)`, `as.factor(year)`, b) %>%
+  spread(`as.factor(year)`, b) %>%
+  rename(`Countries` = `as.factor(country)`)
 
-fixed.2sls.s1 <- lm(fixed100hab ~ fem.lab + log(pop.dens) + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
-                      + as.factor(year) + as.factor(country), 
-               data = un, weights = 1/(se^2))
+# a world `sf` to join with Zipf
+world <- ne_countries(scale = "medium", returnclass = "sf")
+world <- merge(world, zipf.table, by.x = 'iso_a3', by.y = 'Countries', all = T)
 
-mobile.2sls.s1 <- lm(mobile100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
-                      + as.factor(year) + as.factor(country), 
-               data = un, weights = 1/(se^2))
+# yearly maps
+map2000 <- world %>%
+  mutate(eq.n2000 = cut_number(`2000`, 
+                               n = 4,
+                               labels = c("1st quantile", "2nd quantile",
+                                          "3rd quantile", "4th quantile"))) %>%
+  ggplot() +
+  geom_sf(aes(fill = eq.n2000)) + 
+  ggtitle("2000") +
+  theme_minimal() + 
+  theme(legend.position = "top",
+        legend.title = element_blank(),
+        plot.title = element_text(hjust = 0.5))
 
-net.2sls.s1.rse <- coeftest(net.2sls.s1, function(x) vcovHC(x, type="HC0"))
+legend <- get_legend(map2000)
+map2000 <- map2000 + theme(legend.position="none")
 
-bb.2sls.s1.rse <- coeftest(bb.2sls.s1, function(x) vcovHC(x, type="HC0"))
+map2005 <- world %>%
+  mutate(eq.n2005 = cut_number(`2005`, 
+                               n = 4,
+                               labels = c("1st quantile", "2nd quantile",
+                                          "3rd quantile", "4th quantile"))) %>%
+  ggplot() +
+  geom_sf(aes(fill = eq.n2005), show.legend = F) + 
+  ggtitle("2005") +
+  theme_minimal() + 
+  theme(legend.title = element_blank(),
+        plot.title = element_text(hjust = 0.5))
 
-fixed.2sls.s1.rse <- coeftest(fixed.2sls.s1, function(x) vcovHC(x, type="HC0"))
+map2010 <- world %>%
+  mutate(eq.n2010 = cut_number(`2010`, 
+                               n = 4,
+                               labels = c("1st quantile", "2nd quantile",
+                                          "3rd quantile", "4th quantile"))) %>%
+  ggplot() +
+  geom_sf(aes(fill = eq.n2010), show.legend = F) + 
+  ggtitle("2010") +
+  theme_minimal() +
+  theme(legend.title = element_blank(),
+        plot.title = element_text(hjust = 0.5))
 
-mobile.2sls.s1.rse <- coeftest(mobile.2sls.s1, function(x) vcovHC(x, type="HC0"))
+map2015 <- world %>%
+  mutate(eq.n2015 = cut_number(`2015`, 
+                               n = 4,
+                               labels = c("1st quantile", "2nd quantile",
+                                          "3rd quantile", "4th quantile"))) %>%
+  ggplot() +
+  geom_sf(aes(fill = eq.n2015), show.legend = F) + 
+  ggtitle("2015") +
+  theme_minimal() +
+  theme(legend.title = element_blank(),
+        plot.title = element_text(hjust = 0.5))
 
-stargazer(net.2sls.s1, bb.2sls.s1, fixed.2sls.s1, mobile.2sls.s1, type = "latex",
-          se = list(net.2sls.s1.rse[,"Std. Error"], bb.2sls.s1.rse[,"Std. Error"],
-                    fixed.2sls.s1.rse[,"Std. Error"], mobile.2sls.s1.rse[,"Std. Error"]),
-          header = FALSE,
-          no.space=TRUE,
-          title = "First stage regressions for Table 5 \\label{2sls.global.s1}",
-          font.size = "footnotesize",
-          column.sep.width = "1pt",
-          # dep.var.labels=c("Internet users per 100 hab. (log)",
-          #                 "Broadband users per 100 hab. (log)",
-          #                 "Mobile phone users per 100 hab. (log)",
-          #                 "Fixed phone users per 100 hab. (log)"),
-          column.labels=c('Internet users',
-                           "Broadband users",
-                           "Mobile phone users",
-                           "Fixed phone users"),
-          dep.var.labels.include = FALSE,
-          #column.separate = c(1,1,1,1),
-          covariate.labels=c("Female labour force (\\%)",
-                             "Population density (log)",
-                             "Government expenditure (\\% GDP)",
-                             "Trade (\\% of GDP)",
-                             "Non agriculture value added (\\% GDP)",
-                             "GDP growth",
-                             "GDP per capita (log)",
-                             "Population (log)"),
-          single.row = FALSE, df = FALSE,
-          omit = c("country", "year"),
-          omit.stat = c("rsq", "f"),
-          add.lines = list(
-            # c(rownames(summ.net.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
-            #                  round(summ.net.2sls$diagnostics[1, "statistic"], 2),
-            #                  round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
-            #                  round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
-            #                  round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)),
-            #                c(rownames(summ.net.2sls$diagnostics)[2],
-            #                  round(summ.net.2sls$diagnostics[2, "statistic"], 2),
-            #                  round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
-            #                  round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
-            #                  round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)),
-            #                c("P-value",
-            #                  round(summ.net.2sls$diagnostics[2, "p-value"], 2),
-            #                  round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
-            #                  round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
-            #                  round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)),
-                           # c(rownames(summ.net.2sls$diagnostics)[3],
-                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
-                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
-                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
-                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
-                           # c("P-value",
-                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
-                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
-                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
-                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
-                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
-                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
-          notes = c("Robust Std. Errors in parenthesis"))
-
-net.test <- car::linearHypothesis(net.2sls.s1, c("fem.lab=0"))
-bb.test <- car::linearHypothesis(bb.2sls.s1, c("fem.lab=0"))
-mob.test <- car::linearHypothesis(fixed.2sls.s1, c("fem.lab=0"))
-fixed.test <- car::linearHypothesis(mobile.2sls.s1, c("fem.lab=0"))
-
-tests <- data.frame("F" = c(net.test$F[2], bb.test$F[2], mob.test$F[2], fixed.test$F[2]),
-                    "p" = c(net.test$`Pr(>F)`[2], bb.test$`Pr(>F)`[2], mob.test$`Pr(>F)`[2], fixed.test$`Pr(>F)`[2]),
-                    row.names = c("internet users", "broadband users", "mobile phone users", "fixed telephony users"))
-
-stargazer(tests, 
-          summary = F, 
-          type = "latex",
-          digits = 3,
-          title = "F tests for IV (female labour force (\\%)) from first stage regressions for Table 5 \\label{2sls.global.s1.f}",
-          header = FALSE
-)
+lay <- rbind(c(1,2),
+             c(3,4),
+             c(5,5))
+grid.arrange(map2000, map2005, map2010, map2015, legend,
+            nrow=3, 
+            heights=c(2.3, 2.3, 0.3),
+            layout_matrix = lay)
 ```
 
-\newpage
+The Pareto exponents are plotted below as an interactive line graph.
 
-## US model
+```{r, echo=FALSE, warning=FALSE, fig.width = 8}
 
-```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}
+zipf.line <- zipf.table %>% 
+  pivot_longer(!Countries, names_to = "year", values_to = "zipf") %>% 
+  ggplot(aes(x=year, y=zipf, group = Countries, colour = Countries),
+         show.legend = F) + 
+  geom_line() + guides(colour=FALSE) + xlab("Year") +
+  ylab("Pareto exponent")
 
-model_iv1.s1 <- lm(hh.net.share ~ bachelors2005 + 
-                     log(pop2013) + un13 + white2013share + 
-                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
-                   data = us)
-
-model_iv2.s1 <- lm(hh.net.share ~ bachelors2005 + commuting.minutes2005 + 
-                     log(pop2013) + un13 + white2013share + 
-                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
-                   data = us)
-
-rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0")) # rob = rse, naming inconsistency
-rob.model_iv2.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))
-
-stargazer(model_iv1.s1, model_iv2.s1, 
-          type = "latex",
-          title = "First stage regressions for Table 8 \\label{2sls.us.s1}",
-          header = FALSE,
-          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]), 
-          dep.var.labels=c("\\% of households w. internet 2013"),
-          covariate.labels=c("bachelors degree per hab. 2005",
-                             "commute in minutes 2005",
-                             "population 2013 (log)",
-                             "\\% of unemployment 2013",
-                             "\\% of white population 2013",
-                             "income 2013 (log)",
-                             "population density 2013",
-                             "\\% of employment in service 2011",
-                             "commute in minutes 2012"),
-          single.row = FALSE,
-          df = FALSE, 
-          omit.stat = c("rsq", "f"), 
-          notes = c("Robust Std. Error in parenthesis"))
-
-model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("bachelors2005=0"))
-model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("bachelors2005=0","commuting.minutes2005=0"))
-
-tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
-                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
-                    row.names = c("bachelors degree per hab., 2005", "the above + commute in minutes, 2005"))
-
-stargazer(tests, 
-          summary = F, 
-          type = "latex",
-          digits = 3,
-          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
-          header = FALSE
-)
-```
-
-\newpage
-
-## UK model
-
-```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}
-
-model_iv1.s1 <- lm(log(download2011) ~ uni.freq +
-                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
-                     british.pop.share +  pop.dens + work.home.share + services.share
-                     , data = uk, subset = n.tests2011>30)
-
-model_iv2.s1 <- lm(log(download2011) ~ uni.freq + n.tests2011 +
-                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
-                     british.pop.share +  pop.dens + work.home.share + services.share
-                     , data = uk, subset = n.tests2011>30)
-
-rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))
-rob.model_iv2.s1 <- coeftest(model_iv2.s1, function(x) vcovHC(x, type="HC0"))
-
-stargazer(model_iv1.s1, model_iv2.s1,
-          type="latex",
-          summary = F,
-          header = FALSE,
-          no.space=TRUE,
-          title = "First stage regressions for Table 11 \\label{2sls.us.s1}",
-          font.size = "small",
-          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]),
-          dep.var.labels=c("download speed, 2011 (log)"),
-          covariate.labels=c("N. of universities",
-                             "N. of broadband tests, 2011",
-                             "population, 2011 (log)",  
-                             "broadband tests per capita, 2011",
-                             "\\% of unemployment, 2011",
-                             "\\% of British population, 2011",
-                             "population density, 2011",
-                             "\\% of people working from home, 2011",
-                             "employment in service, 2011 (\\%)"),
-          single.row = FALSE,
-          df = FALSE,
-          omit.stat = c("rsq", "f"), 
-          notes = c("Robust Std. Errors"))
-
-# N. of universities",
-# N. of universities, N. of broadband tests, 2011"),
-
-model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("uni.freq=0"))
-model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("uni.freq=0", "n.tests2011=0"))
-
-tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
-                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
-                    row.names = c("N. of universities", "the above + N. of broadband tests, 2011"))
-
-stargazer(tests, 
-          summary = F, 
-          type = "latex",
-          digits = 3,
-          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
-          header = FALSE
-)
+ggplotly(zipf.line,
+         show.legend=F)
 

### 2021-02-01 src/s3.Rmd
visualisations

---
title: "S3 Appendix"
author: ""
output: pdf_document
header-includes:
- \usepackage{multirow}
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/2_final_submission")
  })
---

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(rmarkdown)
library(tidyverse)
library(car)
library(lmtest)
library(plm)
library(stargazer)
library(knitr)
library(rprojroot)

# load data
path <- find_rstudio_root_file()
path.un <- paste0(path, "/data/data_inter/data_for_un_regressions.csv")
path.us <- paste0(path, "/data/data_inter/data_for_usa_regressions.csv")
path.uk <- paste0(path, "/data/data_inter/data_for_uk_regressions.csv")

un <- read_csv(path.un)
us <- read_csv(path.us)
uk <- read_csv(path.uk)
```

# First stage regressions for the 2SLS estimations

## Global model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

net.2sls.s1 <- lm(net100hab ~  fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

bb.2sls.s1 <- lm(bb100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

fixed.2sls.s1 <- lm(fixed100hab ~ fem.lab + log(pop.dens) + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

mobile.2sls.s1 <- lm(mobile100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

net.2sls.s1.rse <- coeftest(net.2sls.s1, function(x) vcovHC(x, type="HC0"))

bb.2sls.s1.rse <- coeftest(bb.2sls.s1, function(x) vcovHC(x, type="HC0"))

fixed.2sls.s1.rse <- coeftest(fixed.2sls.s1, function(x) vcovHC(x, type="HC0"))

mobile.2sls.s1.rse <- coeftest(mobile.2sls.s1, function(x) vcovHC(x, type="HC0"))

stargazer(net.2sls.s1, bb.2sls.s1, fixed.2sls.s1, mobile.2sls.s1, type = "latex",
          se = list(net.2sls.s1.rse[,"Std. Error"], bb.2sls.s1.rse[,"Std. Error"],
                    fixed.2sls.s1.rse[,"Std. Error"], mobile.2sls.s1.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "First stage regressions for Table 5 \\label{2sls.global.s1}",
          font.size = "footnotesize",
          column.sep.width = "1pt",
          # dep.var.labels=c("Internet users per 100 hab. (log)",
          #                 "Broadband users per 100 hab. (log)",
          #                 "Mobile phone users per 100 hab. (log)",
          #                 "Fixed phone users per 100 hab. (log)"),
          column.labels=c('Internet users',
                           "Broadband users",
                           "Mobile phone users",
                           "Fixed phone users"),
          dep.var.labels.include = FALSE,
          #column.separate = c(1,1,1,1),
          covariate.labels=c("Female labour force (\\%)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(
            # c(rownames(summ.net.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
            #                  round(summ.net.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)),
            #                c(rownames(summ.net.2sls$diagnostics)[2],
            #                  round(summ.net.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)),
            #                c("P-value",
            #                  round(summ.net.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis"))

net.test <- car::linearHypothesis(net.2sls.s1, c("fem.lab=0"))
bb.test <- car::linearHypothesis(bb.2sls.s1, c("fem.lab=0"))
mob.test <- car::linearHypothesis(fixed.2sls.s1, c("fem.lab=0"))
fixed.test <- car::linearHypothesis(mobile.2sls.s1, c("fem.lab=0"))

tests <- data.frame("F" = c(net.test$F[2], bb.test$F[2], mob.test$F[2], fixed.test$F[2]),
                    "p" = c(net.test$`Pr(>F)`[2], bb.test$`Pr(>F)`[2], mob.test$`Pr(>F)`[2], fixed.test$`Pr(>F)`[2]),
                    row.names = c("internet users", "broadband users", "mobile phone users", "fixed telephony users"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IV (female labour force (\\%)) from first stage regressions for Table 5 \\label{2sls.global.s1.f}",
          header = FALSE
)
```

\newpage

## US model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

model_iv1.s1 <- lm(hh.net.share ~ bachelors2005 + 
                     log(pop2013) + un13 + white2013share + 
                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
                   data = us)

model_iv2.s1 <- lm(hh.net.share ~ bachelors2005 + commuting.minutes2005 + 
                     log(pop2013) + un13 + white2013share + 
                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
                   data = us)

rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0")) # rob = rse, naming inconsistency
rob.model_iv2.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))

stargazer(model_iv1.s1, model_iv2.s1, 
          type = "latex",
          title = "First stage regressions for Table 8 \\label{2sls.us.s1}",
          header = FALSE,
          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]), 
          dep.var.labels=c("\\% of households w. internet 2013"),
          covariate.labels=c("bachelors degree per hab. 2005",
                             "commute in minutes 2005",
                             "population 2013 (log)",
                             "\\% of unemployment 2013",
                             "\\% of white population 2013",
                             "income 2013 (log)",
                             "population density 2013",
                             "\\% of employment in service 2011",
                             "commute in minutes 2012"),
          single.row = FALSE,
          df = FALSE, 
          omit.stat = c("rsq", "f"), 
          notes = c("Robust Std. Error in parenthesis"))

model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("bachelors2005=0"))
model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("bachelors2005=0","commuting.minutes2005=0"))

tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
                    row.names = c("bachelors degree per hab., 2005", "the above + commute in minutes, 2005"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
          header = FALSE
)
```

\newpage

## UK model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

model_iv1.s1 <- lm(log(download2011) ~ uni.freq +
                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share
                     , data = uk, subset = n.tests2011>30)

model_iv2.s1 <- lm(log(download2011) ~ uni.freq + n.tests2011 +
                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share
                     , data = uk, subset = n.tests2011>30)

rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))
rob.model_iv2.s1 <- coeftest(model_iv2.s1, function(x) vcovHC(x, type="HC0"))

stargazer(model_iv1.s1, model_iv2.s1,
          type="latex",
          summary = F,
          header = FALSE,
          no.space=TRUE,
          title = "First stage regressions for Table 11 \\label{2sls.us.s1}",
          font.size = "small",
          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]),
          dep.var.labels=c("download speed, 2011 (log)"),
          covariate.labels=c("N. of universities",
                             "N. of broadband tests, 2011",
                             "population, 2011 (log)",  
                             "broadband tests per capita, 2011",
                             "\\% of unemployment, 2011",
                             "\\% of British population, 2011",
                             "population density, 2011",
                             "\\% of people working from home, 2011",
                             "employment in service, 2011 (\\%)"),
          single.row = FALSE,
          df = FALSE,
          omit.stat = c("rsq", "f"), 
          notes = c("Robust Std. Errors"))

# N. of universities",
# N. of universities, N. of broadband tests, 2011"),

model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("uni.freq=0"))
model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("uni.freq=0", "n.tests2011=0"))

tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
                    row.names = c("N. of universities", "the above + N. of broadband tests, 2011"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
          header = FALSE
)

```

### 2021-03-23 paper/2_final_submission/tranos_ioannides.Rmd
editing changes

@@
 Hence, face-to-face interactions and the implied knowledge spillovers within cities flourish because of lower transportation and interaction costs [@storper_buzz_2004].
-However, ICT have the capacity to directly affect this process by further reducing transportation cost^[Or, more broader, spatial transmission costs to follow McCann’s terminology [@mccann_globalization_2008]] [@mccann_globalization_2008].
+However, ICT have the capacity to directly affect this process by further reducing transportation cost [@mccann_globalization_2008].
 In essence, the internet and digital communications promote further learning and matching, which are at the heart of the micro-foundations of agglomeration economies identified by Duranton and Puga [@duranton_micro-foundations_2004].
 Web-based applications such as Massive Open Online Courses decrease the need for co-location of actors in order to participate in formal learning activities.
-Furthermore, online social media such as LinkedIn^[LinkedIn has been identified as "the largest professional matchmaker site in the world" [@van_dijck_you_2013], p. 207.
-See also [@gee2019more].]
-can enhance the probability of matching and indeed the quality of matches even within cities.
+Furthermore, online social media such as LinkedIn can enhance the probability of matching and indeed the quality of matches even within cities.
+This is why LinkedIn has been identified as "the largest professional matchmaker site in the world" [[@van_dijck_you_2013] p. 207; [@gee2019more]].
 Therefore, it is natural to ask whether the widespread adoption of ICT could offset the benefits of  immediate physical proximity and result in more dispersed spatial structures, or further reinforce agglomeration  externalities and lead to more concentrated spatial structures.
@@
 
-<!-- <!-- \color{blue} --> -->
 Kolko [@kolko_death_2000] uses internet diffusion data and identifies a clear complementary link between internet usage and city size.
@@
 
-<!-- <!-- \color{blue} --> -->
 Focusing on rural areas, Partridge et al.
@@
 
-<!-- <!-- \color{blue} --> -->
 Moving beyond modelling the direct relationship between ICT and spatial structure,
@@
 
-\color{black}
 Thus, clearly, previous research suggests that the declaration of the “death of distance” has  proven to be premature [@rietveld_transport_2004].
@@
 In addition, although email and instant messaging technologies were widespread in the developed world in early $2000s$, network externalities due to mobile internet and online social media were nowhere close to what we are familiar with today.
-For instance, Facebook users increased from $1$ million in $2004$ to more than $1.5$ billion in $2015$^[[http://newsroom.fb.com/company-info/](http://newsroom.fb.com/company-info/)].
+For instance, Facebook users increased from $1$ million in $2004$ to more than $1.5$ billion in $2015$ [@fb].
 Hence, it might have been premature for the spatial economic effects of ICT adoption to have been materialized by the time that most of the above studies were conducted.
@@
 
-<!-- \color{blue} -->
 The present paper does not address the issue that adoption of ICT may have
@@
 continue in an unencumbered manner; see Dingel and Neiman [@dingel2020many].
-\color{black}
 
@@
 The first step of our methodology is to estimate the
-<!-- \color{blue} -->
 Pareto exponent
-\color{black}
 for a broad sample of countries over time.
-<!-- \color{blue} -->
-The Pareto exponent
-\color{black}
-is one of the most widely used measures of spatial dispersion with numerous applications in urban economics and economic geography ([see for example @black_urban_2003; @frenken_theoretical_2007; @giesen_zipfs_2011; @rauch_cities_2013; @ioannides_zipfs_2003; @ioannides_walled_2017; @nitsch_zipf_2005).
-This an appropriate measure of dispersion because of the extreme heterogeneity of the city size distribution and the very good fits normally obtained with such estimations^[
-<!-- \color{blue} -->
-We also replicate our analysis using other measures of dispersion in Appendix S1. Our results remain consistent.].
-\color{black}
+The Pareto exponent is one of the most widely used measures of spatial dispersion with numerous applications in urban economics and economic geography ([see for example @black_urban_2003; @frenken_theoretical_2007; @giesen_zipfs_2011; @rauch_cities_2013; @ioannides_zipfs_2003; @ioannides_walled_2017; @nitsch_zipf_2005).
+This an appropriate measure of dispersion because of the extreme heterogeneity of the city size distribution and the very good fits normally obtained with such estimations.
+We also replicate our analysis using other measures of dispersion in Appendix S1. Our results remain consistent.
 Skipping details pertaining of the suitability of the approach, which may be found in Ioannides et al.
@@
 
-<!-- \color{blue} -->
 This leads to an estimate of $\zeta$, known as the Zipf coefficient,
@@
 $i$ in country $c$ in year $t$:
-\color{black}
 
@@
 Researchers working in this area must contend with definitional differences as well as differences in availability of different kinds of data sources.
-Definitions of cities differ across countries, for political, administrative and legal reasons^[See [@ioannides2013neighborhoods], Ch.
-8, for issues and pitfalls associated with different definitions.].
-In contrast to the data drawn from Thomas Brinkhoff’s City Population project [@thomas_brinkhoff_city_2014]^[[http://www.citypopulation.de](http://www.citypopulation.de)], which were  used by Soo [@soo_zipfs_2005], Ioannides et al.
+Definitions of cities differ across countries, for political, administrative and legal reasons [see [@ioannides2013neighborhoods], Ch.
+8, for issues and pitfalls associated with different definitions].
+In contrast to the data drawn from Thomas Brinkhoff’s City Population project [@citypop], 
+which were  used by Soo [@soo_zipfs_2005], Ioannides et al.
 [@ioannides2008effect] and Tranos and Ioannides [@tranos2020ict], the multi-country analysis in the present paper is based on the annual population data for urban agglomerations with $300,000$ inhabitants or more from the Department of Economic and Social Affairs of the United Nations [@united_nations_world_2014].
@@
 
-<!-- \color{blue} -->
-
 \begin{align}
@@
 
-\color{black}
 \noindent The model enable us to estimate the effects of a vector $X$ of explanatory variables which are included to account for the spatial structure of country $c$ in year $t$.
@@
 To address a potential omitted variable bias equation \ref{eq3} includes
-<!-- \color{blue} -->
 country ($\theta_c$) and year ($\delta_t$) fixed effects;
-\color{black}
 $e_{ct}$ is the error term.
@@
 Government expenditure as a share of GDP may be a proxy of public investment in some countries and government waste in others.
-
-<!-- \color{blue} -->
 
@@
 
-\color{black}
-
 ```{r, echo=FALSE, warning=FALSE, results='asis'}
@@
 
-In order to capture the micro-dynamics of the urban systems in the two case studies, we follow Batty [@batty_rank_2006] and Havlin [@havlin_distance_1995] and focus on the difference in ranks for individual cities during the study period ^[Changes in ranking of individuals has also been used elsewhere in economics because of its robustness properties.
-C.f.
-Chetty and Hendren [@chetty2018impacts].].
+In order to capture the micro-dynamics of the urban systems in the two case studies, we follow Batty [@batty_rank_2006] and Havlin [@havlin_distance_1995] and focus on the difference in ranks for individual cities during the study period.
+Changes in ranking of individuals have also been used elsewhere in economics because 
+of its robustness properties. C.f. Chetty and Hendren [@chetty2018impacts].
 Contrary to their approach, we are interested in the real difference in rankings rather than in absolute differences, in order to capture whether or a city improves its ranking within the urban system during the study period.
@@
 This section presents the estimation results of \ref{eq3}.
-The LHS variable is $\zeta$, the
-<!-- \color{blue} -->
-Pareto exponent,
-\color{black}
+The LHS variable is $\zeta$, the Pareto exponent,
 as estimated according to the Gabaix and Ibragimov [@gabaix_rank_2011] correction.
@@
 In addition, the observations are weighted with the inverse squared standard error of the estimated
-<!-- \color{blue} -->
-Pareto exponent
-\color{black}
-(see Table \ref{desc.global}) to address potential noise that is carried over from the first part of our identification strategy.
-In regards to the interpretation of the estimated coefficients, given that the
-<!-- \color{blue} -->
-Pareto exponent
-\color{black}
+Pareto exponent (see Table \ref{desc.global}) to address potential noise that is carried over from the first part of our identification strategy.
+In regards to the interpretation of the estimated coefficients, given that the Pareto exponent
 has entered the regression not as an absolute value, but instead as a real number a positive coefficient for a RHS variable indicates an impact towards the decrease of the spatial dispersion of population.
@@
 Endogeneity might be an issue here as spatial structure, which is proxied by the
-<!-- \color{blue} -->
 Pareto exponent,
-\color{black}
 might be affected by another source, which also affects internet penetration.
@@
 We propose here the percentage of female participation in the labor force, which is directly related to digital infrastructure, but not to spatial structure (see Table \ref{desc.global} for the descriptive statistics).
-<!-- \color{blue} -->
 The literature provides empirical and theoretical evidence that ICT are correlated with female participation in labour markets mostly because they enable teleworking [@dettling2017; @arntz2019working; @watson2018digitalisation; @Lechman2020; @Valberg2020]
-\color{black}
 At the same time, female labor participation is not related with spatial structure especially given the diverse sample of countries.
@@
 The exact opposite appears to happen for fixed telephony, which our results indicate that has led to an increase of spatial dispersion of population.
-<!-- \color{blue} -->
 We expect that the IV estimates capture the local average treatment effect
@@
 on the Pareto exponents.
-\color{black}
 Importantly, our identification strategy enables us to address potential reverse causality issues and treat the results of Table \ref{2sls.global} as causal.
@@
 Indeed, for the estimation of the
-<!-- \color{blue} -->
 Pareto exponent
-\color{black}
 we only included urban agglomeration of $300,000$ inhabitants or more due to data availability.
@@
 
-<!-- \color{blue} -->
 It is indeed puzzling that the results on Table \ref{2sls.global} that the
@@
 communicate more.
-\color{black}
 
@@
 The LHS variable we use here is the outcome of equation \ref{eq5}.
-<!-- \color{blue} -->
 We use this variable to estimate \ref{eq4} with OLS.
 Additionally, we employ a normalized form of the dependent variable, so as that
-variable is bounded between $0$ and $1$ and with a mean of  $0.5$^[$$Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.$$].
+variable is bounded between $0$ and $1$ and with a mean of  $0.5$:
+
+\begin{align}
+Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}},\label{eq6}
+\end{align}
+
 This transformation enables us to estimate \ref{eq4}
@@
 two count variables, which may also assume negative values.
-<!-- \color{blue} -->
 The results of the different estimations remain qualitatively the same regardless of the form of the LHS variable we use.
@@
 In addition to the strategy employed for the cross-country analysis -- that is the inclusion of one instrument for which we have strong reasons to believe is uncorrelated with the error term -- data availability allows us to employ a second IV in order to estimate the Sargan over-identification restrictions test (Column 2).
-The main instrument we propose here is Bachelor’s degrees per inhabitant in $2005$^[
-<!-- \color{blue} -->
-Due to differences in scale, time and the endogenous variables, different IVs
-have been unutilized for the different sections of the paper.
-\color{black}].
+The main instrument we propose here is Bachelor’s degrees per inhabitant in $2005$.
+We need to highlight here that due to differences in scale, time and the endogenous variables, 
+different IVs have been utilized for the different sections of the paper. 
 Even if the quality of human capital affect the population growth of a city $10$ years later, our LHS variable adopts a systemic understanding of the US urban system as it measures the relative position of a city within the overall urban system instead of its population growth.
@@
 Specifically, an increase in internet penetration led to an improvement of the position of a city in the US urban hierarchy for the cities and the time period included in our analysis.
-<!-- \color{blue} -->
 Everything else being equal, if a city had experienced an increase of 10 per cent
 of internet penetration, this increase would have improved its relative position by $9$ places.
-\color{black}
 As in the global model reported in [Section 4](#sec4), internet penetration appears to work in favor of agglomeration externalities.
@@
 Universities in England and Wales have been established long time ago and therefore we do not expect them to affect the change in the relative position of built-up areas in $2011$.
-<!-- \color{blue} -->
 The presence of universities and large student populations tend to be correlated
 with the quality of internet infrastructure provision [@GRUBESIC201950].
-\color{black}
 <!-- The correlation coefficient with the LHS variables is smaller than $0.05$. -->
@@
 Moreover, the estimations reported in Table 11 do not suffer from weak instruments as the relevant test is above the rule of thumb value of $10$.
-<!-- \color{blue} -->
 The coefficients derived from OLS (Table \ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are always positive and significant, something which advocates in favour of the effect of broadband speed in enhancing agglomeration forces.
@@
 would have increased its position by $4$ places.
-\color{black}
 
@@
 We believe strongly that the differences lie in the different definitions of urban areas.
-To this date, and in spite of numerous efforts by teams worldwide, there has not been a universally accepted definition, which could moreover be implemented consistently by governments worldwide^[
+To this date, and in spite of numerous efforts by teams worldwide, there has not been a universally accepted definition, which could moreover be implemented consistently by governments worldwide.
 The increasingly popular reliance on high-resolution, pioneered by Rozenfeld et al.
@@
 A promising new development is the EU’s Global Human Settlements Initiative and the associated spatial urban data sets; see Pesaresi et al. [@pesaresi2019ghs].
-].
 Barring the availability of such data, scholars must be careful in drawing conclusions.

### 2021-03-23 paper/2_final_submission/tranos_ioannides.tex
editing changes

@@
 capacity to directly affect this process by further reducing
-transportation cost\footnote{Or, more broader, spatial transmission
-  costs to follow McCann's terminology {[}14{]}} {[}14{]}. In essence,
-the internet and digital communications promote further learning and
-matching, which are at the heart of the micro-foundations of
-agglomeration economies identified by Duranton and Puga {[}15{]}.
-Web-based applications such as Massive Open Online Courses decrease the
-need for co-location of actors in order to participate in formal
-learning activities. Furthermore, online social media such as
-LinkedIn\footnote{LinkedIn has been identified as ``the largest
-  professional matchmaker site in the world'' {[}16{]}, p.~207. See also
-  {[}17{]}.} can enhance the probability of matching and indeed the
-quality of matches even within cities. Therefore, it is natural to ask
+transportation cost {[}14{]}. In essence, the internet and digital
+communications promote further learning and matching, which are at the
+heart of the micro-foundations of agglomeration economies identified by
+Duranton and Puga {[}15{]}. Web-based applications such as Massive Open
+Online Courses decrease the need for co-location of actors in order to
+participate in formal learning activities. Furthermore, online social
+media such as LinkedIn can enhance the probability of matching and
+indeed the quality of matches even within cities. This is why LinkedIn
+has been identified as ``the largest professional matchmaker site in the
+world'' {[}{[}16{]} p.~207; {[}17{]}{]}. Therefore, it is natural to ask
 whether the widespread adoption of ICT could offset the benefits of
@@
 
---\textgreater{} Kolko {[}22{]} uses internet diffusion data and
-identifies a clear complementary link between internet usage and city
-size. Interestingly, he identifies higher internet domain densities in
-remote cities which indicates a substitution effect of the internet for
-longer-distance non-electronic communications. His results are
-consistent for different measures of internet diffusion (internet domain
-density and internet take-up). Sinai and Waldfogel {[}23{]} approach the
-same question from the consumers' point of view and find support for
-complementarity between internet and cities. They study the link between
-market size and locally targeted online content and find more online
-local content in larger markets. However, they also obtain evidence for
-a substitution effect: holding local online content constant, market
-size has a negative effect on individual connectivity. Forman et
-al.~{[}24{]} examines whether commercial internet adoption is higher in
-cities than in rural areas. While the former would indicate a
-complementarity between internet adoption and cities, the latter would
-reflect a substitutability relation, according to which the internet is
-used as a means to offset costs and lack of opportunities related to
+Kolko {[}22{]} uses internet diffusion data and identifies a clear
+complementary link between internet usage and city size. Interestingly,
+he identifies higher internet domain densities in remote cities which
+indicates a substitution effect of the internet for longer-distance
+non-electronic communications. His results are consistent for different
+measures of internet diffusion (internet domain density and internet
+take-up). Sinai and Waldfogel {[}23{]} approach the same question from
+the consumers' point of view and find support for complementarity
+between internet and cities. They study the link between market size and
+locally targeted online content and find more online local content in
+larger markets. However, they also obtain evidence for a substitution
+effect: holding local online content constant, market size has a
+negative effect on individual connectivity. Forman et al.~{[}24{]}
+examines whether commercial internet adoption is higher in cities than
+in rural areas. While the former would indicate a complementarity
+between internet adoption and cities, the latter would reflect a
+substitutability relation, according to which the internet is used as a
+means to offset costs and lack of opportunities related to
 peripheriality. Their results indicate that despite internet adoption by
@@
 
---\textgreater{} Focusing on rural areas, Partridge et al.~{[}28{]} find
-no evidence that rural distance penalties in the US have substantially
-changed since \(1970s\) indicating that technological changes including
-the internet and digital communications have not affected spatial
-structure. Also interesting are the findings of a more recent study by
-Kim and Orazem {[}29{]} on the economic effects of broadband internet in
-rural areas. They identify a positive effect on new firm location
-decisions, with the effect being higher in rural areas with larger
-population and in rural areas adjacent to a metropolitan area. These
-results suggest a complementary relation between the internet and
-agglomeration economies.
-
---\textgreater{} Moving beyond modelling the direct relationship between
-ICT and spatial structure, Bekkerman and Gilpin {[}30{]} focused on the
-role of locally based information resources using a dataset about the US
-libraries during the period \(2000-2008\). Their results suggest that
-internet access increases the demand and the value of locally accessible
+Focusing on rural areas, Partridge et al.~{[}28{]} find no evidence that
+rural distance penalties in the US have substantially changed since
+\(1970s\) indicating that technological changes including the internet
+and digital communications have not affected spatial structure. Also
+interesting are the findings of a more recent study by Kim and Orazem
+{[}29{]} on the economic effects of broadband internet in rural areas.
+They identify a positive effect on new firm location decisions, with the
+effect being higher in rural areas with larger population and in rural
+areas adjacent to a metropolitan area. These results suggest a
+complementary relation between the internet and agglomeration economies.
+
+Moving beyond modelling the direct relationship between ICT and spatial
+structure, Bekkerman and Gilpin {[}30{]} focused on the role of locally
+based information resources using a dataset about the US libraries
+during the period \(2000-2008\). Their results suggest that internet
+access increases the demand and the value of locally accessible
 information, and such complementarities are higher in larger
@@
 cities by focusing on food truck industry in the US.
-
-\color{black}
 
@@
 Facebook users increased from \(1\) million in \(2004\) to more than
-\(1.5\) billion in \(2015\)\footnote{\url{http://newsroom.fb.com/company-info/}}.
-Hence, it might have been premature for the spatial economic effects of
-ICT adoption to have been materialized by the time that most of the
-above studies were conducted.
+\(1.5\) billion in \(2015\) {[}36{]}. Hence, it might have been
+premature for the spatial economic effects of ICT adoption to have been
+materialized by the time that most of the above studies were conducted.
 
@@
 that not all economic activities might continue in an unencumbered
-manner; see Dingel and Neiman {[}36{]}. \color{black}
+manner; see Dingel and Neiman {[}37{]}.
 
@@
 {[}18{]}. The first step of our methodology is to estimate the Pareto
-exponent \color{black} for a broad sample of countries over time. The
-Pareto exponent \color{black} is one of the most widely used measures of
-spatial dispersion with numerous applications in urban economics and
-economic geography ({[}see for example {[}37{]}; {[}38{]}; {[}39{]};
-{[}40{]}; {[}41{]}; {[}42{]}; {[}43{]}). This an appropriate measure of
-dispersion because of the extreme heterogeneity of the city size
-distribution and the very good fits normally obtained with such
-estimations\footnote{We also replicate our analysis using other measures
-  of dispersion in Appendix S1. Our results remain consistent.}.
-\color{black} Skipping details pertaining of the suitability of the
+exponent for a broad sample of countries over time. The Pareto exponent
+is one of the most widely used measures of spatial dispersion with
+numerous applications in urban economics and economic geography ({[}see
+for example {[}38{]}; {[}39{]}; {[}40{]}; {[}41{]}; {[}42{]}; {[}43{]};
+{[}44{]}). This an appropriate measure of dispersion because of the
+extreme heterogeneity of the city size distribution and the very good
+fits normally obtained with such estimations. We also replicate our
+analysis using other measures of dispersion in Appendix S1. Our results
+remain consistent. Skipping details pertaining of the suitability of the
 approach, which may be found in Ioannides et al.~{[}18{]} and Tranos and
@@
 dispersion, equation \ref{eq1} describes the rank of city \(i\) in
-country \(c\) in year \(t\): \color{black}
+country \(c\) in year \(t\):
 
@@
 The estimation of equation \ref{eq2} has typically been performed by
-Ordinary Least Squares (OLS). Gabaix and Ioannides {[}44{]} discuss the
+Ordinary Least Squares (OLS). Gabaix and Ioannides {[}45{]} discuss the
 downward bias of estimates of equation \ref{eq2} using OLS on small
-samples. Gabaix and Ibragimov {[}45{]} propose a practical remedy to
+samples. Gabaix and Ibragimov {[}46{]} propose a practical remedy to
 correct this bias, which we do adopt in this paper: instead of using the
@@
 data sources. Definitions of cities differ across countries, for
-political, administrative and legal reasons\footnote{See {[}46{]}, Ch.
-  8, for issues and pitfalls associated with different definitions.}. In
+political, administrative and legal reasons {[}see {[}47{]}, Ch. 8, for
+issues and pitfalls associated with different definitions{]}. In
 contrast to the data drawn from Thomas Brinkhoff's City Population
-project {[}47{]}\footnote{\url{http://www.citypopulation.de}}, which
-were used by Soo {[}48{]}, Ioannides et al.~{[}18{]} and Tranos and
-Ioannides {[}19{]}, the multi-country analysis in the present paper is
-based on the annual population data for urban agglomerations with
-\(300,000\) inhabitants or more from the Department of Economic and
-Social Affairs of the United Nations {[}9{]}. Despite some criticism
-about the consistency of the urban agglomeration definitions across
-different countries {[}49{]}, this is the only available source for
-yearly, multi-country population data for urban agglomerations
-{[}50--52{]}. The results based on these data are reported and discussed
-in \protect\hyperlink{sec4}{Section 4}.
+project {[}48{]}, which were used by Soo {[}49{]}, Ioannides et
+al.~{[}18{]} and Tranos and Ioannides {[}19{]}, the multi-country
+analysis in the present paper is based on the annual population data for
+urban agglomerations with \(300,000\) inhabitants or more from the
+Department of Economic and Social Affairs of the United Nations {[}9{]}.
+Despite some criticism about the consistency of the urban agglomeration
+definitions across different countries {[}50{]}, this is the only
+available source for yearly, multi-country population data for urban
+agglomerations {[}51--53{]}. The results based on these data are
+reported and discussed in \protect\hyperlink{sec4}{Section 4}.
 
@@
 
-\color{black}
-
 \noindent The model enable us to estimate the effects of a vector \(X\)
@@
 bias equation \ref{eq3} includes country (\(\theta_c\)) and year
-(\(\delta_t\)) fixed effects; \color{black} \(e_{ct}\) is the error
-term. In addition, vector \(X\) includes several control variables,
-whose descriptive statistics together with those for the other variables
-used to estimate \ref{eq3} are reported in Table \ref{desc.global}.
+(\(\delta_t\)) fixed effects; \(e_{ct}\) is the error term. In addition,
+vector \(X\) includes several control variables, whose descriptive
+statistics together with those for the other variables used to estimate
+\ref{eq3} are reported in Table \ref{desc.global}.
 
@@
 the lack of fixed line infrastructure and mobile phone networks are also
-used as the main way to use the internet {[}53,54{]}.
+used as the main way to use the internet {[}54,55{]}.
 
@@
 \end{table}
-
-\color{black}
 
@@
 In order to capture the micro-dynamics of the urban systems in the two
-case studies, we follow Batty {[}55{]} and Havlin {[}56{]} and focus on
-the difference in ranks for individual cities during the study
-period.\footnote{Changes in ranking of individuals has also been used
-  elsewhere in economics because of its robustness properties. C.f.
-  Chetty and Hendren {[}57{]}.} Contrary to their approach, we are
-interested in the real difference in rankings rather than in absolute
-differences, in order to capture whether or a city improves its ranking
-within the urban system during the study period. We then test whether
-our internet-related variable has an effect on the change in ranking.
-Hence, we define the left-hand side (LHS) variable of equation \ref{eq4}
-as follows:
+case studies, we follow Batty {[}56{]} and Havlin {[}57{]} and focus on
+the difference in ranks for individual cities during the study period.
+Changes in ranking of individuals have also been used elsewhere in
+economics because of its robustness properties. C.f. Chetty and Hendren
+{[}58{]}. Contrary to their approach, we are interested in the real
+difference in rankings rather than in absolute differences, in order to
+capture whether or a city improves its ranking within the urban system
+during the study period. We then test whether our internet-related
+variable has an effect on the change in ranking. Hence, we define the
+left-hand side (LHS) variable of equation \ref{eq4} as follows:
 
@@
 This section presents the estimation results of \ref{eq3}. The LHS
-variable is \(\zeta\), the Pareto exponent, \color{black} as estimated
-according to the Gabaix and Ibragimov {[}45{]} correction. The main
-explanatory variables of interest are provided at the top of Table
-\ref{ols.global}; namely internet, broadband, mobile and fixed telephony
-per \(100\) habitants, expressed in natural logarithms. The main
-variables of interest are introduced successively on their own in the
-regressions reported in Table \ref{ols.global}. All regressions include
-country fixed effects to control for unobserved heterogeneity and a time
-trend. In addition, the observations are weighted with the inverse
-squared standard error of the estimated Pareto exponent \color{black}
-(see Table \ref{desc.global}) to address potential noise that is carried
-over from the first part of our identification strategy. In regards to
-the interpretation of the estimated coefficients, given that the Pareto
-exponent \color{black} has entered the regression not as an absolute
-value, but instead as a real number a positive coefficient for a RHS
-variable indicates an impact towards the decrease of the spatial
-dispersion of population. In other words, a positive coefficient
-indicates an effect towards less uniform city sizes that is more
-dispersion of city sizes. The latter is indicative of enhancement of
-agglomeration economies because of the expansion of digital
-technologies.
+variable is \(\zeta\), the Pareto exponent, as estimated according to
+the Gabaix and Ibragimov {[}46{]} correction. The main explanatory
+variables of interest are provided at the top of Table \ref{ols.global};
+namely internet, broadband, mobile and fixed telephony per \(100\)
+habitants, expressed in natural logarithms. The main variables of
+interest are introduced successively on their own in the regressions
+reported in Table \ref{ols.global}. All regressions include country
+fixed effects to control for unobserved heterogeneity and a time trend.
+In addition, the observations are weighted with the inverse squared
+standard error of the estimated Pareto exponent (see Table
+\ref{desc.global}) to address potential noise that is carried over from
+the first part of our identification strategy. In regards to the
+interpretation of the estimated coefficients, given that the Pareto
+exponent has entered the regression not as an absolute value, but
+instead as a real number a positive coefficient for a RHS variable
+indicates an impact towards the decrease of the spatial dispersion of
+population. In other words, a positive coefficient indicates an effect
+towards less uniform city sizes that is more dispersion of city sizes.
+The latter is indicative of enhancement of agglomeration economies
+because of the expansion of digital technologies.
 
@@
 effect. Endogeneity might be an issue here as spatial structure, which
-is proxied by the Pareto exponent, \color{black} might be affected by
-another source, which also affects internet penetration. For instance,
-economic development might affect the concentration of population in
-large cities and at the same time enable more people to go online. If we
-do not address this issue, the coefficient for the main variable of
-interest will capture potential effects that internet penetration has on
-spatial structure, but also potential reverse causality effects that
-spatial structure might generate on internet penetration. To overcome
-this potential problem, Table \ref{2sls.global} reports estimates of
-equation \ref{eq3} using two-stage least squares (2SLS) with
-instrumental variables (IVs). The latter are variables which are
-correlated with our endogenous variables, but do not influence current
-spatial structure. Such an approach will enable us to estimate the
-causal effect -- if any -- of the internet and digital communications on
-spatial structure. At a first stage, our endogenous variables are
-regressed against the IVs. Then, the predicted values of the endogenous
-variables based on the IVs and the other control variables are used
-instead of the endogenous variable to estimate eq. \ref{eq3}. A
-significant effect will verify the causal impact of the internet and
-digital communication usage on spatial structure.
+is proxied by the Pareto exponent, might be affected by another source,
+which also affects internet penetration. For instance, economic
+development might affect the concentration of population in large cities
+and at the same time enable more people to go online. If we do not
+address this issue, the coefficient for the main variable of interest
+will capture potential effects that internet penetration has on spatial
+structure, but also potential reverse causality effects that spatial
+structure might generate on internet penetration. To overcome this
+potential problem, Table \ref{2sls.global} reports estimates of equation
+\ref{eq3} using two-stage least squares (2SLS) with instrumental
+variables (IVs). The latter are variables which are correlated with our
+endogenous variables, but do not influence current spatial structure.
+Such an approach will enable us to estimate the causal effect -- if any
+-- of the internet and digital communications on spatial structure. At a
+first stage, our endogenous variables are regressed against the IVs.
+Then, the predicted values of the endogenous variables based on the IVs
+and the other control variables are used instead of the endogenous
+variable to estimate eq. \ref{eq3}. A significant effect will verify the
+causal impact of the internet and digital communication usage on spatial
+structure.
 
@@
 female participation in labour markets mostly because they enable
-teleworking {[}58--62{]} \color{black} At the same time, female labor
-participation is not related with spatial structure especially given the
-diverse sample of countries. Although one might think that increased
-female labor participation may lead to relocation of households to large
-cities and therefore affect the spatial structure of cities, such a
-process varies a lot within our dataset. Even if this might be true for
-developed countries with mature urban systems for which a location
-within a large urban center might provide opportunities for both male
-and female workers to find jobs, our dataset also includes countries,
-\(41\) per cent of the GDP of which is attributed to primary activities
-(see Table \ref{desc.global}). So, female labor participation might also
-be related to economic activities located outside large urban centers.
-In terms of the relevant tests presented in Table \ref{2sls.global} the
-weak identification first stage F-test exceeds by far all the Stock-Yogo
+teleworking {[}59--63{]} At the same time, female labor participation is
+not related with spatial structure especially given the diverse sample
+of countries. Although one might think that increased female labor
+participation may lead to relocation of households to large cities and
+therefore affect the spatial structure of cities, such a process varies
+a lot within our dataset. Even if this might be true for developed
+countries with mature urban systems for which a location within a large
+urban center might provide opportunities for both male and female
+workers to find jobs, our dataset also includes countries, \(41\) per
+cent of the GDP of which is attributed to primary activities (see Table
+\ref{desc.global}). So, female labor participation might also be related
+to economic activities located outside large urban centers. In terms of
+the relevant tests presented in Table \ref{2sls.global} the weak
+identification first stage F-test exceeds by far all the Stock-Yogo
 critical values for all, but the mobile telephony regression. Hence, we
@@
 population. We expect that the IV estimates capture the local average
-treatment effect {[}63{]} of these labour markets, which are
+treatment effect {[}64{]} of these labour markets, which are
 characterised by more active engagement with the internet, and,
 therefore, experience a larger effect on the Pareto exponents.
-\color{black} Importantly, our identification strategy enables us to
-address potential reverse causality issues and treat the results of
-Table \ref{2sls.global} as causal. Hence, the main finding of our
+Importantly, our identification strategy enables us to address potential
+reverse causality issues and treat the results of Table
+\ref{2sls.global} as causal. Hence, the main finding of our
 multi-country, global analysis based on data for urban agglomerations of
@@
 but are limited to urban agglomeration included in our data. Indeed, for
-the estimation of the Pareto exponent \color{black} we only included
-urban agglomeration of \(300,000\) inhabitants or more due to data
+the estimation of the Pareto exponent we only included urban
+agglomeration of \(300,000\) inhabitants or more due to data
 availability. Hence, the above estimations cannot verify whether such an
@@
 opposite directions. We wish to broach this subject in the context of
-some related literature, namely {[}64{]} and {[}65{]}. The latter
+some related literature, namely {[}65{]} and {[}66{]}. The latter
 supports the notion that ICT favors agglomeration in the sense that in
@@
 communications do not make up for physical distance; individuals nearer
-one another communicate more.\\
-\color{black}
+one another communicate more.
 
@@
 least \(10,000\), but less than \(50,000\) population) level of
-aggregation {[}66{]}. These functional definitions of a city represent
+aggregation {[}67{]}. These functional definitions of a city represent
 in essence labor markets and according to Table \ref{desc.usa} the
@@
 employ a normalized form of the dependent variable, so as that variable
-is bounded between \(0\) and \(1\) and with a mean of \(0.5\)\footnote{\[Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}}.\]}.
+is bounded between \(0\) and \(1\) and with a mean of \(0.5\):
+
+\begin{align}
+Normalized\:difference\:in\:ranks_{i} = \frac{difference\:in\:ranks_{i} + N_{cities}}{2xN_{cities}},\label{eq6}
+\end{align}
+
 This transformation enables us to estimate \ref{eq4} with a
@@
 (Column 2). The main instrument we propose here is Bachelor's degrees
-per inhabitant in \(2005\)\footnote{Due to differences in scale, time
-  and the endogenous variables, different IVs have been unutilized for
-  the different sections of the paper. \color{black}}. Even if the
+per inhabitant in \(2005\). We need to highlight here that due to
+differences in scale, time and the endogenous variables, different IVs
+have been utilized for the different sections of the paper. Even if the
 quality of human capital affect the population growth of a city \(10\)
@@
 increase of 10 per cent of internet penetration, this increase would
-have improved its relative position by \(9\) places. \color{black} As in
-the global model reported in \protect\hyperlink{sec4}{Section 4},
-internet penetration appears to work in favor of agglomeration
-externalities. In regards to the control variables, we can identify a
-negative effect of income, which is consistent with the effect of GDP
-per capita in the global model presented in Table \ref{2sls.global}.
-During the study period, unemployment rates and the share of white
-population negatively affected the ranking of micropolitan and
-metropolitan areas in the US as reflected in the relevant regressions.
+have improved its relative position by \(9\) places. As in the global
+model reported in \protect\hyperlink{sec4}{Section 4}, internet
+penetration appears to work in favor of agglomeration externalities. In
+regards to the control variables, we can identify a negative effect of
+income, which is consistent with the effect of GDP per capita in the
+global model presented in Table \ref{2sls.global}. During the study
+period, unemployment rates and the share of white population negatively
+affected the ranking of micropolitan and metropolitan areas in the US as
+reflected in the relevant regressions.
 
@@
 validity of this data can be found in the work of Riddlesden and
-Singleton {[}67{]}. The point nature of these individual level data
+Singleton {[}68{]}. The point nature of these individual level data
 enables us to aggregate them up to any urban level that we are
@@
 built-up area; larger built-up areas are separated to smaller
-sub-divisions of built-up areas {[}68--70{]}.
+sub-divisions of built-up areas {[}69--71{]}.
 
@@
 correlated with the quality of internet infrastructure provision
-{[}71{]}. \color{black} The addition of the second instrument (absolute
-number of broadband tests in 2011) in column \(2\) results to a Sargan
-test, the null hypothesis of which cannot be rejected. Moreover, the
-estimations reported in Table 11 do not suffer from weak instruments as
-the relevant test is above the rule of thumb value of \(10\). The
-coefficients derived from OLS (Table \ref{ols.uk}) and 2SLS (Table
-\ref{2sls.uk}) are always positive and significant, something which
-advocates in favour of the effect of broadband speed in enhancing
-agglomeration forces. Interestingly, the magnitude of the 2SLS
-coefficient (column 1 in Table \ref{2sls.uk}) is almost 8 times higher
-than the OLS coefficient (column 1 in Table \ref{ols.uk}). We attribute
-the OLS underestimation, which is more or less consistent in all our
-estimations, to the difficulty in directly observing active engagement
-with ICT. Our endogenous variable captures the quality of the internet
-infrastructure as the experienced internet speed. Similarly to the
-endogenous variables of the global and the US model (internet and
-telephony take-up), what these variables do not capture is what
-individuals actually do with ICT {[}72{]}. For example, if we assume a
-bi-modal internet usage for either entertainment or work purposes, one
-would expect that the latter would have a larger effect on spatial
-structure than the former. We believe that our endogenous variables
-capture mostly the former. Therefore, the IV estimations, especially for
-UK case, reveal the local average treatment effect {[}63{]} by capturing
-the effect of these users/households who actively engage with ICT
-(e.g.~work on ICT-related sectors) and, therefore, have a higher effect
-on spatial structure. A back of the envelope calculation indicates that,
-everything else being equal, \(10\) per cent increase in download speed
-for a BUA, would have increased its position by \(4\) places.
-\color{black}
+{[}72{]}. The addition of the second instrument (absolute number of
+broadband tests in 2011) in column \(2\) results to a Sargan test, the
+null hypothesis of which cannot be rejected. Moreover, the estimations
+reported in Table 11 do not suffer from weak instruments as the relevant
+test is above the rule of thumb value of \(10\). The coefficients
+derived from OLS (Table \ref{ols.uk}) and 2SLS (Table \ref{2sls.uk}) are
+always positive and significant, something which advocates in favour of
+the effect of broadband speed in enhancing agglomeration forces.
+Interestingly, the magnitude of the 2SLS coefficient (column 1 in Table
+\ref{2sls.uk}) is almost 8 times higher than the OLS coefficient (column
+1 in Table \ref{ols.uk}). We attribute the OLS underestimation, which is
+more or less consistent in all our estimations, to the difficulty in
+directly observing active engagement with ICT. Our endogenous variable
+captures the quality of the internet infrastructure as the experienced
+internet speed. Similarly to the endogenous variables of the global and
+the US model (internet and telephony take-up), what these variables do
+not capture is what individuals actually do with ICT {[}73{]}. For
+example, if we assume a bi-modal internet usage for either entertainment
+or work purposes, one would expect that the latter would have a larger
+effect on spatial structure than the former. We believe that our
+endogenous variables capture mostly the former. Therefore, the IV
+estimations, especially for UK case, reveal the local average treatment
+effect {[}64{]} by capturing the effect of these users/households who
+actively engage with ICT (e.g.~work on ICT-related sectors) and,
+therefore, have a higher effect on spatial structure. A back of the
+envelope calculation indicates that, everything else being equal, \(10\)
+per cent increase in download speed for a BUA, would have increased its
+position by \(4\) places.
 
@@
 been a universally accepted definition, which could moreover be
-implemented consistently by governments worldwide\footnote{The
-  increasingly popular reliance on high-resolution, pioneered by
-  Rozenfeld et al.~{[}73{]}, and satellite-based lights data, most
-  recently used by Duben and Krause {[}74{]}, for measuring city sizes,
-  are very exciting developments, but they have not yet been accompanied
-  by complementary information on ICT adoption at a comparably granular
-  level. A promising new development is the EU's Global Human
-  Settlements Initiative and the associated spatial urban data sets; see
-  Pesaresi et al.~{[}75{]}.}. Barring the availability of such data,
-scholars must be careful in drawing conclusions. Arguably, the
-measurement of the impact of ICT at the level of agglomerations, on one
-hand, and on changes in rankings, on the other, reported by this paper
-constitute a contribution to the literature.
+implemented consistently by governments worldwide. The increasingly
+popular reliance on high-resolution, pioneered by Rozenfeld et
+al.~{[}74{]}, and satellite-based lights data, most recently used by
+Duben and Krause {[}75{]}, for measuring city sizes, are very exciting
+developments, but they have not yet been accompanied by complementary
+information on ICT adoption at a comparably granular level. A promising
+new development is the EU's Global Human Settlements Initiative and the
+associated spatial urban data sets; see Pesaresi et al.~{[}76{]}.
+Barring the availability of such data, scholars must be careful in
+drawing conclusions. Arguably, the measurement of the impact of ICT at
+the level of agglomerations, on one hand, and on changes in rankings, on
+the other, reported by this paper constitute a contribution to the
+literature.
 
@@
 
+\leavevmode\hypertarget{ref-fb}{}%
+36. Facebook. Facebook company information. 2021. Available:
+\url{http://newsroom.fb.com/company-info/}
+
 \leavevmode\hypertarget{ref-dingel2020many}{}%
-36. Dingel JI, Neiman B. How many jobs can be done at home? National
+37. Dingel JI, Neiman B. How many jobs can be done at home? National
 Bureau of Economic Research; 2020.
@@
 \leavevmode\hypertarget{ref-black_urban_2003}{}%
-37. Black D, Henderson V. Urban evolution in the USA. Journal of
+38. Black D, Henderson V. Urban evolution in the USA. Journal of
 economic geography. 2003;3: 343--372.
@@
 \leavevmode\hypertarget{ref-frenken_theoretical_2007}{}%
-38. Frenken K, Boschma RA. A theoretical framework for evolutionary
+39. Frenken K, Boschma RA. A theoretical framework for evolutionary
 economic geography: Industrial dynamics and urban growth as a branching
@@
 \leavevmode\hypertarget{ref-giesen_zipfs_2011}{}%
-39. Giesen K, Südekum J. Zipf's law for cities in the regions and the
+40. Giesen K, Südekum J. Zipf's law for cities in the regions and the
 country. Journal of economic geography. 2011;11: 667--686.
@@
 \leavevmode\hypertarget{ref-rauch_cities_2013}{}%
-40. Rauch F. Cities as spatial clusters. Journal of economic geography.
+41. Rauch F. Cities as spatial clusters. Journal of economic geography.
 2013; lbt034.
@@
 \leavevmode\hypertarget{ref-ioannides_zipfs_2003}{}%
-41. Ioannides YM, Overman HG. Zipf's law for cities: An empirical
+42. Ioannides YM, Overman HG. Zipf's law for cities: An empirical
 examination. Regional science and urban economics. 2003;33: 127--137.
@@
 \leavevmode\hypertarget{ref-ioannides_walled_2017}{}%
-42. Ioannides YM, Zhang J. Walled cities in late imperial China. Journal
+43. Ioannides YM, Zhang J. Walled cities in late imperial China. Journal
 of Urban Economics. 2017;97: 71--88.
@@
 \leavevmode\hypertarget{ref-nitsch_zipf_2005}{}%
-43. Nitsch V. Zipf zipped. Journal of Urban Economics. 2005;57: 86--100.
+44. Nitsch V. Zipf zipped. Journal of Urban Economics. 2005;57: 86--100.
 
 \leavevmode\hypertarget{ref-gabaix_evolution_2004}{}%
-44. Gabaix X, Ioannides YM. The evolution of city size distributions.
+45. Gabaix X, Ioannides YM. The evolution of city size distributions.
 Handbook of regional and urban economics. 2004;4: 2341--2378.
@@
 \leavevmode\hypertarget{ref-gabaix_rank_2011}{}%
-45. Gabaix X, Ibragimov R. Rank− 1/2: A simple way to improve the OLS
+46. Gabaix X, Ibragimov R. Rank− 1/2: A simple way to improve the OLS
 estimation of tail exponents. Journal of Business \& Economic
@@
 \leavevmode\hypertarget{ref-ioannides2013neighborhoods}{}%
-46. Ioannides Y. From neighborhoods to nations: The economics of social
+47. Ioannides Y. From neighborhoods to nations: The economics of social
 interactions. Princeton University Press; 2013.
 
-\leavevmode\hypertarget{ref-thomas_brinkhoff_city_2014}{}%
-47. Brinkhoff T. City Population. 2014.
+\leavevmode\hypertarget{ref-citypop}{}%
+48. Brinkhoff T. City population. 2021. Available:
+\url{http://www.citypopulation.de}
 
 \leavevmode\hypertarget{ref-soo_zipfs_2005}{}%
-48. Soo KT. Zipf's Law for cities: A cross-country investigation.
+49. Soo KT. Zipf's Law for cities: A cross-country investigation.
 Regional science and urban economics. 2005;35: 239--263.
@@
 \leavevmode\hypertarget{ref-cohen_urban_2004}{}%
-49. Cohen B. Urban growth in developing countries: A review of current
+50. Cohen B. Urban growth in developing countries: A review of current
 trends and a caution regarding existing forecasts. World development.
@@
 \leavevmode\hypertarget{ref-montgomery_urban_2008}{}%
-50. Montgomery MR. The urban transformation of the developing world.
+51. Montgomery MR. The urban transformation of the developing world.
 science. 2008;319: 761--764.
@@
 \leavevmode\hypertarget{ref-chen_absolute_2007}{}%
-51. Chen S, Ravallion M. Absolute poverty measures for the developing
+52. Chen S, Ravallion M. Absolute poverty measures for the developing
 world, 1981--2004. Proceedings of the National Academy of Sciences.
@@
 \leavevmode\hypertarget{ref-decker_global_2007}{}%
-52. Decker EH, Kerkhoff AJ, Moses ME. Global patterns of city size
+53. Decker EH, Kerkhoff AJ, Moses ME. Global patterns of city size
 distributions and their fundamental drivers. PLoS One. 2007;2: e934.
@@
 \leavevmode\hypertarget{ref-donner_research_2008}{}%
-53. Donner J. Research approaches to mobile use in the developing world:
+54. Donner J. Research approaches to mobile use in the developing world:
 A review of the literature. The information society. 2008;24: 140--159.
@@
 \leavevmode\hypertarget{ref-hamilton_are_2003}{}%
-54. Hamilton J. Are main lines and mobile phones substitutes or
+55. Hamilton J. Are main lines and mobile phones substitutes or
 complements? Evidence from Africa. Telecommunications Policy. 2003;27:
@@
 \leavevmode\hypertarget{ref-batty_rank_2006}{}%
-55. Batty M. Rank clocks. Nature. 2006;444: 592--596.
+56. Batty M. Rank clocks. Nature. 2006;444: 592--596.
 
 \leavevmode\hypertarget{ref-havlin_distance_1995}{}%
-56. Havlin S. The distance between Zipf plots. Physica A: Statistical
+57. Havlin S. The distance between Zipf plots. Physica A: Statistical
 Mechanics and its Applications. 1995;216: 148--150.
@@
 \leavevmode\hypertarget{ref-chetty2018impacts}{}%
-57. Chetty R, Hendren N. The impacts of neighborhoods on
+58. Chetty R, Hendren N. The impacts of neighborhoods on
 intergenerational mobility i: Childhood exposure effects. The Quarterly
@@
 \leavevmode\hypertarget{ref-dettling2017}{}%
-58. Dettling LJ. Broadband in the labor market: The impact of
+59. Dettling LJ. Broadband in the labor market: The impact of
 residential high-speed internet on married women's labor force
@@
 \leavevmode\hypertarget{ref-arntz2019working}{}%
-59. Arntz M, Sarra BY, Berlingieri F. Working from home: Heterogeneous
+60. Arntz M, Sarra BY, Berlingieri F. Working from home: Heterogeneous
 effects on hours worked and wages. ZEW-Centre for European Economic
@@
 \leavevmode\hypertarget{ref-watson2018digitalisation}{}%
-60. Watson T, Corliss M, Le M, others. Digitalisation and women's
+61. Watson T, Corliss M, Le M, others. Digitalisation and women's
 workforce participation in the indo-pacific. Australian Journal of
@@
 \leavevmode\hypertarget{ref-Lechman2020}{}%
-61. Lechman E, Popowska M. Enhancing women's engagement in economic
+62. Lechman E, Popowska M. Enhancing women's engagement in economic
 activities through information and communication technology deployment:
@@
 \leavevmode\hypertarget{ref-Valberg2020}{}%
-62. Valberg S. ICT, gender, and the labor market: A cross-country
+63. Valberg S. ICT, gender, and the labor market: A cross-country
 analysis. In: Maiti D, Castellacci F, Melchior A, editors.
@@
 \leavevmode\hypertarget{ref-angrist1995identification}{}%
-63. Angrist JD, Imbens GW. Identification and estimation of local
+64. Angrist JD, Imbens GW. Identification and estimation of local
 average treatment effects. National Bureau of Economic Research; 1995.
@@
 \leavevmode\hypertarget{ref-calabrese2011interplay}{}%
-64. Calabrese F, Smoreda Z, Blondel VD, Ratti C. Interplay between
+65. Calabrese F, Smoreda Z, Blondel VD, Ratti C. Interplay between
 telecommunications and face-to-face interactions: A study using mobile
@@
 \leavevmode\hypertarget{ref-buchel2020cities}{}%
-65. Büchel K, Ehrlich MV. Cities and the structure of social
+66. Büchel K, Ehrlich MV. Cities and the structure of social
 interactions: Evidence from mobile phone data. Journal of Urban
@@
 \leavevmode\hypertarget{ref-file_computer_2014}{}%
-66. File T, Ryan C. Computer and Internet use in the United States:
+67. File T, Ryan C. Computer and Internet use in the United States:
 2013. American Community Survey Reports. 2014;
@@
 \leavevmode\hypertarget{ref-riddlesden_broadband_2014}{}%
-67. Riddlesden D, Singleton AD. Broadband speed equity: A new digital
+68. Riddlesden D, Singleton AD. Broadband speed equity: A new digital
 divide? Applied Geography. 2014;52: 25--33.
@@
 \leavevmode\hypertarget{ref-ons2013}{}%
-68. ONS. 2011 Built-up Areas - Methodology and Guidance. Office for
+69. ONS. 2011 Built-up Areas - Methodology and Guidance. Office for
 National Statistics; 2013 pp. 1--15.
@@
 \leavevmode\hypertarget{ref-johnston_changing_2014}{}%
-69. Johnston R, Poulsen M, Forrest J. The changing ethnic composition of
+70. Johnston R, Poulsen M, Forrest J. The changing ethnic composition of
 urban neighbourhoods in England and Wales, 2001-2011: Creating nations
@@
 \leavevmode\hypertarget{ref-johnston2016ethnic}{}%
-70. Johnston R, Poulsen M, Forrest J. Ethnic residential patterns in
+71. Johnston R, Poulsen M, Forrest J. Ethnic residential patterns in
 urban england and wales, 2001--2011: A system-wide analysis. Tijdschrift
@@
 \leavevmode\hypertarget{ref-GRUBESIC201950}{}%
-71. Grubesic TH, Helderop E, Alizadeh T. Closing information
+72. Grubesic TH, Helderop E, Alizadeh T. Closing information
 asymmetries: A scale agnostic approach for exploring equity implications
@@
 \leavevmode\hypertarget{ref-tranos2020digital}{}%
-72. Tranos E, Kitsos T, Ortega-Argilés R. Digital economy in the uk:
+73. Tranos E, Kitsos T, Ortega-Argilés R. Digital economy in the uk:
 Regional productivity effects of early adoption. Regional Studies.
@@
 \leavevmode\hypertarget{ref-rozenfeld2011area}{}%
-73. Rozenfeld HD, Rybski D, Gabaix X, Makse HA. The area and population
+74. Rozenfeld HD, Rybski D, Gabaix X, Makse HA. The area and population
 of cities: New insights from a different perspective on cities. American
@@
 \leavevmode\hypertarget{ref-duben2019population}{}%
-74. Duben C, Krause M, others. Population, light, and the size
+75. Duben C, Krause M, others. Population, light, and the size
 distribution of cities. 2019.
@@
 \leavevmode\hypertarget{ref-pesaresi2019ghs}{}%
-75. Pesaresi M, Florczyk A, Schiavina M, Melchiorri M, Maffenini L. GHS
+76. Pesaresi M, Florczyk A, Schiavina M, Melchiorri M, Maffenini L. GHS
 settlement grid, updated and refined regio model 2014 in application to

### 2021-03-23 src/s2.Rmd
editing changes

@@
 
-The below maps plot the Pareto exponents for all the countries included in our data 
+The below maps plot the Pareto exponents corrected as per Gabaix and Ibragimov (2011) for all the countries included in our data 
 set for the years 2000, 2005, 2010 and 2010. The quantiles of the yearly exponent 

### 2021-03-26 paper/2_final_submission/tranos_ioannides.Rmd
edits for production

@@
 The next step in our analysis is to estimate equation \ref{eq4} for cities in England and Wales for which we were able to access internet speed micro-data.
-More specifically, we obtained individual speed internet tests from [broadbandspeedchecker.co.uk](broadbandspeedchecker.co.uk).
+More specifically, we obtained individual speed internet tests from [broadbandspeedchecker.co.uk](https://www.broadbandspeedchecker.co.uk/).
 This website enables individuals to directly measure their upload and download internet speed.
@@
 
+# Acknowledgments {.unnumbered}
+
+We are grateful to many audience members at the 2016 Annual Winter Seminar of the German speaking section of the European
+Regional Science Association, the North American Regional Science Council meetings, and Fall 2018 Opportunity and Inclusive
+Growth Conference, Federal Reserve Bank of Minneapolis, MN. We are also grateful to Max Nathan for very helpful comments, and to
+Esteban Rossi-Hansberg and anonymous reviewers, whose suggestions have not been yet fully incorporated in this version. We are also thankful 
+to [broadbandspeedchecker.co.uk](https://www.broadbandspeedchecker.co.uk/) for providing the UK internet speed data.
+Naturally, all errors are ours.
+
 # References {#references .unnumbered}

### 2021-03-26 paper/2_final_submission/tranos_ioannides.tex
edits for production

@@
 speed micro-data. More specifically, we obtained individual speed
-internet tests from \url{broadbandspeedchecker.co.uk}. This website
-enables individuals to directly measure their upload and download
-internet speed. The results of the tests as well as the geo-location of
-the users are recorded by the website operator and were provided to us
-in a fully anonymized manner. More discussion about the nature and the
-validity of this data can be found in the work of Riddlesden and
-Singleton {[}68{]}. The point nature of these individual level data
-enables us to aggregate them up to any urban level that we are
-interested in. Given that all the above analyses focused on functional
-definitions of cities, we adopt here a morphological definition of a
-city or a town for the UK, which enables us to test the effect of
-internet on smaller areas irrespective of being part of a wider urban
-agglomeration. Therefore we aggregate the internet speed data at the
-level of Built-up Areas (BUA) for England and Wales. This is a `bricks
-and mortar' approach which refers to land which is ``irreversibly urban
-in character'' including villages, towns or cities. Some key
-characteristics of these areas include: minimum size of \(20\) hectares;
-areas with less than \(200\) meters between them are linked to a single
-built-up area; larger built-up areas are separated to smaller
+internet tests from
+\href{https://www.broadbandspeedchecker.co.uk/}{broadbandspeedchecker.co.uk}.
+This website enables individuals to directly measure their upload and
+download internet speed. The results of the tests as well as the
+geo-location of the users are recorded by the website operator and were
+provided to us in a fully anonymized manner. More discussion about the
+nature and the validity of this data can be found in the work of
+Riddlesden and Singleton {[}68{]}. The point nature of these individual
+level data enables us to aggregate them up to any urban level that we
+are interested in. Given that all the above analyses focused on
+functional definitions of cities, we adopt here a morphological
+definition of a city or a town for the UK, which enables us to test the
+effect of internet on smaller areas irrespective of being part of a
+wider urban agglomeration. Therefore we aggregate the internet speed
+data at the level of Built-up Areas (BUA) for England and Wales. This is
+a `bricks and mortar' approach which refers to land which is
+``irreversibly urban in character'' including villages, towns or cities.
+Some key characteristics of these areas include: minimum size of \(20\)
+hectares; areas with less than \(200\) meters between them are linked to
+a single built-up area; larger built-up areas are separated to smaller
 sub-divisions of built-up areas {[}69--71{]}.
@@
 
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
+
+We are grateful to many audience members at the 2016 Annual Winter
+Seminar of the German speaking section of the European Regional Science
+Association, the North American Regional Science Council meetings, and
+Fall 2018 Opportunity and Inclusive Growth Conference, Federal Reserve
+Bank of Minneapolis, MN. We are also grateful to Max Nathan for very
+helpful comments, and to Esteban Rossi-Hansberg and anonymous reviewers,
+whose suggestions have not been yet fully incorporated in this version.
+We are also thankful to
+\href{https://www.broadbandspeedchecker.co.uk/}{broadbandspeedchecker.co.uk}
+for providing the UK internet speed data. Naturally, all errors are
+ours.
+
 \hypertarget{references}{%
